```{r load-workspace-if-standalone-03, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
if (!exists("expr_target")) {
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(knitr)
  library(kableExtra)
  wspath <- NULL
  if (requireNamespace("here", quietly = TRUE)) wspath <- here::here("data", "HMA_BCL2_workspace.rds")
  if (is.null(wspath) || !file.exists(wspath)) wspath <- file.path("..", "data", "HMA_BCL2_workspace.rds")
  if (file.exists(wspath)) load(wspath, envir = .GlobalEnv) else stop("Run the master or 01_Data_Loading.Rmd first to create the workspace.")
}
```

## Cohort size and cell lines available per data source

When `data/HMA_BCL2_cohort_assignments.rds` exists (created by knitting **Cell_Line_Assignment_Verification.Rmd** first), Table 1 uses the cleaned, expanded cohort from that verification report. Otherwise it uses fallback lineage assignment from DepMap metadata.

**Cohort** = DepMap cell lines with expression and lineage assignment (same set as in Figure 1). **CRISPR** = cohort cell lines with BCL2 CERES dependency data in DepMap (CERES subpanels in Figure 1); the same cohort is used, but not every cohort line has CERES data, so N (CRISPR) can be less than or equal to N (cohort). **GDSC, PRISM, CTRP** = cohort cell lines with drug sensitivity data in each platform. **Methylation** = cohort cell lines with GSE68379 (Illumina HumanMethylation450) data matched to DepMap (Figures 15–16). A cell line may appear in multiple sources.

Below: **Table 1** (cell lines per data source) is printed first, then **one combined 2×2 figure** with **BCL2** expression (top left) and **BCL2** CERES (top right), **DNMT1** expression (bottom left) and **DNMT1** CERES (bottom right) on the **same** wide lineage x-axis, strata, and fill colors as the BCL2 expression panel (DepMap). Lineage tick labels appear on **both** bottom-row panels; the top row omits x tick labels. Cell lines and `lab`/`subtype` assignments follow the **BCL2 expression** wide figure so DNMT1 panels show the same lineages in the same order with the same box colors.

```{r table-1-cell-lines-by-source, include = TRUE, cache = FALSE, echo = FALSE, results = "asis"}
# Table 1: Cell lines available per data source (merged; includes CRISPR so readers see it may differ from cohort)
# Colorectal lineages excluded from this table only (per prior reporting).
lineage_levels_t1 <- lineage_levels[!grepl("Colorectal", lineage_levels, fixed = TRUE)]
expr_target_t1 <- expr_target %>%
  dplyr::filter(!grepl("Colorectal", as.character(.data$lineage_group), fixed = TRUE)) %>%
  dplyr::mutate(lineage_group = factor(as.character(.data$lineage_group), levels = lineage_levels_t1))
t1_cohort <- tibble(lineage_group = factor(lineage_levels_t1, levels = lineage_levels_t1)) %>%
  left_join(expr_target_t1 %>% dplyr::count(lineage_group, name = "N_cohort"), by = "lineage_group") %>%
  mutate(N_cohort = replace_na(N_cohort, 0L)) %>%
  arrange(lineage_group)

# CRISPR: cohort lines with BCL2 CERES data (same definition as CERES subpanels in Figure 1)
t1_crispr <- if (exists("crispr") && exists("target_ids") && nrow(crispr) > 0 && length(target_ids) > 0) {
  crispr %>%
    filter(grepl("^BCL2\\b", gene_name), depmap_id %in% target_ids) %>%
    left_join(expr_target %>% select(depmap_id, lineage_group), by = "depmap_id") %>%
    filter(!is.na(lineage_group)) %>%
    dplyr::count(lineage_group, name = "N_CRISPR")
} else {
  expr_target %>% distinct(lineage_group) %>% mutate(N_CRISPR = 0L) %>% select(lineage_group, N_CRISPR)
}

t1_gdsc <- if (exists("merged_aza") && nrow(merged_aza) > 0 && "lineage_group" %in% names(merged_aza)) {
  merged_aza %>% filter(!is.na(lineage_group)) %>% dplyr::count(lineage_group, name = "N_GDSC")
} else { expr_target %>% distinct(lineage_group) %>% mutate(N_GDSC = 0L) %>% select(lineage_group, N_GDSC) }
t1_prism <- if (exists("pair_df") && nrow(pair_df) > 0 && "lineage_group" %in% names(pair_df)) {
  pair_df %>% filter(!is.na(lineage_group)) %>% dplyr::count(lineage_group, name = "N_PRISM")
} else { expr_target %>% distinct(lineage_group) %>% mutate(N_PRISM = 0L) %>% select(lineage_group, N_PRISM) }
t1_ctrp <- if (exists("ctrp_long") && nrow(ctrp_long) > 0) {
  ctrp_long %>% distinct(depmap_id, lineage_group) %>% dplyr::count(lineage_group, name = "N_CTRP")
} else { expr_target %>% distinct(lineage_group) %>% mutate(N_CTRP = 0L) %>% select(lineage_group, N_CTRP) }

# Methylation (GSE68379): cohort lines with global methylation data (same match as 09_DNMTi_OnTarget_Methylation)
t1_methylation <- NULL
t1_global_meth_depmap_ids <- character(0)
meth_global_path <- if (requireNamespace("here", quietly = TRUE)) here::here("data", "GSE68379_global_methylation.csv") else file.path("data", "GSE68379_global_methylation.csv")
alt_global <- file.path("..", "data", "GSE68379_global_methylation.csv")
for (f in c(meth_global_path, alt_global)) {
  if (file.exists(f) && file.size(f) > 0 && exists("meta") && exists("expr_target")) {
    global_meth_raw <- tryCatch(readr::read_csv(f, show_col_types = FALSE), error = function(e) NULL)
    if (!is.null(global_meth_raw) && nrow(global_meth_raw) > 0) {
      global_meth <- global_meth_raw %>%
        filter(!is.na(global_mean_beta), global_mean_beta >= 0, global_mean_beta <= 1) %>%
        mutate(
          cell_line_clean = gsub("_AVG\\.Beta$", "", as.character(cell_line)),
          cell_line_stem = gsub("-", "", gsub("_.*$", "", cell_line_clean))
        )
      meta_stem <- meta %>% mutate(cell_line_stem = gsub("-", "", gsub("_.*$", "", cell_line)))
      global_meth_depmap <- global_meth %>%
        inner_join(meta_stem %>% select(depmap_id, cell_line_stem), by = "cell_line_stem") %>%
        distinct(depmap_id, .keep_all = TRUE)
      t1_global_meth_depmap_ids <- unique(as.character(global_meth_depmap$depmap_id))
      meth_lineage <- expr_target_t1 %>%
        select(depmap_id, lineage_group) %>%
        inner_join(global_meth_depmap %>% select(depmap_id), by = "depmap_id") %>%
        dplyr::count(lineage_group, name = "N_methylation")
      t1_methylation <- tibble(lineage_group = factor(lineage_levels_t1, levels = lineage_levels_t1)) %>%
        left_join(meth_lineage, by = "lineage_group") %>%
        mutate(N_methylation = replace_na(N_methylation, 0L))
      break
    }
  }
}
if (is.null(t1_methylation)) {
  t1_methylation <- expr_target_t1 %>% distinct(lineage_group) %>% mutate(N_methylation = 0L) %>% select(lineage_group, N_methylation)
}

t1_body <- t1_cohort %>%
  left_join(t1_crispr, by = "lineage_group") %>%
  left_join(t1_gdsc, by = "lineage_group") %>%
  left_join(t1_prism, by = "lineage_group") %>%
  left_join(t1_ctrp, by = "lineage_group") %>%
  left_join(t1_methylation, by = "lineage_group") %>%
  mutate(across(starts_with("N_"), ~replace(., is.na(.), 0L)))

# Table 1: omit strata with "Unspecified" in the lineage label (all masters; full cohort unchanged for figures)
t1_body <- dplyr::filter(t1_body, !grepl("Unspecified", as.character(.data$lineage_group), fixed = TRUE))

t1_total <- t1_body %>% summarise(
  lineage_group = "Total",
  N_cohort = sum(N_cohort),
  N_CRISPR = sum(N_CRISPR, na.rm = TRUE),
  N_GDSC = sum(N_GDSC),
  N_PRISM = sum(N_PRISM),
  N_CTRP = sum(N_CTRP),
  N_methylation = sum(N_methylation, na.rm = TRUE)
)
t1_df <- dplyr::bind_rows(t1_body, t1_total) %>%
  mutate(Lineage = gsub(" \\(control\\)", "", as.character(lineage_group))) %>%
  select(Lineage, `N (cohort)` = N_cohort, `N (CRISPR)` = N_CRISPR, `N (GDSC)` = N_GDSC, `N (PRISM)` = N_PRISM, `N (CTRP)` = N_CTRP, `N (methylation)` = N_methylation)

t1_kable <- t1_df %>%
  kable(caption = "Table 1. Cell lines available per data source. Cohort = DepMap expression/lineage (Figure 1). CRISPR = cohort lines with BCL2 CERES data (CERES subpanels in Figure 1). GDSC, PRISM, CTRP = cohort lines with drug sensitivity in each platform. Methylation = cohort lines with GSE68379 (Illumina 450k) data matched to DepMap. A line may appear in multiple sources.") %>%
  kable_styling(bootstrap_options = c("bordered", "condensed"), full_width = FALSE)
cat("\n\n")
print(t1_kable)
cat("\n\n")
```

```{r bcl2i-table-42, echo = FALSE, results = "asis"}
# Table 2: BCL2 inhibitors and DNMTi in the dataset — types, mechanisms, and N (cohort cell lines) in GDSC, PRISM, CTRP
# 4.1 BCL2 inhibitors
bcl2i_simple <- tibble::tibble(
  Compound = c("Venetoclax (ABT-199, GDC-0199)", "Navitoclax (ABT-263)", "Obatoclax", "ABT-737"),
  `Target / mechanism` = c(
    "BCL2-selective; FDA-approved (AML, CLL)",
    "BCL2 + BCL-xL; thrombocytopenia risk",
    "Pan-BH3 mimetic (MCL1, BCL2, BCL-xL)",
    "BCL2 + BCL-xL; first-in-class BH3 mimetic"
  ),
  `N (GDSC)` = NA_integer_,
  `N (PRISM)` = NA_integer_,
  `N (CTRP)` = NA_integer_
)
cohort_ids <- if (exists("target_ids")) target_ids else character(0)
# PRISM: cohort cell lines with that compound
if (exists("drug_sens") && exists("bcl2i_prism") && nrow(bcl2i_prism) > 0 && length(cohort_ids) > 0) {
  n_prism <- drug_sens %>% filter(depmap_id %in% cohort_ids, name %in% bcl2i_prism$name) %>%
    group_by(name) %>% summarise(n = n_distinct(depmap_id), .groups = "drop")
  bcl2i_simple <- bcl2i_simple %>%
    mutate(
      `N (PRISM)` = c(
        sum(n_prism$n[grepl("venetoclax|ABT-199|GDC-0199", n_prism$name, ignore.case = TRUE)], na.rm = TRUE),
        sum(n_prism$n[grepl("navitoclax|ABT-263", n_prism$name, ignore.case = TRUE)], na.rm = TRUE),
        sum(n_prism$n[grepl("obatoclax", n_prism$name, ignore.case = TRUE)], na.rm = TRUE),
        sum(n_prism$n[grepl("ABT-737", n_prism$name, ignore.case = TRUE)], na.rm = TRUE)
      )
    )
}
# GDSC: cohort cell lines with that BCL2i (from gdsc_raw if available)
if (exists("gdsc_raw") && nrow(gdsc_raw) > 0 && exists("meta") && length(cohort_ids) > 0) {
  nm <- names(gdsc_raw)
  col_drug <- nm[tolower(nm) == "drug_name"][1]; if (length(col_drug) == 0) col_drug <- "DRUG_NAME"
  col_cell <- nm[tolower(nm) == "cell_line_name"][1]; if (length(col_cell) == 0) col_cell <- "CELL_LINE_NAME"
  gdsc_drugs <- unique(gdsc_raw[[col_drug]])
  bcl2i_gdsc_names <- gdsc_drugs[grepl("venetoclax|ABT-199|GDC-0199|navitoclax|ABT-263|obatoclax|ABT-737", gdsc_drugs, ignore.case = TRUE)]
  if (length(bcl2i_gdsc_names) > 0) {
    meta_stem <- meta %>% mutate(cell_line_stem = gsub("-", "", gsub("_.*$", "", cell_line)))
    gdsc_bcl2i <- gdsc_raw %>%
      filter(.data[[col_drug]] %in% bcl2i_gdsc_names) %>%
      mutate(cell_line_stem = gsub("_.*$", "", gsub("-", "", as.character(.data[[col_cell]]))))
    gdsc_bcl2i <- gdsc_bcl2i %>%
      inner_join(meta_stem %>% select(depmap_id, cell_line_stem), by = "cell_line_stem") %>%
      filter(depmap_id %in% cohort_ids)
    if (nrow(gdsc_bcl2i) > 0) {
      n_gdsc <- gdsc_bcl2i %>%
        mutate(drug_name = as.character(.data[[col_drug]])) %>%
        group_by(drug_name) %>% summarise(n = n_distinct(depmap_id), .groups = "drop")
      bcl2i_simple <- bcl2i_simple %>%
        mutate(
          `N (GDSC)` = c(
            sum(n_gdsc$n[grepl("venetoclax|ABT-199|GDC-0199", n_gdsc$drug_name, ignore.case = TRUE)], na.rm = TRUE),
            sum(n_gdsc$n[grepl("navitoclax|ABT-263", n_gdsc$drug_name, ignore.case = TRUE)], na.rm = TRUE),
            sum(n_gdsc$n[grepl("obatoclax", n_gdsc$drug_name, ignore.case = TRUE)], na.rm = TRUE),
            sum(n_gdsc$n[grepl("ABT-737", n_gdsc$drug_name, ignore.case = TRUE)], na.rm = TRUE)
          )
        )
    }
  }
}
# CTRP: cohort cell lines with any BCL2 inhibitor (CTRP is summarized by drug_class)
if (exists("ctrp_long") && nrow(ctrp_long) > 0 && length(cohort_ids) > 0) {
  n_ctrp_bcl2i <- ctrp_long %>%
    filter(drug_class == "BCL2 inhibitor", depmap_id %in% cohort_ids) %>%
    distinct(depmap_id) %>% nrow()
  bcl2i_simple <- bcl2i_simple %>% mutate(`N (CTRP)` = n_ctrp_bcl2i)
}
# 4.2 DNMTi (hypomethylating agents / HMA): DNMT1 inhibitors in the dataset
dnmti_simple <- tibble::tibble(
  Compound = c("Decitabine (5-aza-2'-deoxycytidine)", "Azacitidine (5-azacytidine)"),
  `Target / mechanism` = c(
    "DNMT1 inhibitor; cytidine analog (DNA incorporation)",
    "DNMT1 inhibitor; cytidine analog (RNA/DNA incorporation)"
  ),
  `N (GDSC)` = NA_integer_,
  `N (PRISM)` = NA_integer_,
  `N (CTRP)` = NA_integer_
)
cohort_ids_tbl <- if (exists("target_ids")) target_ids else character(0)
if (exists("drug_sens") && nrow(drug_sens) > 0 && length(cohort_ids_tbl) > 0) {
  aza_names <- unique(drug_sens$name)[grepl("azacitidine|decitabine|5-aza|dacogen", unique(drug_sens$name), ignore.case = TRUE)]
  if (length(aza_names) > 0) {
    n_prism_aza <- drug_sens %>% filter(depmap_id %in% cohort_ids_tbl, name %in% aza_names) %>%
      group_by(name) %>% summarise(n = n_distinct(depmap_id), .groups = "drop")
    dnmti_simple <- dnmti_simple %>%
      mutate(
        `N (PRISM)` = c(
          sum(n_prism_aza$n[grepl("decitabine|dacogen|5-aza-2", n_prism_aza$name, ignore.case = TRUE)], na.rm = TRUE),
          sum(n_prism_aza$n[grepl("azacitidine|5-azacytidine|5-aza[^2]|vidaza", n_prism_aza$name, ignore.case = TRUE)], na.rm = TRUE)
        )
      )
  }
}
if (exists("gdsc_raw") && nrow(gdsc_raw) > 0 && length(cohort_ids_tbl) > 0 && exists("meta")) {
  nm <- names(gdsc_raw)
  col_drug <- nm[tolower(nm) == "drug_name"][1]; if (length(col_drug) == 0) col_drug <- "DRUG_NAME"
  col_cell <- nm[tolower(nm) == "cell_line_name"][1]; if (length(col_cell) == 0) col_cell <- "CELL_LINE_NAME"
  gdsc_drugs <- unique(gdsc_raw[[col_drug]])
  aza_gdsc <- gdsc_drugs[grepl("azacitidine|decitabine|5-aza|dacogen|vidaza", gdsc_drugs, ignore.case = TRUE)]
  if (length(aza_gdsc) > 0) {
    meta_stem <- meta %>% mutate(cell_line_stem = gsub("-", "", gsub("_.*$", "", cell_line)))
    gdsc_aza <- gdsc_raw %>% filter(.data[[col_drug]] %in% aza_gdsc) %>%
      mutate(cell_line_stem = gsub("_.*$", "", gsub("-", "", as.character(.data[[col_cell]]))))
    gdsc_aza <- gdsc_aza %>% inner_join(meta_stem %>% select(depmap_id, cell_line_stem), by = "cell_line_stem") %>%
      filter(depmap_id %in% cohort_ids_tbl)
    if (nrow(gdsc_aza) > 0) {
      n_gdsc_aza <- gdsc_aza %>% mutate(drug_name = as.character(.data[[col_drug]])) %>%
        group_by(drug_name) %>% summarise(n = n_distinct(depmap_id), .groups = "drop")
      dnmti_simple <- dnmti_simple %>%
        mutate(
          `N (GDSC)` = c(
            sum(n_gdsc_aza$n[grepl("decitabine|dacogen|5-aza-2", n_gdsc_aza$drug_name, ignore.case = TRUE)], na.rm = TRUE),
            sum(n_gdsc_aza$n[grepl("azacitidine|5-azacytidine|vidaza", n_gdsc_aza$drug_name, ignore.case = TRUE)], na.rm = TRUE)
          )
        )
    }
  }
}
if (exists("ctrp_long") && nrow(ctrp_long) > 0 && length(cohort_ids_tbl) > 0) {
  n_ctrp_aza <- ctrp_long %>%
    filter(drug_class == "HMA", depmap_id %in% cohort_ids_tbl) %>%
    distinct(depmap_id) %>% nrow()
  if (n_ctrp_aza > 0) dnmti_simple <- dnmti_simple %>% mutate(`N (CTRP)` = n_ctrp_aza)
}
# 4.3 EZH2 inhibitors (Supp table 2)
ezh2i_simple <- tibble::tibble(
  Compound = c("Tazemetostat (EPZ-6438)"),
  `Target / mechanism` = c("EZH2 (PRC2) inhibitor; H3K27me3; FDA-approved R/R follicular lymphoma, epithelioid sarcoma"),
  `N (GDSC)` = NA_integer_,
  `N (PRISM)` = NA_integer_,
  `N (CTRP)` = NA_integer_
)
# N (GDSC) for EZH2i: use merged_ezh2i (built in 01_Data_Loading) when available so table matches actual analysis cohort
if (exists("merged_ezh2i") && is.data.frame(merged_ezh2i) && nrow(merged_ezh2i) > 0 && "depmap_id" %in% names(merged_ezh2i)) {
  ezh2i_simple <- ezh2i_simple %>% mutate(`N (GDSC)` = as.integer(n_distinct(merged_ezh2i$depmap_id)))
} else if (exists("gdsc_raw") && nrow(gdsc_raw) > 0 && exists("meta") && length(cohort_ids_tbl) > 0) {
  nm <- names(gdsc_raw)
  col_drug <- nm[tolower(nm) == "drug_name"][1]; if (length(col_drug) == 0) col_drug <- "DRUG_NAME"
  col_cell <- nm[tolower(nm) == "cell_line_name"][1]; if (length(col_cell) == 0) col_cell <- "CELL_LINE_NAME"
  gdsc_drugs <- unique(gdsc_raw[[col_drug]])
  ezh2_gdsc_names <- gdsc_drugs[grepl("tazemetostat|EPZ-6438|EZH2|GSK126|GSK343|valemetostat", gdsc_drugs, ignore.case = TRUE)]
  if (length(ezh2_gdsc_names) > 0) {
    meta_stem <- meta %>% mutate(cell_line_stem = gsub("-", "", gsub("_.*$", "", cell_line)))
    gdsc_ezh2 <- gdsc_raw %>%
      filter(.data[[col_drug]] %in% ezh2_gdsc_names) %>%
      mutate(cell_line_stem = gsub("_.*$", "", gsub("-", "", as.character(.data[[col_cell]]))))
    gdsc_ezh2 <- gdsc_ezh2 %>%
      inner_join(meta_stem %>% select(depmap_id, cell_line_stem), by = "cell_line_stem") %>%
      filter(depmap_id %in% cohort_ids_tbl)
    if (nrow(gdsc_ezh2) > 0) ezh2i_simple <- ezh2i_simple %>% mutate(`N (GDSC)` = as.integer(n_distinct(gdsc_ezh2$depmap_id)))
  }
}
# 4.4 HER2-directed small molecules (same strict agent list as ERBB2 / locus-triple GDSC merge)
her2i_simple <- tibble::tibble(
  Compound = c(
    "Lapatinib (Tykerb)", "Tucatinib (Tukysa)", "Afatinib (Gilotrif)", "Neratinib (Nerlynx)",
    "Pyrotinib", "Poziotinib", "Sapitinib"
  ),
  `Target / mechanism` = c(
    "Reversible ATP-competitive HER2/EGFR tyrosine kinase inhibitor",
    "Reversible HER2-selective tyrosine kinase inhibitor",
    "Irreversible pan-ErbB (HER1/HER2/HER4) tyrosine kinase inhibitor",
    "Irreversible pan-ErbB tyrosine kinase inhibitor",
    "Irreversible pan-ErbB tyrosine kinase inhibitor",
    "Irreversible pan-ErbB tyrosine kinase inhibitor",
    "Dual SRC and HER2 inhibitor"
  ),
  `N (GDSC)` = NA_integer_,
  `N (PRISM)` = NA_integer_,
  `N (CTRP)` = 0L
)
if (exists("drug_sens") && nrow(drug_sens) > 0 && length(cohort_ids_tbl) > 0) {
  h2_prism_pat <- "lapatinib|tucatinib|afatinib|neratinib|pyrotinib|poziotinib|sapitinib"
  n_prism_h2 <- drug_sens %>%
    filter(depmap_id %in% cohort_ids_tbl, grepl(h2_prism_pat, name, ignore.case = TRUE)) %>%
    group_by(name) %>%
    summarise(n = n_distinct(depmap_id), .groups = "drop")
  if (nrow(n_prism_h2) > 0) {
    her2i_simple <- her2i_simple %>%
      mutate(
        `N (PRISM)` = c(
          sum(n_prism_h2$n[grepl("lapatinib", n_prism_h2$name, ignore.case = TRUE)], na.rm = TRUE),
          sum(n_prism_h2$n[grepl("tucatinib", n_prism_h2$name, ignore.case = TRUE)], na.rm = TRUE),
          sum(n_prism_h2$n[grepl("afatinib", n_prism_h2$name, ignore.case = TRUE)], na.rm = TRUE),
          sum(n_prism_h2$n[grepl("neratinib", n_prism_h2$name, ignore.case = TRUE)], na.rm = TRUE),
          sum(n_prism_h2$n[grepl("pyrotinib", n_prism_h2$name, ignore.case = TRUE)], na.rm = TRUE),
          sum(n_prism_h2$n[grepl("poziotinib", n_prism_h2$name, ignore.case = TRUE)], na.rm = TRUE),
          sum(n_prism_h2$n[grepl("sapitinib", n_prism_h2$name, ignore.case = TRUE)], na.rm = TRUE)
        )
      )
  }
}
if (exists("gdsc_raw") && nrow(gdsc_raw) > 0 && exists("meta") && length(cohort_ids_tbl) > 0) {
  nm <- names(gdsc_raw)
  col_drug <- nm[tolower(nm) == "drug_name"][1]; if (length(col_drug) == 0) col_drug <- "DRUG_NAME"
  col_cell <- nm[tolower(nm) == "cell_line_name"][1]; if (length(col_cell) == 0) col_cell <- "CELL_LINE_NAME"
  gdsc_drugs <- unique(gdsc_raw[[col_drug]])
  h2_gdsc_names <- gdsc_drugs[grepl("lapatinib|tucatinib|afatinib|neratinib|pyrotinib|poziotinib|sapitinib", gdsc_drugs, ignore.case = TRUE)]
  if (length(h2_gdsc_names) > 0) {
    meta_stem <- meta %>% mutate(cell_line_stem = gsub("-", "", gsub("_.*$", "", cell_line)))
    gdsc_h2 <- gdsc_raw %>%
      filter(.data[[col_drug]] %in% h2_gdsc_names) %>%
      mutate(cell_line_stem = gsub("_.*$", "", gsub("-", "", as.character(.data[[col_cell]]))))
    gdsc_h2 <- gdsc_h2 %>%
      inner_join(meta_stem %>% select(depmap_id, cell_line_stem), by = "cell_line_stem") %>%
      filter(depmap_id %in% cohort_ids_tbl)
    if (nrow(gdsc_h2) > 0) {
      n_gdsc_h2 <- gdsc_h2 %>%
        mutate(drug_name = as.character(.data[[col_drug]])) %>%
        group_by(drug_name) %>%
        summarise(n = n_distinct(depmap_id), .groups = "drop")
      her2i_simple <- her2i_simple %>%
        mutate(
          `N (GDSC)` = c(
            sum(n_gdsc_h2$n[grepl("lapatinib", n_gdsc_h2$drug_name, ignore.case = TRUE)], na.rm = TRUE),
            sum(n_gdsc_h2$n[grepl("tucatinib", n_gdsc_h2$drug_name, ignore.case = TRUE)], na.rm = TRUE),
            sum(n_gdsc_h2$n[grepl("afatinib", n_gdsc_h2$drug_name, ignore.case = TRUE)], na.rm = TRUE),
            sum(n_gdsc_h2$n[grepl("neratinib", n_gdsc_h2$drug_name, ignore.case = TRUE)], na.rm = TRUE),
            sum(n_gdsc_h2$n[grepl("pyrotinib", n_gdsc_h2$drug_name, ignore.case = TRUE)], na.rm = TRUE),
            sum(n_gdsc_h2$n[grepl("poziotinib", n_gdsc_h2$drug_name, ignore.case = TRUE)], na.rm = TRUE),
            sum(n_gdsc_h2$n[grepl("sapitinib", n_gdsc_h2$drug_name, ignore.case = TRUE)], na.rm = TRUE)
          )
        )
    }
  }
}
# NA counts displayed as 0; drop compounds with no cohort coverage in any platform (0 / NA / 0 → 0,0,0 → omit)
.table2_coalesce_drop_all_zero <- function(df) {
  if (nrow(df) == 0L) return(df)
  df %>%
    mutate(
      `N (GDSC)` = as.integer(dplyr::coalesce(`N (GDSC)`, 0L)),
      `N (PRISM)` = as.integer(dplyr::coalesce(`N (PRISM)`, 0L)),
      `N (CTRP)` = as.integer(dplyr::coalesce(`N (CTRP)`, 0L))
    ) %>%
    filter(`N (GDSC)` > 0L | `N (PRISM)` > 0L | `N (CTRP)` > 0L)
}
bcl2i_simple <- .table2_coalesce_drop_all_zero(bcl2i_simple)
dnmti_simple <- .table2_coalesce_drop_all_zero(dnmti_simple)
ezh2i_simple <- .table2_coalesce_drop_all_zero(ezh2i_simple)
her2i_simple <- .table2_coalesce_drop_all_zero(her2i_simple)
t2_kable <- dplyr::bind_rows(
  bcl2i_simple %>% mutate(Section = "4.1 BCL2 inhibitors"),
  dnmti_simple %>% mutate(Section = "4.2 DNMTi (HMA)"),
  ezh2i_simple %>% mutate(Section = "4.3 EZH2 inhibitors"),
  her2i_simple %>% mutate(Section = "4.4 HER2-directed (small molecules)")
) %>%
  select(Section, Compound, `Target / mechanism`, `N (GDSC)`, `N (PRISM)`, `N (CTRP)`) %>%
  knitr::kable(caption = "Table 2 (Supp table 2). B-cell lymphoma 2 inhibitors (BCL2i; 4.1), DNA methyltransferase inhibitors (DNMTi; 4.2), enhancer of zeste homolog 2 inhibitors (EZH2i; 4.3), and HER2-directed small molecules (4.4) in the dataset: compound, target/mechanism, and number of cohort cell lines tested in Genomics of Drug Sensitivity in Cancer (GDSC), Profiling Relative Inhibition Simultaneously in Mixtures (PRISM), and Cancer Therapeutics Response Portal (CTRP). Missing platform counts are shown as 0. Compounds with zero cell lines in all three sources are omitted (e.g. agents not in the PRISM subset and absent from GDSC for this cohort). DNMTi = hypomethylating agents (e.g. decitabine, azacitidine). EZH2i = EZH2 (PRC2) inhibitors (e.g. tazemetostat). CTRP extracts in this pipeline include HMA and BCL2i only, so N (CTRP) for HER2i is 0. HER2 agents align with the strict GDSC list used for ERBB2 concordance (lapatinib, tucatinib, afatinib, neratinib, pyrotinib, poziotinib, sapitinib).") %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered", "condensed"), full_width = FALSE)
if (!exists("hma_bcl2_tnbc_lymph_master") || !isTRUE(hma_bcl2_tnbc_lymph_master)) {
  print(t2_kable)
}
```

### BCL2 expression and BH3-related mitochondrial proteins composition is similar in AML and some stem-like solid tumor lineages

In Figure 1, **each point is one cell line**; the x-axis shows lineage with N per group (e.g. ExPulm-SCNEC N=5). The box summarizes the distribution; overlaid points show the actual sample size. **Red dots** = NEPC (ExPulm-SCNEC from prostate); other points are gray. The same red/gray scheme is used for ExPulm-SCNEC on **expression** panels (BCL2 and DNMT1 RNA). **Wide CERES** panels include only cell lines with non-missing dependency scores (same x-axis order as expression; lines without CERES are omitted from those two panels). CERES tick **N** counts plotted scores per stratum (may differ from expression **N**).

```{r fig1-tnbc-lymph-note, echo = FALSE, results = "asis", eval = exists("cohort_extended_hem") && isTRUE(cohort_extended_hem) && !(exists("hma_bcl2_tnbc_lymph_master") && isTRUE(hma_bcl2_tnbc_lymph_master))}
# Main HMA_BCL2 Master: this chunk is skipped (TNBC folded into Breast; no lymphoma). Extended master only.
tnbc_note_txt <- if (exists("hma_bcl2_tnbc_lymph_master") && isTRUE(hma_bcl2_tnbc_lymph_master)) {
  "**TNBC** (localized and metastatic; **Table 1** omits any lineage with **Unspecified** in the name; unspecified TNBC is also omitted from the cohort summary below) and **lymphoma** strata "
} else {
  "**TNBC** (Localized / Metastatic / Unspecified TNBC) and **lymphoma** strata "
}
knitr::asis_output(paste(
  tnbc_note_txt,
  "(**DLBCL/MCL/CLL** vs **Lymphoma (other)** (comparator lymphoid bucket; not the DLBCL/MCL/CLL stratum); histology-based proxies, ",
  "not DepMap treatment flags) appear on the same axis when present in the extended cohort."
))
```

```{r tnbc-lymph-cohort-summary, echo = FALSE, results = "asis", eval = exists("cohort_extended_hem") && isTRUE(cohort_extended_hem) && !(exists("hma_bcl2_tnbc_lymph_master") && isTRUE(hma_bcl2_tnbc_lymph_master))}
lg_tnbc_summ <- if (exists("lineage_levels_tnbc")) {
  if (exists("hma_bcl2_tnbc_lymph_master") && isTRUE(hma_bcl2_tnbc_lymph_master)) {
    setdiff(lineage_levels_tnbc, "Unspecified TNBC")
  } else {
    lineage_levels_tnbc
  }
} else {
  c("Localized TNBC", "Metastatic TNBC", "Unspecified TNBC")
}
lg_lymph_summ <- c("DLBCL/MCL/CLL", "Lymphoma (other)")
if (exists("expr_target") && nrow(expr_target) > 0L && "lineage_group" %in% names(expr_target)) {
  summ_tl <- expr_target %>%
    dplyr::filter(as.character(.data$lineage_group) %in% c(lg_tnbc_summ, lg_lymph_summ)) %>%
    dplyr::count(.data$lineage_group, name = "N_cell_lines") %>%
    dplyr::arrange(.data$lineage_group)
  if (nrow(summ_tl) > 0L) {
    knitr::kable(
      summ_tl,
      caption = "Cohort: TNBC and lymphoma strata (N cell lines in analysis cohort, same as Table 1 / Figures 1–2). Lymphoma buckets are histology-based proxies for BCL2i-associated types vs other lymphoid malignancies in DepMap text (not treatment exposure)."
    ) %>%
      kableExtra::kable_styling(bootstrap_options = c("bordered", "condensed"), full_width = FALSE)
  } else {
    knitr::asis_output("*No TNBC or lymphoma lines in the extended cohort (check 01_Data_Loading_TNBC_Lymphoma / verification RDS).*")
  }
} else {
  knitr::asis_output("*expr_target not available for TNBC/lymphoma summary.*")
}
```

```{r bcl2-distributions, include = FALSE, cache = FALSE, fig.width = 10, fig.height = 5.5, fig.cap = "Figure 1. BCL2 expression by lineage (single x-axis). N per lineage on x-axis. Red points = NEPC lines only (ExPulm-SCNEC from prostate); other lines are not overlaid."}
# Figure 1/2 composites: axis titles, tick labels, and rotated lineage names (ggplot pt size; +4 pt vs prior default)
.f1_fig12_text_pt <- 24
.f1_lineage_axis_x <- function(angle) ggplot2::element_text(angle = angle, hjust = 1, size = .f1_fig12_text_pt)
.f1_fig12_axis_theme <- function(angle_x = 45) {
  ggplot2::theme(
    text = ggplot2::element_text(size = .f1_fig12_text_pt),
    axis.text.x = .f1_lineage_axis_x(angle_x),
    axis.text.y = ggplot2::element_text(size = .f1_fig12_text_pt),
    axis.title = ggplot2::element_text(size = .f1_fig12_text_pt),
    plot.title = ggplot2::element_text(size = .f1_fig12_text_pt),
    legend.text = ggplot2::element_text(size = .f1_fig12_text_pt),
    strip.text = ggplot2::element_text(size = .f1_fig12_text_pt)
  )
}
# annotate(..., size) uses mm; slightly larger than strict pt→mm so bracket p-values read like axis text.
.f1_fig12_bracket_label_mm <- .f1_fig12_text_pt * (25.4 / 72.27) * 1.28
# Top plot.margin (pt) when ggsignif-style brackets run; keep patchwork row themes in sync with two-tier value.
.f1_fig12_bracket_margin_top_2tier <- 120L
.f1_fig12_bracket_margin_top_1tier <- 58L
# Wide-axis: shorten metastatic (m*) and pooled localized (cl*) tick labels (e.g. "Colorectal — localized" -> "clColorectal", "mBreast").
.f1_axis_tok_tissue <- function(w) {
  w <- trimws(as.character(w))
  if (!nzchar(w)) {
    return("")
  }
  if (grepl("^[A-Z0-9 /-]+$", w) && nchar(w) <= 12L) {
    return(w)
  }
  paste0(toupper(substr(w, 1L, 1L)), tolower(substring(w, 2L)))
}
.f1_axis_short_met_label <- function(x) {
  if (length(x) < 1L) {
    return(x)
  }
  vapply(x, function(s0) {
    s <- as.character(s0)
    if (!nzchar(s)) {
      return(s0)
    }
    suf <- ""
    base <- s
    if (grepl(" N=", s, fixed = TRUE)) {
      parts <- strsplit(s, " N=", fixed = TRUE)[[1L]]
      base <- parts[[1L]]
      suf <- paste0(" N=", paste(parts[-1L], collapse = " N="))
    }
    out <- if (grepl(" — metastatic$", base)) {
      tissue <- sub(" — metastatic$", "", base)
      paste0("m", .f1_axis_tok_tissue(tissue))
    } else if (grepl("^Metastatic\\s+", base, ignore.case = TRUE)) {
      rest <- trimws(sub("^Metastatic\\s+", "", base, ignore.case = TRUE))
      paste0("m", .f1_axis_tok_tissue(rest))
    } else if (grepl(" — localized$", base)) {
      tissue <- sub(" — localized$", "", base)
      paste0("cl", .f1_axis_tok_tissue(tissue))
    } else if (grepl("^Localized\\s+", base, ignore.case = TRUE)) {
      rest <- trimws(sub("^Localized\\s+", "", base, ignore.case = TRUE))
      paste0("cl", .f1_axis_tok_tissue(rest))
    } else {
      base
    }
    paste0(out, suf)
  }, character(1L), USE.NAMES = FALSE)
}
# Wide Figure 1 CERES/DNMT1 CERES: tick labels use counts of plotted (finite) dependency rows per `lab`; NEPC with N=0 is blank (no false cohort N).
.f1_ceres_wide_x_labeller <- function(n_plotted_by_lab) {
  function(x) {
    vapply(x, function(s0) {
      s <- as.character(s0)
      if (!nzchar(s)) {
        return("")
      }
      parts <- strsplit(s, " N=", fixed = TRUE)[[1L]]
      base <- parts[[1L]]
      nc <- unname(n_plotted_by_lab[s])
      if (length(nc) != 1L || is.na(nc)) {
        nc <- 0L
      }
      nc <- as.integer(nc)
      if (nc == 0L && grepl("^NEPC\\b", base)) {
        return("")
      }
      s_new <- paste0(base, " N=", nc)
      .f1_axis_short_met_label(s_new)
    }, character(1L), USE.NAMES = FALSE)
  }
}
# #region agent log
.agent_dbg_ndjson <- function(runId, hypothesisId, location, message, data = list()) {
  if (!requireNamespace("jsonlite", quietly = TRUE)) {
    return(invisible(NULL))
  }
  payload <- list(
    sessionId = "75d08b",
    runId = runId,
    hypothesisId = hypothesisId,
    location = location,
    message = message,
    data = data,
    timestamp = as.integer(as.numeric(Sys.time()) * 1000)
  )
  tryCatch(
    cat(jsonlite::toJSON(payload, auto_unbox = TRUE, null = "null"), "\n", sep = "", file = "/Users/Alex/R/.cursor/debug-75d08b.log", append = TRUE),
    error = function(e) invisible(NULL)
  )
}
# #endregion
# Dashed comparator rule on wide lineage plots (bright red, thick; NEPC point overlay stays #E41A1C)
f1_rule_line_colour <- "#FF2222"
f1_rule_line_linewidth <- 1.4
# Comparison groups for BCL2; subtype = lineage_group (Localized/Metastatic pairs ordered adjacent in plot).
cohort_ext_hem <- exists("cohort_extended_hem") && isTRUE(cohort_extended_hem)
cohort_n <- expr_target %>% dplyr::count(lineage_group, name = "n") %>%
  mutate(subtype = as.character(lineage_group))
# Lung: use expr_target so we get SCLC, Localized NSCLC, Metastatic NSCLC (not misclassified NSCLC as SCLC)
bcl2_lung <- expr_meta %>% filter(lineage_lower == "lung", !is.na(BCL2)) %>%
  inner_join(expr_target %>% filter(lineage_group %in% c("SCLC", "Localized NSCLC", "Metastatic NSCLC")) %>% select(depmap_id, lineage_group), by = "depmap_id") %>%
  mutate(subtype = as.character(lineage_group), tissue = "Lung")
bcl2_breast <- expr_target %>% filter(grepl("Breast", as.character(lineage_group)), !is.na(BCL2)) %>%
  mutate(subtype = as.character(lineage_group), tissue = "Breast")
bcl2_prostate <- expr_target %>% filter(lineage_group %in% c("Localized prostate", "Metastatic prostate"), !is.na(BCL2)) %>%
  mutate(subtype = as.character(lineage_group), tissue = "Prostate")
bcl2_aml <- expr_meta %>% filter(depmap_id %in% aml_lines, !is.na(BCL2)) %>%
  distinct(depmap_id, .keep_all = TRUE) %>%
  mutate(subtype = "AML (control)")
bcl2_urothelial <- expr_target %>% filter(grepl("Urothelial", as.character(lineage_group)), !is.na(BCL2)) %>%
  mutate(subtype = as.character(lineage_group), tissue = "Urothelial")
bcl2_renal <- expr_target %>% filter(grepl("Renal cell", as.character(lineage_group), fixed = TRUE), !is.na(BCL2)) %>%
  mutate(subtype = as.character(lineage_group), tissue = "Renal cell")
bcl2_colorectal <- expr_target %>% filter(grepl("Colorectal", as.character(lineage_group), fixed = TRUE), !is.na(BCL2)) %>%
  mutate(subtype = as.character(lineage_group), tissue = "Colorectal")
bcl2_melanoma <- expr_target %>% filter(grepl("Melanoma", as.character(lineage_group), fixed = TRUE), !is.na(BCL2)) %>%
  mutate(subtype = as.character(lineage_group), tissue = "Melanoma")
bcl2_tnbc <- expr_target %>%
  dplyr::filter(
    cohort_ext_hem,
    grepl("TNBC", as.character(.data$lineage_group), fixed = TRUE),
    !is.na(.data$BCL2)
  ) %>%
  dplyr::mutate(subtype = as.character(.data$lineage_group), tissue = "TNBC")
bcl2_lymph <- expr_target %>%
  dplyr::filter(
    as.character(.data$lineage_group) %in% c(
      "DLBCL/MCL/CLL", "Lymphoma (other)"
    ),
    !is.na(.data$BCL2)
  ) %>%
  dplyr::mutate(subtype = as.character(.data$lineage_group), tissue = "Lymphoma")
nb_lineages_f1 <- c("Neuroblastoma", "Localized Neuroblastoma", "Metastatic Neuroblastoma", "Unspecified Neuroblastoma")
bcl2_neuro <- expr_target %>%
  dplyr::filter(as.character(.data$lineage_group) %in% nb_lineages_f1, !is.na(.data$BCL2)) %>%
  dplyr::mutate(subtype = as.character(.data$lineage_group), tissue = "Neuroblastoma", is_nepc = FALSE)
# #region agent log
tryCatch({
  nb_src <- expr_target %>%
    dplyr::filter(grepl("Neuroblastoma", as.character(.data$lineage_group), ignore.case = TRUE))
  .agent_dbg_ndjson(
    runId = "iter2",
    hypothesisId = "H6",
    location = "03_Rmd:bcl2_neuro_source",
    message = "Neuroblastoma rows in expr_target and BCL2 availability",
    data = list(
      nb_src_n = nrow(nb_src),
      nb_src_lineages = unique(as.character(nb_src$lineage_group)),
      nb_src_bcl2_non_na = sum(!is.na(nb_src$BCL2)),
      nb_in_nb_lineages_f1 = sum(as.character(nb_src$lineage_group) %in% nb_lineages_f1),
      bcl2_neuro_n = nrow(bcl2_neuro)
    )
  )
}, error = function(e) invisible(NULL))
# #endregion
# DepMap soft_tissue lines + PNS non-NBL (both labeled Sarcoma strata in 01)
bcl2_sarcoma <- expr_target %>%
  dplyr::filter(grepl("Sarcoma", as.character(.data$lineage_group), fixed = TRUE), !is.na(.data$BCL2)) %>%
  dplyr::mutate(subtype = as.character(.data$lineage_group), tissue = "Sarcoma", is_nepc = FALSE)
# ExPulm-SCNEC: use expr_target so we plot exactly the cohort that defines n; red dot = NEPC (prostate), others gray
bcl2_expulm_ne <- expr_target %>%
  filter(lineage_group == "ExPulm-SCNEC") %>%
  select(depmap_id, dplyr::any_of("lineage_expulm_detail")) %>%
  left_join(expr_meta %>% select(depmap_id, BCL2), by = "depmap_id") %>%
  filter(!is.na(BCL2)) %>%
  distinct(depmap_id, .keep_all = TRUE) %>%
  mutate(
    subtype = "ExPulm-SCNEC",
    tissue = "ExPulm-SCNEC",
    is_nepc = (dplyr::coalesce(lineage_expulm_detail, "") == "NEPC")
  ) %>%
  dplyr::filter(.data$is_nepc %in% TRUE) %>%
  select(-dplyr::any_of("lineage_expulm_detail"))
bcl2_plot_df <- bind_rows(
  bcl2_lung %>% mutate(is_nepc = FALSE),
  bcl2_breast %>% mutate(tissue = "Breast", is_nepc = FALSE),
  bcl2_prostate %>% mutate(tissue = "Prostate", is_nepc = FALSE),
  bcl2_aml %>% mutate(tissue = "AML", is_nepc = FALSE),
  bcl2_urothelial %>% mutate(is_nepc = FALSE),
  bcl2_renal %>% mutate(is_nepc = FALSE),
  bcl2_colorectal %>% mutate(is_nepc = FALSE),
  bcl2_melanoma %>% mutate(is_nepc = FALSE),
  bcl2_tnbc %>% dplyr::mutate(is_nepc = FALSE),
  bcl2_lymph %>% dplyr::mutate(is_nepc = FALSE),
  bcl2_neuro,
  bcl2_sarcoma,
  bcl2_expulm_ne
) %>% dplyr::filter(!is.na(.data$subtype)) %>%
  dplyr::mutate(is_nepc = dplyr::coalesce(.data$is_nepc, FALSE)) %>%
  distinct(depmap_id, tissue, subtype, .keep_all = TRUE)  # one point per cell line (avoids duplicate rows from joins)
# Order tissue so SCLC (Lung), ExPulm-SCNEC, AML appear next to each other; Localized/Metastatic pairs adjacent within each tissue
tissue_order_f1 <- c(
  "Lung", "ExPulm-SCNEC", "AML", "Neuroblastoma", "Breast",
  if (cohort_ext_hem) "TNBC",
  "Lymphoma",
  "Prostate", "Urothelial", "Renal cell", "Colorectal", "Melanoma", "Sarcoma"
)
# Wide Figure 1 x-axis order (shared with CERES + DNMT1): stem-like (left-of-line) block first, then remainder by lineage_levels within tissue.
# Neuroblastoma before Lung so NB strata sit left of the red dashed rule (drawn before first NSCLC); Lung NSCLC comparators stay right of the line.
f1_tissue_axis_order <- c(
  "ExPulm-SCNEC", "Lymphoma", "AML", "Neuroblastoma", "Lung", "Breast",
  if (cohort_ext_hem) "TNBC",
  "Prostate", "Urothelial", "Renal cell", "Colorectal", "Melanoma", "Sarcoma"
)
bcl2_plot_df <- bcl2_plot_df %>%
  dplyr::mutate(
    tissue = factor(as.character(tissue), levels = intersect(tissue_order_f1, unique(as.character(tissue)))),
    subtype = factor(
      dplyr::if_else(
        as.character(.data$subtype) %in% c("Localized Neuroblastoma", "Metastatic Neuroblastoma"),
        "Neuroblastoma",
        as.character(.data$subtype)
      ),
      levels = unique(c(as.character(lineage_levels), "Neuroblastoma"))
    )
  )
# #region agent log
tryCatch({
  nb_rows_fac <- bcl2_plot_df %>% dplyr::filter(as.character(.data$tissue) == "Neuroblastoma")
  .agent_dbg_ndjson(
    runId = "iter4",
    hypothesisId = "H9",
    location = "03_Rmd:bcl2_subtype_factor_cast",
    message = "Neuroblastoma rows after subtype pooling and factor cast",
    data = list(
      lineage_levels_has_nb = any(as.character(lineage_levels) == "Neuroblastoma"),
      nb_rows_total = nrow(nb_rows_fac),
      nb_subtype_na = sum(is.na(nb_rows_fac$subtype)),
      nb_subtype_values = unique(as.character(nb_rows_fac$subtype))
    )
  )
}, error = function(e) invisible(NULL))
# #endregion
bcl2_plot_n <- bcl2_plot_df %>% group_by(tissue, subtype) %>% summarise(n_plot = n(), .groups = "drop")
bcl2_n_labels <- bcl2_plot_n %>%
  left_join(cohort_n %>% select(subtype, n_cohort = n), by = "subtype") %>%
  mutate(n_cell = if_else(!is.na(n_cohort), n_cohort, n_plot)) %>%
  mutate(lab = paste0(gsub(" \\(control\\)", "", as.character(subtype)), " N=", n_cell)) %>%
  select(tissue, subtype, lab, n_cell)
# Extended TNBC/Lymph master: require N > 5 per stratum except anchor stem-like lineages (SCLC, ExPulm-SCNEC, AML)
hma_ext_fig <- exists("hma_bcl2_tnbc_lymph_master") && isTRUE(hma_bcl2_tnbc_lymph_master)
plastic_anchors <- if (exists("lineage_plastic_ref_canonical")) {
  as.character(lineage_plastic_ref_canonical)
} else {
  c("SCLC", "ExPulm-SCNEC", "AML (control)")
}
# When no large-N localized stratum will be rescued later, promote one high-N metastatic bucket through the n>5 filter
loc_pre <- bcl2_n_labels %>%
  dplyr::mutate(st = as.character(.data$subtype)) %>%
  dplyr::filter(grepl("^Localized\\b", .data$st)) %>%
  dplyr::filter(!grepl("Neuroblastoma", .data$st, fixed = TRUE)) %>%
  dplyr::filter(!grepl("NSCLC", .data$st, ignore.case = TRUE)) %>%
  dplyr::filter(.data$n_cell >= 20L)
if (nrow(loc_pre) == 0L) {
  meta_one <- bcl2_n_labels %>%
    dplyr::mutate(st = as.character(.data$subtype)) %>%
    dplyr::filter(grepl("^Metastatic\\b", .data$st)) %>%
    dplyr::filter(.data$n_cell >= 25L) %>%
    dplyr::filter(!(.data$st %in% plastic_anchors)) %>%
    dplyr::arrange(dplyr::desc(.data$n_cell), .data$st)
  if (nrow(meta_one) > 0L) {
    plastic_anchors <- unique(c(plastic_anchors, as.character(meta_one$st[[1L]])))
  }
}
if (hma_ext_fig) {
  # #region agent log
  tryCatch({
    nb_lbl_pre <- bcl2_n_labels %>% dplyr::filter(grepl("Neuroblastoma", as.character(.data$subtype), ignore.case = TRUE))
    .agent_dbg_ndjson(
      runId = "iter3",
      hypothesisId = "H8",
      location = "03_Rmd:bcl2_n_labels_pre_hma_filter",
      message = "Neuroblastoma labels before hma_ext_fig n>5/plastic_anchors filter",
      data = list(
        n_nb_lbl_pre = nrow(nb_lbl_pre),
        nb_lbl_pre_subtypes = unique(as.character(nb_lbl_pre$subtype)),
        nb_lbl_pre_ncell = as.integer(nb_lbl_pre$n_cell)
      )
    )
  }, error = function(e) invisible(NULL))
  # #endregion
  nb_lineage_lv <- intersect(
    c("Neuroblastoma", "Unspecified Neuroblastoma"),
    if (exists("lineage_levels")) as.character(lineage_levels) else character(0)
  )
  sarcoma_lineage_lv <- intersect(
    c("Localized Sarcoma", "Metastatic Sarcoma", "Unspecified Sarcoma"),
    if (exists("lineage_levels")) as.character(lineage_levels) else character(0)
  )
  plastic_anchors <- unique(c(
    plastic_anchors,
    "DLBCL/MCL/CLL",
    nb_lineage_lv,
    sarcoma_lineage_lv
  ))
  bcl2_n_labels <- bcl2_n_labels %>%
    dplyr::filter(.data$n_cell > 5L | as.character(.data$subtype) %in% plastic_anchors) %>%
    dplyr::filter(
      !grepl("Unspecified", as.character(.data$subtype), fixed = TRUE) |
        grepl("Unspecified Neuroblastoma", as.character(.data$subtype), fixed = TRUE) |
        grepl("Unspecified Sarcoma", as.character(.data$subtype), fixed = TRUE)
    )
  # #region agent log
  tryCatch({
    nb_lbl_post <- bcl2_n_labels %>% dplyr::filter(grepl("Neuroblastoma", as.character(.data$subtype), ignore.case = TRUE))
    .agent_dbg_ndjson(
      runId = "iter3",
      hypothesisId = "H8",
      location = "03_Rmd:bcl2_n_labels_post_hma_filter",
      message = "Neuroblastoma labels after hma_ext_fig n>5/plastic_anchors filter",
      data = list(
        n_nb_lbl_post = nrow(nb_lbl_post),
        nb_lbl_post_subtypes = unique(as.character(nb_lbl_post$subtype)),
        plastic_anchors_nb = as.character(plastic_anchors[grepl("Neuroblastoma", as.character(plastic_anchors), ignore.case = TRUE)])
      )
    )
  }, error = function(e) invisible(NULL))
  # #endregion
}
# Include all cohort subtypes in Figure 1 unless extended master filtered above
bcl2_keep <- bcl2_n_labels
bcl2_plot_df <- bcl2_plot_df %>% semi_join(bcl2_keep, by = c("tissue", "subtype"))
bcl2_n_labels <- bcl2_n_labels %>% semi_join(bcl2_keep, by = c("tissue", "subtype"))
# One additional localized lineage-stable comparator for Fig. 1 wide panels.
# Prefer Localized prostate only when it has N > 5 in each subpanel (BCL2 expr, BCL2 CERES, DNMT1 expr, DNMT1 CERES);
# otherwise choose another localized stable subtype with the largest cross-panel minimum N.
f1_extra_locmeta_keep <- NA_character_
if (nrow(bcl2_n_labels) > 0L) {
  cand_loc_f1 <- bcl2_n_labels %>%
    dplyr::mutate(st = as.character(.data$subtype)) %>%
    dplyr::filter(grepl("^Localized\\b", .data$st)) %>%
    dplyr::filter(!grepl("Neuroblastoma", .data$st, fixed = TRUE)) %>%
    dplyr::filter(!grepl("NSCLC", .data$st, ignore.case = TRUE)) %>%
    dplyr::arrange(dplyr::desc(.data$n_cell), .data$st)
  if (nrow(cand_loc_f1) > 0L && exists("expr_target")) {
    cand_lv <- unique(as.character(cand_loc_f1$st))
    n_bcl2_expr <- expr_target %>%
      dplyr::filter(as.character(.data$lineage_group) %in% cand_lv, !is.na(.data$BCL2)) %>%
      dplyr::count(.data$lineage_group, name = "n_bcl2_expr")
    n_dnmt1_expr <- if ("DNMT1" %in% names(expr_target)) {
      expr_target %>%
        dplyr::filter(as.character(.data$lineage_group) %in% cand_lv, !is.na(.data$DNMT1)) %>%
        dplyr::count(.data$lineage_group, name = "n_dnmt1_expr")
    } else if (exists("expr_meta") && nrow(expr_meta) > 0L && "DNMT1" %in% names(expr_meta)) {
      expr_target %>%
        dplyr::select(.data$depmap_id, .data$lineage_group) %>%
        dplyr::inner_join(expr_meta %>% dplyr::select(.data$depmap_id, .data$DNMT1), by = "depmap_id") %>%
        dplyr::filter(as.character(.data$lineage_group) %in% cand_lv, !is.na(.data$DNMT1)) %>%
        dplyr::count(.data$lineage_group, name = "n_dnmt1_expr")
    } else {
      tibble::tibble(lineage_group = cand_lv, n_dnmt1_expr = 0L)
    }
    n_bcl2_ceres <- tibble::tibble(lineage_group = cand_lv, n_bcl2_ceres = 0L)
    if (exists("crispr") && nrow(crispr) > 0L && exists("target_ids") && length(target_ids) > 0L) {
      n_bcl2_ceres <- crispr %>%
        dplyr::filter(grepl("^BCL2\\b", .data$gene_name), .data$depmap_id %in% target_ids, is.finite(.data$dependency)) %>%
        dplyr::inner_join(expr_target %>% dplyr::select(.data$depmap_id, .data$lineage_group), by = "depmap_id") %>%
        dplyr::filter(as.character(.data$lineage_group) %in% cand_lv) %>%
        dplyr::count(.data$lineage_group, name = "n_bcl2_ceres")
    }
    n_dnmt1_ceres <- tibble::tibble(lineage_group = cand_lv, n_dnmt1_ceres = 0L)
    if (exists("crispr_dnmt1") && nrow(crispr_dnmt1) > 0L) {
      n_dnmt1_ceres <- crispr_dnmt1 %>%
        dplyr::filter(as.character(.data$lineage_group) %in% cand_lv, is.finite(.data$dependency)) %>%
        dplyr::count(.data$lineage_group, name = "n_dnmt1_ceres")
    }
    cand_rank <- cand_loc_f1 %>%
      dplyr::transmute(st = as.character(.data$st), n_label = as.integer(.data$n_cell)) %>%
      dplyr::left_join(n_bcl2_expr %>% dplyr::transmute(st = as.character(.data$lineage_group), n_bcl2_expr = as.integer(.data$n_bcl2_expr)), by = "st") %>%
      dplyr::left_join(n_bcl2_ceres %>% dplyr::transmute(st = as.character(.data$lineage_group), n_bcl2_ceres = as.integer(.data$n_bcl2_ceres)), by = "st") %>%
      dplyr::left_join(n_dnmt1_expr %>% dplyr::transmute(st = as.character(.data$lineage_group), n_dnmt1_expr = as.integer(.data$n_dnmt1_expr)), by = "st") %>%
      dplyr::left_join(n_dnmt1_ceres %>% dplyr::transmute(st = as.character(.data$lineage_group), n_dnmt1_ceres = as.integer(.data$n_dnmt1_ceres)), by = "st") %>%
      dplyr::mutate(
        n_bcl2_expr = dplyr::coalesce(.data$n_bcl2_expr, 0L),
        n_bcl2_ceres = dplyr::coalesce(.data$n_bcl2_ceres, 0L),
        n_dnmt1_expr = dplyr::coalesce(.data$n_dnmt1_expr, 0L),
        n_dnmt1_ceres = dplyr::coalesce(.data$n_dnmt1_ceres, 0L),
        n_min_4panels = pmin(.data$n_bcl2_expr, .data$n_bcl2_ceres, .data$n_dnmt1_expr, .data$n_dnmt1_ceres)
      ) %>%
      dplyr::arrange(dplyr::desc(.data$n_min_4panels), dplyr::desc(.data$n_label), .data$st)
    pick_prostate <- cand_rank %>%
      dplyr::filter(.data$st == "Localized prostate", .data$n_min_4panels > 5L)
    if (nrow(pick_prostate) > 0L) {
      f1_extra_locmeta_keep <- "Localized prostate"
    } else {
      pick_alt <- cand_rank %>% dplyr::filter(.data$n_min_4panels > 5L)
      if (nrow(pick_alt) > 0L) {
        f1_extra_locmeta_keep <- as.character(pick_alt$st[[1L]])
      } else if (nrow(cand_loc_f1) > 0L && is.finite(cand_loc_f1$n_cell[[1]]) && cand_loc_f1$n_cell[[1]] >= 20L) {
        f1_extra_locmeta_keep <- cand_loc_f1$st[[1L]]
      }
    }
  } else if (nrow(cand_loc_f1) > 0L && is.finite(cand_loc_f1$n_cell[[1]]) && cand_loc_f1$n_cell[[1]] >= 20L) {
    f1_extra_locmeta_keep <- cand_loc_f1$st[[1L]]
  }
}
bcl2_n_labels <- bcl2_n_labels %>% dplyr::select(tissue, subtype, lab)
# Figures 1–2: drop localized non-stem-like comparator buckets (keep localized neuroblastoma + localized NSCLC only),
# then drop selected strata entirely from these figures (localized TNBC / localized sarcoma / Lymphoma (other)).
.f1_exclude_lineage_from_fig12_plots <- function(lg_chr) {
  # Keep comparator set consistent across BCL2 and DNMT1 expression/CERES panels.
  # Keep metastatic TNBC on Figure 1 wide-axis (shown as mTNBC); only exclude localized TNBC.
  drop_set <- c("Localized TNBC", "Localized Sarcoma", "Lymphoma (other)")
  if (exists("f1_extra_locmeta_keep", inherits = TRUE)) {
    ex <- as.character(get("f1_extra_locmeta_keep", inherits = TRUE))
    if (length(ex) == 1L && !is.na(ex) && nzchar(ex)) {
      drop_set <- unique(c(drop_set, "Metastatic prostate"))
    }
  }
  as.character(lg_chr) %in% drop_set
}
.f1_keep_localized_subtype_chr <- function(st_chr) {
  ex <- NA_character_
  if (exists("f1_extra_locmeta_keep", inherits = TRUE)) {
    ex <- as.character(get("f1_extra_locmeta_keep", inherits = TRUE))
    if (length(ex) != 1L || is.na(ex) || !nzchar(ex)) ex <- NA_character_
  }
  grepl("Neuroblastoma", st_chr, fixed = TRUE) |
    grepl("\\bNSCLC\\b", st_chr, ignore.case = TRUE) |
    (!is.na(ex) & st_chr == ex)
}
{
  st_drop <- as.character(bcl2_plot_df$subtype)
  keep_loc_st <- .f1_keep_localized_subtype_chr(st_drop)
  drop_loc_st <- grepl("^Localized\\b", st_drop) & !keep_loc_st
  bcl2_plot_df <- dplyr::filter(bcl2_plot_df, !drop_loc_st)
  st_drop2 <- as.character(bcl2_n_labels$subtype)
  keep_loc_st2 <- .f1_keep_localized_subtype_chr(st_drop2)
  drop_loc_st2 <- grepl("^Localized\\b", st_drop2) & !keep_loc_st2
  bcl2_n_labels <- dplyr::filter(bcl2_n_labels, !drop_loc_st2)
}
bcl2_plot_df <- dplyr::filter(bcl2_plot_df, !.f1_exclude_lineage_from_fig12_plots(as.character(.data$subtype)))
bcl2_n_labels <- dplyr::filter(bcl2_n_labels, !.f1_exclude_lineage_from_fig12_plots(as.character(.data$subtype)))
# Remove malformed lineage group labels before wide-axis pooling/order to avoid NA buckets.
bcl2_plot_df <- bcl2_plot_df %>%
  dplyr::filter(
    !is.na(.data$subtype),
    nzchar(trimws(as.character(.data$subtype))),
    toupper(trimws(as.character(.data$subtype))) != "NA"
  )
bcl2_n_labels <- bcl2_n_labels %>%
  dplyr::filter(
    !is.na(.data$subtype),
    nzchar(trimws(as.character(.data$subtype))),
    toupper(trimws(as.character(.data$subtype))) != "NA"
  )
# #region agent log
tryCatch({
  keep_counts <- bcl2_plot_df %>%
    dplyr::mutate(subtype_chr = as.character(.data$subtype)) %>%
    dplyr::filter(grepl("TNBC|Melanoma|Sarcoma", .data$subtype_chr, ignore.case = TRUE)) %>%
    dplyr::count(.data$subtype_chr, name = "n_expr")
  .agent_dbg_ndjson(
    runId = "pre-fix-debug2",
    hypothesisId = "H3",
    location = "03_Rmd:bcl2_keep_focus_subtypes",
    message = "BCL2 kept lineage focus counts after figure filters",
    data = list(
      n_rows = nrow(bcl2_plot_df),
      focus_subtypes = as.character(keep_counts$subtype_chr),
      focus_counts = as.integer(keep_counts$n_expr),
      extra_locmeta_keep = if (exists("f1_extra_locmeta_keep", inherits = TRUE)) as.character(get("f1_extra_locmeta_keep", inherits = TRUE)) else NA_character_
    )
  )
}, error = function(e) invisible(NULL))
# #endregion
# #region agent log
tryCatch({
  nb_plot <- bcl2_plot_df %>% dplyr::filter(grepl("Neuroblastoma", as.character(.data$subtype), ignore.case = TRUE))
  nb_lbl <- bcl2_n_labels %>% dplyr::filter(grepl("Neuroblastoma", as.character(.data$subtype), ignore.case = TRUE))
  .agent_dbg_ndjson(
    runId = "iter2",
    hypothesisId = "H7",
    location = "03_Rmd:bcl2_post_filters",
    message = "Neuroblastoma rows after localized/exclude filters",
    data = list(
      nb_plot_n = nrow(nb_plot),
      nb_plot_subtypes = unique(as.character(nb_plot$subtype)),
      nb_label_n = nrow(nb_lbl),
      nb_label_subtypes = unique(as.character(nb_lbl$subtype))
    )
  )
}, error = function(e) invisible(NULL))
# #endregion
if (nrow(bcl2_plot_df) >= 2) {
  bcl2_plot_df <- bcl2_plot_df %>% left_join(bcl2_n_labels %>% dplyr::select(tissue, subtype, lab), by = c("tissue", "subtype"))
  # Single x-axis: all lineages in one order (tissue then subtype), like CERES plot below
  lab_levels_f1 <- bcl2_plot_df %>% distinct(tissue, subtype, lab) %>%
    arrange(tissue, match(as.character(subtype), lineage_levels)) %>% pull(lab)
  bcl2_plot_df <- bcl2_plot_df %>% mutate(lab = factor(lab, levels = unique(as.character(lab_levels_f1))))
  # Final dedup: one point per cell line
  bcl2_plot_df <- bcl2_plot_df %>% distinct(depmap_id, lab, .keep_all = TRUE)
  bcl2_plot_df <- bcl2_plot_df %>% filter(!is.na(lab), nzchar(trimws(as.character(lab))))
  # Extended master: if >10 x-axis groups, pool non-anchor strata within tissue to localized vs metastatic
  if (hma_ext_fig && length(unique(as.character(bcl2_plot_df$lab))) > 10L) {
    st <- as.character(bcl2_plot_df$subtype)
    tt <- as.character(bcl2_plot_df$tissue)
    pooled_axis_preview <- dplyr::case_when(
      st %in% plastic_anchors ~ st,
      grepl("Neuroblastoma", st, fixed = TRUE) & !grepl("^Unspecified", st, fixed = TRUE) ~ "Neuroblastoma",
      grepl("^Localized\\b", st) ~ paste0(tt, " — localized"),
      grepl("^Metastatic\\b", st) ~ paste0(tt, " — metastatic"),
      TRUE ~ st
    )
    # #region agent log
    tryCatch({
      .agent_dbg_ndjson(
        runId = "pre-fix-debug2",
        hypothesisId = "H1",
        location = "03_Rmd:bcl2_pooling_preview",
        message = "BCL2 pooled-axis preview before filtering",
        data = list(
          n_rows = nrow(bcl2_plot_df),
          n_na_pooled = sum(is.na(pooled_axis_preview)),
          n_blank_pooled = sum(!is.na(pooled_axis_preview) & !nzchar(trimws(as.character(pooled_axis_preview)))),
          pooled_head = utils::head(unique(as.character(pooled_axis_preview)), 10L)
        )
      )
    }, error = function(e) invisible(NULL))
    # #endregion
    bcl2_plot_df <- bcl2_plot_df %>%
      dplyr::mutate(
        # Pool localized + metastatic neuroblastoma to one axis bucket (do not split to "Neuroblastoma — localized/metastatic")
        pooled_axis = dplyr::case_when(
          st %in% plastic_anchors ~ st,
          grepl("Neuroblastoma", st, fixed = TRUE) & !grepl("^Unspecified", st, fixed = TRUE) ~ "Neuroblastoma",
          grepl("^Localized\\b", st) ~ paste0(tt, " — localized"),
          grepl("^Metastatic\\b", st) ~ paste0(tt, " — metastatic"),
          TRUE ~ st
        )
      ) %>%
      dplyr::filter(!is.na(.data$pooled_axis), nzchar(trimws(as.character(.data$pooled_axis))))
    n_pool <- bcl2_plot_df %>% dplyr::count(.data$pooled_axis, name = "nn")
    bcl2_plot_df <- bcl2_plot_df %>%
      dplyr::left_join(n_pool, by = "pooled_axis") %>%
      dplyr::mutate(lab = paste0(gsub(" \\(control\\)", "", as.character(.data$pooled_axis)), " N=", .data$nn))
    ord_l <- bcl2_plot_df %>%
      dplyr::distinct(.data$tissue, .data$pooled_axis, .data$lab) %>%
      dplyr::arrange(.data$tissue, .data$pooled_axis) %>%
      dplyr::pull(.data$lab)
    bcl2_plot_df <- bcl2_plot_df %>%
      dplyr::mutate(lab = factor(.data$lab, levels = unique(ord_l))) %>%
      dplyr::select(-dplyr::any_of(c("pooled_axis", "nn")))
  }
  # Extended master: drop pooled labels with "Unspecified" except neuroblastoma / sarcoma strata
  if (hma_ext_fig && nrow(bcl2_plot_df) > 0L) {
    bcl2_plot_df <- bcl2_plot_df %>%
      dplyr::filter(
        !grepl("Unspecified", as.character(.data$lab), fixed = TRUE) |
          grepl("Unspecified Neuroblastoma", as.character(.data$lab), fixed = TRUE) |
          grepl("Unspecified Sarcoma", as.character(.data$lab), fixed = TRUE)
      )
  }
  # X-axis: fixed tumor order (f1_tissue_axis_order × lineage_levels); CERES wide inherits levels(bcl2_plot_df$lab); DNMT1 wide matches below
  f1_ord_tbl <- bcl2_plot_df %>%
    dplyr::distinct(.data$tissue, .data$subtype, .data$lab) %>%
    dplyr::mutate(
      ti_rnk = match(as.character(.data$tissue), f1_tissue_axis_order),
      st_rnk = match(as.character(.data$subtype), lineage_levels)
    ) %>%
    dplyr::mutate(
      ti_rnk = dplyr::coalesce(as.integer(.data$ti_rnk), 999L),
      st_rnk = dplyr::coalesce(as.integer(.data$st_rnk), 999L)
    )
  ord_labs_f1 <- f1_ord_tbl %>%
    dplyr::arrange(.data$ti_rnk, .data$st_rnk, as.character(.data$lab)) %>%
    dplyr::pull(.data$lab) %>%
    unique()
  extra_labs <- setdiff(unique(as.character(bcl2_plot_df$lab)), ord_labs_f1)
  if (length(extra_labs) > 0L) {
    ord_labs_f1 <- c(ord_labs_f1, sort(extra_labs))
  }
  # Lymphoma (other): comparator arm — x-axis at end of comparator block (strictly right of red vline before NSCLC)
  f1_lymph_other_subtype <- "Lymphoma (other)"
  lo_labs_f1 <- bcl2_plot_df %>%
    dplyr::filter(as.character(.data$subtype) == f1_lymph_other_subtype) %>%
    dplyr::distinct(.data$lab) %>%
    dplyr::pull(.data$lab) %>%
    unique() %>%
    as.character()
  if (length(lo_labs_f1) > 0L && length(ord_labs_f1) > 0L) {
    lo_in <- intersect(lo_labs_f1, ord_labs_f1)
    if (length(lo_in) > 0L) {
      is_comp_f1 <- grepl("NSCLC", ord_labs_f1, fixed = TRUE) |
        grepl("Lung — localized", ord_labs_f1, fixed = TRUE) |
        grepl("Lung — metastatic", ord_labs_f1, fixed = TRUE)
      j_comp <- suppressWarnings(min(which(is_comp_f1)))
      if (is.finite(j_comp) && j_comp > 1L) {
        left_vec <- ord_labs_f1[seq_len(j_comp - 1L)]
        right_vec <- ord_labs_f1[j_comp:length(ord_labs_f1)]
        # Append Lymphoma (other) after NSCLC/comparator block so it is strictly right of the red vline
        ord_labs_f1 <- c(setdiff(left_vec, lo_in), setdiff(right_vec, lo_in), lo_in)
      } else {
        nb_labs_pre <- unique(as.character(bcl2_plot_df$lab[as.character(bcl2_plot_df$tissue) == "Neuroblastoma"]))
        idx_nb <- which(ord_labs_f1 %in% nb_labs_pre)
        if (length(idx_nb) > 0L) {
          k_nb <- max(idx_nb)
          left_vec <- ord_labs_f1[seq_len(k_nb)]
          right_vec <- ord_labs_f1[(k_nb + 1L):length(ord_labs_f1)]
          ord_labs_f1 <- c(setdiff(left_vec, lo_in), setdiff(right_vec, lo_in), lo_in)
        }
      }
    }
  }
  bcl2_plot_df <- bcl2_plot_df %>%
    dplyr::mutate(lab = factor(as.character(.data$lab), levels = ord_labs_f1))
  # Wide-axis display: if the only ExPulm-SCNEC stratum is entirely NEPC-labeled lines, show "NEPC" on the x-axis (DepMap subtype stays ExPulm-SCNEC).
  if (nrow(bcl2_plot_df) > 0L && all(c("subtype", "is_nepc", "lab") %in% names(bcl2_plot_df))) {
    ex_i <- as.character(bcl2_plot_df$subtype) == "ExPulm-SCNEC"
    if (any(ex_i) && all(bcl2_plot_df$is_nepc[ex_i] %in% TRUE)) {
      lev_ex <- levels(bcl2_plot_df$lab)
      lev_ex2 <- sub("ExPulm-SCNEC", "NEPC", lev_ex, fixed = TRUE)
      bcl2_plot_df <- bcl2_plot_df %>%
        dplyr::mutate(lab = factor(sub("ExPulm-SCNEC", "NEPC", as.character(.data$lab), fixed = TRUE), levels = lev_ex2))
    }
  }
  .f1_sanitize_lab_rows <- function(df) {
    if (is.null(df) || !is.data.frame(df) || !"lab" %in% names(df) || nrow(df) < 1L) {
      return(df)
    }
    lab_chr <- as.character(df$lab)
    base_chr <- trimws(sub("\\s*N=\\s*.*$", "", lab_chr))
    bad_lab <- is.na(df$lab) |
      !nzchar(trimws(lab_chr)) |
      grepl("^NA\\b", lab_chr) |
      !nzchar(base_chr) |
      grepl("^N=\\s*\\d+$", trimws(lab_chr))
    if (any(bad_lab)) {
      df <- dplyr::filter(df, !bad_lab)
    }
    if (nrow(df) > 0L && is.factor(df$lab)) {
      df$lab <- factor(as.character(df$lab), levels = intersect(levels(df$lab), unique(as.character(df$lab))))
    }
    df
  }
  # Drop empty / malformed wide-axis buckets (e.g. "NA N=…") after pooling + joins.
  bcl2_plot_df <- .f1_sanitize_lab_rows(bcl2_plot_df)
  # Keep TNBC metastatic stratum (mTNBC) when present; only malformed labels are removed above.
  # #region agent log
  tryCatch({
    nb_axis_rows <- bcl2_plot_df %>% dplyr::filter(grepl("Neuroblastoma", as.character(.data$lab), fixed = TRUE))
    .agent_dbg_ndjson(
      runId = "post-fix",
      hypothesisId = "H1",
      location = "03_Rmd:bcl2_axis_final",
      message = "Neuroblastoma presence on expression axis",
      data = list(
        n_rows_total = nrow(bcl2_plot_df),
        n_nb_axis_rows = nrow(nb_axis_rows),
        nb_axis_labs = unique(as.character(nb_axis_rows$lab)),
        nb_axis_subtypes = unique(as.character(nb_axis_rows$subtype))
      )
    )
  }, error = function(e) invisible(NULL))
  # #endregion
  # Wide axis vline + stem-like arm: prefer placing line so Localized/Metastatic NSCLC are strictly to the right; else NB-based fallback
  f1_enriched_lineages <- character(0)
  f1_vline_x <- NA_real_
  f1_wilcox_left_aml_p <- NA_real_
  f1_p_subtitle <- NULL
  lv_f1 <- levels(bcl2_plot_df$lab)
  # Comparator arm to the right of the dashed line: explicit NSCLC strata and pooled lung localized/metastatic (no "NSCLC" in label text)
  ns_idx_f1 <- which(
    grepl("NSCLC", lv_f1, fixed = TRUE) |
      grepl("Lung — localized", lv_f1, fixed = TRUE) |
      grepl("Lung — metastatic", lv_f1, fixed = TRUE)
  )
  nb_lab_set_f1 <- unique(as.character(bcl2_plot_df$lab[as.character(bcl2_plot_df$tissue) == "Neuroblastoma"]))
  nb_idx_f1 <- which(lv_f1 %in% nb_lab_set_f1)
  left_labs_f1 <- character(0)
  if (length(ns_idx_f1) > 0L && length(lv_f1) > 0L) {
    k_before_nsclc <- min(ns_idx_f1) - 1L
    if (k_before_nsclc >= 1L) {
      f1_vline_x <- min(ns_idx_f1) - 0.5
      left_labs_f1 <- lv_f1[seq_len(k_before_nsclc)]
    }
  } else if (length(nb_idx_f1) > 0L && length(lv_f1) > 0L) {
    k_right_nb <- max(nb_idx_f1)
    f1_vline_x <- k_right_nb + 0.5
    left_labs_f1 <- lv_f1[seq_len(k_right_nb)]
  }
  if (length(left_labs_f1) > 0L) {
    f1_enriched_lineages <- bcl2_plot_df %>%
      dplyr::filter(as.character(.data$lab) %in% left_labs_f1) %>%
      dplyr::distinct(.data$subtype) %>%
      dplyr::pull(.data$subtype) %>%
      unique() %>%
      as.character()
    # Comparator bucket: never part of stem-like / left-of-line arm
    f1_enriched_lineages <- setdiff(f1_enriched_lineages, "Lymphoma (other)")
    # NSCLC is a non-stem-like lung comparator (right of dashed rule): never part of the stem-like arm
    f1_enriched_lineages <- f1_enriched_lineages[!grepl("NSCLC", f1_enriched_lineages, fixed = TRUE)]
    f1_enriched_lineages <- f1_enriched_lineages[!.f1_exclude_lineage_from_fig12_plots(f1_enriched_lineages)]
    aml_b2_f1 <- dplyr::filter(bcl2_plot_df, as.character(.data$subtype) == "AML (control)") %>% dplyr::pull(.data$BCL2)
    enr_b2_f1 <- dplyr::filter(
      bcl2_plot_df,
      as.character(.data$lab) %in% left_labs_f1,
      as.character(.data$subtype) != "AML (control)"
    ) %>%
      dplyr::pull(.data$BCL2)
    enr_b2_f1 <- as.numeric(stats::na.omit(enr_b2_f1))
    aml_b2_f1 <- as.numeric(stats::na.omit(aml_b2_f1))
    if (length(enr_b2_f1) >= 2L && length(aml_b2_f1) >= 2L) {
      wt_f1 <- tryCatch(
        stats::wilcox.test(enr_b2_f1, aml_b2_f1, alternative = "greater", exact = FALSE),
        error = function(e) NULL
      )
      if (!is.null(wt_f1) && !is.na(wt_f1$p.value)) {
        f1_wilcox_left_aml_p <- unname(wt_f1$p.value)
        pl <- if (f1_wilcox_left_aml_p < 0.001) "<0.001" else format.pval(f1_wilcox_left_aml_p, digits = 3)
        line_txt <- if (length(ns_idx_f1) > 0L) {
          "left of localized/metastatic NSCLC strata"
        } else {
          "right of neuroblastoma strata"
        }
        f1_p_subtitle <- paste0("Red dashed line: ", line_txt, ".")
      }
    }
  }
  f1_vline_leftof_nsclc <- length(ns_idx_f1) > 0L && length(left_labs_f1) > 0L
  f1_enriched_lineages_ordered <- if (length(f1_enriched_lineages) > 0L && exists("lineage_levels")) {
    f1_enriched_lineages[order(match(f1_enriched_lineages, as.character(lineage_levels)))]
  } else {
    character(0)
  }
  # Tighter y window for wide BCL2 expression (top row): cap upper near ~6 log2(TPM+1) so lineage contrasts read larger.
  f1_bcl2_expr_y_hi_cap <- 6.15
  f1_bcl2_expr_ylim <- local({
    yv <- bcl2_plot_df$BCL2
    yv <- yv[is.finite(yv)]
    if (length(yv) < 1L) {
      return(NULL)
    }
    yr <- range(yv)
    d <- diff(yr)
    if (!is.finite(d) || d < 1e-6) d <- 0.35
    pad <- max(0.1, 0.065 * d)
    y_lo <- max(0, yr[1] - pad)
    y_hi <- min(f1_bcl2_expr_y_hi_cap, yr[2] + pad)
    # Widen window to include every finite NEPC BCL2 value (cap can otherwise hide the red dot).
    yn <- bcl2_plot_df$BCL2[bcl2_plot_df$is_nepc %in% TRUE]
    yn <- yn[is.finite(yn)]
    if (length(yn) > 0L) {
      rng_n <- if (length(yn) >= 2L) diff(range(yn)) else 0
      pad_n <- max(0.1, 0.12 * max(rng_n, max(abs(yn), na.rm = TRUE) * 0.08, 0.22))
      y_lo <- max(0, min(y_lo, min(yn) - pad_n))
      y_hi <- max(y_hi, max(yn) + pad_n)
      y_hi <- min(8.5, y_hi)
    }
    if (!is.finite(y_hi) || y_hi <= y_lo) y_hi <- y_lo + 0.45
    c(y_lo, y_hi)
  })
  .mk_bcl2_expr_plot <- function(dat, vline_x = NA_real_, subtitle_txt = NULL) {
    g <- ggplot(dat, aes(x = lab, y = BCL2, fill = subtype)) +
      geom_boxplot(alpha = 0.7, outlier.shape = NA) +
      ggplot2::geom_point(
        data = dplyr::filter(dat, .data$is_nepc %in% TRUE),
        ggplot2::aes(x = lab, y = BCL2),
        inherit.aes = FALSE,
        color = "#E41A1C",
        position = ggplot2::position_jitter(width = 0.12, height = 0),
        alpha = 0.92,
        size = 2.5
      ) +
      ggplot2::scale_x_discrete(drop = FALSE, labels = .f1_axis_short_met_label) +
      labs(x = "Lineage", y = "BCL2 expression (log2(TPM+1))") +
      theme_minimal() +
      .f1_fig12_axis_theme(45) +
      theme(legend.position = "none")
    if (!is.na(vline_x)) {
      g <- g + ggplot2::geom_vline(
        xintercept = vline_x,
        linetype = "dashed",
        linewidth = f1_rule_line_linewidth,
        alpha = 1,
        colour = f1_rule_line_colour
      )
    }
    g
  }
  # Wide x-axis: NEPC-only point overlay (red); metastatic tick labels shortened via .f1_axis_short_met_label.
  # Bracket p-values are added in chunk `bcl2-fig-table-dependency` after `similarity-panels` defines fig12_* p-values.
  p_bcl2 <- .mk_bcl2_expr_plot(bcl2_plot_df, vline_x = f1_vline_x, subtitle_txt = NULL)
} else {
  p_bcl2 <- NULL
  p_placeholder <- ggplot() + annotate("text", x = 0.5, y = 0.5, label = "Insufficient data for Figure 1 (need ≥2 groups with BCL2). Run Rscript scripts/load_depmap_cache.R then re-knit.", size = 5) + theme_void()
  f1_enriched_lineages_ordered <- character(0)
}
```

```{r similarity-panels, include = FALSE, cache = FALSE}
# Create BCL2 and DNMT1 similarity panels for use as subfigures in Figure 1 (BCL2) and Figure 2 (DNMT1)
lineage_plastic <- if (exists("lineage_plastic_ref_canonical")) lineage_plastic_ref_canonical else c("SCLC", "ExPulm-SCNEC", "AML (control)")
hma_fig1_b2 <- exists("hma_bcl2_tnbc_lymph_master") && isTRUE(hma_bcl2_tnbc_lymph_master)
lymph_bcl2_context_lbl <- "DLBCL/MCL/CLL"
lineage_bcl2_fig1_multi <- lineage_plastic
if (hma_fig1_b2) {
  lineage_bcl2_fig1_multi <- c(lineage_plastic, lymph_bcl2_context_lbl)
}
# Wide Figure 1: lineages left of dashed vertical rule on Fig. 1 wide axis (NSCLC to the right of line when those strata are shown; else NB rule)
if (exists("f1_enriched_lineages") && length(f1_enriched_lineages) > 0L) {
  lineage_bcl2_fig1_multi <- f1_enriched_lineages
}
lineage_bcl2_fig1_multi <- lineage_bcl2_fig1_multi[!grepl("NSCLC", lineage_bcl2_fig1_multi, fixed = TRUE)]
lineage_bcl2_fig1_multi <- lineage_bcl2_fig1_multi[!.f1_exclude_lineage_from_fig12_plots(lineage_bcl2_fig1_multi)]
lbl_b2_enriched_two <- if (exists("f1_enriched_lineages") && length(f1_enriched_lineages) > 0L) {
  "Stem-like"
} else if (hma_fig1_b2) {
  "Stem-like"
} else {
  "Stem-like (SCLC, ExPulm-SCNEC, AML)"
}
lbl_other_two <- "Non-stem-like"
p_three <- p_two <- p_three_dnmt1 <- p_two_dnmt1 <- NULL
kw_txt <- wx_txt <- kw_dnmt1_txt <- wx_dnmt1_txt <- "—"
ni_aml_expr_note <- ""
cap_bcl2_sim <- "Figure 2b. BCL2 and DNMT1 similarity (see subfigures in Figures 1 and 2)."
if (exists("expr_target") && nrow(expr_target) > 0 && "BCL2" %in% names(expr_target)) {
  bcl2_sim <- expr_target %>% filter(!is.na(BCL2)) %>%
    mutate(is_nepc = (lineage_group == "ExPulm-SCNEC" & dplyr::coalesce(lineage_expulm_detail, "") == "NEPC"))
  lg0 <- as.character(bcl2_sim$lineage_group)
  keep_loc <- .f1_keep_localized_subtype_chr(lg0)
  drop_loc <- grepl("^Localized\\b", lg0) & !keep_loc
  bcl2_sim <- dplyr::filter(bcl2_sim, !drop_loc)
  bcl2_sim <- dplyr::filter(bcl2_sim, !.f1_exclude_lineage_from_fig12_plots(as.character(.data$lineage_group)))
  bcl2_three <- bcl2_sim %>% filter(lineage_group %in% lineage_bcl2_fig1_multi)
  bcl2_sim_two <- bcl2_sim %>%
    mutate(
      in_plastic_grp = (.data$lineage_group %in% lineage_bcl2_fig1_multi) &
        !grepl("NSCLC", as.character(.data$lineage_group), fixed = TRUE),
      group_2 = if_else(in_plastic_grp, lbl_b2_enriched_two, lbl_other_two)
    )
  kw_three <- tryCatch(kruskal.test(BCL2 ~ lineage_group, data = bcl2_three), error = function(e) list(p.value = NA_real_))
  lbl_enr_g2 <- lbl_b2_enriched_two
  lbl_oth_g2 <- lbl_other_two
  vals_b2_enr <- dplyr::filter(bcl2_sim_two, .data$group_2 == lbl_enr_g2) %>% dplyr::pull(.data$BCL2)
  vals_b2_oth <- dplyr::filter(bcl2_sim_two, .data$group_2 == lbl_oth_g2) %>% dplyr::pull(.data$BCL2)
  wx_vs_other <- if (length(stats::na.omit(vals_b2_enr)) >= 1L && length(stats::na.omit(vals_b2_oth)) >= 1L) {
    tryCatch(
      stats::wilcox.test(vals_b2_enr, vals_b2_oth, alternative = "greater", exact = FALSE),
      error = function(e) list(p.value = NA_real_)
    )
  } else {
    list(p.value = NA_real_)
  }
  kw_p <- kw_three$p.value
  wx_p <- wx_vs_other$p.value
  wx_txt <- if (is.na(wx_p)) "—" else if (wx_p < 0.001) "<0.001" else format.pval(wx_p, digits = 2)
  n_three <- bcl2_three %>% group_by(lineage_group) %>% summarise(n = n(), .groups = "drop")
  lab_three <- setNames(paste0(gsub(" \\(control\\)", "", as.character(n_three$lineage_group)), " N=", n_three$n), as.character(n_three$lineage_group))
  prio_b <- c("SCLC", "ExPulm-SCNEC", "AML (control)", if (hma_fig1_b2) lymph_bcl2_context_lbl else NULL)
  u_b <- unique(as.character(bcl2_three$lineage_group))
  ord_b <- c(intersect(prio_b, u_b), setdiff(u_b, prio_b))
  bcl2_three <- bcl2_three %>% mutate(lineage_group = factor(as.character(lineage_group), levels = ord_b))
  use_f1_nb_arm <- exists("f1_enriched_lineages") && length(f1_enriched_lineages) > 0L
  ttl_multi <- if (use_f1_nb_arm) {
    paste0(
      "BCL2: lineages left of Fig. 1 vertical rule (n = ", length(ord_b), " groups; ",
      if (exists("f1_vline_leftof_nsclc") && isTRUE(f1_vline_leftof_nsclc)) {
        "NSCLC to the right of dashed line"
      } else {
        "neuroblastoma-based rule if NSCLC absent"
      },
      ")"
    )
  } else if (length(ord_b) >= 4L) {
    if (hma_fig1_b2) "Stem-like lineages + DLBCL/MCL/CLL (four groups)" else "Stem-like lineages (four groups)"
  } else {
    if (hma_fig1_b2) "Stem-like lineages + DLBCL/MCL/CLL (three groups)" else "Stem-like lineages (three groups)"
  }
  aml_bcl2_vals <- dplyr::filter(bcl2_sim, .data$lineage_group == "AML (control)") %>% dplyr::pull(.data$BCL2)
  ni_bits_expr <- character(0)
  if (exists("f1_wilcox_left_aml_p") && !is.na(f1_wilcox_left_aml_p) && use_f1_nb_arm) {
    plf <- if (f1_wilcox_left_aml_p < 0.001) "<0.001" else format.pval(f1_wilcox_left_aml_p, digits = 3)
    ni_sub_bcl2 <- paste(
      strwrap(
        paste0(
          "Same arm as Fig. 1 (left of red dashed line, excluding AML) vs AML — 1-sided Wilcoxon (NI / ≥ AML): p = ", plf, "."
        ),
        width = 92
      ),
      collapse = "\n"
    )
    ni_aml_expr_note <- paste0(
      " Combined BCL2 (left of Fig. 1 line",
      if (exists("f1_vline_leftof_nsclc") && isTRUE(f1_vline_leftof_nsclc)) ", NSCLC right of line" else "",
      ", excl. AML) vs AML: p = ", plf, "."
    )
  } else if (length(aml_bcl2_vals) >= 2L) {
    for (lg in ord_b) {
      if (identical(as.character(lg), "AML (control)")) next
      xv <- dplyr::filter(bcl2_sim, as.character(.data$lineage_group) == as.character(lg)) %>% dplyr::pull(.data$BCL2)
      if (length(stats::na.omit(xv)) < 1L) next
      wt <- tryCatch(
        stats::wilcox.test(xv, aml_bcl2_vals, alternative = "greater", exact = FALSE),
        error = function(e) NULL
      )
      if (is.null(wt) || is.na(wt$p.value)) next
      p1 <- unname(wt$p.value)
      pl <- if (p1 < 0.001) "<0.001" else format.pval(p1, digits = 2)
      short_lg <- gsub(" \\(control\\)", "", as.character(lg))
      ni_bits_expr <- c(ni_bits_expr, paste0(short_lg, " p=", pl))
    }
    ni_sub_bcl2 <- if (length(ni_bits_expr) > 0L) {
      paste(strwrap(paste0("vs AML (1-sided Wilcoxon, stochastic ≥ AML): ", paste(ni_bits_expr, collapse = "; ")), width = 92), collapse = "\n")
    } else {
      NULL
    }
    if (length(ni_bits_expr) > 0L) {
      ni_aml_expr_note <- paste0(" BCL2 RNA vs AML (1-sided): ", paste(ni_bits_expr, collapse = "; "), ".")
    }
  } else {
    ni_sub_bcl2 <- NULL
  }
  # Multi-lineage BCL2 panel: corner label = non-inferior vs AML (1-sided), not Kruskal–Wallis
  bcl2_ni_p <- NA_real_
  if (exists("f1_wilcox_left_aml_p") && !is.na(f1_wilcox_left_aml_p) && use_f1_nb_arm) {
    bcl2_ni_p <- f1_wilcox_left_aml_p
  } else if (length(aml_bcl2_vals) >= 2L && nrow(bcl2_three) > 0L) {
    pool_b2_ni <- dplyr::filter(bcl2_three, as.character(.data$lineage_group) != "AML (control)") %>%
      dplyr::pull(.data$BCL2) %>%
      stats::na.omit() %>%
      as.numeric()
    if (length(pool_b2_ni) >= 2L) {
      wt_ni_b2 <- tryCatch(
        stats::wilcox.test(pool_b2_ni, aml_bcl2_vals, alternative = "greater", exact = FALSE),
        error = function(e) NULL
      )
      if (!is.null(wt_ni_b2) && !is.na(wt_ni_b2$p.value)) {
        bcl2_ni_p <- unname(wt_ni_b2$p.value)
      }
    }
  }
  kw_txt <- if (is.na(bcl2_ni_p)) "—" else if (bcl2_ni_p < 0.001) "<0.001" else format.pval(bcl2_ni_p, digits = 2)
  kw_lab_bcl2 <- if (is.na(bcl2_ni_p)) {
    "NI vs AML: p = —"
  } else {
    paste0("NI vs AML (1-sided): p = ", ifelse(bcl2_ni_p < 0.001, "<0.001", format.pval(bcl2_ni_p, digits = 2)))
  }
  p_three <- ggplot(bcl2_three, aes(x = lineage_group, y = BCL2, fill = lineage_group)) +
    geom_boxplot(alpha = 0.7, outlier.shape = NA) +
    geom_point(aes(color = is_nepc), position = position_jitter(width = 0.12, height = 0), alpha = 0.85, size = 2.5) +
    scale_color_manual(values = c("TRUE" = "#E41A1C", "FALSE" = "gray45"), guide = "none") +
    annotate("text", x = Inf, y = Inf, label = kw_lab_bcl2, hjust = 1.05, vjust = 1.5, size = 5) +
    scale_x_discrete(labels = lab_three) +
    labs(x = "Lineage", y = "BCL2 expression (log2(TPM+1))", title = ttl_multi, subtitle = ni_sub_bcl2) +
    theme_minimal() + .f1_fig12_axis_theme(25) + theme(legend.position = "none", plot.subtitle = element_text(size = 20, lineheight = 1.05))
  n_two <- bcl2_sim_two %>% group_by(group_2) %>% summarise(n = n(), .groups = "drop")
  lab_two <- setNames(paste0(as.character(n_two$group_2), " N=", n_two$n), as.character(n_two$group_2))
  wx_lab_bcl2 <- if (is.na(wx_vs_other$p.value)) {
    "1-sided vs stable: p = —"
  } else {
    paste0(
      "1-sided vs stable: p = ",
      ifelse(wx_vs_other$p.value < 0.001, "<0.001", format.pval(wx_vs_other$p.value, digits = 2))
    )
  }
  p_two <- ggplot(bcl2_sim_two, aes(x = group_2, y = BCL2, fill = group_2)) +
    geom_boxplot(alpha = 0.7, outlier.shape = NA) +
    geom_point(aes(color = is_nepc), position = position_jitter(width = 0.12, height = 0), alpha = 0.85, size = 2.5) +
    scale_color_manual(values = c("TRUE" = "#E41A1C", "FALSE" = "gray45"), guide = "none") +
    annotate("text", x = Inf, y = Inf, label = wx_lab_bcl2, hjust = 1.05, vjust = 1.5, size = 5) +
    scale_x_discrete(labels = lab_two) +
    labs(x = "", y = "BCL2 expression (log2(TPM+1))", title = "BCL2 (two groups)") +
    theme_minimal() + .f1_fig12_axis_theme(15) + theme(legend.position = "none")
  cap_bcl2_sim <- paste0(
    "Figure 2b. BCL2 and DNMT1 similarity. BCL2 multi-lineage: NI vs AML (1-sided) p = ", kw_txt,
    "; two-group Wilcoxon (stem-like > non-stem-like, 1-sided) p = ", wx_txt, "). DNMT1 (see Figure 2 subfigures).",
    if (nzchar(ni_aml_expr_note)) ni_aml_expr_note else ""
  )
  if (exists("expr_meta_dnmt1") && nrow(expr_meta_dnmt1) > 0 && "DNMT1" %in% names(expr_meta_dnmt1)) {
    dnmt1_sim <- expr_target %>%
      mutate(is_nepc = (lineage_group == "ExPulm-SCNEC" & dplyr::coalesce(lineage_expulm_detail, "") == "NEPC")) %>%
      left_join(expr_meta_dnmt1 %>% select(depmap_id, DNMT1), by = "depmap_id") %>%
      filter(!is.na(DNMT1))
    lg0d <- as.character(dnmt1_sim$lineage_group)
    keep_loc_d <- .f1_keep_localized_subtype_chr(lg0d)
    drop_loc_d <- grepl("^Localized\\b", lg0d) & !keep_loc_d
    dnmt1_sim <- dplyr::filter(dnmt1_sim, !drop_loc_d)
    dnmt1_sim <- dplyr::filter(dnmt1_sim, !.f1_exclude_lineage_from_fig12_plots(as.character(.data$lineage_group)))
    lineage_dnmt1_multi <- lineage_bcl2_fig1_multi
    dnmt1_three <- dnmt1_sim %>% filter(lineage_group %in% lineage_dnmt1_multi)
    lbl_dnmt1_enriched_two <- lbl_b2_enriched_two
    dnmt1_sim_two <- dnmt1_sim %>%
      mutate(
        in_plastic_dnmt1 = (.data$lineage_group %in% lineage_dnmt1_multi) &
          !grepl("NSCLC", as.character(.data$lineage_group), fixed = TRUE),
        group_2 = if_else(in_plastic_dnmt1, lbl_dnmt1_enriched_two, lbl_other_two)
      )
    kw_three_dnmt1 <- tryCatch(kruskal.test(DNMT1 ~ lineage_group, data = dnmt1_three), error = function(e) list(p.value = NA_real_))
    wx_vs_other_dnmt1 <- tryCatch(wilcox.test(DNMT1 ~ group_2, data = dnmt1_sim_two), error = function(e) list(p.value = NA_real_))
    # Same arm as Fig. 1 (left of red line, excl. AML) vs AML: 1-sided Wilcoxon, DNMT1 stochastically ≤ AML
    dnmt1_ni_p <- NA_real_
    if (use_f1_nb_arm && exists("f1_enriched_lineages") && length(f1_enriched_lineages) > 0L) {
      pool_dnmt1_ni <- dnmt1_sim %>%
        dplyr::filter(as.character(.data$lineage_group) %in% as.character(f1_enriched_lineages)) %>%
        dplyr::filter(as.character(.data$lineage_group) != "AML (control)") %>%
        dplyr::filter(!grepl("NSCLC", as.character(.data$lineage_group), fixed = TRUE))
      aml_dnmt1_ni <- dnmt1_sim %>% dplyr::filter(.data$lineage_group == "AML (control)")
      pd1 <- stats::na.omit(pool_dnmt1_ni$DNMT1)
      ad1 <- stats::na.omit(aml_dnmt1_ni$DNMT1)
      if (length(pd1) >= 2L && length(ad1) >= 2L) {
        wt_dnmt1_ni <- tryCatch(
          stats::wilcox.test(pd1, ad1, alternative = "less", exact = FALSE),
          error = function(e) NULL
        )
        if (!is.null(wt_dnmt1_ni) && !is.na(wt_dnmt1_ni$p.value)) {
          dnmt1_ni_p <- unname(wt_dnmt1_ni$p.value)
        }
      }
    }
    kw_dnmt1_txt <- if (!is.na(dnmt1_ni_p)) {
      ifelse(dnmt1_ni_p < 0.001, "<0.001", format.pval(dnmt1_ni_p, digits = 2))
    } else if (is.na(kw_three_dnmt1$p.value)) {
      "—"
    } else if (kw_three_dnmt1$p.value < 0.001) {
      "<0.001"
    } else {
      format.pval(kw_three_dnmt1$p.value, digits = 2)
    }
    wx_dnmt1_txt <- if (is.na(wx_vs_other_dnmt1$p.value)) "—" else if (wx_vs_other_dnmt1$p.value < 0.001) "<0.001" else format.pval(wx_vs_other_dnmt1$p.value, digits = 2)
    kw_lab_dnmt1 <- if (!is.na(dnmt1_ni_p)) {
      paste0(
        "NI vs AML (1-sided, ≤ AML): p = ",
        ifelse(dnmt1_ni_p < 0.001, "<0.001", format.pval(dnmt1_ni_p, digits = 2))
      )
    } else if (is.na(kw_three_dnmt1$p.value)) {
      "p = —"
    } else {
      paste0("p = ", ifelse(kw_three_dnmt1$p.value < 0.001, "<0.001", format.pval(kw_three_dnmt1$p.value, digits = 2)))
    }
    n_three_d <- dnmt1_three %>% group_by(lineage_group) %>% summarise(n = n(), .groups = "drop")
    lab_three_d <- setNames(paste0(gsub(" \\(control\\)", "", as.character(n_three_d$lineage_group)), " N=", n_three_d$n), as.character(n_three_d$lineage_group))
    p_three_dnmt1 <- ggplot(dnmt1_three, aes(x = lineage_group, y = DNMT1, fill = lineage_group)) +
      geom_boxplot(alpha = 0.7, outlier.shape = NA) +
      geom_point(aes(color = is_nepc), position = position_jitter(width = 0.12, height = 0), alpha = 0.85, size = 2.5) +
      scale_color_manual(values = c("TRUE" = "#E41A1C", "FALSE" = "gray45"), guide = "none") +
      annotate("text", x = Inf, y = Inf, label = kw_lab_dnmt1, hjust = 1.05, vjust = 1.5, size = 5) +
      scale_x_discrete(labels = lab_three_d) +
      labs(
        x = "Lineage", y = "DNMT1 expression (log2(TPM+1))",
        title = if (length(unique(dnmt1_three$lineage_group)) >= 4L) "DNMT1 (four lineages)" else "DNMT1 (three lineages)"
      ) +
      theme_minimal() + .f1_fig12_axis_theme(25) + theme(legend.position = "none")
    n_two_d <- dnmt1_sim_two %>% group_by(group_2) %>% summarise(n = n(), .groups = "drop")
    lab_two_d <- setNames(paste0(as.character(n_two_d$group_2), " N=", n_two_d$n), as.character(n_two_d$group_2))
    wx_lab_dnmt1 <- if (is.na(wx_vs_other_dnmt1$p.value)) "p = —" else paste0("p = ", ifelse(wx_vs_other_dnmt1$p.value < 0.001, "<0.001", format.pval(wx_vs_other_dnmt1$p.value, digits = 2)))
    p_two_dnmt1 <- ggplot(dnmt1_sim_two, aes(x = group_2, y = DNMT1, fill = group_2)) +
      geom_boxplot(alpha = 0.7, outlier.shape = NA) +
      geom_point(aes(color = is_nepc), position = position_jitter(width = 0.12, height = 0), alpha = 0.85, size = 2.5) +
      scale_color_manual(values = c("TRUE" = "#E41A1C", "FALSE" = "gray45"), guide = "none") +
      annotate("text", x = Inf, y = Inf, label = wx_lab_dnmt1, hjust = 1.05, vjust = 1.5, size = 5) +
      scale_x_discrete(labels = lab_two_d) +
      labs(x = "", y = "DNMT1 expression (log2(TPM+1))", title = "DNMT1 (two groups)") +
      theme_minimal() + .f1_fig12_axis_theme(15) + theme(legend.position = "none")
    cap_bcl2_sim <- paste0(
      "Figure 2b. BCL2 and DNMT1 similarity. BCL2 multi-lineage: NI vs AML (1-sided) p = ", kw_txt, "; two-group (1-sided) p = ", wx_txt, ". DNMT1 multi-lineage: ",
      if (!is.na(dnmt1_ni_p)) {
        paste0("NI vs AML (1-sided, DNMT1 ≤ AML) p = ", kw_dnmt1_txt)
      } else {
        paste0("Kruskal–Wallis p = ", kw_dnmt1_txt)
      },
      "; two-group Wilcoxon p = ", wx_dnmt1_txt, ").",
      if (nzchar(ni_aml_expr_note)) ni_aml_expr_note else ""
    )
  }
}
# Expose numeric p-values for wide-figure bracket annotations (Figures 1–2)
fig12_bcl2_ni_p <- if (exists("bcl2_ni_p")) bcl2_ni_p else NA_real_
fig12_bcl2_two_p <- if (exists("wx_vs_other") && !is.null(wx_vs_other)) wx_vs_other$p.value else NA_real_
fig12_dnmt1_ni_p <- if (exists("dnmt1_ni_p")) dnmt1_ni_p else NA_real_
fig12_dnmt1_two_p <- if (exists("wx_vs_other_dnmt1")) wx_vs_other_dnmt1$p.value else NA_real_
```

```{r ceres-similarity-panels, include = FALSE, cache = FALSE}
# BCL2 and DNMT1 CERES similarity panels (stem-like lineages + two-group) and broad DNMT1 CERES for Figure 2
lineage_plastic_ceres <- if (exists("lineage_plastic_ref_canonical")) lineage_plastic_ref_canonical else c("SCLC", "ExPulm-SCNEC", "AML (control)")
lbl_b2_ceres_two <- if (exists("lbl_b2_enriched_two")) lbl_b2_enriched_two else "Stem-like (SCLC, ExPulm-SCNEC, AML)"
lineage_bcl2_ceres_multi <- if (exists("lineage_bcl2_fig1_multi")) lineage_bcl2_fig1_multi else lineage_plastic_ceres
ni_aml_ceres_note <- ""
p_three_bcl2_ceres <- p_two_bcl2_ceres <- p_three_dnmt1_ceres <- p_two_dnmt1_ceres <- p_dnmt1_ceres <- NULL
dnmt1_ceres_plot_df <- NULL
# BCL2 CERES: build crispr_bcl2 (same as 08) then multi-lineage + two panels (aligned with BCL2 expression Figure 1)
if (exists("crispr") && nrow(crispr) > 0 && exists("target_ids") && length(target_ids) > 0 && exists("expr_target")) {
  crispr_bcl2 <- crispr %>%
    filter(grepl("^BCL2\\b", gene_name), depmap_id %in% target_ids) %>%
    select(depmap_id, cell_line, dependency) %>%
    left_join(expr_target %>% select(depmap_id, lineage_group, dplyr::any_of("lineage_expulm_detail")), by = "depmap_id") %>%
    filter(!is.na(lineage_group)) %>%
    mutate(is_nepc = (lineage_group == "ExPulm-SCNEC" & dplyr::coalesce(lineage_expulm_detail, "") == "NEPC"))
  lg0c <- as.character(crispr_bcl2$lineage_group)
  keep_loc_c <- .f1_keep_localized_subtype_chr(lg0c)
  drop_loc_c <- grepl("^Localized\\b", lg0c) & !keep_loc_c
  crispr_bcl2 <- dplyr::filter(crispr_bcl2, !drop_loc_c)
  crispr_bcl2 <- dplyr::filter(crispr_bcl2, !.f1_exclude_lineage_from_fig12_plots(as.character(.data$lineage_group)))
  if (nrow(crispr_bcl2) >= 2) {
    bcl2_three_ceres <- crispr_bcl2 %>%
      dplyr::filter(.data$lineage_group %in% lineage_bcl2_ceres_multi, is.finite(.data$dependency))
    bcl2_sim_two_ceres <- crispr_bcl2 %>%
      mutate(
        in_plastic_ceres = (.data$lineage_group %in% lineage_bcl2_ceres_multi) &
          !grepl("NSCLC", as.character(.data$lineage_group), fixed = TRUE),
        group_2 = if_else(in_plastic_ceres, lbl_b2_ceres_two, lbl_other_two)
      ) %>%
      dplyr::filter(is.finite(.data$dependency))
    kw_bcl2 <- tryCatch(kruskal.test(dependency ~ lineage_group, data = bcl2_three_ceres), error = function(e) list(p.value = NA_real_))
    wx_bcl2 <- tryCatch(wilcox.test(dependency ~ group_2, data = bcl2_sim_two_ceres), error = function(e) list(p.value = NA_real_))
    n_three_bc <- bcl2_three_ceres %>% group_by(lineage_group) %>% summarise(n = n(), .groups = "drop")
    lab_three_bc <- setNames(paste0(gsub(" \\(control\\)", "", as.character(n_three_bc$lineage_group)), " N=", n_three_bc$n), as.character(n_three_bc$lineage_group))
    lymph_lbl_c <- "DLBCL/MCL/CLL"
    u_c <- unique(as.character(bcl2_three_ceres$lineage_group))
    prio_c <- c("SCLC", "ExPulm-SCNEC", "AML (control)")
    if (lymph_lbl_c %in% u_c) prio_c <- c(prio_c, lymph_lbl_c)
    ord_c <- c(intersect(prio_c, u_c), setdiff(u_c, prio_c))
    bcl2_three_ceres <- bcl2_three_ceres %>% mutate(lineage_group = factor(as.character(lineage_group), levels = ord_c))
    ttl_ceres_multi <- if (length(ord_c) >= 4L) "BCL2 CERES (four lineages)" else "BCL2 CERES (three lineages)"
    aml_ceres_vals <- dplyr::filter(crispr_bcl2, .data$lineage_group == "AML (control)", is.finite(.data$dependency)) %>%
      dplyr::pull(.data$dependency)
    # Single pooled NI p: all non-AML lineages in Fig. 1 arm (or multi-panel lineages) vs AML; greater = stochastically more positive CERES (not more dependent)
    bcl2_ceres_ni_p <- NA_real_
    lg_pool_ceres <- if (exists("f1_enriched_lineages") && length(f1_enriched_lineages) > 0L) {
      as.character(f1_enriched_lineages)
    } else {
      as.character(lineage_bcl2_ceres_multi)
    }
    lg_pool_ceres <- lg_pool_ceres[!grepl("NSCLC", lg_pool_ceres, fixed = TRUE)]
    lg_pool_ceres <- lg_pool_ceres[!.f1_exclude_lineage_from_fig12_plots(lg_pool_ceres)]
    pool_bcl2_ceres <- crispr_bcl2 %>%
      dplyr::filter(as.character(.data$lineage_group) %in% lg_pool_ceres) %>%
      dplyr::filter(as.character(.data$lineage_group) != "AML (control)") %>%
      dplyr::filter(is.finite(.data$dependency))
    pbc <- stats::na.omit(pool_bcl2_ceres$dependency)
    abc <- stats::na.omit(aml_ceres_vals)
    if (length(pbc) >= 2L && length(abc) >= 2L) {
      wt_bcer <- tryCatch(
        stats::wilcox.test(pbc, abc, alternative = "greater", exact = FALSE),
        error = function(e) NULL
      )
      if (!is.null(wt_bcer) && !is.na(wt_bcer$p.value)) {
        bcl2_ceres_ni_p <- unname(wt_bcer$p.value)
      }
    }
    kw_lab_bcl2 <- if (!is.na(bcl2_ceres_ni_p)) {
      paste0(
        "NI vs AML (1-sided, less BCL2-dependent: combined non-AML lineages, CERES ≥ AML): p = ",
        ifelse(bcl2_ceres_ni_p < 0.001, "<0.001", format.pval(bcl2_ceres_ni_p, digits = 2))
      )
    } else if (is.na(kw_bcl2$p.value)) {
      "p = —"
    } else {
      paste0("p = ", ifelse(kw_bcl2$p.value < 0.001, "<0.001", format.pval(kw_bcl2$p.value, digits = 2)))
    }
    ni_sub_bcl2_ceres <- NULL
    if (!is.na(bcl2_ceres_ni_p)) {
      pln <- if (bcl2_ceres_ni_p < 0.001) "<0.001" else format.pval(bcl2_ceres_ni_p, digits = 2)
      ni_aml_ceres_note <- paste0(
        " BCL2 CERES vs AML (1-sided Wilcoxon, alternative: combined non-AML ≥ AML; more positive CERES = less BCL2-dependent): p = ",
        pln, "."
      )
    }
    p_three_bcl2_ceres <- ggplot(bcl2_three_ceres, aes(x = lineage_group, y = dependency, fill = lineage_group)) +
      geom_boxplot(alpha = 0.7, outlier.shape = NA) +
      geom_hline(yintercept = 0, linetype = 2) +
      geom_point(aes(color = is_nepc), position = position_jitter(width = 0.12, height = 0), alpha = 0.85, size = 2.5) +
      scale_color_manual(values = c("TRUE" = "#E41A1C", "FALSE" = "gray45"), guide = "none") +
      annotate("text", x = Inf, y = Inf, label = kw_lab_bcl2, hjust = 1.05, vjust = 1.5, size = 5) +
      scale_x_discrete(labels = lab_three_bc) +
      labs(x = "Lineage", y = "BCL2 CERES", title = ttl_ceres_multi, subtitle = ni_sub_bcl2_ceres) +
      theme_minimal() + .f1_fig12_axis_theme(25) + theme(legend.position = "none", plot.subtitle = element_text(size = 20, lineheight = 1.05))
    n_two_bc <- bcl2_sim_two_ceres %>% group_by(group_2) %>% summarise(n = n(), .groups = "drop")
    lab_two_bc <- setNames(paste0(as.character(n_two_bc$group_2), " N=", n_two_bc$n), as.character(n_two_bc$group_2))
    wx_lab_bcl2 <- if (is.na(wx_bcl2$p.value)) "p = —" else paste0("p = ", ifelse(wx_bcl2$p.value < 0.001, "<0.001", format.pval(wx_bcl2$p.value, digits = 2)))
    p_two_bcl2_ceres <- ggplot(bcl2_sim_two_ceres, aes(x = group_2, y = dependency, fill = group_2)) +
      geom_boxplot(alpha = 0.7, outlier.shape = NA) +
      geom_hline(yintercept = 0, linetype = 2) +
      geom_point(aes(color = is_nepc), position = position_jitter(width = 0.12, height = 0), alpha = 0.85, size = 2.5) +
      scale_color_manual(values = c("TRUE" = "#E41A1C", "FALSE" = "gray45"), guide = "none") +
      annotate("text", x = Inf, y = Inf, label = wx_lab_bcl2, hjust = 1.05, vjust = 1.5, size = 5) +
      scale_x_discrete(labels = lab_two_bc) +
      labs(x = "", y = "BCL2 CERES", title = "BCL2 CERES (two groups)") +
      theme_minimal() + .f1_fig12_axis_theme(15) + theme(legend.position = "none")
  }
}
fig12_bcl2_ceres_ni_p <- if (exists("bcl2_ceres_ni_p")) bcl2_ceres_ni_p else NA_real_
fig12_bcl2_ceres_two_p <- if (exists("wx_bcl2")) wx_bcl2$p.value else NA_real_
# DNMT1 CERES: three/two panels + broad by-lineage plot for Figure 2
if (exists("crispr_dnmt1") && nrow(crispr_dnmt1) > 0 && exists("expr_target")) {
  key_lineages_dnmt1 <- c("SCLC", "ExPulm-SCNEC", "AML (control)")
  tnbc_lymph_dnmt1_ceres <- c(
    if (exists("cohort_extended_hem") && isTRUE(cohort_extended_hem)) {
      c(
        if (exists("lineage_levels_tnbc")) {
          setdiff(lineage_levels_tnbc, "Unspecified TNBC")
        } else {
          c("Localized TNBC", "Metastatic TNBC")
        },
        "DLBCL/MCL/CLL"
      )
    } else {
      character(0)
    },
    "Neuroblastoma", "Unspecified Neuroblastoma"
  )
  tnbc_lymph_dnmt1_ceres <- setdiff(
    tnbc_lymph_dnmt1_ceres,
    c("Localized TNBC", "Localized Sarcoma", "Lymphoma (other)")
  )
  crispr_dnmt1 <- crispr_dnmt1 %>%
    left_join(expr_target %>% select(depmap_id, dplyr::any_of("lineage_expulm_detail")), by = "depmap_id") %>%
    mutate(is_nepc = (lineage_group == "ExPulm-SCNEC" & dplyr::coalesce(lineage_expulm_detail, "") == "NEPC"))
  lg0dn <- as.character(crispr_dnmt1$lineage_group)
  keep_loc_dn <- .f1_keep_localized_subtype_chr(lg0dn)
  drop_loc_dn <- grepl("^Localized\\b", lg0dn) & !keep_loc_dn
  crispr_dnmt1 <- dplyr::filter(crispr_dnmt1, !drop_loc_dn)
  crispr_dnmt1 <- dplyr::filter(crispr_dnmt1, !.f1_exclude_lineage_from_fig12_plots(as.character(.data$lineage_group)))
  lineage_dnmt1_ceres_multi <- if (exists("lineage_bcl2_ceres_multi")) lineage_bcl2_ceres_multi else lineage_plastic_ceres
  lbl_dnmt1_ceres_two <- if (exists("lbl_b2_enriched_two")) lbl_b2_enriched_two else "Stem-like (SCLC, ExPulm-SCNEC, AML)"
  lbl_dnmt1_ceres_other <- if (exists("lbl_other_two")) lbl_other_two else "Non-stem-like"
  dnmt1_three_ceres <- crispr_dnmt1 %>%
    dplyr::filter(.data$lineage_group %in% lineage_dnmt1_ceres_multi, is.finite(.data$dependency))
  dnmt1_sim_two_ceres <- crispr_dnmt1 %>%
    mutate(
      in_plastic_dnmt1_ceres = (.data$lineage_group %in% lineage_dnmt1_ceres_multi) &
        !grepl("NSCLC", as.character(.data$lineage_group), fixed = TRUE),
      group_2 = if_else(in_plastic_dnmt1_ceres, lbl_dnmt1_ceres_two, lbl_dnmt1_ceres_other)
    ) %>%
    dplyr::filter(is.finite(.data$dependency))
  if (nrow(dnmt1_three_ceres) >= 2) {
    kw_dnmt1 <- tryCatch(kruskal.test(dependency ~ lineage_group, data = dnmt1_three_ceres), error = function(e) list(p.value = NA_real_))
    wx_dnmt1 <- tryCatch(wilcox.test(dependency ~ group_2, data = dnmt1_sim_two_ceres), error = function(e) list(p.value = NA_real_))
    dnmt1_ceres_ni_p <- NA_real_
    if (exists("f1_enriched_lineages") && length(f1_enriched_lineages) > 0L) {
      pool_dc <- crispr_dnmt1 %>%
        dplyr::filter(as.character(.data$lineage_group) %in% as.character(f1_enriched_lineages)) %>%
        dplyr::filter(as.character(.data$lineage_group) != "AML (control)") %>%
        dplyr::filter(!grepl("NSCLC", as.character(.data$lineage_group), fixed = TRUE)) %>%
        dplyr::filter(is.finite(.data$dependency))
      aml_dc <- crispr_dnmt1 %>%
        dplyr::filter(.data$lineage_group == "AML (control)", is.finite(.data$dependency))
      pdc <- stats::na.omit(pool_dc$dependency)
      adc <- stats::na.omit(aml_dc$dependency)
      if (length(pdc) >= 2L && length(adc) >= 2L) {
        wt_dc <- tryCatch(
          stats::wilcox.test(pdc, adc, alternative = "greater", exact = FALSE),
          error = function(e) NULL
        )
        if (!is.null(wt_dc) && !is.na(wt_dc$p.value)) {
          dnmt1_ceres_ni_p <- unname(wt_dc$p.value)
        }
      }
    }
    kw_lab_dnmt1 <- if (!is.na(dnmt1_ceres_ni_p)) {
      paste0(
        "NI vs AML (1-sided, less DNMT1-dependent: combined non-AML lineages, CERES ≥ AML): p = ",
        ifelse(dnmt1_ceres_ni_p < 0.001, "<0.001", format.pval(dnmt1_ceres_ni_p, digits = 2))
      )
    } else if (is.na(kw_dnmt1$p.value)) {
      "p = —"
    } else {
      paste0("p = ", ifelse(kw_dnmt1$p.value < 0.001, "<0.001", format.pval(kw_dnmt1$p.value, digits = 2)))
    }
    n_three_dc <- dnmt1_three_ceres %>% group_by(lineage_group) %>% summarise(n = n(), .groups = "drop")
    lab_three_dc <- setNames(paste0(gsub(" \\(control\\)", "", as.character(n_three_dc$lineage_group)), " N=", n_three_dc$n), as.character(n_three_dc$lineage_group))
    p_three_dnmt1_ceres <- ggplot(dnmt1_three_ceres, aes(x = lineage_group, y = dependency, fill = lineage_group)) +
      geom_boxplot(alpha = 0.7, outlier.shape = NA) +
      geom_hline(yintercept = 0, linetype = 2) +
      geom_point(aes(color = is_nepc), position = position_jitter(width = 0.12, height = 0), alpha = 0.85, size = 2.5) +
      scale_color_manual(values = c("TRUE" = "#E41A1C", "FALSE" = "gray45"), guide = "none") +
      annotate("text", x = Inf, y = Inf, label = kw_lab_dnmt1, hjust = 1.05, vjust = 1.5, size = 5) +
      scale_x_discrete(labels = lab_three_dc) +
      labs(
        x = "Lineage", y = "DNMT1 CERES",
        title = if (length(unique(dnmt1_three_ceres$lineage_group)) >= 4L) {
          "DNMT1 CERES (four lineages)"
        } else {
          "DNMT1 CERES (three lineages)"
        }
      ) +
      theme_minimal() + .f1_fig12_axis_theme(25) + theme(legend.position = "none")
    n_two_dc <- dnmt1_sim_two_ceres %>% group_by(group_2) %>% summarise(n = n(), .groups = "drop")
    lab_two_dc <- setNames(paste0(as.character(n_two_dc$group_2), " N=", n_two_dc$n), as.character(n_two_dc$group_2))
    wx_lab_dnmt1 <- if (is.na(wx_dnmt1$p.value)) "p = —" else paste0("p = ", ifelse(wx_dnmt1$p.value < 0.001, "<0.001", format.pval(wx_dnmt1$p.value, digits = 2)))
    p_two_dnmt1_ceres <- ggplot(dnmt1_sim_two_ceres, aes(x = group_2, y = dependency, fill = group_2)) +
      geom_boxplot(alpha = 0.7, outlier.shape = NA) +
      geom_hline(yintercept = 0, linetype = 2) +
      geom_point(aes(color = is_nepc), position = position_jitter(width = 0.12, height = 0), alpha = 0.85, size = 2.5) +
      scale_color_manual(values = c("TRUE" = "#E41A1C", "FALSE" = "gray45"), guide = "none") +
      annotate("text", x = Inf, y = Inf, label = wx_lab_dnmt1, hjust = 1.05, vjust = 1.5, size = 5) +
      scale_x_discrete(labels = lab_two_dc) +
      labs(x = "", y = "DNMT1 CERES", title = "DNMT1 CERES (two groups)") +
      theme_minimal() + .f1_fig12_axis_theme(15) + theme(legend.position = "none")
  }
  # Broad DNMT1 CERES by lineage (for Figure 2 bottom-left)
  dnmt1_crispr_n <- crispr_dnmt1 %>% group_by(lineage_group) %>% summarise(n = n(), .groups = "drop")
  dnmt1_keep <- dnmt1_crispr_n %>%
    dplyr::filter(
      n >= 5 |
        .data$lineage_group == "ExPulm-SCNEC" |
        as.character(.data$lineage_group) %in% tnbc_lymph_dnmt1_ceres
    ) %>%
    dplyr::pull(.data$lineage_group)
  if (exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df) && nrow(bcl2_plot_df) > 0L && "subtype" %in% names(bcl2_plot_df)) {
    axis_sub_d <- unique(as.character(bcl2_plot_df$subtype))
    in_dn <- unique(as.character(crispr_dnmt1$lineage_group))
    dnmt1_keep <- union(dnmt1_keep, intersect(axis_sub_d, in_dn))
  }
  # #region agent log
  tryCatch({
    .agent_dbg_ndjson(
      runId = "post-fix",
      hypothesisId = "H4",
      location = "03_Rmd:dnmt1_keep",
      message = "DNMT1 keep set and neuroblastoma availability pre-join",
      data = list(
        dnmt1_keep = as.character(dnmt1_keep),
        in_dn_nb = sum(grepl("Neuroblastoma", as.character(crispr_dnmt1$lineage_group), fixed = TRUE))
      )
    )
  }, error = function(e) invisible(NULL))
  # #endregion
  crispr_dnmt1_plot <- crispr_dnmt1 %>% dplyr::filter(lineage_group %in% dnmt1_keep)
  p_dnmt1_ceres <- NULL
  if (nrow(crispr_dnmt1_plot) > 0) {
    if (exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df) &&
          nrow(bcl2_plot_df) > 0L &&
          all(c("depmap_id", "lab", "subtype") %in% names(bcl2_plot_df))) {
      axis_dn <- bcl2_plot_df %>%
        dplyr::distinct(.data$depmap_id, .keep_all = TRUE) %>%
        dplyr::select(.data$depmap_id, .data$lab, .data$subtype, .data$is_nepc)
      crispr_d_dep <- crispr_dnmt1_plot %>%
        dplyr::select(.data$depmap_id, .data$dependency)
      ddj <- axis_dn %>%
        dplyr::left_join(crispr_d_dep, by = "depmap_id")
      if (nrow(ddj) >= 2L) {
        lev_dn <- levels(bcl2_plot_df$lab)
        ddj <- ddj %>%
          dplyr::filter(!is.na(.data$lab), nzchar(trimws(as.character(.data$lab))))
        ddj$lab <- factor(as.character(ddj$lab), levels = lev_dn)
        ddj <- ddj %>% dplyr::filter(!is.na(.data$lab))
        ddj <- dplyr::filter(ddj, is.finite(.data$dependency))
        # #region agent log
        tryCatch({
          nb_ddj <- ddj %>% dplyr::filter(grepl("Neuroblastoma", as.character(.data$lab), fixed = TRUE))
          .agent_dbg_ndjson(
            runId = "post-fix",
            hypothesisId = "H5",
            location = "03_Rmd:dnmt1_ddj_finite",
            message = "DNMT1 wide CERES rows after finite filter",
            data = list(
              n_rows = nrow(ddj),
              n_nb_rows = nrow(nb_ddj),
              nb_labs = unique(as.character(nb_ddj$lab))
            )
          )
        }, error = function(e) invisible(NULL))
        # #endregion
        n_dnmt1_ceres_by_lab <- setNames(rep(0L, length(lev_dn)), lev_dn)
        if (nrow(ddj) > 0L && "lab" %in% names(ddj)) {
          cn_d1 <- ddj %>% dplyr::count(.data$lab, name = "nn")
          for (ri in seq_len(nrow(cn_d1))) {
            k <- as.character(cn_d1$lab[[ri]])
            if (k %in% names(n_dnmt1_ceres_by_lab)) {
              n_dnmt1_ceres_by_lab[[k]] <- as.integer(cn_d1$nn[[ri]])
            }
          }
        }
        f1_dnmt1_ceres_labels_fun <- .f1_ceres_wide_x_labeller(n_dnmt1_ceres_by_lab)
        if (nrow(ddj) >= 2L) {
          f1_dnmt1_ceres_ylim <- local({
            yv <- ddj$dependency
            yv <- yv[is.finite(yv)]
            yn <- ddj$dependency[ddj$is_nepc %in% TRUE]
            yn <- yn[is.finite(yn)]
            if (length(yv) < 2L) {
              # Still define limits so y=0 reference and any finite NEPC overlay are visible.
              pad0 <- 0.14
              if (length(yn) >= 1L) {
                y_lo0 <- min(min(yn) - pad0, -pad0)
                y_hi0 <- max(max(yn) + pad0, pad0)
                return(c(max(-4.8, y_lo0), min(1.2, y_hi0)))
              }
              return(c(-0.35, 0.35))
            }
            qs <- stats::quantile(yv, c(0.04, 0.96), names = FALSE, na.rm = TRUE)
            d <- diff(qs)
            if (!is.finite(d) || d < 0.18) d <- max(0.42, diff(range(yv)))
            pad <- max(0.07, 0.14 * d)
            y_lo <- max(-3.15, qs[[1L]] - pad)
            y_hi <- min(0.62, qs[[2L]] + pad)
            # Always include finite NEPC overlay CERES so coord_cartesian does not clip the red points.
            if (length(yn) > 0L) {
              rng_n <- if (length(yn) >= 2L) diff(range(yn)) else 0
              pad_n <- max(0.08, 0.12 * max(rng_n, max(abs(yn), na.rm = TRUE) * 0.12, 0.2))
              y_lo <- min(y_lo, min(yn) - pad_n)
              y_hi <- max(y_hi, max(yn) + pad_n)
              y_lo <- max(-4.6, y_lo)
              y_hi <- min(1.15, y_hi)
            }
            # CERES panels: always span y=0 (dashed reference).
            pad0 <- 0.12
            y_lo <- min(y_lo, -pad0)
            y_hi <- max(y_hi, pad0)
            y_lo <- max(-4.85, y_lo)
            y_hi <- min(1.25, y_hi)
            if (!is.finite(y_hi) || y_hi <= y_lo) {
              return(NULL)
            }
            c(y_lo, y_hi)
          })
          kw_dnmt1_broad <- tryCatch(
            stats::kruskal.test(dependency ~ lab, data = ddj),
            error = function(e) list(p.value = NA_real_)
          )
          p_kw_dnmt1 <- if (!is.null(kw_dnmt1_broad$p.value) && !is.na(kw_dnmt1_broad$p.value)) {
            format.pval(kw_dnmt1_broad$p.value, digits = 2)
          } else {
            "—"
          }
          vx_ddj <- if (exists("f1_vline_x")) f1_vline_x else NA_real_
          p_dnmt1_ceres <- ggplot(ddj, aes(x = lab, y = dependency, fill = subtype)) +
            geom_boxplot(alpha = 0.7, outlier.shape = NA, na.rm = TRUE) +
            geom_hline(yintercept = 0, linetype = 2) +
            ggplot2::geom_point(
              data = dplyr::filter(ddj, .data$is_nepc %in% TRUE),
              ggplot2::aes(x = lab, y = dependency),
              inherit.aes = FALSE,
              color = "#E41A1C",
              position = ggplot2::position_jitter(width = 0.12, height = 0),
              alpha = 0.92,
              size = 2.5,
              na.rm = TRUE
            )
          if (!is.na(vx_ddj)) {
            p_dnmt1_ceres <- p_dnmt1_ceres + ggplot2::geom_vline(
              xintercept = vx_ddj,
              linetype = "dashed",
              linewidth = f1_rule_line_linewidth,
              alpha = 1,
              colour = f1_rule_line_colour
            )
          }
          p_dnmt1_ceres <- p_dnmt1_ceres +
            ggplot2::scale_x_discrete(drop = FALSE, labels = f1_dnmt1_ceres_labels_fun) +
            labs(x = "Lineage", y = "DNMT1 CERES") +
            theme_minimal() +
            .f1_fig12_axis_theme(45) +
            theme(legend.position = "none")
          dnmt1_ceres_plot_df <- ddj
        }
      }
    }
    if (is.null(p_dnmt1_ceres)) {
      u_d <- unique(as.character(crispr_dnmt1_plot$lineage_group))
      prio_d <- c(key_lineages_dnmt1, tnbc_lymph_dnmt1_ceres)
      ord_d <- c(intersect(prio_d, u_d), setdiff(u_d, prio_d))
      crispr_dnmt1_plot <- crispr_dnmt1_plot %>%
        mutate(lineage_group = factor(as.character(lineage_group), levels = ord_d))
      crispr_dnmt1_plot_fin <- dplyr::filter(crispr_dnmt1_plot, is.finite(.data$dependency))
      kw_dnmt1_broad <- tryCatch(kruskal.test(dependency ~ lineage_group, data = crispr_dnmt1_plot_fin), error = function(e) list(p.value = NA))
      p_kw_dnmt1 <- if (!is.null(kw_dnmt1_broad$p.value)) format.pval(kw_dnmt1_broad$p.value, digits = 2) else "—"
      lbl_dnmt1 <- crispr_dnmt1_plot_fin %>% group_by(lineage_group) %>% summarise(n = n(), .groups = "drop")
      lab_dnmt1_ceres <- setNames(paste0(gsub(" \\(control\\)", "", as.character(lbl_dnmt1$lineage_group)), " N=", lbl_dnmt1$n), as.character(lbl_dnmt1$lineage_group))
      p_dnmt1_ceres <- ggplot(crispr_dnmt1_plot_fin, aes(x = lineage_group, y = dependency, fill = lineage_group)) +
        geom_boxplot(alpha = 0.7, outlier.shape = NA) +
        geom_hline(yintercept = 0, linetype = 2) +
        geom_point(aes(color = is_nepc), position = position_jitter(width = 0.12, height = 0), alpha = 0.85, size = 2.5) +
        scale_color_manual(values = c("TRUE" = "#E41A1C", "FALSE" = "gray45"), guide = "none") +
        scale_x_discrete(labels = lab_dnmt1_ceres) +
        labs(x = "Lineage", y = "DNMT1 CERES") +
        theme_minimal() + .f1_fig12_axis_theme(45) + theme(legend.position = "none")
      dnmt1_ceres_plot_df <- dplyr::transmute(
        crispr_dnmt1_plot_fin,
        lab = as.character(.data$lineage_group),
        subtype = as.character(.data$lineage_group),
        dependency = .data$dependency
      )
      vx_dfb <- if (exists("f1_vline_x")) f1_vline_x else NA_real_
      if (!is.na(vx_dfb) && exists("bcl2_plot_df") && nrow(bcl2_plot_df) > 0L &&
            length(ord_d) == length(levels(bcl2_plot_df$lab))) {
        p_dnmt1_ceres <- p_dnmt1_ceres + ggplot2::geom_vline(
          xintercept = vx_dfb,
          linetype = "dashed",
          linewidth = f1_rule_line_linewidth,
          alpha = 1,
          colour = f1_rule_line_colour
        )
      }
    }
  }
}
fig12_dnmt1_ceres_ni_p <- if (exists("dnmt1_ceres_ni_p")) dnmt1_ceres_ni_p else NA_real_
fig12_dnmt1_ceres_two_p <- if (exists("wx_dnmt1")) wx_dnmt1$p.value else NA_real_
if (exists("cap_bcl2_sim") && exists("ni_aml_ceres_note") && nzchar(ni_aml_ceres_note)) {
  cap_bcl2_sim <- paste0(cap_bcl2_sim, ni_aml_ceres_note)
}
```

```{r tnbc-master-lineage-def-t2-then-figures, echo = FALSE, results = "asis", eval = exists("cohort_extended_hem") && isTRUE(cohort_extended_hem) && exists("hma_bcl2_tnbc_lymph_master") && isTRUE(hma_bcl2_tnbc_lymph_master)}
knitr::asis_output(paste0(
  "### Lymphoma strata: BCL2-treated / stem-like proxy\n\n",
  "Lymphoma lines are classified from DepMap disease, subtype, and molecular text (not treatment exposure). ",
  "DLBCL/MCL/CLL includes chronic lymphocytic leukemia / small lymphocytic lymphoma (CLL/SLL; venetoclax FDA-indication proxy) and B-NHL labels commonly studied with BCL2-pathway therapy in trials ",
  "(e.g. mantle cell, follicular, DLBCL and related high-grade B-cell patterns, marginal-zone/MALT, Waldenström/lymphoplasmacytic; full regex patterns are in DNMTi_BCL2_Child_Expanded.Rmd). ",
  "Lymphoma (other) pools remaining lymphoid lines in the extended cohort that do not match those patterns.\n\n",
  "TNBC: localized and metastatic strata appear on the same axis when present. Table 1 omits lineages with \"Unspecified\" in the name; unspecified TNBC is omitted from the cohort summary table below. ",
  "TNBC and lymphoma buckets (DLBCL/MCL/CLL vs Lymphoma (other)) appear on the x-axis in Figures 1–2 when present.\n\n"
))
lg_tnbc_summ <- if (exists("lineage_levels_tnbc")) {
  setdiff(lineage_levels_tnbc, "Unspecified TNBC")
} else {
  c("Localized TNBC", "Metastatic TNBC")
}
lg_lymph_summ <- c("DLBCL/MCL/CLL", "Lymphoma (other)")
if (exists("expr_target") && nrow(expr_target) > 0L && "lineage_group" %in% names(expr_target)) {
  summ_tl <- expr_target %>%
    dplyr::filter(as.character(.data$lineage_group) %in% c(lg_tnbc_summ, lg_lymph_summ)) %>%
    dplyr::count(.data$lineage_group, name = "N_cell_lines") %>%
    dplyr::arrange(.data$lineage_group)
  if (nrow(summ_tl) > 0L) {
    print(
      knitr::kable(
        summ_tl,
        caption = "Cohort: TNBC and lymphoma strata (N cell lines in analysis cohort, same as Table 1 / Figures 1–2). Lymphoma buckets are histology-based proxies (not treatment exposure)."
      ) %>%
        kableExtra::kable_styling(bootstrap_options = c("bordered", "condensed"), full_width = FALSE)
    )
  } else {
    knitr::asis_output("*No TNBC or lymphoma lines in the extended cohort (check 01_Data_Loading_TNBC_Lymphoma / verification RDS).*\n\n")
  }
} else {
  knitr::asis_output("*expr_target not available for TNBC/lymphoma summary.*\n\n")
}
if (exists("t2_kable")) {
  print(t2_kable)
  cat("\n\n")
}
```

```{r bcl2-fig-table-dependency, cache = FALSE, fig.show = "hide"}
# Build BCL2 dependency plot (CERES) for Figure 1 composite
p_bcl2_ceres <- NULL
if (exists("crispr") && nrow(crispr) > 0 && exists("target_ids") && length(target_ids) > 0 && exists("expr_target")) {
  crispr_bcl2_f1 <- crispr %>%
    filter(grepl("^BCL2\\b", gene_name), depmap_id %in% target_ids) %>%
    select(depmap_id, dependency) %>%
    left_join(expr_target %>% select(depmap_id, lineage_group, dplyr::any_of("lineage_expulm_detail")), by = "depmap_id") %>%
    filter(!is.na(lineage_group)) %>%
    mutate(is_nepc = (lineage_group == "ExPulm-SCNEC" & dplyr::coalesce(lineage_expulm_detail, "") == "NEPC"))
  crispr_bcl2_f1 <- crispr_bcl2_f1 %>%
    dplyr::filter(
      !grepl("^Localized\\b", as.character(.data$lineage_group)) |
        .f1_keep_localized_subtype_chr(as.character(.data$lineage_group))
    )
  crispr_bcl2_f1 <- dplyr::filter(crispr_bcl2_f1, !.f1_exclude_lineage_from_fig12_plots(as.character(.data$lineage_group)))
  if (nrow(crispr_bcl2_f1) > 0) {
    # Lymphoma + neuroblastoma: always allow in wide CERES (n<5 anchor), like expression panel
    tnbc_lymph_ceres <- c(
      if (exists("cohort_extended_hem") && isTRUE(cohort_extended_hem)) {
        if (exists("lineage_levels_tnbc")) {
          setdiff(lineage_levels_tnbc, "Unspecified TNBC")
        } else {
          c("Localized TNBC", "Metastatic TNBC")
        }
      } else {
        character(0)
      },
      "DLBCL/MCL/CLL",
      "Neuroblastoma", "Unspecified Neuroblastoma",
      "Localized Sarcoma", "Metastatic Sarcoma", "Unspecified Sarcoma"
    )
    tnbc_lymph_ceres <- setdiff(
      tnbc_lymph_ceres,
      c("Localized TNBC", "Localized Sarcoma", "Lymphoma (other)")
    )
    c_n <- crispr_bcl2_f1 %>% group_by(lineage_group) %>% summarise(n = n(), .groups = "drop")
    keep_l <- c_n %>%
      dplyr::filter(
        n >= 5 |
          as.character(.data$lineage_group) == "ExPulm-SCNEC" |
          as.character(.data$lineage_group) %in% tnbc_lymph_ceres
      ) %>%
      dplyr::pull(.data$lineage_group)
    # Keep CRISPR rows for any lineage present on the Figure 1 expression axis (avoids empty CERES slots when n < 5).
    if (exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df) && nrow(bcl2_plot_df) > 0L && "subtype" %in% names(bcl2_plot_df)) {
      axis_sub <- unique(as.character(bcl2_plot_df$subtype))
      in_crispr <- unique(as.character(crispr_bcl2_f1$lineage_group))
      keep_l <- union(keep_l, intersect(axis_sub, in_crispr))
    }
    # #region agent log
    tryCatch({
      .agent_dbg_ndjson(
        runId = "post-fix",
        hypothesisId = "H2",
        location = "03_Rmd:bcl2_keep_l",
        message = "BCL2 CERES lineage keep set and neuroblastoma availability",
        data = list(
          keep_l = as.character(keep_l),
          in_crispr_nb = sum(grepl("Neuroblastoma", as.character(crispr_bcl2_f1$lineage_group), fixed = TRUE)),
          axis_nb_lab = unique(as.character(bcl2_plot_df$lab[grepl("Neuroblastoma", as.character(bcl2_plot_df$lab), fixed = TRUE)]))
        )
      )
    }, error = function(e) invisible(NULL))
    # #endregion
    crispr_bcl2_f1 <- crispr_bcl2_f1 %>% filter(lineage_group %in% keep_l)
    if (nrow(crispr_bcl2_f1) > 0) {
      use_bcl2_axis <- exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df) &&
        nrow(bcl2_plot_df) > 0L &&
        all(c("lab", "subtype", "depmap_id") %in% names(bcl2_plot_df))
      if (use_bcl2_axis) {
        bcl2_axis_c <- bcl2_plot_df %>%
          dplyr::distinct(.data$depmap_id, .keep_all = TRUE) %>%
          dplyr::select(.data$depmap_id, .data$lab, .data$subtype, .data$is_nepc)
        crispr_dep <- crispr_bcl2_f1 %>%
          dplyr::select(.data$depmap_id, .data$dependency)
        # Same cell lines as expression (left join) so every lineage slot can draw CERES when values exist.
        ceres_f1_df <- bcl2_axis_c %>%
          dplyr::left_join(crispr_dep, by = "depmap_id")
        if (nrow(ceres_f1_df) >= 2L) {
          ceres_f1_df <- ceres_f1_df %>% dplyr::filter(!is.na(.data$lab), nzchar(trimws(as.character(.data$lab))))
          lev_lab <- levels(bcl2_plot_df$lab)
          if (length(lev_lab) > 0L) {
            ceres_f1_df$lab <- factor(as.character(ceres_f1_df$lab), levels = lev_lab)
          }
          ceres_f1_df <- dplyr::filter(ceres_f1_df, is.finite(.data$dependency))
          # #region agent log
          tryCatch({
            nb_ceres_rows <- ceres_f1_df %>% dplyr::filter(grepl("Neuroblastoma", as.character(.data$lab), fixed = TRUE))
            .agent_dbg_ndjson(
              runId = "post-fix",
              hypothesisId = "H3",
              location = "03_Rmd:bcl2_ceres_f1_df",
              message = "BCL2 wide CERES rows after finite filter",
              data = list(
                n_rows = nrow(ceres_f1_df),
                n_nb_rows = nrow(nb_ceres_rows),
                nb_labs = unique(as.character(nb_ceres_rows$lab))
              )
            )
          }, error = function(e) invisible(NULL))
          # #endregion
          n_ceres_f1_by_lab <- setNames(rep(0L, length(lev_lab)), lev_lab)
          if (nrow(ceres_f1_df) > 0L && "lab" %in% names(ceres_f1_df)) {
            cn_b1 <- ceres_f1_df %>% dplyr::count(.data$lab, name = "nn")
            for (ri in seq_len(nrow(cn_b1))) {
              k <- as.character(cn_b1$lab[[ri]])
              if (k %in% names(n_ceres_f1_by_lab)) {
                n_ceres_f1_by_lab[[k]] <- as.integer(cn_b1$nn[[ri]])
              }
            }
          }
          f1_bcl2_ceres_labels_fun <- .f1_ceres_wide_x_labeller(n_ceres_f1_by_lab)
          if (nrow(ceres_f1_df) >= 2L) {
            # Top-row BCL2 CERES: slightly tighter y than default limits so contrasts read larger (still shows ~0 reference).
            f1_bcl2_ceres_ylim <- local({
              yv <- ceres_f1_df$dependency
              yv <- yv[is.finite(yv)]
              yn <- ceres_f1_df$dependency[ceres_f1_df$is_nepc %in% TRUE]
              yn <- yn[is.finite(yn)]
              if (length(yv) < 2L) {
                pad0 <- 0.14
                if (length(yn) >= 1L) {
                  y_lo0 <- min(min(yn) - pad0, -pad0)
                  y_hi0 <- max(max(yn) + pad0, pad0)
                  return(c(max(-4.8, y_lo0), min(1.2, y_hi0)))
                }
                return(c(-0.35, 0.35))
              }
              qs <- stats::quantile(yv, c(0.04, 0.96), names = FALSE, na.rm = TRUE)
              d <- diff(qs)
              if (!is.finite(d) || d < 0.18) d <- max(0.42, diff(range(yv)))
              pad <- max(0.07, 0.14 * d)
              y_lo <- max(-3.15, qs[[1L]] - pad)
              y_hi <- min(0.62, qs[[2L]] + pad)
              if (length(yn) > 0L) {
                rng_n <- if (length(yn) >= 2L) diff(range(yn)) else 0
                pad_n <- max(0.08, 0.12 * max(rng_n, max(abs(yn), na.rm = TRUE) * 0.12, 0.2))
                y_lo <- min(y_lo, min(yn) - pad_n)
                y_hi <- max(y_hi, max(yn) + pad_n)
                y_lo <- max(-4.6, y_lo)
                y_hi <- min(1.15, y_hi)
              }
              pad0 <- 0.12
              y_lo <- min(y_lo, -pad0)
              y_hi <- max(y_hi, pad0)
              y_lo <- max(-4.85, y_lo)
              y_hi <- min(1.25, y_hi)
              if (!is.finite(y_hi) || y_hi <= y_lo) {
                return(NULL)
              }
              c(y_lo, y_hi)
            })
            .mk_bcl2_ceres_wide <- function(dat, vline_x = NA_real_, labels_fun = .f1_axis_short_met_label) {
              g <- ggplot(dat, aes(x = lab, y = dependency, fill = subtype)) +
                geom_boxplot(alpha = 0.7, outlier.shape = NA, na.rm = TRUE) +
                geom_hline(yintercept = 0, linetype = 2) +
                ggplot2::geom_point(
                  data = dplyr::filter(dat, .data$is_nepc %in% TRUE),
                  ggplot2::aes(x = lab, y = dependency),
                  inherit.aes = FALSE,
                  color = "#E41A1C",
                  position = ggplot2::position_jitter(width = 0.12, height = 0),
                  alpha = 0.92,
                  size = 2.5,
                  na.rm = TRUE
                ) +
                ggplot2::scale_x_discrete(drop = FALSE, labels = labels_fun) +
                labs(x = "Lineage", y = "BCL2 CERES") +
                theme_minimal() +
                .f1_fig12_axis_theme(45) +
                theme(legend.position = "none")
              if (!is.na(vline_x)) {
                g <- g + ggplot2::geom_vline(
                  xintercept = vline_x,
                  linetype = "dashed",
                  linewidth = f1_rule_line_linewidth,
                  alpha = 1,
                  colour = f1_rule_line_colour
                )
              }
              g
            }
            vx_ceres <- if (exists("f1_vline_x")) f1_vline_x else NA_real_
            p_bcl2_ceres <- .mk_bcl2_ceres_wide(ceres_f1_df, vline_x = vx_ceres, labels_fun = f1_bcl2_ceres_labels_fun)
            bcl2_ceres_plot_df <- ceres_f1_df
          }
        }
      }
      if (is.null(p_bcl2_ceres)) {
        key_l <- c("SCLC", "ExPulm-SCNEC", "AML (control)")
        u_lg <- unique(as.character(crispr_bcl2_f1$lineage_group))
        prio_lg <- c(key_l, tnbc_lymph_ceres)
        ord_lg <- c(intersect(prio_lg, u_lg), setdiff(u_lg, prio_lg))
        crispr_bcl2_f1 <- crispr_bcl2_f1 %>%
          mutate(lineage_group = factor(as.character(lineage_group), levels = ord_lg))
        crispr_bcl2_f1 <- dplyr::filter(crispr_bcl2_f1, is.finite(.data$dependency))
        if (nrow(crispr_bcl2_f1) >= 2L) {
          lbl_bcl2 <- crispr_bcl2_f1 %>% group_by(lineage_group) %>% summarise(n = n(), .groups = "drop")
          lab_bcl2_ceres <- setNames(paste0(gsub(" \\(control\\)", "", as.character(lbl_bcl2$lineage_group)), " N=", lbl_bcl2$n), as.character(lbl_bcl2$lineage_group))
          p_bcl2_ceres <- ggplot(crispr_bcl2_f1, aes(x = lineage_group, y = dependency, fill = lineage_group)) +
            geom_boxplot(alpha = 0.7, outlier.shape = NA) +
            geom_hline(yintercept = 0, linetype = 2) +
            geom_point(aes(color = is_nepc), position = position_jitter(width = 0.12, height = 0), alpha = 0.85, size = 2.5) +
            scale_color_manual(values = c("TRUE" = "#E41A1C", "FALSE" = "gray45"), guide = "none") +
            scale_x_discrete(labels = lab_bcl2_ceres) +
            labs(x = "Lineage", y = "BCL2 CERES") +
            theme_minimal() + .f1_fig12_axis_theme(45) + theme(legend.position = "none")
          vx_fb <- if (exists("f1_vline_x")) f1_vline_x else NA_real_
          if (!is.na(vx_fb) && exists("bcl2_plot_df") && nrow(bcl2_plot_df) > 0L &&
                length(ord_lg) == length(levels(bcl2_plot_df$lab))) {
            p_bcl2_ceres <- p_bcl2_ceres + ggplot2::geom_vline(
              xintercept = vx_fb,
              linetype = "dashed",
              linewidth = f1_rule_line_linewidth,
              alpha = 1,
              colour = f1_rule_line_colour
            )
          }
          bcl2_ceres_plot_df <- dplyr::transmute(
            crispr_bcl2_f1,
            lab = as.character(.data$lineage_group),
            subtype = as.character(.data$lineage_group),
            dependency = .data$dependency
          )
        }
      }
    }
  }
}
.fig12_signif_brackets_wide <- function(p, y_num, ni_p, two_p, vline_x, lv_chr, aml_subtype = "AML (control)", plot_df, ylim = NULL) {
  if (is.null(p)) return(p)
  if (!is.null(ylim) && length(ylim) == 2L && all(is.finite(ylim))) {
    ymin <- ylim[[1L]]
    ymax <- ylim[[2L]]
  } else {
    ymax <- suppressWarnings(max(y_num, na.rm = TRUE))
    ymin <- suppressWarnings(min(y_num, na.rm = TRUE))
  }
  if (!is.finite(ymax) || !is.finite(ymin)) return(p)
  # Anchor brackets to observed y (box tops), not full ylim span — avoids "low" brackets when ylim has headroom (e.g. BCL2 expression cap).
  y_num_f <- suppressWarnings(as.numeric(y_num))
  y_num_f <- y_num_f[is.finite(y_num_f)]
  y_dat_max <- if (length(y_num_f)) max(y_num_f) else ymax
  y_dat_min <- if (length(y_num_f)) min(y_num_f) else ymin
  y_dat_max <- min(y_dat_max, ymax)
  y_dat_min <- max(y_dat_min, ymin)
  d_win <- ymax - ymin
  d_dat <- y_dat_max - y_dat_min
  if (!is.finite(d_dat) || d_dat < 1e-6 * d_win) d_dat <- 0.06 * d_win
  tick_h <- max(0.014 * d_win, 0.03 * d_dat)
  lab_dy <- max(0.02 * d_win, 0.028 * d_dat)
  # NI tier just above data; stem-like tier stacked above NI (less total height than two independent offsets from ymax).
  y_hi <- y_dat_max + 0.08 * d_dat + 0.025 * d_win
  gap_tiers <- max(0.035 * d_win, 0.05 * d_dat, tick_h + lab_dy * 0.35)
  y_top <- y_hi + lab_dy + gap_tiers + 0.02 * d_dat
  fmt_p <- function(pv) {
    if (length(pv) != 1L || is.na(pv)) return("—")
    if (pv < 0.001) return("<0.001")
    format.pval(pv, digits = 2)
  }
  sig_sfx <- function(pv) {
    if (length(pv) != 1L || is.na(pv)) return("")
    if (pv < 0.01) return("**")
    if (pv < 0.05) return("*")
    ""
  }
  lv <- lv_chr
  if (length(lv) < 1L) return(p)
  aml_lab_chr <- dplyr::filter(plot_df, as.character(.data$subtype) == aml_subtype) %>%
    dplyr::distinct(.data$lab) %>%
    dplyr::pull(.data$lab) %>%
    as.character()
  aml_lab <- intersect(aml_lab_chr, lv)
  if (length(aml_lab) < 1L) {
    aml_guess <- lv[vapply(lv, function(lb) grepl("^AML\\b", lb), logical(1L))]
    aml_lab <- head(aml_guess, 1L)
  }
  if (length(aml_lab) < 1L) return(p)
  aml_lab <- aml_lab[[1]]
  aml_i <- match(aml_lab, lv)
  if (is.na(aml_i)) return(p)
  # NI-vs-AML bracket should stop strictly left of the red divider (do not include first box to the right, e.g., NSCLC).
  left_i <- if (!is.na(vline_x)) seq_len(max(0L, floor(vline_x))) else integer(0)
  left_i <- setdiff(left_i, aml_i)
  left_i <- left_i[is.finite(left_i) & left_i >= 1L]
  # Top margin: two p-value tiers + large annotate() mm need generous headroom (patchwork row theme must match).
  two_tier <- !is.na(two_p) && !is.na(vline_x) && length(lv_chr) >= 1L
  bracket_plot_margin_top <- as.numeric(if (isTRUE(two_tier)) {
    get0(".f1_fig12_bracket_margin_top_2tier", ifnotfound = 120L, inherits = TRUE)
  } else {
    get0(".f1_fig12_bracket_margin_top_1tier", ifnotfound = 58L, inherits = TRUE)
  })
  # Right margin: keep consistent across all four Fig.1 patchwork panels so column panels align (annotate p-values sit in the right gutter).
  bracket_plot_margin_r <- 36
  if (!is.null(ylim) && length(ylim) == 2L && all(is.finite(ylim))) {
    p2 <- p + ggplot2::coord_cartesian(ylim = ylim, clip = "off") +
      ggplot2::theme(plot.margin = ggplot2::margin(t = bracket_plot_margin_top, r = bracket_plot_margin_r, b = 6, l = 6))
  } else {
    p2 <- p + ggplot2::coord_cartesian(clip = "off") +
      ggplot2::theme(plot.margin = ggplot2::margin(t = bracket_plot_margin_top, r = bracket_plot_margin_r, b = 6, l = 6))
  }
  add_bracket_segments <- function(p_in, x0, x1, y0, y1) {
    p_in +
      ggplot2::annotate(
        "segment",
        x = x0, xend = x1, y = y0, yend = y0,
        linewidth = 0.35,
        color = "gray20"
      ) +
      ggplot2::annotate(
        "segment",
        x = x0, xend = x0, y = y0, yend = y1,
        linewidth = 0.35,
        color = "gray20"
      ) +
      ggplot2::annotate(
        "segment",
        x = x1, xend = x1, y = y0, yend = y1,
        linewidth = 0.35,
        color = "gray20"
      )
  }
  xc <- function(i) mean(c(i - 0.25, i + 0.25))
  br_mm <- if (exists(".f1_fig12_bracket_label_mm", inherits = TRUE)) get(".f1_fig12_bracket_label_mm", inherits = TRUE) else 10.5
  if (!is.na(ni_p) && length(left_i) >= 1L) {
    xmin_plastic <- min(left_i)
    p2 <- add_bracket_segments(p2, xmin_plastic, max(left_i), y_hi, y_hi - tick_h)
    p2 <- p2 +
      ggplot2::annotate(
        "text",
        x = mean(c(xc(xmin_plastic), xc(max(left_i)))),
        y = y_hi + lab_dy,
        label = paste0("NI vs AML: ", fmt_p(ni_p), sig_sfx(ni_p)),
        size = br_mm
      )
  }
  if (!is.na(two_p) && !is.na(vline_x)) {
    # Span all lineage boxes (leftmost through rightmost) so the bracket visually covers the full comparator axis.
    xr <- seq_len(length(lv))
    if (length(xr) >= 1L) {
      p2 <- add_bracket_segments(p2, min(xr), max(xr), y_top, y_top - tick_h)
      p2 <- p2 +
        ggplot2::annotate(
          "text",
          x = mean(c(xc(min(xr)), xc(max(xr)))),
          y = y_top + lab_dy,
          label = paste0("Stem-like vs other: ", fmt_p(two_p), sig_sfx(two_p)),
          size = br_mm
        )
    }
  }
  p2
}
# Wide BCL2 expression: bracket p-values (requires fig12_* from `similarity-panels`, which runs before this chunk).
if (!is.null(p_bcl2) && exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df) && nrow(bcl2_plot_df) > 0L && "BCL2" %in% names(bcl2_plot_df)) {
  ni_p_e <- if (exists("fig12_bcl2_ni_p")) fig12_bcl2_ni_p else NA_real_
  two_p_e <- if (exists("fig12_bcl2_two_p")) fig12_bcl2_two_p else NA_real_
  vx_e <- if (exists("f1_vline_x")) f1_vline_x else NA_real_
  lv_e <- as.character(levels(bcl2_plot_df$lab))
  p_bcl2 <- .fig12_signif_brackets_wide(
    p_bcl2,
    y_num = bcl2_plot_df$BCL2,
    ni_p = ni_p_e,
    two_p = two_p_e,
    vline_x = vx_e,
    lv_chr = lv_e,
    plot_df = bcl2_plot_df,
    ylim = if (exists("f1_bcl2_expr_ylim", inherits = TRUE)) f1_bcl2_expr_ylim else NULL
  )
  fig_path_b2 <- if (requireNamespace("here", quietly = TRUE)) here::here("figures", "BCL2_rationale_Fig1.png") else "figures/BCL2_rationale_Fig1.png"
  rmd_fig_b2 <- if (requireNamespace("here", quietly = TRUE)) here::here("rmarkdown", "figures") else file.path("rmarkdown", "figures")
  if (dir.exists(dirname(fig_path_b2))) {
    tryCatch(ggsave(fig_path_b2, p_bcl2, width = 12, height = 6.5, dpi = 150), error = function(e) NULL)
    tryCatch(ggsave(sub("\\.png$", ".pdf", fig_path_b2), p_bcl2, width = 12, height = 6.5, device = "pdf"), error = function(e) NULL)
  }
  if (dir.exists(rmd_fig_b2)) {
    tryCatch(ggsave(file.path(rmd_fig_b2, "BCL2_rationale_Fig1.png"), p_bcl2, width = 12, height = 6.5, dpi = 150), error = function(e) NULL)
    tryCatch(ggsave(file.path(rmd_fig_b2, "BCL2_rationale_Fig1.pdf"), p_bcl2, width = 12, height = 6.5, device = "pdf"), error = function(e) NULL)
  }
  if (exists(".save_expanded_pdf") && ((exists("dnmti_bcl2_expanded_vector_pdf") && isTRUE(dnmti_bcl2_expanded_vector_pdf)) || (exists("hma_bcl2_master_vector_pdf") && isTRUE(hma_bcl2_master_vector_pdf)))) {
    .save_expanded_pdf(p_bcl2, "03_BCL2_rationale_Fig1_wide", 12, 6.5)
  }
}
if (exists("p_bcl2_ceres") && !is.null(p_bcl2_ceres) && exists("bcl2_ceres_plot_df") && is.data.frame(bcl2_ceres_plot_df) && nrow(bcl2_ceres_plot_df) > 0L) {
  ni_p_c <- if (exists("fig12_bcl2_ceres_ni_p")) fig12_bcl2_ceres_ni_p else NA_real_
  two_p_c <- if (exists("fig12_bcl2_ceres_two_p")) fig12_bcl2_ceres_two_p else NA_real_
  vx_c2 <- if (exists("f1_vline_x")) f1_vline_x else NA_real_
  lv_c <- as.character(levels(bcl2_plot_df$lab))
  p_bcl2_ceres <- .fig12_signif_brackets_wide(
    p_bcl2_ceres,
    y_num = bcl2_ceres_plot_df$dependency,
    ni_p = ni_p_c,
    two_p = two_p_c,
    vline_x = vx_c2,
    lv_chr = lv_c,
    plot_df = bcl2_ceres_plot_df,
    ylim = if (exists("f1_bcl2_ceres_ylim", inherits = TRUE)) f1_bcl2_ceres_ylim else NULL
  )
}
if (exists("p_dnmt1_ceres") && !is.null(p_dnmt1_ceres) && exists("dnmt1_ceres_plot_df") && is.data.frame(dnmt1_ceres_plot_df) && nrow(dnmt1_ceres_plot_df) > 0L &&
      exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df) && "lab" %in% names(bcl2_plot_df)) {
  ni_p_dn <- if (exists("fig12_dnmt1_ceres_ni_p")) fig12_dnmt1_ceres_ni_p else NA_real_
  two_p_dn <- if (exists("fig12_dnmt1_ceres_two_p")) fig12_dnmt1_ceres_two_p else NA_real_
  vx_dn <- if (exists("f1_vline_x")) f1_vline_x else NA_real_
  lv_dn <- as.character(levels(bcl2_plot_df$lab))
  p_dnmt1_ceres <- .fig12_signif_brackets_wide(
    p_dnmt1_ceres,
    y_num = dnmt1_ceres_plot_df$dependency,
    ni_p = ni_p_dn,
    two_p = two_p_dn,
    vline_x = vx_dn,
    lv_chr = lv_dn,
    plot_df = dnmt1_ceres_plot_df,
    ylim = if (exists("f1_dnmt1_ceres_ylim", inherits = TRUE)) f1_dnmt1_ceres_ylim else NULL
  )
}
# BCL2 wide panels are stacked with DNMT1 in chunk `fig-bcl2-dnmt1-lineage-combined` (single figure for the child report).
```

### DNMT1 panels (stacked with BCL2 in one figure below)

```{r dnmt1-distributions, cache = FALSE, fig.show = "hide"}
# DNMT1 wide panels (same cohort and x-axis as BCL2); rendered with BCL2 in `fig-bcl2-dnmt1-lineage-combined`.
cohort_ext_hem_f2 <- exists("cohort_extended_hem") && isTRUE(cohort_extended_hem)
dnmt1_fig2_ok <- exists("expr_target") && (
  (exists("expr_meta_dnmt1") && is.data.frame(expr_meta_dnmt1) && nrow(expr_meta_dnmt1) > 0L &&
    sum(!is.na(expr_meta_dnmt1$DNMT1)) >= 2L) ||
  (exists("expr_meta") && is.data.frame(expr_meta) && "DNMT1" %in% names(expr_meta) && sum(!is.na(expr_meta$DNMT1)) >= 2L)
)
p_dnmt1 <- NULL
if (dnmt1_fig2_ok) {
  # Use full expression matrix for DNMT1 when available so DNMT1 expression keeps the same lymphoma strata as BCL2 in Figure 1
  # (expr_meta_dnmt1 can be a narrowed subset in some workspaces).
  dnmt1_expr_src <- if (exists("expr_meta") && is.data.frame(expr_meta) && "DNMT1" %in% names(expr_meta)) {
    expr_meta
  } else {
    expr_meta_dnmt1
  }
  dnmt1_with_lineage <- dnmt1_expr_src %>%
    left_join(expr_target %>% select(depmap_id, lineage_group), by = "depmap_id") %>%
    filter(!is.na(lineage_group), !is.na(DNMT1))
  dnmt1_with_lineage <- dnmt1_with_lineage %>%
    mutate(
      subtype = as.character(lineage_group),
      tissue = dplyr::case_when(
        lineage_lower == "lung" ~ "Lung",
        grepl("ExPulm|NEPC|NE/small", subtype) ~ "ExPulm-SCNEC",
        as.character(lineage_group) == "AML (control)" ~ "AML",
        grepl("Lymphoma", as.character(.data$lineage_group), fixed = TRUE) |
          as.character(.data$lineage_group) == "DLBCL/MCL/CLL" ~ "Lymphoma",
        grepl("Neuroblastoma", subtype, fixed = TRUE) ~ "Neuroblastoma",
        grepl("TNBC", subtype, fixed = TRUE) ~ "TNBC",
        grepl("Breast", subtype) ~ "Breast",
        grepl("prostate", subtype, ignore.case = TRUE) ~ "Prostate",
        grepl("Urothelial", subtype, fixed = TRUE) ~ "Urothelial",
        grepl("Renal cell", subtype, fixed = TRUE) ~ "Renal cell",
        grepl("Colorectal", subtype, fixed = TRUE) ~ "Colorectal",
        grepl("Melanoma", subtype, fixed = TRUE) ~ "Melanoma",
        grepl("Sarcoma", subtype, fixed = TRUE) ~ "Sarcoma",
        TRUE ~ NA_character_
      )
    ) %>% filter(!is.na(tissue))
  # Lung: use expr_target so SCLC, Localized NSCLC, Metastatic NSCLC (Localized/Metastatic adjacent in plot)
  dnmt1_lung <- dnmt1_expr_src %>% filter(lineage_lower == "lung", !is.na(DNMT1)) %>%
    inner_join(expr_target %>% filter(lineage_group %in% c("SCLC", "Localized NSCLC", "Metastatic NSCLC")) %>% select(depmap_id, lineage_group), by = "depmap_id") %>%
    mutate(subtype = as.character(lineage_group), tissue = "Lung")
  dnmt1_plot_df_f2 <- dnmt1_with_lineage %>% filter(tissue != "Lung") %>% bind_rows(dnmt1_lung) %>%
    distinct(depmap_id, tissue, subtype, .keep_all = TRUE)  # one point per cell line
  tissue_order_f2 <- if (exists("f1_tissue_axis_order")) {
    f1_tissue_axis_order
  } else {
    c(
      "Lung", "ExPulm-SCNEC", "AML", "Neuroblastoma", "Breast",
      if (cohort_ext_hem_f2) "TNBC",
      "Lymphoma",
      "Prostate", "Urothelial", "Renal cell", "Colorectal", "Melanoma", "Sarcoma"
    )
  }
  dnmt1_plot_df_f2 <- dnmt1_plot_df_f2 %>%
    mutate(
      tissue = factor(as.character(tissue), levels = intersect(tissue_order_f2, unique(as.character(tissue)))),
      subtype = factor(
        dplyr::if_else(
          as.character(.data$subtype) %in% c("Localized Neuroblastoma", "Metastatic Neuroblastoma"),
          "Neuroblastoma",
          as.character(.data$subtype)
        ),
        levels = unique(c(as.character(lineage_levels), "Neuroblastoma"))
      )
    )
  dnmt1_n_f2 <- dnmt1_plot_df_f2 %>% group_by(tissue, subtype) %>% summarise(n = n(), .groups = "drop")
  hma_ext_f2 <- exists("hma_bcl2_tnbc_lymph_master") && isTRUE(hma_bcl2_tnbc_lymph_master)
  plastic_anchors_f2 <- if (exists("lineage_plastic_ref_canonical")) {
    as.character(lineage_plastic_ref_canonical)
  } else {
    c("SCLC", "ExPulm-SCNEC", "AML (control)")
  }
  if (hma_ext_f2) {
    nb_f2 <- intersect(
      c("Neuroblastoma", "Unspecified Neuroblastoma"),
      if (exists("lineage_levels")) as.character(lineage_levels) else character(0)
    )
    sarcoma_f2 <- intersect(
      c("Localized Sarcoma", "Metastatic Sarcoma", "Unspecified Sarcoma"),
      if (exists("lineage_levels")) as.character(lineage_levels) else character(0)
    )
    plastic_anchors_f2 <- unique(c(
      plastic_anchors_f2,
      "DLBCL/MCL/CLL",
      nb_f2,
      sarcoma_f2
    ))
    dnmt1_n_f2 <- dnmt1_n_f2 %>%
      dplyr::filter(.data$n > 5L | as.character(.data$subtype) %in% plastic_anchors_f2) %>%
      dplyr::filter(
        !grepl("Unspecified", as.character(.data$subtype), fixed = TRUE) |
          grepl("Unspecified Neuroblastoma", as.character(.data$subtype), fixed = TRUE) |
          grepl("Unspecified Sarcoma", as.character(.data$subtype), fixed = TRUE)
      )
  }
  # Match BCL2 wide Fig. 1 anchors (incl. optional high-N metastatic promote from bcl2-distributions)
  if (exists("plastic_anchors") && length(plastic_anchors) > 0L) {
    plastic_anchors_f2 <- unique(c(plastic_anchors_f2, as.character(plastic_anchors)))
  }
  dnmt1_keep_f2 <- dnmt1_n_f2
  dnmt1_plot_df_f2 <- dnmt1_plot_df_f2 %>% semi_join(dnmt1_keep_f2, by = c("tissue", "subtype"))
  st_dnf <- as.character(dnmt1_plot_df_f2$subtype)
  keep_loc_dnf <- .f1_keep_localized_subtype_chr(st_dnf)
  drop_loc_dnf <- grepl("^Localized\\b", st_dnf) & !keep_loc_dnf
  dnmt1_plot_df_f2 <- dplyr::filter(dnmt1_plot_df_f2, !drop_loc_dnf)
  st_kf2 <- as.character(dnmt1_keep_f2$subtype)
  keep_loc_kf2 <- .f1_keep_localized_subtype_chr(st_kf2)
  drop_loc_kf2 <- grepl("^Localized\\b", st_kf2) & !keep_loc_kf2
  dnmt1_keep_f2 <- dplyr::filter(dnmt1_keep_f2, !drop_loc_kf2)
  dnmt1_plot_df_f2 <- dplyr::filter(dnmt1_plot_df_f2, !.f1_exclude_lineage_from_fig12_plots(as.character(.data$subtype)))
  dnmt1_keep_f2 <- dplyr::filter(dnmt1_keep_f2, !.f1_exclude_lineage_from_fig12_plots(as.character(.data$subtype)))
  # #region agent log
  tryCatch({
    keep_counts2 <- dnmt1_plot_df_f2 %>%
      dplyr::mutate(subtype_chr = as.character(.data$subtype)) %>%
      dplyr::filter(grepl("TNBC|Melanoma|Sarcoma", .data$subtype_chr, ignore.case = TRUE)) %>%
      dplyr::count(.data$subtype_chr, name = "n_expr")
    .agent_dbg_ndjson(
      runId = "pre-fix-debug2",
      hypothesisId = "H3",
      location = "03_Rmd:dnmt1_keep_focus_subtypes",
      message = "DNMT1 kept lineage focus counts after figure filters",
      data = list(
        n_rows = nrow(dnmt1_plot_df_f2),
        focus_subtypes = as.character(keep_counts2$subtype_chr),
        focus_counts = as.integer(keep_counts2$n_expr)
      )
    )
  }, error = function(e) invisible(NULL))
  # #endregion
  dnmt1_n_lab_f2 <- dnmt1_plot_df_f2 %>% group_by(tissue, subtype) %>% summarise(n = n(), .groups = "drop") %>% mutate(lab = paste0(gsub(" \\(control\\)", "", as.character(subtype)), " N=", n))
  dnmt1_plot_df_f2 <- dnmt1_plot_df_f2 %>% left_join(dnmt1_n_lab_f2 %>% select(tissue, subtype, lab), by = c("tissue", "subtype"))
  dnmt1_plot_df_f2 <- dnmt1_plot_df_f2 %>%
    dplyr::left_join(expr_target %>% dplyr::select(.data$depmap_id, dplyr::any_of("lineage_expulm_detail")), by = "depmap_id") %>%
    dplyr::mutate(is_nepc = (.data$subtype == "ExPulm-SCNEC" & dplyr::coalesce(.data$lineage_expulm_detail, "") == "NEPC")) %>%
    dplyr::mutate(
      lab = dplyr::if_else(
        as.character(.data$subtype) == "ExPulm-SCNEC" & .data$is_nepc,
        sub("ExPulm-SCNEC", "NEPC", as.character(.data$lab), fixed = TRUE),
        as.character(.data$lab)
      )
    )
  # Order x-axis so Localized X and Metastatic X are adjacent within each tissue
  lab_levels_f2 <- dnmt1_plot_df_f2 %>% distinct(tissue, subtype, lab) %>%
    arrange(tissue, match(as.character(subtype), lineage_levels)) %>% pull(lab)
  dnmt1_plot_df_f2 <- dnmt1_plot_df_f2 %>% mutate(lab = factor(lab, levels = lab_levels_f2))
  if (nrow(dnmt1_plot_df_f2) >= 2) {
    if (hma_ext_f2 && length(unique(as.character(dnmt1_plot_df_f2$lab))) > 10L) {
      st2 <- as.character(dnmt1_plot_df_f2$subtype)
      tt2 <- as.character(dnmt1_plot_df_f2$tissue)
      pooled_axis_preview2 <- dplyr::case_when(
        st2 %in% plastic_anchors_f2 ~ st2,
        grepl("Neuroblastoma", st2, fixed = TRUE) & !grepl("^Unspecified", st2, fixed = TRUE) ~ "Neuroblastoma",
        grepl("^Localized\\b", st2) ~ paste0(tt2, " — localized"),
        grepl("^Metastatic\\b", st2) ~ paste0(tt2, " — metastatic"),
        TRUE ~ st2
      )
      # #region agent log
      tryCatch({
        .agent_dbg_ndjson(
          runId = "pre-fix-debug2",
          hypothesisId = "H1",
          location = "03_Rmd:dnmt1_pooling_preview",
          message = "DNMT1 pooled-axis preview before filtering",
          data = list(
            n_rows = nrow(dnmt1_plot_df_f2),
            n_na_pooled = sum(is.na(pooled_axis_preview2)),
            n_blank_pooled = sum(!is.na(pooled_axis_preview2) & !nzchar(trimws(as.character(pooled_axis_preview2)))),
            pooled_head = utils::head(unique(as.character(pooled_axis_preview2)), 10L)
          )
        )
      }, error = function(e) invisible(NULL))
      # #endregion
      dnmt1_plot_df_f2 <- dnmt1_plot_df_f2 %>%
        dplyr::mutate(
          pooled_axis = dplyr::case_when(
            st2 %in% plastic_anchors_f2 ~ st2,
            grepl("Neuroblastoma", st2, fixed = TRUE) & !grepl("^Unspecified", st2, fixed = TRUE) ~ "Neuroblastoma",
            grepl("^Localized\\b", st2) ~ paste0(tt2, " — localized"),
            grepl("^Metastatic\\b", st2) ~ paste0(tt2, " — metastatic"),
            TRUE ~ st2
          )
        ) %>%
        dplyr::filter(!is.na(.data$pooled_axis), nzchar(trimws(as.character(.data$pooled_axis))))
      n_pool2 <- dnmt1_plot_df_f2 %>% dplyr::count(.data$pooled_axis, name = "nn")
      dnmt1_plot_df_f2 <- dnmt1_plot_df_f2 %>%
        dplyr::left_join(n_pool2, by = "pooled_axis") %>%
        dplyr::mutate(lab = paste0(gsub(" \\(control\\)", "", as.character(.data$pooled_axis)), " N=", .data$nn))
      ord2 <- dnmt1_plot_df_f2 %>%
        dplyr::distinct(.data$tissue, .data$pooled_axis, .data$lab) %>%
        dplyr::arrange(.data$tissue, .data$pooled_axis) %>%
        dplyr::pull(.data$lab)
      dnmt1_plot_df_f2 <- dnmt1_plot_df_f2 %>%
        dplyr::mutate(lab = factor(.data$lab, levels = unique(ord2))) %>%
        dplyr::select(-dplyr::any_of(c("pooled_axis", "nn")))
    }
    if (hma_ext_f2 && nrow(dnmt1_plot_df_f2) > 0L) {
      dnmt1_plot_df_f2 <- dnmt1_plot_df_f2 %>%
        dplyr::filter(
          !grepl("Unspecified", as.character(.data$lab), fixed = TRUE) |
            grepl("Unspecified Neuroblastoma", as.character(.data$lab), fixed = TRUE) |
            grepl("Unspecified Sarcoma", as.character(.data$lab), fixed = TRUE)
        )
    }
    dnmt1_plot_df_f2 <- .f1_sanitize_lab_rows(dnmt1_plot_df_f2)
    # Keep TNBC metastatic stratum (mTNBC) for DNMT1 to match BCL2 wide-axis lineage boxes.
    # Match Figure 1 wide x-axis (BCL2 expression) lab order when available
    if (exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df) && "lab" %in% names(bcl2_plot_df)) {
      lev_ref_dn <- levels(bcl2_plot_df$lab)
      if (length(lev_ref_dn) > 0L) {
        # Strictly follow BCL2 wide-axis levels for DNMT1 expression to keep left-column boxes/colors aligned.
        dnmt1_plot_df_f2 <- dnmt1_plot_df_f2 %>%
          dplyr::filter(as.character(.data$lab) %in% lev_ref_dn)
        lev_full_dn <- lev_ref_dn
        dnmt1_plot_df_f2 <- dnmt1_plot_df_f2 %>%
          dplyr::mutate(lab = factor(as.character(.data$lab), levels = lev_full_dn))
      }
    }
    # Bottom-row DNMT1 expression: cap upper y (log2 TPM+1) near ~9 to zoom contrasts.
    f1_dnmt1_expr_y_hi_cap <- 9
    f1_dnmt1_expr_ylim <- local({
      yv <- dnmt1_plot_df_f2$DNMT1
      yv <- yv[is.finite(yv)]
      if (length(yv) < 1L) {
        return(NULL)
      }
      yr <- range(yv)
      d <- diff(yr)
      if (!is.finite(d) || d < 1e-6) d <- 0.45
      pad <- max(0.12, 0.065 * d)
      y_lo <- max(2.5, yr[1] - pad)
      y_hi <- min(f1_dnmt1_expr_y_hi_cap, yr[2] + pad)
      yn <- dnmt1_plot_df_f2$DNMT1[dnmt1_plot_df_f2$is_nepc %in% TRUE]
      yn <- yn[is.finite(yn)]
      if (length(yn) > 0L) {
        rng_n <- if (length(yn) >= 2L) diff(range(yn)) else 0
        pad_n <- max(0.12, 0.12 * max(rng_n, max(abs(yn), na.rm = TRUE) * 0.06, 0.28))
        y_lo <- min(y_lo, min(yn) - pad_n)
        y_hi <- max(y_hi, max(yn) + pad_n)
        y_lo <- max(1.5, y_lo)
        y_hi <- min(12, y_hi)
      }
      if (!is.finite(y_hi) || y_hi <= y_lo) y_hi <- y_lo + 0.55
      c(y_lo, y_hi)
    })
    .mk_dnmt1_expr_plot <- function(dat, vline_x = NA_real_) {
      g <- ggplot(dat, aes(x = lab, y = DNMT1, fill = subtype)) +
        geom_boxplot(alpha = 0.7, outlier.shape = NA) +
        ggplot2::geom_point(
          data = dplyr::filter(dat, .data$is_nepc %in% TRUE),
          ggplot2::aes(x = lab, y = DNMT1),
          inherit.aes = FALSE,
          color = "#E41A1C",
          position = ggplot2::position_jitter(width = 0.12, height = 0),
          alpha = 0.92,
          size = 2.5
        ) +
        ggplot2::scale_x_discrete(drop = FALSE, labels = .f1_axis_short_met_label) +
        labs(x = "Lineage", y = "DNMT1 expression (log2(TPM+1))") +
        theme_minimal() +
        .f1_fig12_axis_theme(45) +
        theme(legend.position = "none")
      if (!is.na(vline_x)) {
        g <- g + ggplot2::geom_vline(
          xintercept = vline_x,
          linetype = "dashed",
          linewidth = f1_rule_line_linewidth,
          alpha = 1,
          colour = f1_rule_line_colour
        )
      }
      g
    }
    vx_dnmt1_expr <- if (exists("f1_vline_x")) f1_vline_x else NA_real_
    p_dnmt1 <- .mk_dnmt1_expr_plot(dnmt1_plot_df_f2, vline_x = vx_dnmt1_expr)
    if (exists(".fig12_signif_brackets_wide") && exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df) && nrow(dnmt1_plot_df_f2) > 0L) {
      ni_p_d1 <- if (exists("fig12_dnmt1_ni_p")) fig12_dnmt1_ni_p else NA_real_
      two_p_d1 <- if (exists("fig12_dnmt1_two_p")) fig12_dnmt1_two_p else NA_real_
      lv_d1 <- as.character(levels(bcl2_plot_df$lab))
      p_dnmt1 <- .fig12_signif_brackets_wide(
        p_dnmt1,
        y_num = dnmt1_plot_df_f2$DNMT1,
        ni_p = ni_p_d1,
        two_p = two_p_d1,
        vline_x = vx_dnmt1_expr,
        lv_chr = lv_d1,
        plot_df = dnmt1_plot_df_f2,
        ylim = if (exists("f1_dnmt1_expr_ylim", inherits = TRUE)) f1_dnmt1_expr_ylim else NULL
      )
    }
    # DNMT1 expression + CERES are stacked with BCL2 in chunk `fig-bcl2-dnmt1-lineage-combined`.
  }
}
```

```{r fig-bcl2-dnmt1-lineage-combined, cache = FALSE, fig.width = 18, fig.height = 26.5, fig.cap = "Figure 1. BCL2 and DNMT1 expression and genetic dependency across lineages (one composite figure). RNA-seq and CRISPR dependency are from DepMap. Expression is log2(TPM + 1); dependency is CERES (more negative = stronger apparent dependency on the named gene). Layout: 2×2 grid on the same wide lineage x-axis and strata when DNMT1 data are available—top left BCL2 expression, top right BCL2 CERES, bottom left DNMT1 expression, bottom right DNMT1 CERES. Expression panels use the full cohort per `lab`/`subtype`; **wide CERES** panels plot only cell lines with non-missing CERES (same x-axis order and fill colors as BCL2 expression; lines without dependency scores are omitted from the CERES panels). The **N** in each CERES tick label counts plotted dependency scores in that stratum (expression cohort **N** may differ); NEPC-only strata with no plotted CERES have a blank tick label. Lineage tick labels are drawn on **both** bottom-row panels; the top row omits x tick labels. Metastatic strata on the x-axis use an **m** prefix (e.g. mBreast for metastatic breast). Bracketed p-values on the BCL2 panels: pooled non-AML left-of-line arm vs AML (1-sided Wilcoxon; expression stochastic ≥ AML; CERES combined non-AML ≥ AML) and stem-like vs non-stem-like to the right of the dashed rule (expression: stem-like > other). DNMT1 panels use the same comparator logic as in the prior two-figure layout (expression stochastic ≤ AML vs pooled non-AML when that arm applies; CERES combined non-AML ≥ AML; stem-like vs non-stem-like two-group Wilcoxon). On **expression** panels, **points** are drawn for **NEPC** lines only (red); other lines are not overlaid as points. NSCLC strata are non-stem-like comparators for two-group p-values. Abbreviations: AML, acute myeloid leukemia; BCL2, B-cell lymphoma 2; CERES, CRISPR Essentiality Ranked by Expression of effect and Specificity; DepMap, Cancer Dependency Map; DNMT1, DNA methyltransferase 1; ExPulm-SCNEC, extrapulmonary small cell neuroendocrine carcinoma; NEPC, neuroendocrine prostate cancer; NSCLC, non-small cell lung cancer; SCLC, small cell lung cancer; TPM, transcripts per million."}
if (!exists("f1_rule_line_colour")) f1_rule_line_colour <- "#FF2222"
if (!exists("f1_rule_line_linewidth")) f1_rule_line_linewidth <- 1.4
fig_dir_lineage <- if (requireNamespace("here", quietly = TRUE)) here::here("figures") else "figures"
# Canonical wide lineage order and subtype fill = BCL2 expression panel (all four grid panels match these)
lev_f1_grid <- if (exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df) && is.factor(bcl2_plot_df$lab)) {
  levels(bcl2_plot_df$lab)
} else {
  character(0)
}
f1_fill_subtype_levels <- if (exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df) && is.factor(bcl2_plot_df$subtype)) {
  levels(bcl2_plot_df$subtype)
} else if (exists("lineage_levels")) {
  as.character(lineage_levels)
} else {
  character(0)
}
# Per-level colors: golden-ratio hue steps so *consecutive* subtype factor levels are visually far apart
# (sequential hues cluster red/orange/yellow; this mixes cool blues, magentas, greens, etc. along the axis).
.f1_distinct_subtype_palette <- function(level_chr) {
  k <- length(level_chr)
  if (k < 1L) {
    return(character(0))
  }
  if (k == 1L) {
    return(stats::setNames("#5B9BD5", level_chr))
  }
  golden <- 0.6180339887498948482045868343656
  ii <- seq_len(k) - 1L
  h <- (0.11 + ii * golden) %% 1
  s <- 0.70 + 0.24 * ((ii + 1L) %% 3L) / 2
  v <- 0.84 + 0.14 * ((ii + 2L) %% 4L) / 3
  cols <- grDevices::hsv(h = h, s = pmin(0.94, s), v = pmin(0.98, v))
  stats::setNames(cols, level_chr)
}
f1_fill_palette <- if (length(f1_fill_subtype_levels) > 0L) {
  .f1_distinct_subtype_palette(f1_fill_subtype_levels)
} else {
  character(0)
}
.f1_wide_fill_scale <- if (length(f1_fill_palette) > 0L) {
  ggplot2::scale_fill_manual(
    values = f1_fill_palette,
    limits = f1_fill_subtype_levels,
    drop = FALSE,
    na.value = "grey88"
  )
} else {
  NULL
}
.f1_relevel_lab_subtype_df <- function(df) {
  if (is.null(df) || !is.data.frame(df) || length(lev_f1_grid) < 1L) {
    return(df)
  }
  if ("lab" %in% names(df)) {
    df$lab <- factor(as.character(df$lab), levels = lev_f1_grid)
  }
  if ("subtype" %in% names(df) && length(f1_fill_subtype_levels) > 0L) {
    df$subtype <- factor(as.character(df$subtype), levels = f1_fill_subtype_levels)
  }
  df
}
.f1_harmonize_plot_df_to_bcl2 <- function(df) {
  if (is.null(df) || !is.data.frame(df) || !"depmap_id" %in% names(df) || is.null(bcl2_f1_id_map) || nrow(bcl2_f1_id_map) < 1L) {
    return(df)
  }
  drop <- intersect(c("lab", "subtype", "is_nepc"), names(df))
  base <- if (length(drop)) dplyr::select(df, -dplyr::all_of(drop)) else df
  out <- dplyr::inner_join(base, bcl2_f1_id_map, by = "depmap_id")
  .f1_relevel_lab_subtype_df(out)
}
.f1_patch_plot_data <- function(p, df) {
  if (is.null(p) || is.null(df) || !nrow(df)) {
    return(p)
  }
  tryCatch(ggplot2::`%+%`(p, df), error = function(e) p)
}
.theme_f1_hide_lineage_x <- ggplot2::theme(
  axis.text.x = ggplot2::element_blank(),
  axis.ticks.x = ggplot2::element_blank(),
  axis.title.x = ggplot2::element_blank()
)
# Extra vertical gap in the 2×2 composite (t must be ≥ .f1_fig12_bracket_margin_top_2tier so labels are not clipped).
.theme_f1_patchwork_top_row <- ggplot2::theme(
  plot.margin = ggplot2::margin(
    t = get0(".f1_fig12_bracket_margin_top_2tier", ifnotfound = 120L, inherits = TRUE),
    r = 36,
    b = 26,
    l = 6
  )
)
.theme_f1_patchwork_bottom_row <- ggplot2::theme(
  plot.margin = ggplot2::margin(
    t = get0(".f1_fig12_bracket_margin_top_2tier", ifnotfound = 120L, inherits = TRUE) + 8L,
    r = 36,
    b = 14,
    l = 6
  )
)
if (exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df) && nrow(bcl2_plot_df) > 0L && length(lev_f1_grid) > 0L) {
  bcl2_plot_df <- .f1_relevel_lab_subtype_df(bcl2_plot_df)
  if (exists("p_bcl2") && !is.null(p_bcl2)) {
    p_bcl2 <- .f1_patch_plot_data(p_bcl2, bcl2_plot_df)
  }
}
bcl2_f1_id_map <- if (exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df) && nrow(bcl2_plot_df) > 0L &&
  all(c("depmap_id", "lab", "subtype", "is_nepc") %in% names(bcl2_plot_df))) {
  bcl2_plot_df %>%
    dplyr::distinct(.data$depmap_id, .keep_all = TRUE) %>%
    dplyr::transmute(
      .data$depmap_id,
      lab = .data$lab,
      subtype = factor(as.character(.data$subtype), levels = f1_fill_subtype_levels),
      is_nepc = .data$is_nepc
    )
} else {
  NULL
}
if (exists("bcl2_ceres_plot_df") && is.data.frame(bcl2_ceres_plot_df) && "depmap_id" %in% names(bcl2_ceres_plot_df)) {
  bcl2_ceres_plot_df <- .f1_harmonize_plot_df_to_bcl2(bcl2_ceres_plot_df)
  if (exists("p_bcl2_ceres") && !is.null(p_bcl2_ceres)) {
    p_bcl2_ceres <- .f1_patch_plot_data(p_bcl2_ceres, bcl2_ceres_plot_df)
  }
}
if (exists("dnmt1_plot_df_f2") && is.data.frame(dnmt1_plot_df_f2) && "depmap_id" %in% names(dnmt1_plot_df_f2)) {
  dnmt1_plot_df_f2 <- .f1_harmonize_plot_df_to_bcl2(dnmt1_plot_df_f2)
  if (exists("p_dnmt1") && !is.null(p_dnmt1)) {
    p_dnmt1 <- .f1_patch_plot_data(p_dnmt1, dnmt1_plot_df_f2)
  }
}
if (exists("dnmt1_ceres_plot_df") && is.data.frame(dnmt1_ceres_plot_df) && "depmap_id" %in% names(dnmt1_ceres_plot_df)) {
  dnmt1_ceres_plot_df <- .f1_harmonize_plot_df_to_bcl2(dnmt1_ceres_plot_df)
  if (exists("p_dnmt1_ceres") && !is.null(p_dnmt1_ceres)) {
    p_dnmt1_ceres <- .f1_patch_plot_data(p_dnmt1_ceres, dnmt1_ceres_plot_df)
  }
}
if (!is.null(.f1_wide_fill_scale)) {
  if (exists("p_bcl2") && !is.null(p_bcl2)) p_bcl2 <- p_bcl2 + .f1_wide_fill_scale
  if (exists("p_bcl2_ceres") && !is.null(p_bcl2_ceres)) p_bcl2_ceres <- p_bcl2_ceres + .f1_wide_fill_scale
  if (exists("p_dnmt1") && !is.null(p_dnmt1)) p_dnmt1 <- p_dnmt1 + .f1_wide_fill_scale
  if (exists("p_dnmt1_ceres") && !is.null(p_dnmt1_ceres)) p_dnmt1_ceres <- p_dnmt1_ceres + .f1_wide_fill_scale
}
# #region agent log
tryCatch({
  b_lv <- if (exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df) && is.factor(bcl2_plot_df$lab)) as.character(levels(bcl2_plot_df$lab)) else character(0)
  d_lv <- if (exists("dnmt1_plot_df_f2") && is.data.frame(dnmt1_plot_df_f2) && is.factor(dnmt1_plot_df_f2$lab)) as.character(levels(dnmt1_plot_df_f2$lab)) else character(0)
  b_key <- if (exists("bcl2_plot_df") && is.data.frame(bcl2_plot_df)) {
    bcl2_plot_df %>%
      dplyr::distinct(.data$lab, .data$subtype) %>%
      dplyr::mutate(lab = as.character(.data$lab), subtype = as.character(.data$subtype))
  } else {
    data.frame(lab = character(0), subtype = character(0))
  }
  d_key <- if (exists("dnmt1_plot_df_f2") && is.data.frame(dnmt1_plot_df_f2)) {
    dnmt1_plot_df_f2 %>%
      dplyr::distinct(.data$lab, .data$subtype) %>%
      dplyr::mutate(lab = as.character(.data$lab), subtype = as.character(.data$subtype))
  } else {
    data.frame(lab = character(0), subtype = character(0))
  }
  subtype_mismatch <- dplyr::full_join(b_key, d_key, by = "lab", suffix = c("_bcl2", "_dnmt1")) %>%
    dplyr::filter(!is.na(.data$lab) & .data$subtype_bcl2 != .data$subtype_dnmt1)
  focus_labs <- c("mMelanoma", "mSarcoma", "mTNBC", "NEPC")
  .agent_dbg_ndjson(
    runId = "pre-fix-debug2",
    hypothesisId = "H2",
    location = "03_Rmd:fig1_axis_alignment_snapshot",
    message = "Expression panel axis/subtype alignment snapshot",
    data = list(
      bcl2_level_count = length(b_lv),
      dnmt1_level_count = length(d_lv),
      same_level_order = identical(b_lv, d_lv),
      bcl2_only_levels = setdiff(b_lv, d_lv),
      dnmt1_only_levels = setdiff(d_lv, b_lv),
      bcl2_has_na_like = sum(grepl("^NA\\b|^\\s*$", b_lv)),
      dnmt1_has_na_like = sum(grepl("^NA\\b|^\\s*$", d_lv)),
      focus_in_bcl2 = b_lv[grepl(paste0("^(", paste(focus_labs, collapse = "|"), ")"), b_lv)],
      focus_in_dnmt1 = d_lv[grepl(paste0("^(", paste(focus_labs, collapse = "|"), ")"), d_lv)],
      subtype_mismatch_n = nrow(subtype_mismatch),
      subtype_mismatch_head = utils::head(paste0(subtype_mismatch$lab, ": ", subtype_mismatch$subtype_bcl2, " vs ", subtype_mismatch$subtype_dnmt1), 12L),
      bcl2_levels = b_lv,
      dnmt1_levels = d_lv,
      bcl2_tnbc_levels = b_lv[grepl("TNBC", b_lv, ignore.case = TRUE)],
      dnmt1_tnbc_levels = d_lv[grepl("TNBC", d_lv, ignore.case = TRUE)],
      bcl2_empty_base_levels = b_lv[!nzchar(trimws(sub("\\s*N=\\s*.*$", "", b_lv)))],
      dnmt1_empty_base_levels = d_lv[!nzchar(trimws(sub("\\s*N=\\s*.*$", "", d_lv)))]
    )
  )
}, error = function(e) invisible(NULL))
# #endregion
# Bottom row: show lineage labels on both panels (top row hides x labels only)
p_br_dnmt <- if (exists("p_dnmt1_ceres") && !is.null(p_dnmt1_ceres)) {
  p_dnmt1_ceres
} else if (requireNamespace("patchwork", quietly = TRUE)) {
  patchwork::plot_spacer()
} else {
  NULL
}
has_dnmt_bottom <- isTRUE(dnmt1_fig2_ok) && exists("p_dnmt1") && !is.null(p_dnmt1)
use_grid_2x2 <- has_dnmt_bottom && requireNamespace("patchwork", quietly = TRUE) &&
  (!is.null(p_bcl2) || !is.null(p_bcl2_ceres))
if (use_grid_2x2) {
  p_tl <- if (!is.null(p_bcl2)) {
    p_bcl2 + .theme_f1_hide_lineage_x + .theme_f1_patchwork_top_row
  } else {
    patchwork::plot_spacer()
  }
  p_tr <- if (!is.null(p_bcl2_ceres)) {
    p_bcl2_ceres + .theme_f1_hide_lineage_x + .theme_f1_patchwork_top_row
  } else {
    patchwork::plot_spacer()
  }
  p_bl <- if (!is.null(p_dnmt1)) p_dnmt1 + .theme_f1_patchwork_bottom_row else patchwork::plot_spacer()
  p_br <- if (!is.null(p_br_dnmt)) p_br_dnmt + .theme_f1_patchwork_bottom_row else patchwork::plot_spacer()
  p_lineage_all <- (p_tl | p_tr) / (p_bl | p_br) +
    patchwork::plot_layout(widths = c(1, 1), heights = c(1, 1))
  print(p_lineage_all)
  h_pdf <- 26.5
  if (dir.exists(fig_dir_lineage)) {
    tryCatch(
      ggplot2::ggsave(
        file.path(fig_dir_lineage, "Figure1_BCL2_DNMT1_expression_and_dependency.pdf"),
        p_lineage_all,
        width = 18,
        height = h_pdf,
        device = "pdf"
      ),
      error = function(e) NULL
    )
    tryCatch(
      ggplot2::ggsave(
        file.path(fig_dir_lineage, "Figure1_BCL2_expression_and_dependency.pdf"),
        p_lineage_all,
        width = 18,
        height = h_pdf,
        device = "pdf"
      ),
      error = function(e) NULL
    )
  }
  if (exists(".save_expanded_pdf") && ((exists("dnmti_bcl2_expanded_vector_pdf") && isTRUE(dnmti_bcl2_expanded_vector_pdf)) || (exists("hma_bcl2_master_vector_pdf") && isTRUE(hma_bcl2_master_vector_pdf)))) {
    .save_expanded_pdf(p_lineage_all, "03_Figure1_BCL2_DNMT1_lineage_combined", 18, h_pdf)
  }
} else if (requireNamespace("patchwork", quietly = TRUE) && (!is.null(p_bcl2) || !is.null(p_bcl2_ceres))) {
  # BCL2-only or partial: single row, two columns (lineage labels on both panels)
  p_l <- if (!is.null(p_bcl2)) p_bcl2 else patchwork::plot_spacer()
  p_r <- if (!is.null(p_bcl2_ceres)) p_bcl2_ceres else patchwork::plot_spacer()
  p_lineage_all <- p_l | p_r
  print(p_lineage_all)
  h_pdf <- 12
  if (dir.exists(fig_dir_lineage)) {
    tryCatch(
      ggplot2::ggsave(
        file.path(fig_dir_lineage, "Figure1_BCL2_DNMT1_expression_and_dependency.pdf"),
        p_lineage_all,
        width = 18,
        height = h_pdf,
        device = "pdf"
      ),
      error = function(e) NULL
    )
    tryCatch(
      ggplot2::ggsave(
        file.path(fig_dir_lineage, "Figure1_BCL2_expression_and_dependency.pdf"),
        p_lineage_all,
        width = 18,
        height = h_pdf,
        device = "pdf"
      ),
      error = function(e) NULL
    )
  }
  if (exists(".save_expanded_pdf") && ((exists("dnmti_bcl2_expanded_vector_pdf") && isTRUE(dnmti_bcl2_expanded_vector_pdf)) || (exists("hma_bcl2_master_vector_pdf") && isTRUE(hma_bcl2_master_vector_pdf)))) {
    .save_expanded_pdf(p_lineage_all, "03_Figure1_BCL2_DNMT1_lineage_combined", 18, h_pdf)
  }
} else {
  plots_lineage_stack <- list()
  if (!is.null(p_bcl2)) plots_lineage_stack <- c(plots_lineage_stack, list(p_bcl2))
  if (!is.null(p_bcl2_ceres)) plots_lineage_stack <- c(plots_lineage_stack, list(p_bcl2_ceres))
  nl <- length(plots_lineage_stack)
  if (nl > 0L) {
    for (pl in plots_lineage_stack) print(pl)
  } else {
    p_miss <- if (exists("p_placeholder")) {
      p_placeholder
    } else {
      ggplot2::ggplot() +
        ggplot2::theme_void() +
        ggplot2::labs(title = "Insufficient data for lineage expression/CERES figure.")
    }
    print(p_miss)
  }
}
```

```{r interpretation-03-bcl2-rationale, echo = FALSE, results = "asis", eval = !exists("hma_bcl2_tnbc_lymph_master") || !isTRUE(hma_bcl2_tnbc_lymph_master)}
knitr::asis_output("**Interpretation**: Stem-like subtypes (SCLC, ExPulm-SCNEC) may show higher BCL2 expression than well-differentiated solid tumor comparators (NSCLC, localized prostate), with some cell lines expressing similar BCL2 RNA expression as AML. These neuroendocrine/small cell populations are the primary rationale for testing DNMTi + BCL2 inhibitor in solid tumors.")
```

```{r ne-lineages-summary-table, cache = FALSE}
# Table 3: per-lineage (all sensitive lineages, same cohort N as two-group) + two-group summaries + Fig. 2b multi-lineage p row
# CRISPR dependency long (cohort lines); N(CERES) can be < expression N
crispr_bcl2_t3 <- NULL
if (exists("crispr") && nrow(crispr) > 0L && exists("target_ids") && length(target_ids) > 0L && exists("expr_target")) {
  crispr_bcl2_t3 <- crispr %>%
    dplyr::filter(grepl("^BCL2\\b", .data$gene_name), .data$depmap_id %in% target_ids) %>%
    dplyr::left_join(expr_target %>% dplyr::select(.data$depmap_id, .data$lineage_group), by = "depmap_id") %>%
    dplyr::filter(!is.na(.data$lineage_group)) %>%
    dplyr::transmute(.data$depmap_id, .data$lineage_group, BCL2_ceres = .data$dependency)
}
crispr_dnmt1_t3 <- NULL
if (exists("crispr_dnmt1") && nrow(crispr_dnmt1) > 0L && exists("expr_target")) {
  ids_dnmt1 <- if (exists("target_ids") && length(target_ids) > 0L) {
    target_ids
  } else {
    unique(expr_target$depmap_id)
  }
  crispr_dnmt1_t3 <- crispr_dnmt1 %>%
    dplyr::filter(.data$depmap_id %in% ids_dnmt1) %>%
    dplyr::filter(!is.na(.data$lineage_group)) %>%
    dplyr::transmute(.data$depmap_id, .data$lineage_group, DNMT1_ceres = .data$dependency)
}
has_dnmt1_tbl <- exists("expr_meta_dnmt1") && is.data.frame(expr_meta_dnmt1) &&
  "DNMT1" %in% names(expr_meta_dnmt1) && nrow(expr_meta_dnmt1) > 0L
has_dnmt1_ceres_tbl <- !is.null(crispr_dnmt1_t3) && nrow(crispr_dnmt1_t3) > 0L
fmt_p_scalar <- function(pv) {
  if (length(pv) != 1L || is.na(pv)) {
    return("—")
  }
  if (pv < 0.001) {
    return("<0.001")
  }
  format.pval(pv, digits = 2)
}
# Multi-lineage BCL2 / DNMT1 CERES p as on Figure 2b corner labels (NI vs AML when computed; else Kruskal)
t3_p_bcl2_ceres_fig2b_multi <- if (exists("bcl2_ceres_ni_p") && !is.na(bcl2_ceres_ni_p)) {
  fmt_p_scalar(bcl2_ceres_ni_p)
} else if (exists("kw_bcl2") && inherits(kw_bcl2, "htest") && length(kw_bcl2$p.value) == 1L && !is.na(kw_bcl2$p.value)) {
  fmt_p_scalar(unname(kw_bcl2$p.value))
} else {
  "—"
}
t3_p_dnmt1_ceres_fig2b_multi <- if (exists("dnmt1_ceres_ni_p") && !is.na(dnmt1_ceres_ni_p)) {
  fmt_p_scalar(dnmt1_ceres_ni_p)
} else if (exists("kw_dnmt1") && inherits(kw_dnmt1, "htest") && length(kw_dnmt1$p.value) == 1L && !is.na(kw_dnmt1$p.value)) {
  fmt_p_scalar(unname(kw_dnmt1$p.value))
} else {
  "—"
}
# Same cohort as two-group summaries (per-lineage N sums to stem-like N)
lineage_plastic_t3 <- if (exists("lineage_plastic_ref_canonical")) lineage_plastic_ref_canonical else c("SCLC", "ExPulm-SCNEC", "AML (control)")
hma_t3_ext <- exists("hma_bcl2_tnbc_lymph_master") && isTRUE(hma_bcl2_tnbc_lymph_master)
t3_use_f1_nb <- exists("f1_enriched_lineages") && length(f1_enriched_lineages) > 0L
if (t3_use_f1_nb) {
  lineage_plastic_t3 <- f1_enriched_lineages
} else if (hma_t3_ext) {
  lineage_plastic_t3 <- unique(c(lineage_plastic_t3, "DLBCL/MCL/CLL"))
}
plastic_lbl_t3 <- "Stem-like"
other_lbl_t3 <- "Non-stem-like"
expr_two_data <- expr_target
if (exists("expr_meta_dnmt1") && "DNMT1" %in% names(expr_meta_dnmt1) && nrow(expr_meta_dnmt1) > 0L) {
  expr_two_data <- expr_two_data %>% dplyr::left_join(expr_meta_dnmt1 %>% dplyr::select(.data$depmap_id, .data$DNMT1), by = "depmap_id")
}
if ("DNMT1" %in% names(expr_two_data)) {
  expr_two_data <- expr_two_data %>% dplyr::filter(!is.na(.data$BCL2) | !is.na(.data$DNMT1))
} else {
  expr_two_data <- expr_two_data %>% dplyr::filter(!is.na(.data$BCL2))
}
if (exists("f1_enriched_lineages_ordered") && length(f1_enriched_lineages_ordered) > 0L) {
  t3_f1_lineages_ordered <- f1_enriched_lineages_ordered[
    as.character(f1_enriched_lineages_ordered) %in% as.character(lineage_plastic_t3)
  ]
} else {
  t3_prio <- c("SCLC", "ExPulm-SCNEC", "AML (control)")
  if (isTRUE(hma_t3_ext)) {
    t3_prio <- c(t3_prio, "DLBCL/MCL/CLL")
  }
  t3_f1_lineages_ordered <- c(
    intersect(t3_prio, as.character(lineage_plastic_t3)),
    setdiff(as.character(lineage_plastic_t3), t3_prio)
  )
}
tbl_ne_summary <- NULL
# Stem-like lineages, per lineage (Figure 1 order; includes SCLC, ExPulm-SCNEC, AML when in arm)
if (length(t3_f1_lineages_ordered) > 0L) {
  t3_f1_dat <- expr_two_data %>%
    dplyr::filter(as.character(.data$lineage_group) %in% as.character(t3_f1_lineages_ordered))
  if (nrow(t3_f1_dat) > 0L) {
      sum_f1 <- t3_f1_dat %>%
        dplyr::group_by(.data$lineage_group) %>%
        dplyr::summarise(
          N = dplyr::n(),
          `BCL2 median` = round(stats::median(.data$BCL2, na.rm = TRUE), 3),
          `BCL2 IQR` = paste0(
            round(stats::quantile(.data$BCL2, 0.25, na.rm = TRUE), 2), "–",
            round(stats::quantile(.data$BCL2, 0.75, na.rm = TRUE), 2)
          ),
          .groups = "drop"
        )
      if (has_dnmt1_tbl && "DNMT1" %in% names(t3_f1_dat) && sum(!is.na(t3_f1_dat$DNMT1)) > 0L) {
        d1f <- t3_f1_dat %>%
          dplyr::group_by(.data$lineage_group) %>%
          dplyr::summarise(
            `DNMT1 median` = round(stats::median(.data$DNMT1, na.rm = TRUE), 3),
            `DNMT1 IQR` = paste0(
              round(stats::quantile(.data$DNMT1, 0.25, na.rm = TRUE), 2), "–",
              round(stats::quantile(.data$DNMT1, 0.75, na.rm = TRUE), 2)
            ),
            .groups = "drop"
          )
        sum_f1 <- sum_f1 %>% dplyr::left_join(d1f, by = "lineage_group")
      }
      sum_f1 <- sum_f1 %>%
        dplyr::mutate(`lg_ord` = match(as.character(.data$lineage_group), as.character(t3_f1_lineages_ordered))) %>%
        dplyr::arrange(.data$`lg_ord`) %>%
        dplyr::mutate(Lineage = gsub(" \\(control\\)", "", as.character(.data$lineage_group))) %>%
        dplyr::select(-.data$lineage_group, -.data$`lg_ord`)
      # CERES per lineage (sensitive arm)
      if (!is.null(crispr_bcl2_t3) && nrow(crispr_bcl2_t3) > 0L) {
        ceres_f1_b <- crispr_bcl2_t3 %>%
          dplyr::filter(as.character(.data$lineage_group) %in% as.character(t3_f1_lineages_ordered)) %>%
          dplyr::group_by(.data$lineage_group) %>%
          dplyr::summarise(
            `BCL2 CERES median` = round(stats::median(.data$BCL2_ceres, na.rm = TRUE), 3),
            `BCL2 CERES IQR` = paste0(
              round(stats::quantile(.data$BCL2_ceres, 0.25, na.rm = TRUE), 2), "–",
              round(stats::quantile(.data$BCL2_ceres, 0.75, na.rm = TRUE), 2)
            ),
            .groups = "drop"
          ) %>%
          dplyr::mutate(Lineage = gsub(" \\(control\\)", "", as.character(.data$lineage_group))) %>%
          dplyr::select(-.data$lineage_group)
        sum_f1 <- sum_f1 %>% dplyr::left_join(ceres_f1_b, by = "Lineage")
      }
      if (!is.null(crispr_dnmt1_t3) && nrow(crispr_dnmt1_t3) > 0L) {
        ceres_f1_d <- crispr_dnmt1_t3 %>%
          dplyr::filter(as.character(.data$lineage_group) %in% as.character(t3_f1_lineages_ordered)) %>%
          dplyr::group_by(.data$lineage_group) %>%
          dplyr::summarise(
            `DNMT1 CERES median` = round(stats::median(.data$DNMT1_ceres, na.rm = TRUE), 3),
            `DNMT1 CERES IQR` = paste0(
              round(stats::quantile(.data$DNMT1_ceres, 0.25, na.rm = TRUE), 2), "–",
              round(stats::quantile(.data$DNMT1_ceres, 0.75, na.rm = TRUE), 2)
            ),
            .groups = "drop"
          ) %>%
          dplyr::mutate(Lineage = gsub(" \\(control\\)", "", as.character(.data$lineage_group))) %>%
          dplyr::select(-.data$lineage_group)
        sum_f1 <- sum_f1 %>% dplyr::left_join(ceres_f1_d, by = "Lineage")
      }
      if (!"BCL2 CERES median" %in% names(sum_f1)) {
        sum_f1 <- sum_f1 %>% dplyr::mutate(`BCL2 CERES median` = NA_real_, `BCL2 CERES IQR` = "")
      }
      if (has_dnmt1_ceres_tbl && !"DNMT1 CERES median" %in% names(sum_f1)) {
        sum_f1 <- sum_f1 %>% dplyr::mutate(`DNMT1 CERES median` = NA_real_, `DNMT1 CERES IQR` = "")
      }
      if (has_dnmt1_tbl && !"DNMT1 median" %in% names(sum_f1)) {
        sum_f1 <- sum_f1 %>% dplyr::mutate(`DNMT1 median` = NA_real_, `DNMT1 IQR` = "")
      }
      sum_f1 <- sum_f1 %>%
        dplyr::mutate(
          `p (BCL2)` = "",
          `p (BCL2 CERES)` = "",
          `p (DNMT1)` = if (has_dnmt1_tbl) "" else NA_character_,
          `p (DNMT1 CERES)` = if (has_dnmt1_ceres_tbl) "" else NA_character_
        )
      if (!has_dnmt1_tbl) {
        sum_f1 <- sum_f1 %>% dplyr::mutate(`p (DNMT1)` = NULL)
      }
      if (!has_dnmt1_ceres_tbl) {
        sum_f1 <- sum_f1 %>% dplyr::mutate(`p (DNMT1 CERES)` = NULL)
      }
      hdr_f1 <- tibble::tibble(
        Lineage = "Stem-like lineages (per lineage)",
        N = NA_integer_,
        `BCL2 median` = NA_real_,
        `BCL2 IQR` = "",
        `BCL2 CERES median` = NA_real_,
        `BCL2 CERES IQR` = ""
      )
      if (has_dnmt1_tbl) {
        hdr_f1 <- hdr_f1 %>% tibble::add_column(`DNMT1 median` = NA_real_, `DNMT1 IQR` = "", .after = "BCL2 IQR")
      }
      if (has_dnmt1_ceres_tbl) {
        hdr_f1 <- hdr_f1 %>% tibble::add_column(`DNMT1 CERES median` = NA_real_, `DNMT1 CERES IQR` = "", .after = "BCL2 CERES IQR")
      }
      hdr_f1 <- hdr_f1 %>%
        dplyr::mutate(
          `p (BCL2)` = "",
          `p (BCL2 CERES)` = "",
          `p (DNMT1)` = if (has_dnmt1_tbl) "" else NA_character_,
          `p (DNMT1 CERES)` = if (has_dnmt1_ceres_tbl) "" else NA_character_
        )
      if (!has_dnmt1_tbl) {
        hdr_f1 <- hdr_f1 %>% dplyr::mutate(`p (DNMT1)` = NULL)
      }
      if (!has_dnmt1_ceres_tbl) {
        hdr_f1 <- hdr_f1 %>% dplyr::mutate(`p (DNMT1 CERES)` = NULL)
      }
      n_sensitive_two <- nrow(dplyr::filter(
        expr_two_data,
        as.character(.data$lineage_group) %in% as.character(lineage_plastic_t3)
      ))
      stopifnot(sum(sum_f1$N) == n_sensitive_two)
      tbl_ne_summary <- dplyr::bind_rows(tbl_ne_summary, hdr_f1, sum_f1)
  }
}
# Two-group: stem-like vs non-stem-like (expr_two_data already built above)
rows_two_group <- NULL
if (exists("expr_two_data") && nrow(expr_two_data) > 0L) {
  expr_two_data <- expr_two_data %>%
    dplyr::mutate(group_2 = dplyr::if_else(
      as.character(.data$lineage_group) %in% as.character(lineage_plastic_t3),
      plastic_lbl_t3,
      other_lbl_t3
    ))
  rows_two_group <- expr_two_data %>%
    dplyr::group_by(.data$group_2) %>%
    dplyr::summarise(
      N = dplyr::n(),
      BCL2_median = round(stats::median(.data$BCL2, na.rm = TRUE), 3),
      BCL2_IQR = paste0(
        round(stats::quantile(.data$BCL2, 0.25, na.rm = TRUE), 2), "–",
        round(stats::quantile(.data$BCL2, 0.75, na.rm = TRUE), 2)
      ),
      .groups = "drop"
    )
  if ("DNMT1" %in% names(expr_two_data) && sum(!is.na(expr_two_data$DNMT1)) > 0L) {
    dnmt1_two <- expr_two_data %>%
      dplyr::group_by(.data$group_2) %>%
      dplyr::summarise(
        DNMT1_median = round(stats::median(.data$DNMT1, na.rm = TRUE), 3),
        DNMT1_IQR = paste0(
          round(stats::quantile(.data$DNMT1, 0.25, na.rm = TRUE), 2), "–",
          round(stats::quantile(.data$DNMT1, 0.75, na.rm = TRUE), 2)
        ),
        .groups = "drop"
      )
    rows_two_group <- rows_two_group %>% dplyr::left_join(dnmt1_two, by = "group_2")
  } else if (has_dnmt1_tbl) {
    rows_two_group <- rows_two_group %>% dplyr::mutate(DNMT1_median = NA_real_, DNMT1_IQR = "")
  }
  # Two-group CERES summaries
  if (!is.null(crispr_bcl2_t3) && nrow(crispr_bcl2_t3) > 0L) {
    ceres_two_b <- crispr_bcl2_t3 %>%
      dplyr::inner_join(expr_two_data %>% dplyr::select(.data$depmap_id, .data$group_2), by = "depmap_id") %>%
      dplyr::group_by(.data$group_2) %>%
      dplyr::summarise(
        `BCL2 CERES median` = round(stats::median(.data$BCL2_ceres, na.rm = TRUE), 3),
        `BCL2 CERES IQR` = paste0(
          round(stats::quantile(.data$BCL2_ceres, 0.25, na.rm = TRUE), 2), "–",
          round(stats::quantile(.data$BCL2_ceres, 0.75, na.rm = TRUE), 2)
        ),
        .groups = "drop"
      )
    rows_two_group <- rows_two_group %>% dplyr::left_join(ceres_two_b, by = "group_2")
  }
  if (has_dnmt1_ceres_tbl && !is.null(crispr_dnmt1_t3) && nrow(crispr_dnmt1_t3) > 0L) {
    ceres_two_d <- crispr_dnmt1_t3 %>%
      dplyr::inner_join(expr_two_data %>% dplyr::select(.data$depmap_id, .data$group_2), by = "depmap_id") %>%
      dplyr::group_by(.data$group_2) %>%
      dplyr::summarise(
        `DNMT1 CERES median` = round(stats::median(.data$DNMT1_ceres, na.rm = TRUE), 3),
        `DNMT1 CERES IQR` = paste0(
          round(stats::quantile(.data$DNMT1_ceres, 0.25, na.rm = TRUE), 2), "–",
          round(stats::quantile(.data$DNMT1_ceres, 0.75, na.rm = TRUE), 2)
        ),
        .groups = "drop"
      )
    rows_two_group <- rows_two_group %>% dplyr::left_join(ceres_two_d, by = "group_2")
  }
  if (!"BCL2 CERES median" %in% names(rows_two_group)) {
    rows_two_group <- rows_two_group %>% dplyr::mutate(`BCL2 CERES median` = NA_real_, `BCL2 CERES IQR` = "")
  }
  if (has_dnmt1_ceres_tbl && !"DNMT1 CERES median" %in% names(rows_two_group)) {
    rows_two_group <- rows_two_group %>% dplyr::mutate(`DNMT1 CERES median` = NA_real_, `DNMT1 CERES IQR` = "")
  }
  rows_two_group <- rows_two_group %>%
    dplyr::mutate(
      Lineage = dplyr::if_else(.data$group_2 == plastic_lbl_t3, "Stem-like", "Non-stem-like"),
      `p (BCL2)` = "",
      `p (BCL2 CERES)` = "",
      `p (DNMT1)` = if (has_dnmt1_tbl) "" else NA_character_,
      `p (DNMT1 CERES)` = if (has_dnmt1_ceres_tbl) "" else NA_character_
    ) %>%
    dplyr::select(-.data$group_2)
  if (!has_dnmt1_tbl) {
    rows_two_group <- rows_two_group %>% dplyr::mutate(`p (DNMT1)` = NULL)
  }
  if (!has_dnmt1_ceres_tbl) {
    rows_two_group <- rows_two_group %>% dplyr::mutate(`p (DNMT1 CERES)` = NULL)
  }
  rows_two_group <- rows_two_group %>%
    dplyr::rename(`BCL2 median` = .data$BCL2_median, `BCL2 IQR` = .data$BCL2_IQR)
  if (has_dnmt1_tbl) {
    rows_two_group <- rows_two_group %>% dplyr::rename(`DNMT1 median` = .data$DNMT1_median, `DNMT1 IQR` = .data$DNMT1_IQR)
  }
  rows_two_group <- rows_two_group %>%
    dplyr::select(dplyr::any_of(c(
      "Lineage", "N", "BCL2 median", "BCL2 IQR", "DNMT1 median", "DNMT1 IQR",
      "BCL2 CERES median", "BCL2 CERES IQR", "DNMT1 CERES median", "DNMT1 CERES IQR",
      "p (BCL2)", "p (DNMT1)", "p (BCL2 CERES)", "p (DNMT1 CERES)"
    )))
}
t3_p_row_shell <- function(lineage_lbl) {
  pr <- tibble::tibble(
    Lineage = lineage_lbl,
    N = NA_integer_,
    `BCL2 median` = NA_real_,
    `BCL2 IQR` = "",
    `BCL2 CERES median` = NA_real_,
    `BCL2 CERES IQR` = ""
  )
  if (has_dnmt1_tbl) {
    pr <- pr %>% tibble::add_column(`DNMT1 median` = NA_real_, `DNMT1 IQR` = "", .after = "BCL2 IQR")
  }
  if (has_dnmt1_ceres_tbl) {
    pr <- pr %>% tibble::add_column(`DNMT1 CERES median` = NA_real_, `DNMT1 CERES IQR` = "", .after = "BCL2 CERES IQR")
  }
  pr
}
# Multi-lineage subpanels: same p text as BCL2/DNMT1 expression and CERES multi-lineage plots (kw_txt / kw_dnmt1_txt; NI or Kruskal for CERES)
p_row_multi <- t3_p_row_shell("p (Fig. 2b multi-lineage)") %>%
  dplyr::mutate(
    `p (BCL2)` = if (exists("kw_txt")) kw_txt else "—",
    `p (DNMT1)` = if (has_dnmt1_tbl) (if (exists("kw_dnmt1_txt")) kw_dnmt1_txt else "—") else NA_character_,
    `p (BCL2 CERES)` = t3_p_bcl2_ceres_fig2b_multi,
    `p (DNMT1 CERES)` = if (has_dnmt1_ceres_tbl) t3_p_dnmt1_ceres_fig2b_multi else NA_character_
  )
if (!has_dnmt1_tbl) {
  p_row_multi <- p_row_multi %>% dplyr::mutate(`p (DNMT1)` = NULL)
}
if (!has_dnmt1_ceres_tbl) {
  p_row_multi <- p_row_multi %>% dplyr::mutate(`p (DNMT1 CERES)` = NULL)
}
if (!is.null(rows_two_group) && nrow(rows_two_group) > 0L) {
  tbl_ne_summary <- dplyr::bind_rows(tbl_ne_summary, rows_two_group, p_row_multi)
} else {
  tbl_ne_summary <- dplyr::bind_rows(tbl_ne_summary, p_row_multi)
}
t3_kable_cols <- c(
  "Lineage", "N", "BCL2 median", "BCL2 IQR", "DNMT1 median", "DNMT1 IQR",
  "BCL2 CERES median", "BCL2 CERES IQR", "DNMT1 CERES median", "DNMT1 CERES IQR",
  "p (BCL2)", "p (DNMT1)", "p (BCL2 CERES)", "p (DNMT1 CERES)"
)
tbl_ne_summary <- tbl_ne_summary %>% dplyr::select(dplyr::any_of(t3_kable_cols))
knitr::kable(tbl_ne_summary, caption = paste0(
  "Table 3. BCL2 and DNMT1 RNA (log2 TPM+1) and CRISPR CERES dependency (median, IQR) in the extended cohort. ",
  "Per-lineage section (when present): each row is one stem-like lineage in Figure 1 order; per-lineage N uses the same expression cohort as the two-group summaries ",
  "(sum of per-lineage N equals N in the stem-like summary row). ",
  "CERES uses DepMap lines with dependency scores (N can be smaller than RNA N). ",
  "Two summary rows compare stem-like vs non-stem-like. ",
  "p (Fig. 2b multi-lineage): same values as the multi-lineage subpanels in Figure 2b—BCL2 and DNMT1 RNA use the annotated expression tests ",
  "(NI vs AML, 1-sided, when shown; else as in figure caption); BCL2 and DNMT1 CERES use the annotated multi-lineage tests ",
  "(NI vs AML, 1-sided, when computed; otherwise Kruskal–Wallis across lineages in that panel)."
))
```

```{r interpretation-03-table3-fig2b, echo = FALSE, results = "asis", eval = !exists("hma_bcl2_tnbc_lymph_master") || !isTRUE(hma_bcl2_tnbc_lymph_master)}
kw_s <- if (exists("kw_txt")) kw_txt else "—"
wx_s <- if (exists("wx_txt")) wx_txt else "—"
kwd_s <- if (exists("kw_dnmt1_txt")) kw_dnmt1_txt else "—"
wxd_s <- if (exists("wx_dnmt1_txt")) wx_dnmt1_txt else "—"
knitr::asis_output(paste0(
  "**Interpretation (Figure 2b subfigures in Figures 1 and 2)**: **BCL2:** Multi-lineage subpanel reports **non-inferior vs AML** (1-sided Wilcoxon, BCL2 stochastically ≥ AML; p = ", kw_s, "); two-group subpanel compares stem-like vs non-stem-like (**1-sided** Wilcoxon, stem-like > non-stem-like; p = ", wx_s, "). **DNMT1:** When the Fig. 1 arm applies, multi-lineage subpanel reports **non-inferior vs AML** (1-sided Wilcoxon, DNMT1 stochastically ≤ AML; p = ", kwd_s, "); otherwise Kruskal–Wallis across lineages (same p text). Two-group Wilcoxon (p = ", wxd_s, ")."
))
```
