---
title: "Understanding Differential Observation Time vs Immortal Time Bias"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(gridExtra)
```

## The Problem: Differential Observation Time

### Scenario: Post-Cystectomy ctDNA Analysis

In your analysis, you're trying to predict which patients will have **positive ctDNA after cystectomy**. However, patients have different follow-up durations:

```{r create_visualization_data}
# Create example data to illustrate the problem
set.seed(123)

# Patient 1: Only 3 months of follow-up
patient1 <- data.frame(
  patient = "Patient A",
  time = seq(0, 3, by = 0.5),
  ctDNA_status = c(0, 0, 0, 0, 0, 0, 0),  # Never became positive
  follow_up_end = 3
)

# Patient 2: 12 months of follow-up
patient2 <- data.frame(
  patient = "Patient B",
  time = seq(0, 12, by = 0.5),
  ctDNA_status = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),  # Never became positive
  follow_up_end = 12
)

# Patient 3: 12 months, became positive at month 8
patient3 <- data.frame(
  patient = "Patient C",
  time = seq(0, 12, by = 0.5),
  ctDNA_status = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1),
  follow_up_end = 12
)

# Combine
all_patients <- bind_rows(patient1, patient2, patient3)
```

```{r plot_differential_observation, fig.width=12, fig.height=6, fig.cap="Differential Observation Time: Patients have different opportunities to develop detectable ctDNA"}
# Create visualization
p1 <- ggplot(all_patients, aes(x = time, y = patient, color = factor(ctDNA_status))) +
  geom_segment(aes(x = 0, xend = follow_up_end, y = patient, yend = patient), 
               color = "gray70", linewidth = 3, alpha = 0.3) +
  geom_point(size = 4, alpha = 0.7) +
  scale_color_manual(values = c("0" = "blue", "1" = "red"),
                     labels = c("0" = "Negative", "1" = "Positive"),
                     name = "ctDNA Status") +
  labs(
    x = "Time from Cystectomy (months)",
    y = "",
    title = "Differential Observation Time Problem",
    subtitle = "Patient A has only 3 months to develop positive ctDNA, while Patients B & C have 12 months"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11),
    legend.position = "bottom"
  ) +
  geom_vline(xintercept = 3, linetype = "dashed", color = "orange", linewidth = 1, alpha = 0.7) +
  geom_vline(xintercept = 6, linetype = "dashed", color = "green", linewidth = 1, alpha = 0.7) +
  annotate("text", x = 3, y = 3.3, label = "3-month\nlandmark", hjust = 0.5, size = 3, color = "orange") +
  annotate("text", x = 6, y = 3.3, label = "6-month\nlandmark", hjust = 0.5, size = 3, color = "green")

p1
```

**The Problem:**
- **Patient A** (3 months follow-up): Never had a chance to develop positive ctDNA after month 3
- **Patient B** (12 months follow-up): Had 4x more opportunity but still negative
- **Patient C** (12 months follow-up): Became positive at month 8

If we use **any post-cystectomy ctDNA** (regardless of timing), we're comparing:
- Patients with 3 months of observation vs patients with 12 months of observation
- This creates **differential opportunity** for the outcome to occur

---

## How This Differs from Immortal Time Bias

### Immortal Time Bias: A Period Where Outcome CANNOT Occur

```{r create_immortal_time_example}
# Example: Studying "time from treatment to death"
# But only including patients who survived long enough to receive treatment

immortal_time_example <- data.frame(
  scenario = c("Patient Selection", "Immortal Time Period", "Follow-up"),
  start = c(0, 0, 3),
  end = c(0, 3, 12),
  status = c("Alive (required for study entry)", "Cannot die (immortal)", "Can die")
)

p2 <- ggplot(immortal_time_example, aes(x = start, xend = end, y = scenario, yend = scenario)) +
  geom_segment(linewidth = 4, alpha = 0.7, color = c("green", "red", "blue")) +
  geom_text(aes(x = (start + end)/2, label = status), 
            vjust = -1.5, size = 3.5, fontface = "bold") +
  labs(
    x = "Time",
    y = "",
    title = "Immortal Time Bias",
    subtitle = "A period where the outcome (death) CANNOT occur due to study design"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", color = "red"),
    plot.subtitle = element_text(size = 11),
    axis.text.y = element_text(size = 10, face = "bold")
  ) +
  xlim(-1, 13)

p2
```

**Key Difference:**

| Aspect | Differential Observation Time | Immortal Time Bias |
|--------|-------------------------------|-------------------|
| **Mechanism** | Different **opportunities** for outcome | Outcome **cannot occur** during a period |
| **Cause** | Varying follow-up durations | Study design/exclusion criteria |
| **Example** | 3 months vs 12 months follow-up | Only including patients who survived to receive treatment |
| **Bias Type** | Length bias / Observation time bias | Selection bias / Time-dependent bias |
| **Solution** | Landmark analysis (fix observation time) | Proper time-zero definition / Landmark analysis |

---

## Side-by-Side Comparison

```{r comparison_plot, fig.width=14, fig.height=8}
# Create comparison visualization

# Panel 1: Differential Observation Time
comparison_data1 <- data.frame(
  patient = rep(c("Short FU\n(3 mo)", "Long FU\n(12 mo)"), each = 13),
  time = rep(seq(0, 12, by = 1), 2),
  can_develop_outcome = c(
    rep(1, 4), rep(0, 9),  # Short FU: can develop in months 0-3, then observation ends
    rep(1, 13)  # Long FU: can develop at any time
  )
)

p3 <- ggplot(comparison_data1, aes(x = time, y = patient, fill = factor(can_develop_outcome))) +
  geom_tile(width = 0.8, height = 0.8, alpha = 0.7) +
  scale_fill_manual(values = c("0" = "lightgray", "1" = "lightblue"),
                    labels = c("0" = "No observation", "1" = "Can develop outcome"),
                    name = "") +
  labs(
    x = "Time (months)",
    y = "",
    title = "Differential Observation Time",
    subtitle = "Different OPPORTUNITIES for outcome to occur"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    legend.position = "bottom"
  ) +
  geom_vline(xintercept = 3, linetype = "dashed", color = "red", linewidth = 1)

# Panel 2: Immortal Time Bias
comparison_data2 <- data.frame(
  period = c("Before Treatment", "Immortal Period", "After Treatment"),
  start = c(0, 0, 3),
  end = c(0, 3, 12),
  outcome_possible = c(0, 0, 1)  # 0 = cannot occur, 1 = can occur
)

p4 <- ggplot(comparison_data2, aes(x = start, xend = end, y = period, yend = period)) +
  geom_segment(aes(color = factor(outcome_possible)), linewidth = 6, alpha = 0.8) +
  scale_color_manual(values = c("0" = "red", "1" = "green"),
                     labels = c("0" = "Outcome CANNOT occur", "1" = "Outcome CAN occur"),
                     name = "") +
  labs(
    x = "Time",
    y = "",
    title = "Immortal Time Bias",
    subtitle = "A period where outcome CANNOT occur"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    legend.position = "bottom"
  ) +
  xlim(-0.5, 12.5)

grid.arrange(p3, p4, ncol = 2)
```

---

## How Landmark Analysis Solves the Differential Observation Time Problem

```{r landmark_solution, fig.width=12, fig.height=6}
# Show how landmark analysis fixes the problem
landmark_data <- data.frame(
  patient = rep(c("Patient A", "Patient B", "Patient C"), each = 13),
  time = rep(seq(0, 12, by = 1), 3),
  has_ctDNA_measurement = c(
    c(1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0),  # Patient A: measured at 0, 1, 2, 3 months
    c(1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0),  # Patient B: measured at 0, 6 months
    c(1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0)   # Patient C: measured at 0, 6 months
  ),
  ctDNA_positive = c(
    c(0, 0, 0, 0, NA, NA, NA, NA, NA, NA, NA, NA, NA),
    c(0, NA, NA, NA, NA, 0, NA, NA, NA, NA, NA, NA, NA),
    c(0, NA, NA, NA, NA, 1, NA, NA, NA, NA, NA, NA, NA)
  )
)

# Filter to only show measurements
landmark_plot_data <- landmark_data %>%
  filter(has_ctDNA_measurement == 1)

p5 <- ggplot(landmark_data, aes(x = time, y = patient)) +
  # Show observation periods as background
  geom_rect(data = data.frame(patient = c("Patient A", "Patient B", "Patient C"),
                              xmin = c(0, 0, 0), xmax = c(3, 12, 12),
                              ymin = c(0.7, 1.7, 2.7), ymax = c(1.3, 2.3, 3.3)),
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "gray90", alpha = 0.3, inherit.aes = FALSE) +
  # Mark 6-month landmark
  geom_vline(xintercept = 6, linetype = "dashed", color = "green", linewidth = 2, alpha = 0.7) +
  # Show measurements
  geom_point(data = landmark_plot_data,
             aes(color = factor(ctDNA_positive), size = has_ctDNA_measurement), 
             alpha = 0.8) +
  scale_color_manual(values = c("0" = "blue", "1" = "red"),
                     labels = c("0" = "Negative", "1" = "Positive"),
                     name = "ctDNA Status",
                     na.value = "gray", drop = FALSE) +
  scale_size_continuous(range = c(3, 6), guide = "none") +
  labs(
    x = "Time from Cystectomy (months)",
    y = "",
    title = "Landmark Analysis Solution: 6-Month Landmark",
    subtitle = "All patients evaluated at the SAME timepoint (6 months), ensuring equal observation opportunity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", color = "green"),
    plot.subtitle = element_text(size = 11),
    legend.position = "right"
  ) +
  annotate("text", x = 6, y = 3.5, label = "6-Month\nLandmark\n(Equal observation time)", 
           hjust = 0.5, size = 4, color = "green", fontface = "bold")

p5
```

**Key Benefits of Landmark Analysis:**
1. **Equal observation time**: All patients are evaluated at the same timepoint (e.g., 6 months)
2. **Fair comparison**: Patients with different total follow-up durations are now comparable
3. **Reduces bias**: Eliminates the advantage/disadvantage of having longer/shorter follow-up

---

## Summary

**Differential Observation Time (Your Issue):**
- Patients have **different opportunities** to develop the outcome
- Solution: **Landmark analysis** - evaluate everyone at the same timepoint

**Immortal Time Bias:**
- There's a period where the outcome **cannot occur** due to study design
- Solution: Proper time-zero definition, landmark analysis, or time-dependent analysis

**Both can be addressed with landmark analysis**, but the underlying mechanisms differ!
