---
title: "Patients Missing Bajorin Risk Factors"
date: "`r Sys.Date()`"
output: html_document
---

# Patients Missing Bajorin Risk Factors

This document identifies patients who were included in the final analysis cohorts for **Question 1 (Baseline ctDNA Prognosis)** and **Question 2 (On-Treatment ctDNA Response)** but did not have Bajorin risk factors (KPS and visceral metastases) available.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "")
knitr::opts_knit$set(root.dir = "~/R")
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Load shared setup
source("ctDNA_Research_Questions_Shared_Setup.R")

# Load required libraries
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(stringr)
  library(readxl)
  library(janitor)
  library(lubridate)
  library(knitr)
  library(kableExtra)
})
```

```{r load_data, include=FALSE}
# Load minimal data
data_list <- load_minimal_data(include_survival = TRUE)
dat_combined <- data_list$dat_combined
ctdna_long <- data_list$ctdna_long
survival_data <- data_list$survival_data
```

```{r load_bajorin_data, include=FALSE}
# ====================================================================================
# Load Bajorin Risk Factors Data (KPS and Visceral Metastases)
# ====================================================================================
bajorin_data <- tryCatch({
  bajorin_raw <- read_excel("ctDNA_Bajorin_Lexi.xlsx", sheet = 1, skip = 1)
  
  cat("Bajorin file loaded:\n")
  cat("  Total rows:", nrow(bajorin_raw), "\n")
  cat("  Total columns:", ncol(bajorin_raw), "\n")
  
  # Check if we have enough columns (need at least 6: Patient ID, Therapy Line, KPS, Visceral Mets)
  if (nrow(bajorin_raw) > 0 && ncol(bajorin_raw) >= 6) {
    # Check column types
    cat("  Column 2 (Patient ID) class:", class(bajorin_raw[[2]]), "\n")
    cat("  Column 4 (Therapy Line) class:", class(bajorin_raw[[4]]), "\n")
    cat("  Column 5 (KPS) class:", class(bajorin_raw[[5]]), "\n")
    cat("  Column 6 (Visceral Mets) class:", class(bajorin_raw[[6]]), "\n")
    
    bajorin_clean <- tibble(
      patient_id_raw = bajorin_raw[[2]],  # Column 2: Patient ID
      therapy_line_raw = bajorin_raw[[4]],  # Column 4: Therapy Line
      kps_raw = bajorin_raw[[5]],  # Column 5: KPS (not column 6!)
      visceral_mets_raw = bajorin_raw[[6]]  # Column 6: Visceral Mets (not column 7!)
    ) %>%
      # Convert to character for processing
      mutate(
        patient_id_raw = as.character(patient_id_raw),
        therapy_line_raw = as.character(therapy_line_raw),
        kps_raw = as.character(kps_raw),
        visceral_mets_raw = as.character(visceral_mets_raw)
      ) %>%
      filter(!str_detect(tolower(patient_id_raw), "patient|id|screener|mrn|therapy")) %>%
      filter(!is.na(patient_id_raw), patient_id_raw != "", patient_id_raw != "NA") %>%
      mutate(
        # Convert patient_id to character - handle both numeric and character inputs
        # If numeric (like 22, 179), convert to character; if already character, keep as is
        # Remove any decimal points if Excel stored as numeric (e.g., 22.0 -> 22)
        patient_id_char = as.character(patient_id_raw),
        # Normalize: if it's a pure number (possibly with .0), convert to integer string
        patient_id_char = ifelse(!is.na(suppressWarnings(as.numeric(patient_id_char))) & 
                                 str_detect(patient_id_char, "^\\.?\\d+\\.?0*$"), 
                                 as.character(as.integer(suppressWarnings(as.numeric(patient_id_char)))), 
                                 patient_id_char),
        patient_id = str_trim(patient_id_char),
        therapy_line = case_when(
          str_detect(toupper(str_trim(therapy_line_raw)), "^1L$") ~ 1,
          str_detect(toupper(str_trim(therapy_line_raw)), "^2L$") ~ 2,
          str_detect(toupper(str_trim(therapy_line_raw)), "^3L$") ~ 3,
          str_detect(toupper(str_trim(therapy_line_raw)), "^4L\\+") ~ 4,
          str_detect(str_trim(therapy_line_raw), "^\\d+") ~ as.numeric(str_extract(str_trim(therapy_line_raw), "^\\d+")),
          # Also handle if therapy_line_raw is already a number
          !is.na(suppressWarnings(as.numeric(str_trim(therapy_line_raw)))) ~ as.numeric(str_trim(therapy_line_raw)),
          TRUE ~ NA_real_
        ),
        kps_binary = case_when(
          str_trim(kps_raw) == "1" | (!is.na(suppressWarnings(as.numeric(kps_raw))) & as.numeric(kps_raw) == 1) ~ 1,
          str_trim(kps_raw) == "0" | (!is.na(suppressWarnings(as.numeric(kps_raw))) & as.numeric(kps_raw) == 0) ~ 0,
          TRUE ~ NA_real_
        ),
        visceral_mets_binary = case_when(
          str_detect(tolower(str_trim(visceral_mets_raw)), "^yes$|^y$") ~ 1,
          str_detect(tolower(str_trim(visceral_mets_raw)), "^no$|^n$") ~ 0,
          TRUE ~ NA_real_
        )
      ) %>%
      filter(!is.na(patient_id), !is.na(therapy_line), patient_id != "") %>%
      select(patient_id, therapy_line, kps_binary, visceral_mets_binary)
    
    cat("Loaded Bajorin data:\n")
    cat("  Total observations:", nrow(bajorin_clean), "\n")
    cat("  Unique patients:", length(unique(bajorin_clean$patient_id)), "\n")
    cat("  Observations with KPS:", sum(!is.na(bajorin_clean$kps_binary)), "\n")
    cat("  Observations with Visceral Mets:", sum(!is.na(bajorin_clean$visceral_mets_binary)), "\n")
    cat("  Observations with both:", sum(!is.na(bajorin_clean$kps_binary) & !is.na(bajorin_clean$visceral_mets_binary)), "\n")
    cat("  Sample patient IDs:", paste(head(unique(bajorin_clean$patient_id), 10), collapse = ", "), "\n")
    
    # Show any rows that were filtered out
    total_before_filter <- nrow(tibble(
      patient_id_raw = as.character(bajorin_raw[[2]]),
      therapy_line_raw = as.character(bajorin_raw[[4]])
    ) %>% filter(!str_detect(tolower(patient_id_raw), "patient|id|screener|mrn|therapy")))
    
    if (nrow(bajorin_clean) < total_before_filter) {
      cat("  Warning:", total_before_filter - nrow(bajorin_clean), "rows filtered out (missing patient_id or therapy_line)\n")
    }
    
    bajorin_clean
  } else {
    tibble(patient_id = character(0), therapy_line = numeric(0), 
           kps_binary = numeric(0), visceral_mets_binary = numeric(0))
  }
}, error = function(e) {
  cat("Warning: Could not load Bajorin data file:", e$message, "\n")
  tibble(patient_id = character(0), therapy_line = numeric(0), 
         kps_binary = numeric(0), visceral_mets_binary = numeric(0))
})
```

```{r normalize_patient_ids_main_data, include=FALSE}
# Normalize patient IDs in main dataset to match Bajorin data format
# This ensures consistent matching when joining
dat_combined <- dat_combined %>%
  mutate(
    patient_id_orig = patient_id,
    patient_id_char = as.character(patient_id),
    # Normalize: if it's a pure number (possibly with .0), convert to integer string
    patient_id_char = ifelse(!is.na(suppressWarnings(as.numeric(patient_id_char))) & 
                             str_detect(patient_id_char, "^\\.?\\d+\\.?0*$"), 
                             as.character(as.integer(suppressWarnings(as.numeric(patient_id_char)))), 
                             patient_id_char),
    patient_id = str_trim(patient_id_char)
  ) %>%
  select(-patient_id_char, -patient_id_orig)

# Also normalize in ctdna_long
ctdna_long <- ctdna_long %>%
  mutate(
    patient_id_char = as.character(patient_id),
    patient_id_char = ifelse(!is.na(suppressWarnings(as.numeric(patient_id_char))) & 
                             str_detect(patient_id_char, "^\\.?\\d+\\.?0*$"), 
                             as.character(as.integer(suppressWarnings(as.numeric(patient_id_char)))), 
                             patient_id_char),
    patient_id = str_trim(patient_id_char)
  ) %>%
  select(-patient_id_char)
```

```{r define_universe_q1, include=FALSE}
# ====================================================================================
# QUESTION 1: Baseline ctDNA Prognosis Analysis
# Replicate universe and outcomes definition from Question1_Baseline_ctDNA_Prognosis.Rmd
# ====================================================================================

# Extract therapy start dates and settings
therapy_start_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("systemic.*therapy.*\\d+.*start.*date", ignore_case = TRUE))]
therapy_setting_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("systemic.*therapy.*\\d+.*treatment.*setting", ignore_case = TRUE))]

# Get all palliative therapy lines
all_palliative_therapy_lines <- tibble(patient_id = character(0), therapy_line = integer(0), therapy_start_date = as.Date(character(0)))

if (length(therapy_start_cols) > 0 && length(therapy_setting_cols) > 0) {
  therapy_dates_with_lines <- dat_combined %>%
    select(patient_id, any_of(therapy_start_cols)) %>%
    pivot_longer(-patient_id, names_to = "therapy_var", values_to = "start_date_raw") %>%
    mutate(
      start_date = as_date_safe(start_date_raw),
      therapy_line = as.integer(str_extract(therapy_var, "\\d+"))
    ) %>%
    filter(!is.na(start_date), !is.na(therapy_line)) %>%
    select(patient_id, therapy_line, therapy_start_date = start_date)
  
  therapy_settings_with_lines <- dat_combined %>%
    select(patient_id, any_of(therapy_setting_cols)) %>%
    pivot_longer(-patient_id, names_to = "setting_var", values_to = "setting_raw") %>%
    mutate(
      setting = as.character(setting_raw),
      therapy_line = as.integer(str_extract(setting_var, "\\d+"))
    ) %>%
    filter(!is.na(therapy_line)) %>%
    select(patient_id, therapy_line, setting)
  
  all_palliative_therapy_lines <- therapy_dates_with_lines %>%
    left_join(therapy_settings_with_lines, by = c("patient_id", "therapy_line")) %>%
    mutate(
      setting_lower = tolower(setting),
      is_palliative = str_detect(setting_lower, "palliative|metastatic|1l.*metastatic|2l.*metastatic|3l.*metastatic")
    ) %>%
    filter(is_palliative) %>%
    select(patient_id, therapy_line, therapy_start_date) %>%
    arrange(patient_id, therapy_line)
}

# Identify patients with confirmed metastatic disease
clinical_state_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("clinical_disease_state.*metastatic|clinical_disease_state.*first.*ct", ignore_case = TRUE))]
metastatic_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("metastatic_disease_ever_diagnosed|metastatic.*yes.*no", ignore_case = TRUE))]
m_stage_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("pathology_m_stage", ignore_case = TRUE))]

first_palliative_therapy <- all_palliative_therapy_lines %>%
  group_by(patient_id) %>%
  summarise(first_palliative_start = min(therapy_start_date, na.rm = TRUE), .groups = "drop")

metastatic_status <- dat_combined %>%
  select(patient_id, 
         any_of(clinical_state_cols),
         any_of(metastatic_cols),
         any_of(m_stage_cols)) %>%
  mutate(
    metastatic_at_ctdna = if (length(clinical_state_cols) > 0) {
      str_detect(tolower(as.character(.data[[clinical_state_cols[1]]])), "metastatic|advanced")
    } else {
      NA
    },
    metastatic_ever_diagnosed = if (length(metastatic_cols) > 0) {
      case_when(
        str_detect(tolower(as.character(.data[[metastatic_cols[1]]])), "yes|y|1") ~ TRUE,
        str_detect(tolower(as.character(.data[[metastatic_cols[1]]])), "no|n|0") ~ FALSE,
        TRUE ~ NA
      )
    } else {
      NA
    },
    m_stage_metastatic = if (length(m_stage_cols) > 0) {
      str_detect(toupper(as.character(.data[[m_stage_cols[1]]])), "M1")
    } else {
      NA
    }
  ) %>%
  left_join(first_palliative_therapy, by = "patient_id") %>%
  mutate(
    has_palliative_therapy = !is.na(first_palliative_start),
    has_metastatic_disease = case_when(
      !is.na(metastatic_at_ctdna) & metastatic_at_ctdna == TRUE ~ TRUE,
      has_palliative_therapy == TRUE ~ TRUE,
      !is.na(m_stage_metastatic) & m_stage_metastatic == TRUE ~ TRUE,
      !is.na(metastatic_ever_diagnosed) & metastatic_ever_diagnosed == TRUE ~ TRUE,
      TRUE ~ NA
    )
  ) %>%
  filter(has_metastatic_disease == TRUE) %>%
  select(patient_id)

# Create universe for Question 1
universe_q1 <- all_palliative_therapy_lines %>%
  inner_join(metastatic_status, by = "patient_id") %>%
  left_join(ctdna_long, by = "patient_id") %>%
  filter(
    !is.na(ctdna_date),
    ctdna_date <= therapy_start_date,
    ctdna_date >= (therapy_start_date - days(21))
  ) %>%
  arrange(patient_id, therapy_line, desc(ctdna_date)) %>%
  group_by(patient_id, therapy_line, therapy_start_date) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  select(patient_id, therapy_line, therapy_start_date, 
         ctdna_date, ctdna_value) %>%
  rename(
    ctdna_before_line_date = ctdna_date,
    ctdna_before_line_value = ctdna_value
  ) %>%
  mutate(time_zero_date = therapy_start_date)
```

```{r extract_outcomes_q1, include=FALSE}
# Extract outcomes for Question 1 (with survival data filtering)
death_date_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("^death_date$|death.*date$", ignore_case = TRUE)) & 
                                       !str_detect(names(dat_combined), regex("cause|reason", ignore_case = TRUE))]
survival_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("last.*follow.*up.*date|follow.*up.*date", ignore_case = TRUE)) & 
                                     !str_detect(names(dat_combined), regex("death", ignore_case = TRUE))]
alive_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("^alive_yes_no$|alive.*yes.*no", ignore_case = TRUE))]
imaging_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("obvious.*disease.*imaging|disease.*present.*imaging", ignore_case = TRUE))]

outcomes_q1 <- universe_q1 %>%
  left_join(
    dat_combined %>%
      select(patient_id, all_of(c(death_date_cols, survival_cols, alive_cols, imaging_cols))),
    by = "patient_id"
  ) %>%
  mutate(
    death_date = if (length(death_date_cols) > 0) {
      as_date_safe(.data[[death_date_cols[1]]])
    } else {
      as.Date(NA)
    },
    last_fu_date = if (length(survival_cols) > 0) {
      as_date_safe(.data[[survival_cols[1]]])
    } else {
      as.Date(NA)
    },
    alive_status = if (length(alive_cols) > 0) {
      case_when(
        str_detect(tolower(as.character(.data[[alive_cols[1]]])), "yes|y|1") ~ "Yes",
        str_detect(tolower(as.character(.data[[alive_cols[1]]])), "no|n|0") ~ "No",
        TRUE ~ NA_character_
      )
    } else {
      NA_character_
    },
    imaging_disease_present = if (length(imaging_cols) > 0) {
      case_when(
        str_detect(tolower(as.character(.data[[imaging_cols[1]]])), "yes|y|1") ~ "Yes",
        str_detect(tolower(as.character(.data[[imaging_cols[1]]])), "no|n|0") ~ "No",
        TRUE ~ NA_character_
      )
    } else {
      NA_character_
    },
    event_or_censor_date = case_when(
      !is.na(death_date) ~ death_date,
      !is.na(last_fu_date) ~ last_fu_date,
      TRUE ~ as.Date(NA)
    ),
    mfs_time = as.numeric(event_or_censor_date - time_zero_date),
    mfs_event = case_when(
      !is.na(death_date) ~ 1,
      alive_status == "No" ~ 1,
      alive_status == "Yes" & imaging_disease_present == "Yes" ~ 1,
      alive_status == "Yes" & imaging_disease_present == "No" ~ 0,
      alive_status == "Yes" & is.na(imaging_disease_present) ~ NA_real_,
      TRUE ~ NA_real_
    ),
    imaging_required = case_when(
      !is.na(death_date) ~ FALSE,
      alive_status == "No" ~ FALSE,
      alive_status == "Yes" ~ TRUE,
      TRUE ~ TRUE
    )
  ) %>%
  filter(!is.na(mfs_time), mfs_time > 0, 
         !(imaging_required == TRUE & is.na(imaging_disease_present)))

# Join Bajorin data for Question 1
# First, ensure patient_id formats match (convert both to character and trim)
# The main dataset uses patient_id from record_id_linking_key which may be numeric or character
# Normalize both to character, removing any decimal points from numeric conversions
outcomes_q1 <- outcomes_q1 %>%
  mutate(
    patient_id_char = as.character(patient_id),
    # Normalize: if it's a pure number (possibly with .0), convert to integer string
    patient_id_char = ifelse(!is.na(suppressWarnings(as.numeric(patient_id_char))) & 
                             str_detect(patient_id_char, "^\\.?\\d+\\.?0*$"), 
                             as.character(as.integer(suppressWarnings(as.numeric(patient_id_char)))), 
                             patient_id_char),
    patient_id = str_trim(patient_id_char)
  ) %>%
  select(-patient_id_char)

# Bajorin data should already be character, but ensure it's trimmed and normalized
bajorin_data <- bajorin_data %>%
  mutate(
    patient_id_char = as.character(patient_id),
    # Normalize: if it's a pure number (possibly with .0), convert to integer string
    patient_id_char = ifelse(!is.na(suppressWarnings(as.numeric(patient_id_char))) & 
                             str_detect(patient_id_char, "^\\.?\\d+\\.?0*$"), 
                             as.character(as.integer(suppressWarnings(as.numeric(patient_id_char)))), 
                             patient_id_char),
    patient_id = str_trim(patient_id_char)
  ) %>%
  select(-patient_id_char)

# IMPORTANT: Normalize therapy lines >= 4 to 4 for matching
# Bajorin data converts "4L+" to therapy_line = 4, but analysis may have therapy_line = 4, 5, 6, etc.
# All therapy lines >= 4 should match with Bajorin data therapy_line = 4
bajorin_data_normalized <- bajorin_data %>%
  mutate(therapy_line_match = ifelse(therapy_line >= 4, 4, therapy_line))

outcomes_q1_normalized <- outcomes_q1 %>%
  mutate(therapy_line_match = ifelse(therapy_line >= 4, 4, therapy_line))

# Diagnostic: Check patient ID matching
cat("=== DIAGNOSTIC: Patient ID Matching for Question 1 ===\n")
cat("Total outcomes_q1 observations:", nrow(outcomes_q1), "\n")
cat("Total bajorin_data observations:", nrow(bajorin_data), "\n")
cat("Unique patient IDs in outcomes_q1:", length(unique(outcomes_q1$patient_id)), "\n")
cat("Unique patient IDs in bajorin_data:", length(unique(bajorin_data$patient_id)), "\n")

# Check overlap with normalized therapy lines
outcomes_patient_lines <- outcomes_q1_normalized %>%
  select(patient_id, therapy_line_match) %>%
  distinct()
bajorin_patient_lines_normalized <- bajorin_data_normalized %>%
  select(patient_id, therapy_line_match) %>%
  distinct()

matched_lines <- inner_join(outcomes_patient_lines, bajorin_patient_lines_normalized, 
                            by = c("patient_id", "therapy_line_match"))
cat("Matched patient-line observations (with normalized therapy lines):", nrow(matched_lines), "\n")
cat("Unmatched patient-line observations in outcomes:", 
    nrow(outcomes_patient_lines) - nrow(matched_lines), "\n")

# Show sample of unmatched
unmatched <- anti_join(outcomes_patient_lines, bajorin_patient_lines_normalized, 
                       by = c("patient_id", "therapy_line_match"))
if (nrow(unmatched) > 0) {
  cat("Sample unmatched patient IDs:", 
      paste(head(unique(unmatched$patient_id), 10), collapse = ", "), "\n")
}

outcomes_q1_with_bajorin <- outcomes_q1_normalized %>%
  left_join(
    bajorin_data_normalized %>% select(patient_id, therapy_line_match, kps_binary, visceral_mets_binary),
    by = c("patient_id", "therapy_line_match")
  ) %>%
  select(-therapy_line_match)

# Identify patients missing Bajorin risk factors in Question 1
missing_bajorin_q1 <- outcomes_q1_with_bajorin %>%
  filter(is.na(kps_binary) | is.na(visceral_mets_binary)) %>%
  mutate(
    missing_kps = is.na(kps_binary),
    missing_visceral_mets = is.na(visceral_mets_binary),
    missing_both = is.na(kps_binary) & is.na(visceral_mets_binary)
  ) %>%
  select(patient_id, therapy_line, missing_kps, missing_visceral_mets, missing_both) %>%
  distinct()
```

```{r define_universe_q2, include=FALSE}
# ====================================================================================
# QUESTION 2: On-Treatment ctDNA Response Analysis
# Replicate landmark analysis setup from Question2_Landmark_Analysis.Rmd
# ====================================================================================

# Use the same universe as Question 1
universe_q2 <- universe_q1

# Extract outcomes for Question 2 (same as Question 1)
outcomes_q2 <- outcomes_q1

# Landmark analysis: find on-treatment ctDNA within 6-16 weeks
landmark_window_start_days <- 6 * 7
landmark_window_end_days <- 16 * 7

landmark_data_q2 <- outcomes_q2 %>%
  select(patient_id, therapy_line, time_zero_date, ctdna_before_line_date, 
         ctdna_before_line_value, mfs_time, mfs_event) %>%
  left_join(
    ctdna_long %>%
      select(patient_id, ctdna_date, ctdna_value),
    by = "patient_id"
  ) %>%
  mutate(
    days_from_therapy_start = as.numeric(ctdna_date - time_zero_date),
    in_landmark_window = days_from_therapy_start >= landmark_window_start_days & 
                         days_from_therapy_start <= landmark_window_end_days
  ) %>%
  filter(in_landmark_window) %>%
  arrange(patient_id, therapy_line, desc(days_from_therapy_start)) %>%
  group_by(patient_id, therapy_line) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  mutate(
    baseline_ctdna = ctdna_before_line_value,
    on_tx_ctdna = ctdna_value,
    landmark_time = days_from_therapy_start
  ) %>%
  select(patient_id, therapy_line, time_zero_date, landmark_time,
         baseline_ctdna, on_tx_ctdna)
```

```{r extract_landmark_outcomes_q2, include=FALSE}
# Perform landmark analysis: only include patients who survived to landmark
# Join landmark_data_q2 with outcomes_q2 to get mfs_time and mfs_event
# This matches the exact logic from Question2_Landmark_Analysis.Rmd
landmark_outcomes_q2 <- landmark_data_q2 %>%
  left_join(
    outcomes_q2 %>% 
      select(patient_id, therapy_line, time_zero_date, mfs_time, mfs_event, imaging_disease_present),
    by = c("patient_id", "therapy_line", "time_zero_date")
  ) %>%
  left_join(
    dat_combined %>%
      select(patient_id, all_of(c(death_date_cols, alive_cols))),
    by = "patient_id"
  ) %>%
  mutate(
    death_date = if (length(death_date_cols) > 0) {
      as_date_safe(.data[[death_date_cols[1]]])
    } else {
      as.Date(NA)
    },
    alive_status = if (length(alive_cols) > 0) {
      case_when(
        str_detect(tolower(as.character(.data[[alive_cols[1]]])), "yes|y|1") ~ "Yes",
        str_detect(tolower(as.character(.data[[alive_cols[1]]])), "no|n|0") ~ "No",
        TRUE ~ NA_character_
      )
    } else {
      NA_character_
    },
    # Calculate landmark survival variables AFTER joining mfs_time from outcomes_q2
    survived_to_landmark = !is.na(mfs_time) & mfs_time >= landmark_time,
    mfs_time_from_landmark = mfs_time - landmark_time,
    mfs_event_from_landmark = mfs_event,
    # Determine if imaging data is required (only for alive patients)
    imaging_required = case_when(
      !is.na(death_date) ~ FALSE,
      alive_status == "No" ~ FALSE,
      alive_status == "Yes" ~ TRUE,
      TRUE ~ TRUE
    )
  ) %>%
  filter(survived_to_landmark, 
         !is.na(mfs_time_from_landmark) & mfs_time_from_landmark > 0,
         !(imaging_required == TRUE & is.na(imaging_disease_present)))

# Join Bajorin data for Question 2
# Ensure patient_id formats match (convert to character and trim, normalize numeric)
landmark_outcomes_q2 <- landmark_outcomes_q2 %>%
  mutate(
    patient_id_char = as.character(patient_id),
    # Normalize: if it's a pure number (possibly with .0), convert to integer string
    patient_id_char = ifelse(!is.na(suppressWarnings(as.numeric(patient_id_char))) & 
                             str_detect(patient_id_char, "^\\.?\\d+\\.?0*$"), 
                             as.character(as.integer(suppressWarnings(as.numeric(patient_id_char)))), 
                             patient_id_char),
    patient_id = str_trim(patient_id_char)
  ) %>%
  select(-patient_id_char)

# Normalize therapy lines for Question 2 as well
landmark_outcomes_q2_normalized <- landmark_outcomes_q2 %>%
  mutate(therapy_line_match = ifelse(therapy_line >= 4, 4, therapy_line))

# Diagnostic: Check patient ID matching for Question 2
cat("\n=== DIAGNOSTIC: Patient ID Matching for Question 2 ===\n")
cat("Total landmark_outcomes_q2 observations:", nrow(landmark_outcomes_q2), "\n")

landmark_patient_lines <- landmark_outcomes_q2_normalized %>%
  select(patient_id, therapy_line_match) %>%
  distinct()

matched_lines_q2 <- inner_join(landmark_patient_lines, bajorin_patient_lines_normalized, 
                               by = c("patient_id", "therapy_line_match"))
cat("Matched patient-line observations (with normalized therapy lines):", nrow(matched_lines_q2), "\n")
cat("Unmatched patient-line observations in landmark outcomes:", 
    nrow(landmark_patient_lines) - nrow(matched_lines_q2), "\n")

# Show sample of unmatched
unmatched_q2 <- anti_join(landmark_patient_lines, bajorin_patient_lines_normalized, 
                         by = c("patient_id", "therapy_line_match"))
if (nrow(unmatched_q2) > 0) {
  cat("Sample unmatched patient IDs:", 
      paste(head(unique(unmatched_q2$patient_id), 10), collapse = ", "), "\n")
}

landmark_outcomes_q2_with_bajorin <- landmark_outcomes_q2_normalized %>%
  left_join(
    bajorin_data_normalized %>% select(patient_id, therapy_line_match, kps_binary, visceral_mets_binary),
    by = c("patient_id", "therapy_line_match")
  ) %>%
  select(-therapy_line_match)

# Identify patients missing Bajorin risk factors in Question 2
missing_bajorin_q2 <- landmark_outcomes_q2_with_bajorin %>%
  filter(is.na(kps_binary) | is.na(visceral_mets_binary)) %>%
  mutate(
    missing_kps = is.na(kps_binary),
    missing_visceral_mets = is.na(visceral_mets_binary),
    missing_both = is.na(kps_binary) & is.na(visceral_mets_binary)
  ) %>%
  select(patient_id, therapy_line, missing_kps, missing_visceral_mets, missing_both) %>%
  distinct()
```

## Diagnostic: Patient ID Matching

```{r diagnostic_matching, echo=FALSE}
# Check specific patient IDs mentioned by user
check_patients <- c("179", "22", "250", "343")

cat("### Checking Specific Patient IDs\n\n")
for (pid in check_patients) {
  pid_char <- str_trim(as.character(pid))  # Ensure consistent format
  cat("**Patient ID:", pid_char, "**\n")
  
  # Check in outcomes_q1 (after formatting)
  in_q1 <- outcomes_q1 %>%
    filter(patient_id == pid_char) %>%
    select(patient_id, therapy_line) %>%
    distinct()
  
  # Check in bajorin_data (after formatting)
  in_bajorin <- bajorin_data %>%
    filter(patient_id == pid_char) %>%
    select(patient_id, therapy_line, kps_binary, visceral_mets_binary) %>%
    distinct()
  
  # Check the actual join result
  joined_check <- outcomes_q1 %>%
    filter(patient_id == pid_char) %>%
    left_join(
      bajorin_data %>% select(patient_id, therapy_line, kps_binary, visceral_mets_binary),
      by = c("patient_id", "therapy_line")
    ) %>%
    select(patient_id, therapy_line, kps_binary, visceral_mets_binary) %>%
    distinct()
  
  # Check what patient_id values actually exist (for debugging)
  q1_ids <- outcomes_q1 %>%
    filter(str_detect(str_trim(as.character(patient_id)), paste0("^", pid_char, "$"))) %>%
    select(patient_id) %>%
    distinct()
  
  bajorin_ids <- bajorin_data %>%
    filter(str_detect(str_trim(as.character(patient_id)), paste0("^", pid_char, "$"))) %>%
    select(patient_id) %>%
    distinct()
  
  cat("- In Question 1 analysis:", if(nrow(in_q1) > 0) paste("Yes (", nrow(in_q1), "therapy lines)", sep="") else "No", "\n")
  if (nrow(q1_ids) > 0 && nrow(in_q1) == 0) {
    cat("  - Note: Found similar IDs in Q1:", paste(head(unique(q1_ids$patient_id), 5), collapse=", "), "\n")
  }
  
  cat("- In Bajorin data:", if(nrow(in_bajorin) > 0) paste("Yes (", nrow(in_bajorin), "therapy lines)", sep="") else "No", "\n")
  if (nrow(bajorin_ids) > 0 && nrow(in_bajorin) == 0) {
    cat("  - Note: Found similar IDs in Bajorin:", paste(head(unique(bajorin_ids$patient_id), 5), collapse=", "), "\n")
  }
  
  if (nrow(in_q1) > 0) {
    cat("- Therapy lines in analysis:", paste(in_q1$therapy_line, collapse=", "), "\n")
  }
  if (nrow(in_bajorin) > 0) {
    cat("- Therapy lines in Bajorin data:", paste(in_bajorin$therapy_line, collapse=", "), "\n")
    cat("- KPS values:", paste(ifelse(is.na(in_bajorin$kps_binary), "NA", in_bajorin$kps_binary), collapse=", "), "\n")
    cat("- Visceral Mets values:", paste(ifelse(is.na(in_bajorin$visceral_mets_binary), "NA", in_bajorin$visceral_mets_binary), collapse=", "), "\n")
  }
  
  # Show join result
  if (nrow(joined_check) > 0) {
    cat("- After join - therapy lines with Bajorin data:", 
        paste(joined_check$therapy_line[!is.na(joined_check$kps_binary) | !is.na(joined_check$visceral_mets_binary)], collapse=", "), "\n")
    cat("- After join - therapy lines missing Bajorin data:", 
        paste(joined_check$therapy_line[is.na(joined_check$kps_binary) | is.na(joined_check$visceral_mets_binary)], collapse=", "), "\n")
  }
  
  # Check in missing list (after formatting)
  in_missing_q1 <- missing_bajorin_q1 %>%
    filter(str_trim(as.character(patient_id)) == pid_char)
  
  if (nrow(in_missing_q1) > 0) {
    cat("- Listed as missing Bajorin data:", "Yes", "\n")
    cat("- Missing KPS:", ifelse(any(in_missing_q1$missing_kps), "Yes", "No"), "\n")
    cat("- Missing Visceral Mets:", ifelse(any(in_missing_q1$missing_visceral_mets), "Yes", "No"), "\n")
    
    # Check if it's a therapy line mismatch
    if (nrow(in_q1) > 0 && nrow(in_bajorin) > 0) {
      q1_lines <- sort(unique(in_q1$therapy_line))
      bajorin_lines <- sort(unique(in_bajorin$therapy_line))
      if (!identical(q1_lines, bajorin_lines)) {
        cat("- **WARNING: Therapy line mismatch!**\n")
        cat("  - Analysis has lines:", paste(q1_lines, collapse=", "), "\n")
        cat("  - Bajorin has lines:", paste(bajorin_lines, collapse=", "), "\n")
      }
    }
  } else if (nrow(in_q1) > 0) {
    cat("- Listed as missing Bajorin data:", "No (has complete Bajorin data)", "\n")
  }
  cat("\n")
}
```

## Summary Statistics

```{r summary_statistics}
cat("### Question 1: Baseline ctDNA Prognosis Analysis\n\n")
cat("**Total patient-line observations in final analysis:**", nrow(outcomes_q1), "\n")
cat("**Unique patients in final analysis:**", length(unique(outcomes_q1$patient_id)), "\n\n")

cat("**Patients missing Bajorin risk factors:**\n")
cat("- Total patient-line observations missing KPS or visceral mets:", nrow(missing_bajorin_q1), "\n")
cat("- Unique patients missing KPS or visceral mets:", length(unique(missing_bajorin_q1$patient_id)), "\n")
cat("- Missing KPS only:", sum(missing_bajorin_q1$missing_kps & !missing_bajorin_q1$missing_visceral_mets), "\n")
cat("- Missing visceral mets only:", sum(!missing_bajorin_q1$missing_kps & missing_bajorin_q1$missing_visceral_mets), "\n")
cat("- Missing both:", sum(missing_bajorin_q1$missing_both), "\n\n")

cat("### Question 2: On-Treatment ctDNA Response Analysis\n\n")
cat("**Total patient-line observations in final landmark analysis:**", nrow(landmark_outcomes_q2), "\n")
cat("**Unique patients in final landmark analysis:**", length(unique(landmark_outcomes_q2$patient_id)), "\n\n")

cat("**Patients missing Bajorin risk factors:**\n")
cat("- Total patient-line observations missing KPS or visceral mets:", nrow(missing_bajorin_q2), "\n")
cat("- Unique patients missing KPS or visceral mets:", length(unique(missing_bajorin_q2$patient_id)), "\n")
cat("- Missing KPS only:", sum(missing_bajorin_q2$missing_kps & !missing_bajorin_q2$missing_visceral_mets), "\n")
cat("- Missing visceral mets only:", sum(!missing_bajorin_q2$missing_kps & missing_bajorin_q2$missing_visceral_mets), "\n")
cat("- Missing both:", sum(missing_bajorin_q2$missing_both), "\n")
```

## Question 1: Patients Missing Bajorin Risk Factors

```{r q1_missing_bajorin_table}
missing_bajorin_q1_detailed <- missing_bajorin_q1 %>%
  arrange(patient_id, therapy_line) %>%
  mutate(
    `Therapy Line` = case_when(
      therapy_line == 1 ~ "1L",
      therapy_line == 2 ~ "2L",
      therapy_line == 3 ~ "3L",
      therapy_line >= 4 ~ "4L+",
      TRUE ~ as.character(therapy_line)
    ),
    `Missing KPS` = ifelse(missing_kps, "Yes", "No"),
    `Missing Visceral Mets` = ifelse(missing_visceral_mets, "Yes", "No"),
    `Missing Both` = ifelse(missing_both, "Yes", "No")
  ) %>%
  select(`Patient ID` = patient_id, `Therapy Line`, `Missing KPS`, `Missing Visceral Mets`, `Missing Both`)

if (nrow(missing_bajorin_q1_detailed) > 0) {
  missing_bajorin_q1_detailed %>%
    kable(caption = paste0("Question 1: Patients Missing Bajorin Risk Factors (N = ", 
                          length(unique(missing_bajorin_q1_detailed$`Patient ID`)), 
                          " unique patients, ", nrow(missing_bajorin_q1_detailed), 
                          " patient-line observations)")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
} else {
  cat("**No patients missing Bajorin risk factors in Question 1 analysis.**\n")
}
```

## Question 2: Patients Missing Bajorin Risk Factors

```{r q2_missing_bajorin_table}
missing_bajorin_q2_detailed <- missing_bajorin_q2 %>%
  arrange(patient_id, therapy_line) %>%
  mutate(
    `Therapy Line` = case_when(
      therapy_line == 1 ~ "1L",
      therapy_line == 2 ~ "2L",
      therapy_line == 3 ~ "3L",
      therapy_line >= 4 ~ "4L+",
      TRUE ~ as.character(therapy_line)
    ),
    `Missing KPS` = ifelse(missing_kps, "Yes", "No"),
    `Missing Visceral Mets` = ifelse(missing_visceral_mets, "Yes", "No"),
    `Missing Both` = ifelse(missing_both, "Yes", "No")
  ) %>%
  select(`Patient ID` = patient_id, `Therapy Line`, `Missing KPS`, `Missing Visceral Mets`, `Missing Both`)

if (nrow(missing_bajorin_q2_detailed) > 0) {
  missing_bajorin_q2_detailed %>%
    kable(caption = paste0("Question 2: Patients Missing Bajorin Risk Factors (N = ", 
                          length(unique(missing_bajorin_q2_detailed$`Patient ID`)), 
                          " unique patients, ", nrow(missing_bajorin_q2_detailed), 
                          " patient-line observations)")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
} else {
  cat("**No patients missing Bajorin risk factors in Question 2 analysis.**\n")
}
```

## Combined List

```{r combined_missing_bajorin_table}
# Combine both analyses
missing_bajorin_combined <- bind_rows(
  missing_bajorin_q1_detailed %>% mutate(Analysis = "Question 1: Baseline ctDNA Prognosis"),
  missing_bajorin_q2_detailed %>% mutate(Analysis = "Question 2: On-Treatment ctDNA Response")
) %>%
  select(Analysis, `Patient ID`, `Therapy Line`, `Missing KPS`, `Missing Visceral Mets`, `Missing Both`) %>%
  arrange(Analysis, `Patient ID`, `Therapy Line`)

if (nrow(missing_bajorin_combined) > 0) {
  missing_bajorin_combined %>%
    kable(caption = "Combined List: Patients Missing Bajorin Risk Factors in Question 1 and/or Question 2") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
} else {
  cat("**No patients missing Bajorin risk factors in either analysis.**\n")
}
```

```{r export_csv, include=FALSE}
# Export to CSV
if (nrow(missing_bajorin_combined) > 0) {
  write.csv(missing_bajorin_combined, "Patients_Missing_Bajorin_Risk_Factors.csv", row.names = FALSE)
  cat("Results exported to: Patients_Missing_Bajorin_Risk_Factors.csv\n")
} else {
  cat("No patients missing Bajorin risk factors - no CSV file created.\n")
}
```

