---
title: "ctDNA Threshold Before Cystectomy Predicting Post-Cystectomy Positive ctDNA"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "")
knitr::opts_knit$set(root.dir = "~/R")
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Load shared setup
source("ctDNA_Research_Questions_Shared_Setup.R")

# Load required libraries
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(stringr)
  library(readxl)
  library(janitor)
  library(lubridate)
  library(ggplot2)
  library(purrr)
  library(knitr)
  library(kableExtra)
})

# Check and load pROC
if (!require(pROC, quietly = TRUE)) {
  install.packages("pROC", repos = "https://cloud.r-project.org")
  library(pROC)
}

# Check and load cutpointR
if (!require(cutpointr, quietly = TRUE)) {
  install.packages("cutpointr", repos = "https://cloud.r-project.org")
  library(cutpointr)
}
```

```{r load_data, include=FALSE}
# Load minimal data
data_list <- load_minimal_data(include_survival = FALSE)
dat_combined <- data_list$dat_combined
ctdna_long <- data_list$ctdna_long
```

## Research Questions

```{r research_questions_top, echo=FALSE}
cat("### Primary Research Question\n\n")
cat("**Is pre-cystectomy ctDNA associated with post-cystectomy molecular residual disease, such that ctDNA measured before surgery can identify patients at higher risk of persistent ctDNA after cystectomy?**\n\n")
cat("### Secondary (Methodologic) Question\n\n")
cat("Can a quantitative pre-cystectomy ctDNA threshold stratify risk of post-cystectomy ctDNA positivity at a prespecified post-operative landmark, informing identification of patients who may warrant earlier adjuvant therapy or consideration of additional systemic treatment prior to surgery?\n\n")
```

## Methods

```{r methods_section, echo=FALSE}
cat("# Methods\n\n")
cat("## Study Population\n\n")
cat("This analysis included patients who underwent cystectomy and had ctDNA levels available both before and after cystectomy. Pre-cystectomy ctDNA measurements were required to be within 60 days prior to surgery. The primary analysis used the first post-cystectomy ctDNA measurement from 4 weeks (30 days) after surgery onward, within 1 year of cystectomy. A landmark analysis was also performed at the 3-month timepoint to account for different follow-up durations.\n\n")

cat("## Landmark Analysis: Prespecified Framework\n\n")
cat("The 3-month landmark analysis was prespecified with the following framework:\n\n")
cat("**Primary Population:** All cystectomy patients (mixed cohort), with neoadjuvant chemotherapy (NAC) status adjusted for in models or reported in stratified analyses.\n\n")
cat("**Primary Endpoint:** Post-cystectomy ctDNA positivity at the 3-month landmark timepoint (90 days post-surgery, with ±30 day window).\n\n")
cat("**Primary Role of Threshold:** Risk stratification and anticipating post-operative minimal residual disease (MRD), **not** as a decision tool for surgery vs. additional therapy. The threshold serves to identify patients at higher risk of post-operative ctDNA positivity, which may inform monitoring intensity and adjuvant therapy considerations, but does not determine surgical candidacy.\n\n")

cat("## Clinical Context and Rationale for Threshold Selection\n\n")
cat("In the post-neoadjuvant, pre-cystectomy setting, ctDNA functions as a decision biomarker to identify patients for whom surgery alone may be insufficient. In this context, **false-negative results carry greater clinical harm than false-positive results**, as they may miss patients who would benefit from additional therapy before or after surgery. False negatives may lead to missed opportunities for treatment escalation and worse outcomes. False positives, while concerning, are more acceptable as patients still proceed to cystectomy and may benefit from additional monitoring or adjuvant therapy.\n\n")
cat("Therefore, test performance was interpreted with **primary emphasis on sensitivity and negative predictive value**, while specificity and positive predictive value were considered secondary. Sensitivity protects patients from being undertreated; specificity protects patients from being overtreated — and in this setting, undertreatment is the greater risk.\n\n")

cat("## Threshold Identification\n\n")
cat("The optimal threshold was identified to **maximize sensitivity** while maintaining acceptable specificity, meeting the following criteria:\n")
cat("- **Sensitivity ≥ 80%** (primary constraint)\n")
cat("- **Specificity ≥ 50%** (secondary constraint)\n\n")
cat("Among all thresholds meeting these criteria, the **highest threshold** was selected. This approach prioritizes maximizing sensitivity (capturing patients who need additional therapy) while maintaining reasonable specificity. If no threshold met both criteria, the analysis fell back to the highest threshold with sensitivity ≥ 80%, or if that was not available, the Youden criterion (maximizing sensitivity + specificity - 1) was used.\n\n")

cat("## Association Analysis\n\n")
cat("In addition to threshold-based analysis, logistic regression was performed to assess the association between pre-cystectomy ctDNA (as a continuous predictor, log-transformed) and post-cystectomy ctDNA positivity. Models included neoadjuvant chemotherapy status as a covariate or interaction term. Results are reported as odds ratios per doubling (or per log unit) of pre-cystectomy ctDNA, along with area under the ROC curve (AUC) and calibration metrics.\n\n")

cat("## Bootstrap Internal Validation\n\n")
cat("Bootstrap resampling (n = 1000 iterations) was performed separately for:\n")
cat("1. **Primary Analysis:** First post-cystectomy ctDNA from 4 weeks onward (within 1 year)\n")
cat("2. **3-Month Landmark Analysis:** ctDNA measurements at the 3-month landmark timepoint\n\n")
cat("For each bootstrap iteration, stratified resampling was performed to maintain the class distribution (positive vs. negative post-cystectomy ctDNA). The optimal threshold was identified on each bootstrap sample using the same criteria (sensitivity ≥ 80% and specificity ≥ 50%), and diagnostic measures were calculated. Bootstrap-validated measures (median/mean with 95% confidence intervals) and optimism estimates (apparent - bootstrap mean) were reported.\n\n")

cat("## Diagnostic Measures\n\n")
cat("The following diagnostic measures were calculated:\n")
cat("- **Sensitivity:** Proportion of patients with positive post-cystectomy ctDNA who had pre-cystectomy ctDNA above the threshold\n")
cat("- **Specificity:** Proportion of patients with negative post-cystectomy ctDNA who had pre-cystectomy ctDNA below the threshold\n")
cat("- **Positive Predictive Value (PPV):** Proportion of patients with pre-cystectomy ctDNA above the threshold who had positive post-cystectomy ctDNA\n")
cat("- **Negative Predictive Value (NPV):** Proportion of patients with pre-cystectomy ctDNA below the threshold who had negative post-cystectomy ctDNA\n")
cat("- **Area Under the ROC Curve (AUC):** Overall diagnostic accuracy of the test\n\n")

cat("## Sensitivity and Specificity Types\n\n")
cat("**Cumulative Sensitivity:** The proportion of patients who have positive ctDNA at or before the landmark timepoint among all patients who eventually develop positive post-cystectomy ctDNA. This captures all positive cases until the given time point.\n\n")
cat("**Static Specificity:** The proportion of patients who have persistently negative ctDNA until the landmark timepoint among all patients who remain negative at the landmark. This represents cases that are persistently negative until the given time point.\n\n")
```

```{r define_universe, include=FALSE}
# ==============================
# Universe: Patients who underwent cystectomy and had ctDNA levels available 
# both before and after cystectomy (within 1 year)
# ==============================

# Extract cystectomy date (surgery_1_date or similar)
surgery_date_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("surgery.*1.*date|cystectomy.*date|first.*surgery.*date", ignore_case = TRUE))]

if (length(surgery_date_cols) == 0) {
  # Try alternative names
  surgery_date_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("definitive.*surgery|radical.*cystectomy", ignore_case = TRUE))]
}

# Get cystectomy patients
cystectomy_patients <- dat_combined %>%
  select(patient_id, all_of(surgery_date_cols)) %>%
  mutate(
    cystectomy_date = if (length(surgery_date_cols) > 0) {
      as_date_safe(.data[[surgery_date_cols[1]]])
    } else {
      as.Date(NA)
    }
  ) %>%
  filter(!is.na(cystectomy_date)) %>%
  select(patient_id, cystectomy_date)

# Get ctDNA before cystectomy (within 60 days)
ctdna_pre_cystectomy <- ctdna_long %>%
  inner_join(cystectomy_patients, by = "patient_id") %>%
  filter(
    ctdna_date < cystectomy_date,  # ctDNA must be before cystectomy
    ctdna_date >= (cystectomy_date - days(60))  # ctDNA within 60 days before cystectomy
  ) %>%
  arrange(patient_id, desc(ctdna_date)) %>%
  group_by(patient_id) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  select(patient_id, pre_cystectomy_ctdna_date = ctdna_date, pre_cystectomy_ctdna_value = ctdna_value)

# Get FIRST ctDNA after cystectomy (within 1 year, but excluding first 30 days to avoid false positives)
# Clinical recommendation: Start checking ctDNA after 4 weeks (30 days) post-surgery due to false positives
# This is the FIRST post-cystectomy ctDNA measurement from 4 weeks onward
ctdna_post_cystectomy <- ctdna_long %>%
  inner_join(cystectomy_patients, by = "patient_id") %>%
  filter(
    ctdna_date > cystectomy_date + days(30),  # Exclude first 30 days (4 weeks) to avoid false positives
    ctdna_date <= cystectomy_date + days(365)  # Within 1 year
  ) %>%
  arrange(patient_id, ctdna_date) %>%
  group_by(patient_id) %>%
  slice_head(n = 1) %>%  # Take FIRST post-cystectomy ctDNA from 30 days onward
  ungroup() %>%
  select(patient_id, post_cystectomy_ctdna_date = ctdna_date, post_cystectomy_ctdna_value = ctdna_value)

# Get neoadjuvant therapy status
# First, try direct neoadjuvant column
neoadjuvant_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("neoadjuvant.*systemic.*therapy|neoadj.*systemic", ignore_case = TRUE))]

# Also check systemic therapy treatment setting columns
setting_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("systemic.*therapy.*\\d+.*treatment_setting|systemic.*therapy.*\\d+.*setting", ignore_case = TRUE))]

cat("**Neoadjuvant Therapy Identification:**\n")
cat("Direct neoadjuvant columns found:", length(neoadjuvant_cols), "\n")
if (length(neoadjuvant_cols) > 0) {
  cat("Column names:", paste(neoadjuvant_cols, collapse = ", "), "\n")
}
cat("Treatment setting columns found:", length(setting_cols), "\n")
if (length(setting_cols) > 0) {
  cat("First few setting columns:", paste(head(setting_cols, 3), collapse = ", "), "\n")
}

neoadjuvant_status <- dat_combined %>%
  select(patient_id, all_of(c(neoadjuvant_cols, setting_cols))) %>%
  mutate(
    # Method 1: Direct neoadjuvant column (handle both character and numeric)
    # Only count as "yes" if explicitly "yes", "y", "1", "true", or numeric 1
    # Do NOT count "N/A", "No", or missing as "yes"
    neoadj_from_direct = if (length(neoadjuvant_cols) > 0) {
      rowSums(across(all_of(neoadjuvant_cols), ~ {
        val <- as.character(.)
        val_clean <- tolower(str_trim(val))
        # Check for "yes", "y", "1", "true", or numeric 1
        # Explicitly exclude "n/a", "na", "no", "false", "0"
        (val_clean %in% c("yes", "y", "1", "true")) | 
        (!is.na(suppressWarnings(as.numeric(val))) & as.numeric(val) == 1)
      }), na.rm = TRUE) > 0
    } else {
      FALSE
    },
    # Method 2: From treatment setting columns (PRIMARY METHOD)
    # This takes precedence - if ANY treatment setting says "neoadjuvant", patient received it
    # This is important because direct column may say "N/A" but treatment settings may indicate neoadjuvant
    neoadj_from_settings = if (length(setting_cols) > 0) {
      rowSums(across(all_of(setting_cols), ~ {
        val <- tolower(str_trim(as.character(.)))
        # Look for "neo" or "neoadjuvant" in treatment setting
        str_detect(val, "neo|neoadjuvant") & !is.na(val) & val != "" & 
        !str_detect(val, "no.*neo|not.*neo")  # Exclude "no neoadjuvant" type entries
      }), na.rm = TRUE) > 0
    } else {
      FALSE
    },
    # Combine both methods: If EITHER method indicates neoadjuvant, patient received it
    # Treatment settings take precedence (can override "N/A" in direct column)
    neoadjuvant_therapy = neoadj_from_direct | neoadj_from_settings
  ) %>%
  select(patient_id, neoadjuvant_therapy, neoadj_from_direct, neoadj_from_settings)

# Diagnostic output
cat("\n**Neoadjuvant Therapy Identification Summary:**\n")
cat("Total patients in dat_combined:", nrow(dat_combined), "\n")
cat("Patients with neoadjuvant therapy (direct column):", sum(neoadjuvant_status$neoadj_from_direct, na.rm = TRUE), "\n")
cat("Patients with neoadjuvant therapy (from settings):", sum(neoadjuvant_status$neoadj_from_settings, na.rm = TRUE), "\n")
cat("Patients with neoadjuvant therapy (combined):", sum(neoadjuvant_status$neoadjuvant_therapy, na.rm = TRUE), "\n")
cat("Patients without neoadjuvant therapy:", sum(!neoadjuvant_status$neoadjuvant_therapy, na.rm = TRUE), "\n")
cat("Patients with missing neoadjuvant status:", sum(is.na(neoadjuvant_status$neoadjuvant_therapy)), "\n\n")

neoadjuvant_status <- neoadjuvant_status %>% select(patient_id, neoadjuvant_therapy)

# Define universe: patients with both pre- and post-cystectomy ctDNA
# Join with cystectomy dates to calculate time from surgery
universe <- ctdna_pre_cystectomy %>%
  inner_join(ctdna_post_cystectomy, by = "patient_id") %>%
  left_join(cystectomy_patients, by = "patient_id") %>%
  left_join(neoadjuvant_status, by = "patient_id") %>%
  mutate(
    days_from_surgery_to_ctdna = as.numeric(post_cystectomy_ctdna_date - cystectomy_date),
    months_from_surgery_to_ctdna = days_from_surgery_to_ctdna / 30.44,
    post_cystectomy_positive = ifelse(!is.na(post_cystectomy_ctdna_value) & post_cystectomy_ctdna_value > 0, 1, 0),
    has_post_cystectomy_ctdna = TRUE,  # All patients in universe have post-cystectomy ctDNA by definition
    neoadjuvant_therapy = ifelse(is.na(neoadjuvant_therapy), FALSE, neoadjuvant_therapy),
    neoadjuvant_group = ifelse(neoadjuvant_therapy, "Neoadjuvant Therapy", "No Neoadjuvant Therapy")
  )

cat("Universe size: n =", nrow(universe), "\n")
cat("Patients with positive post-cystectomy ctDNA: n =", sum(universe$post_cystectomy_positive, na.rm = TRUE), "\n")
cat("Patients with negative post-cystectomy ctDNA: n =", sum(universe$post_cystectomy_positive == 0, na.rm = TRUE), "\n")
cat("Time from surgery to FIRST post-cystectomy ctDNA measurement (median, IQR):", 
    round(median(universe$days_from_surgery_to_ctdna, na.rm = TRUE), 0), "days (",
    round(quantile(universe$days_from_surgery_to_ctdna, 0.25, na.rm = TRUE), 0), "-",
    round(quantile(universe$days_from_surgery_to_ctdna, 0.75, na.rm = TRUE), 0), ")\n")
cat("**Note:** Post-cystectomy ctDNA measurements are the FIRST post-cystectomy ctDNA from 30 days onwards (excluding first 4 weeks) to avoid false positives.\n")
cat("**Definition:** 'Time from surgery to ctDNA measurement' refers to the time from cystectomy to the FIRST post-cystectomy ctDNA measurement (from 4 weeks/30 days onwards, within 1 year).\n")
cat("Population studied: Patients who underwent cystectomy and had ctDNA levels available both before and after cystectomy (FIRST post-cystectomy ctDNA from 4 weeks/30 days onwards, within 1 year)\n")
```

```{r population_description}
# Display population description at top
cat("## Population Studied\n\n")
cat("**Universe:** Patients who underwent cystectomy and had ctDNA levels available both before and after cystectomy\n\n")
cat("**N =", nrow(universe), "\n\n")
cat("**Inclusion Criteria:**\n")
cat("- Underwent cystectomy\n")
cat("- Had at least one ctDNA measurement before cystectomy (within 60 days prior to surgery)\n")
cat("- Had at least one ctDNA measurement after cystectomy (FIRST post-cystectomy ctDNA from 4 weeks/30 days onwards, within 1 year)\n")
cat("  - **Rationale:** Excluding first 4 weeks (30 days) post-surgery to avoid false positives from surgical trauma\n")
cat("  - **Note:** This analysis uses the FIRST post-cystectomy ctDNA measurement from 30 days onward, not subsequent measurements\n\n")
```

## Data Overview

### Descriptive Characteristics

```{r descriptive_table}
# Brief descriptive table
universe_with_days <- universe %>%
  mutate(
    days_to_cystectomy = as.numeric(cystectomy_date - pre_cystectomy_ctdna_date)
  )

desc_table <- universe_with_days %>%
  summarise(
    N = n(),
    `Pre-Cystectomy ctDNA (median, IQR)` = paste0(
      round(median(pre_cystectomy_ctdna_value, na.rm = TRUE), 2), 
      " (", 
      round(quantile(pre_cystectomy_ctdna_value, 0.25, na.rm = TRUE), 2), 
      "-", 
      round(quantile(pre_cystectomy_ctdna_value, 0.75, na.rm = TRUE), 2), 
      ")"
    ),
    `Days from Pre-Cystectomy ctDNA to Cystectomy (median, IQR)` = paste0(
      round(median(days_to_cystectomy, na.rm = TRUE), 0), 
      " (", 
      round(quantile(days_to_cystectomy, 0.25, na.rm = TRUE), 0), 
      "-", 
      round(quantile(days_to_cystectomy, 0.75, na.rm = TRUE), 0), 
      ")"
    ),
    `Post-Cystectomy Positive` = sum(post_cystectomy_positive, na.rm = TRUE),
    `Post-Cystectomy Negative` = sum(post_cystectomy_positive == 0, na.rm = TRUE)
  ) %>%
  kable(caption = "Descriptive Characteristics of Universe") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)

desc_table
```

### Time from Surgery to First Post-Cystectomy ctDNA Measurement

```{r time_from_surgery_histogram, fig.width=10, fig.height=6, fig.cap="Distribution of Time from Cystectomy to First Post-Cystectomy ctDNA Measurement. Vertical lines indicate potential landmark timepoints (3, 6, 9, 12 months)."}
# Histogram of time from surgery to ctDNA measurement
# This helps identify the best landmark timepoint

# Create plot - ensure it's always the last expression
if (nrow(universe) > 0 && "days_from_surgery_to_ctdna" %in% names(universe)) {
  # Calculate histogram breaks for better visualization
  max_days <- max(universe$days_from_surgery_to_ctdna, na.rm = TRUE)
  hist_breaks <- seq(0, max_days + 30, by = 30)
  
  # Calculate max count for y-axis positioning
  hist_data <- hist(universe$days_from_surgery_to_ctdna, breaks = hist_breaks, plot = FALSE)
  max_count <- max(hist_data$counts, na.rm = TRUE)
  
  # Create the plot
  p_time_hist <- universe %>%
    ggplot(aes(x = days_from_surgery_to_ctdna)) +
    geom_histogram(breaks = hist_breaks, fill = "steelblue", alpha = 0.7, color = "black", closed = "left") +
    # Add vertical lines for potential landmark timepoints
    geom_vline(xintercept = 90, linetype = "dashed", color = "orange", linewidth = 0.8, alpha = 0.7) +
    geom_vline(xintercept = 180, linetype = "dashed", color = "red", linewidth = 1) +
    geom_vline(xintercept = 270, linetype = "dashed", color = "purple", linewidth = 0.8, alpha = 0.7) +
    geom_vline(xintercept = 365, linetype = "dashed", color = "darkgreen", linewidth = 0.8, alpha = 0.7) +
    # Add labels for landmark timepoints
    annotate("text", x = 90, y = max_count * 0.95, label = "3 mo", hjust = 0.5, vjust = 0, color = "orange", size = 4, fontface = "bold") +
    annotate("text", x = 180, y = max_count * 0.95, label = "6 mo", hjust = 0.5, vjust = 0, color = "red", size = 4, fontface = "bold") +
    annotate("text", x = 270, y = max_count * 0.95, label = "9 mo", hjust = 0.5, vjust = 0, color = "purple", size = 4, fontface = "bold") +
    annotate("text", x = 365, y = max_count * 0.95, label = "12 mo", hjust = 0.5, vjust = 0, color = "darkgreen", size = 4, fontface = "bold") +
    labs(
      x = "Days from Surgery to ctDNA Measurement",
      y = "Number of Patients",
      title = "Distribution of Time from Cystectomy to Post-Cystectomy ctDNA Measurement",
      subtitle = "Vertical lines indicate potential landmark timepoints (3, 6, 9, 12 months)"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 12),
      plot.subtitle = element_text(hjust = 0.5, size = 10)
    )
} else {
  # Create empty plot with error message if data not available
  p_time_hist <- ggplot() +
    annotate("text", x = 0.5, y = 0.5, label = "Cannot create histogram:\nUniverse data not available or missing required columns", 
             size = 5, hjust = 0.5) +
    theme_void()
}

# Plot must be the last expression to be rendered by knitr
p_time_hist
```

```{r time_from_surgery_summary}
# Summary statistics for time from surgery
if (nrow(universe) > 0 && "days_from_surgery_to_ctdna" %in% names(universe)) {
  cat("\n**Time from Surgery to First Post-Cystectomy ctDNA Measurement:**\n")
  cat("(This refers to the time from cystectomy to the FIRST ctDNA measurement from 30 days onwards)\n")
  cat("- Median:", round(median(universe$days_from_surgery_to_ctdna, na.rm = TRUE), 0), "days (", 
      round(median(universe$months_from_surgery_to_ctdna, na.rm = TRUE), 1), "months)\n")
  cat("- IQR:", round(quantile(universe$days_from_surgery_to_ctdna, 0.25, na.rm = TRUE), 0), "-",
      round(quantile(universe$days_from_surgery_to_ctdna, 0.75, na.rm = TRUE), 0), "days\n")
  cat("- Range:", round(min(universe$days_from_surgery_to_ctdna, na.rm = TRUE), 0), "-",
      round(max(universe$days_from_surgery_to_ctdna, na.rm = TRUE), 0), "days\n\n")
}
```

```{r landmark_timepoint_analysis}
# Landmark count table removed per updated analytic plan; the 3-month landmark
# is used a priori as the clinically relevant timepoint for risk stratification.
```

```{r landmark_analysis, include=FALSE}
# ====================================================================================
# LANDMARK ANALYSIS SETUP
# ====================================================================================

# Define landmark timepoints
landmark_3mo_days <- 90
landmark_6mo_days <- 180
landmark_window_days <- 30  # ±30 days around landmark for flexibility
# Note: Window means measurements within [landmark - 30, landmark + 30] days are included

# Function to create landmark data for a given landmark timepoint
create_landmark_data <- function(landmark_days_val, landmark_window_days_val) {
  landmark_data <- ctdna_long %>%
    inner_join(cystectomy_patients, by = "patient_id") %>%
    filter(
      ctdna_date > cystectomy_date + days(30),  # Exclude first 30 days (4 weeks) to avoid false positives
      ctdna_date <= cystectomy_date + days(365)  # Still within 1 year
    ) %>%
    mutate(
      days_from_surgery = as.numeric(ctdna_date - cystectomy_date),
      distance_from_landmark = abs(days_from_surgery - landmark_days_val)
    ) %>%
    # For each patient, take the ctDNA measurement closest to landmark timepoint
    arrange(patient_id, distance_from_landmark) %>%
    group_by(patient_id) %>%
    slice_head(n = 1) %>%
    ungroup() %>%
    # Only include patients with ctDNA within landmark window (at or before landmark + window)
    filter(days_from_surgery <= landmark_days_val + landmark_window_days_val) %>%
    mutate(
      landmark_ctdna_value = ctdna_value,
      landmark_ctdna_date = ctdna_date,
      landmark_positive = ifelse(!is.na(landmark_ctdna_value) & landmark_ctdna_value > 0, 1, 0)
    ) %>%
    select(patient_id, cystectomy_date, landmark_ctdna_date, landmark_ctdna_value, 
           days_from_surgery, landmark_positive)
  
  return(landmark_data)
}

# Create landmark data for 3-month and 6-month landmarks
landmark_data_3mo <- create_landmark_data(landmark_3mo_days, landmark_window_days)
landmark_data_6mo <- create_landmark_data(landmark_6mo_days, landmark_window_days)

# For backward compatibility, use 6-month as default
landmark_days <- landmark_6mo_days
landmark_months <- 6
landmark_data <- landmark_data_6mo

# Create landmark universes: patients with pre-cystectomy ctDNA AND landmark ctDNA
landmark_universe_3mo <- ctdna_pre_cystectomy %>%
  inner_join(landmark_data_3mo, by = "patient_id") %>%
  mutate(
    post_cystectomy_positive_landmark = landmark_positive
  )

landmark_universe_6mo <- ctdna_pre_cystectomy %>%
  inner_join(landmark_data_6mo, by = "patient_id") %>%
  mutate(
    post_cystectomy_positive_landmark = landmark_positive
  )

# For backward compatibility
landmark_universe <- landmark_universe_6mo

cat("\n**Landmark Analysis (3 months):**\n")
cat("Landmark universe size: n =", nrow(landmark_universe_3mo), "\n")
cat("Patients with positive ctDNA at landmark: n =", sum(landmark_universe_3mo$post_cystectomy_positive_landmark, na.rm = TRUE), "\n")
cat("Patients with negative ctDNA at landmark: n =", sum(landmark_universe_3mo$post_cystectomy_positive_landmark == 0, na.rm = TRUE), "\n")
cat("Median time from surgery to landmark ctDNA:", round(median(landmark_universe_3mo$days_from_surgery, na.rm = TRUE), 0), "days\n")
cat("IQR:", round(quantile(landmark_universe_3mo$days_from_surgery, 0.25, na.rm = TRUE), 0), "-",
    round(quantile(landmark_universe_3mo$days_from_surgery, 0.75, na.rm = TRUE), 0), "days\n")

```

## Landmark Analysis Rationale

```{r landmark_analysis_display}
recommended_landmark_months <- 3
recommended_landmark_universe <- landmark_universe_3mo
recommended_landmark_n <- nrow(landmark_universe_3mo)

cat("### Landmark Framework\n\n")
cat("For risk stratification, we focus on a **3-month post-operative landmark (90 days, ±30-day window)** so that all patients are evaluated over a comparable follow-up interval.\n\n")
cat("**Landmark Universe for Recommended Analysis:** Patients with:\n")
cat("- Pre-cystectomy ctDNA measurement (within 60 days prior to surgery)\n")
cat("- Post-cystectomy ctDNA measurement within the ±", landmark_window_days, "-day window around the 3-month landmark (60–120 days post-surgery)\n\n", sep = "")
cat("**N =", recommended_landmark_n, " patients**\n\n", sep = "")
```

## Results

### Threshold Analysis

```{r threshold_analysis, echo=FALSE}
# ROC analysis with optimal threshold identification using Youden criterion
# Primary analysis: Use original universe (first ctDNA within 1 year, from 30 days onwards)
analysis_data <- universe

# Landmark analysis: Use landmark universes (ctDNA at 3-month and 6-month landmarks)
if (exists("landmark_universe_3mo") && nrow(landmark_universe_3mo) > 0) {
  landmark_analysis_data_3mo <- landmark_universe_3mo %>%
    mutate(
      post_cystectomy_positive = post_cystectomy_positive_landmark
    )
} else {
  landmark_analysis_data_3mo <- NULL
}

if (exists("landmark_universe_6mo") && nrow(landmark_universe_6mo) > 0) {
  landmark_analysis_data_6mo <- landmark_universe_6mo %>%
    mutate(
      post_cystectomy_positive = post_cystectomy_positive_landmark
    )
} else {
  landmark_analysis_data_6mo <- NULL
}

# For backward compatibility
if (exists("landmark_universe") && nrow(landmark_universe) > 0) {
  landmark_analysis_data <- landmark_universe %>%
    mutate(
      post_cystectomy_positive = post_cystectomy_positive_landmark
    )
} else {
  landmark_analysis_data <- landmark_analysis_data_6mo
}

if (nrow(analysis_data) > 10) {
  # Prepare data for cutpointR
  analysis_data_clean <- analysis_data %>%
    filter(!is.na(pre_cystectomy_ctdna_value) & !is.na(post_cystectomy_positive)) %>%
    select(x = pre_cystectomy_ctdna_value, class = post_cystectomy_positive)
  
  if (nrow(analysis_data_clean) > 10) {
    # First, find the highest threshold with sensitivity >= 80% and specificity >= 50%
    # Calculate ROC curve to get all thresholds
    roc_obj_temp <- pROC::roc(analysis_data_clean$class, analysis_data_clean$x, quiet = TRUE)
    coords_all_temp <- pROC::coords(roc_obj_temp, ret = c("threshold", "sensitivity", "specificity", "ppv", "npv"), transpose = FALSE)
    
    # Filter for thresholds meeting criteria: sensitivity >= 80% and specificity >= 50%
    # Prioritize sensitivity - find highest threshold meeting both criteria
    valid_thresholds <- coords_all_temp %>%
      filter(sensitivity >= 0.80 & specificity >= 0.50) %>%
      arrange(desc(threshold))  # Highest threshold = maximizes specificity while maintaining sensitivity
    
    if (nrow(valid_thresholds) > 0) {
      # Select the highest threshold meeting criteria (maximizes specificity while maintaining sensitivity)
      optimal_threshold_manual <- valid_thresholds$threshold[1]
      optimal_sensitivity_manual <- valid_thresholds$sensitivity[1]
      optimal_specificity_manual <- valid_thresholds$specificity[1]
      optimal_ppv_manual <- valid_thresholds$ppv[1]
      optimal_npv_manual <- valid_thresholds$npv[1]
    } else {
      # If no threshold meets both criteria, use the highest threshold with sensitivity >= 80%
      valid_thresholds_sp <- coords_all_temp %>%
        filter(sensitivity >= 0.80) %>%
        arrange(desc(threshold))
      if (nrow(valid_thresholds_sp) > 0) {
        optimal_threshold_manual <- valid_thresholds_sp$threshold[1]
        optimal_sensitivity_manual <- valid_thresholds_sp$sensitivity[1]
        optimal_specificity_manual <- valid_thresholds_sp$specificity[1]
        optimal_ppv_manual <- valid_thresholds_sp$ppv[1]
        optimal_npv_manual <- valid_thresholds_sp$npv[1]
      } else {
        # Fallback: use Youden criterion
        coords_all_temp <- coords_all_temp %>%
          mutate(youden = sensitivity + specificity - 1) %>%
          arrange(desc(youden))
        optimal_threshold_manual <- coords_all_temp$threshold[1]
        optimal_sensitivity_manual <- coords_all_temp$sensitivity[1]
        optimal_specificity_manual <- coords_all_temp$specificity[1]
        optimal_ppv_manual <- coords_all_temp$ppv[1]
        optimal_npv_manual <- coords_all_temp$npv[1]
      }
    }
    
    # Use optimal threshold from manual calculation
    optimal_threshold_cutpointr <- optimal_threshold_manual
    optimal_sensitivity_cutpointr <- optimal_sensitivity_manual
    optimal_specificity_cutpointr <- optimal_specificity_manual
    optimal_ppv_cutpointr <- optimal_ppv_manual
    optimal_npv_cutpointr <- optimal_npv_manual
    
    # Perform manual bootstrap validation
    # Function to find optimal threshold on a dataset
    find_optimal_threshold <- function(data_subset) {
      if (nrow(data_subset) < 10) return(list(threshold = NA, sensitivity = NA, specificity = NA, ppv = NA, npv = NA))
      
      roc_temp <- tryCatch({
        pROC::roc(data_subset$class, data_subset$x, quiet = TRUE)
      }, error = function(e) NULL)
      
      if (is.null(roc_temp)) return(list(threshold = NA, sensitivity = NA, specificity = NA, ppv = NA, npv = NA))
      
      coords_temp <- tryCatch({
        pROC::coords(roc_temp, ret = c("threshold", "sensitivity", "specificity", "ppv", "npv"), transpose = FALSE)
      }, error = function(e) NULL)
      
      if (is.null(coords_temp) || nrow(coords_temp) == 0) {
        return(list(threshold = NA, sensitivity = NA, specificity = NA, ppv = NA, npv = NA))
      }
      
      # Filter for thresholds meeting criteria: sensitivity >= 80% and specificity >= 50%
      valid_thresholds <- coords_temp %>%
        filter(sensitivity >= 0.80 & specificity >= 0.50) %>%
        arrange(desc(threshold))  # Highest threshold = maximizes specificity while maintaining sensitivity
      
      if (nrow(valid_thresholds) > 0) {
        return(list(
          threshold = valid_thresholds$threshold[1],
          sensitivity = valid_thresholds$sensitivity[1],
          specificity = valid_thresholds$specificity[1],
          ppv = valid_thresholds$ppv[1],
          npv = valid_thresholds$npv[1]
        ))
      } else {
        # Fallback: highest threshold with sensitivity >= 80%
        valid_sp <- coords_temp %>%
          filter(sensitivity >= 0.80) %>%
          arrange(desc(threshold))
        if (nrow(valid_sp) > 0) {
          return(list(
            threshold = valid_sp$threshold[1],
            sensitivity = valid_sp$sensitivity[1],
            specificity = valid_sp$specificity[1],
            ppv = valid_sp$ppv[1],
            npv = valid_sp$npv[1]
          ))
        } else {
          # Final fallback: Youden
          coords_temp <- coords_temp %>%
            mutate(youden = sensitivity + specificity - 1) %>%
            arrange(desc(youden))
          return(list(
            threshold = coords_temp$threshold[1],
            sensitivity = coords_temp$sensitivity[1],
            specificity = coords_temp$specificity[1],
            ppv = coords_temp$ppv[1],
            npv = coords_temp$npv[1]
          ))
        }
      }
    }
    
    # Perform bootstrap (1000 iterations)
    set.seed(123)
    n_boot <- 1000
    boot_results <- list()
    
    for (i in 1:n_boot) {
      # Stratified bootstrap: maintain class distribution
      boot_indices <- c(
        sample(which(analysis_data_clean$class == 1), size = sum(analysis_data_clean$class == 1), replace = TRUE),
        sample(which(analysis_data_clean$class == 0), size = sum(analysis_data_clean$class == 0), replace = TRUE)
      )
      boot_data <- analysis_data_clean[boot_indices, ]
      boot_results[[i]] <- find_optimal_threshold(boot_data)
    }
    
    # Convert to data frame
    boot_df <- do.call(rbind, lapply(boot_results, function(x) {
      data.frame(
        threshold = x$threshold,
        sensitivity = x$sensitivity,
        specificity = x$specificity,
        ppv = x$ppv,
        npv = x$npv
      )
    }))
    
    # Remove rows with all NA and enforce finite thresholds (exclude Inf)
    boot_df_clean <- boot_df %>%
      filter(!is.na(threshold) | !is.na(sensitivity) | !is.na(specificity)) %>%
      filter(is.finite(threshold))
    
    if (nrow(boot_df_clean) > 0) {
      boot_optimal <- list(
        threshold_median = median(boot_df_clean$threshold, na.rm = TRUE),
        threshold_lower = quantile(boot_df_clean$threshold, 0.025, na.rm = TRUE, names = FALSE),
        threshold_upper = quantile(boot_df_clean$threshold, 0.975, na.rm = TRUE, names = FALSE),
        sensitivity_mean = mean(boot_df_clean$sensitivity, na.rm = TRUE),
        sensitivity_lower = quantile(boot_df_clean$sensitivity, 0.025, na.rm = TRUE, names = FALSE),
        sensitivity_upper = quantile(boot_df_clean$sensitivity, 0.975, na.rm = TRUE, names = FALSE),
        specificity_mean = mean(boot_df_clean$specificity, na.rm = TRUE),
        specificity_lower = quantile(boot_df_clean$specificity, 0.025, na.rm = TRUE, names = FALSE),
        specificity_upper = quantile(boot_df_clean$specificity, 0.975, na.rm = TRUE, names = FALSE),
        ppv_mean = mean(boot_df_clean$ppv, na.rm = TRUE),
        ppv_lower = quantile(boot_df_clean$ppv, 0.025, na.rm = TRUE, names = FALSE),
        ppv_upper = quantile(boot_df_clean$ppv, 0.975, na.rm = TRUE, names = FALSE),
        npv_mean = mean(boot_df_clean$npv, na.rm = TRUE),
        npv_lower = quantile(boot_df_clean$npv, 0.025, na.rm = TRUE, names = FALSE),
        npv_upper = quantile(boot_df_clean$npv, 0.975, na.rm = TRUE, names = FALSE),
        optimism_se = optimal_sensitivity_cutpointr - mean(boot_df_clean$sensitivity, na.rm = TRUE),
        optimism_sp = optimal_specificity_cutpointr - mean(boot_df_clean$specificity, na.rm = TRUE),
        optimism_ppv = optimal_ppv_cutpointr - mean(boot_df_clean$ppv, na.rm = TRUE),
        optimism_npv = optimal_npv_cutpointr - mean(boot_df_clean$npv, na.rm = TRUE)
      )
    } else {
      boot_optimal <- NULL
    }
    
    # Also calculate Youden for comparison
    roc_obj <- pROC::roc(analysis_data$post_cystectomy_positive, analysis_data$pre_cystectomy_ctdna_value)
    coords_all <- coords(roc_obj, ret = c("threshold", "sensitivity", "specificity"), transpose = FALSE)
    if (is.data.frame(coords_all) && nrow(coords_all) > 0) {
      coords_all <- coords_all %>%
        mutate(youden = sensitivity + specificity - 1) %>%
        arrange(desc(youden))
      optimal_threshold_youden <- coords_all$threshold[1]
      optimal_sensitivity_youden <- coords_all$sensitivity[1]
      optimal_specificity_youden <- coords_all$specificity[1]
    } else {
      optimal_threshold_youden <- NA
      optimal_sensitivity_youden <- NA
      optimal_specificity_youden <- NA
    }
    
    # Use cutpointR optimal threshold as primary
    optimal_threshold <- optimal_threshold_cutpointr
  
    # Store optimal metrics from cutpointR
    # Ensure all values are properly initialized (not NULL or length 0)
    coords_optimal <- list(
      threshold = if (length(optimal_threshold) == 0 || is.null(optimal_threshold) || (length(optimal_threshold) > 0 && is.na(optimal_threshold))) NA_real_ else optimal_threshold,
      sensitivity = if (length(optimal_sensitivity_cutpointr) == 0 || is.null(optimal_sensitivity_cutpointr) || (length(optimal_sensitivity_cutpointr) > 0 && is.na(optimal_sensitivity_cutpointr))) NA_real_ else optimal_sensitivity_cutpointr,
      specificity = if (length(optimal_specificity_cutpointr) == 0 || is.null(optimal_specificity_cutpointr) || (length(optimal_specificity_cutpointr) > 0 && is.na(optimal_specificity_cutpointr))) NA_real_ else optimal_specificity_cutpointr,
      ppv = if (length(optimal_ppv_cutpointr) == 0 || is.null(optimal_ppv_cutpointr) || (length(optimal_ppv_cutpointr) > 0 && is.na(optimal_ppv_cutpointr))) NA_real_ else optimal_ppv_cutpointr,
      npv = if (length(optimal_npv_cutpointr) == 0 || is.null(optimal_npv_cutpointr) || (length(optimal_npv_cutpointr) > 0 && is.na(optimal_npv_cutpointr))) NA_real_ else optimal_npv_cutpointr
    )
    
    # Also keep ROC object for plotting
    if (!exists("roc_obj")) {
      roc_obj <- pROC::roc(analysis_data$post_cystectomy_positive, analysis_data$pre_cystectomy_ctdna_value)
    }
  } else {
    # If cutpointR fails, fall back to Youden
    roc_obj <- pROC::roc(analysis_data$post_cystectomy_positive, analysis_data$pre_cystectomy_ctdna_value)
    coords_all <- coords(roc_obj, ret = c("threshold", "sensitivity", "specificity"), transpose = FALSE)
    if (is.data.frame(coords_all) && nrow(coords_all) > 0) {
      coords_all <- coords_all %>%
        mutate(youden = sensitivity + specificity - 1) %>%
        arrange(desc(youden))
      optimal_threshold <- coords_all$threshold[1]
      coords_optimal <- list(
        threshold = optimal_threshold,
        sensitivity = coords_all$sensitivity[1],
        specificity = coords_all$specificity[1],
        ppv = NA,
        npv = NA
      )
    } else {
      optimal_threshold <- 0.1
      coords_optimal <- list(threshold = 0.1, sensitivity = NA, specificity = NA, ppv = NA, npv = NA)
    }
    boot_optimal <- NULL
  }
  
  # Find sensitivity-focused thresholds (for high-stakes scenario)
  # Get all coordinates to find thresholds with high sensitivity
  all_coords <- coords(roc_obj, ret = c("threshold", "sensitivity", "specificity", "ppv", "npv"), transpose = FALSE)
  
  # Find threshold with sensitivity ≥ 95% (use lowest threshold that achieves this)
  coords_sens95_df <- all_coords %>%
    filter(sensitivity >= 0.95) %>%
    arrange(threshold) %>%  # Lower threshold = more sensitive
    slice_head(n = 1)
  
  if (nrow(coords_sens95_df) > 0) {
    coords_sens95 <- list(
      threshold = coords_sens95_df$threshold[1],
      sensitivity = coords_sens95_df$sensitivity[1],
      specificity = coords_sens95_df$specificity[1],
      ppv = coords_sens95_df$ppv[1],
      npv = coords_sens95_df$npv[1]
    )
  } else {
    coords_sens95 <- NULL
  }
  
  # Find threshold with sensitivity ≥ 90% (use lowest threshold that achieves this)
  coords_sens90_df <- all_coords %>%
    filter(sensitivity >= 0.90) %>%
    arrange(threshold) %>%  # Lower threshold = more sensitive
    slice_head(n = 1)
  
  if (nrow(coords_sens90_df) > 0) {
    coords_sens90 <- list(
      threshold = coords_sens90_df$threshold[1],
      sensitivity = coords_sens90_df$sensitivity[1],
      specificity = coords_sens90_df$specificity[1],
      ppv = coords_sens90_df$ppv[1],
      npv = coords_sens90_df$npv[1]
    )
  } else {
    coords_sens90 <- NULL
  }
  
} else {
  cat("Insufficient data for ROC analysis (need at least 10 patients with post-cystectomy ctDNA)\n")
}
```

```{r threshold_method_display, echo=FALSE}
# Display optimal threshold results
if (exists("coords_optimal") && !is.null(coords_optimal) && length(coords_optimal) > 0) {
  cat("### Primary Analysis: Optimal Threshold Results (Apparent/Overfitted)\n\n")
  cat("**Note:** The threshold and diagnostic measures shown below are **apparent values** identified on the original sample. These are overfitted and should be interpreted with caution. The bootstrap-validated measures (shown below) provide more realistic estimates.\n\n")
  
  if (!is.null(coords_optimal$threshold) && length(coords_optimal$threshold) > 0 && !is.na(coords_optimal$threshold)) {
    cat("**Identified Optimal Threshold (Apparent):**", round(coords_optimal$threshold, 3), "\n\n")
  }
  cat("**Diagnostic Measures at This Threshold (Apparent/Overfitted):**\n\n")
  if (!is.null(coords_optimal$sensitivity) && length(coords_optimal$sensitivity) > 0 && !is.na(coords_optimal$sensitivity)) {
    cat("- Sensitivity =", round(coords_optimal$sensitivity * 100, 1), "%\n")
  }
  if (!is.null(coords_optimal$specificity) && length(coords_optimal$specificity) > 0 && !is.na(coords_optimal$specificity)) {
    cat("- Specificity =", round(coords_optimal$specificity * 100, 1), "%\n")
  }
  if (!is.null(coords_optimal$ppv) && length(coords_optimal$ppv) > 0 && !is.na(coords_optimal$ppv)) {
    cat("- PPV =", round(coords_optimal$ppv * 100, 1), "%\n")
  }
  if (!is.null(coords_optimal$npv) && length(coords_optimal$npv) > 0 && !is.na(coords_optimal$npv)) {
    cat("- NPV =", round(coords_optimal$npv * 100, 1), "%\n")
  }
  cat("\n")
}
```

```{r bootstrap_validation_display, echo=FALSE}
# Display bootstrap validation results for primary analysis
if (exists("boot_optimal") && !is.null(boot_optimal)) {
  # Check if bootstrap results are valid (not all NA)
  has_valid_results <- !is.na(boot_optimal$threshold_median) || 
                       !is.na(boot_optimal$sensitivity_mean) || 
                       !is.na(boot_optimal$specificity_mean)
  
  if (has_valid_results) {
    cat("### Bootstrap Internal Validation Results (Primary Analysis)\n\n")
    cat("**Method:** Bootstrap resampling (n = 1000 iterations) was performed on the primary analysis sample to estimate the optimism in diagnostic measures. On each bootstrap sample, the optimal threshold was identified using the same criteria (sensitivity ≥ 80% and specificity ≥ 50%), and the median threshold across all bootstrap samples is reported.\n\n")
    cat("**Note:** The bootstrap-validated threshold (median across bootstrap samples) may differ from the apparent threshold identified on the original sample. This is expected and reflects the variability in threshold selection across different samples. The bootstrap-validated threshold is generally more reliable for future patients.\n\n")
    cat("**Bootstrap-Validated Diagnostic Measures (95% CI):**\n\n")
    
    # Threshold
    if (!is.na(boot_optimal$threshold_median)) {
      cat("- **Threshold:**", round(boot_optimal$threshold_median, 3))
      if (!is.na(boot_optimal$threshold_lower) && !is.na(boot_optimal$threshold_upper)) {
        cat(" (", round(boot_optimal$threshold_lower, 3), "-", round(boot_optimal$threshold_upper, 3), ")\n")
      } else {
        cat("\n")
      }
    } else {
      cat("- **Threshold:** NA\n")
    }
    
    # Sensitivity
    if (!is.na(boot_optimal$sensitivity_mean)) {
      cat("- **Sensitivity:**", round(boot_optimal$sensitivity_mean * 100, 1), "%")
      if (!is.na(boot_optimal$sensitivity_lower) && !is.na(boot_optimal$sensitivity_upper)) {
        cat(" (", round(boot_optimal$sensitivity_lower * 100, 1), "-", round(boot_optimal$sensitivity_upper * 100, 1), "%)\n")
      } else {
        cat("\n")
      }
    } else {
      cat("- **Sensitivity:** NA\n")
    }
    
    # Specificity
    if (!is.na(boot_optimal$specificity_mean)) {
      cat("- **Specificity:**", round(boot_optimal$specificity_mean * 100, 1), "%")
      if (!is.na(boot_optimal$specificity_lower) && !is.na(boot_optimal$specificity_upper)) {
        cat(" (", round(boot_optimal$specificity_lower * 100, 1), "-", round(boot_optimal$specificity_upper * 100, 1), "%)\n")
      } else {
        cat("\n")
      }
    } else {
      cat("- **Specificity:** NA\n")
    }
    
    # PPV
    if (!is.na(boot_optimal$ppv_mean)) {
      cat("- **PPV:**", round(boot_optimal$ppv_mean * 100, 1), "%")
      if (!is.na(boot_optimal$ppv_lower) && !is.na(boot_optimal$ppv_upper)) {
        cat(" (", round(boot_optimal$ppv_lower * 100, 1), "-", round(boot_optimal$ppv_upper * 100, 1), "%)\n")
      } else {
        cat("\n")
      }
    } else {
      cat("- **PPV:** NA\n")
    }
    
    # NPV
    if (!is.na(boot_optimal$npv_mean)) {
      cat("- **NPV:**", round(boot_optimal$npv_mean * 100, 1), "%")
      if (!is.na(boot_optimal$npv_lower) && !is.na(boot_optimal$npv_upper)) {
        cat(" (", round(boot_optimal$npv_lower * 100, 1), "-", round(boot_optimal$npv_upper * 100, 1), "%)\n")
      } else {
        cat("\n")
      }
    } else {
      cat("- **NPV:** NA\n")
    }
    
    cat("\n")
    cat("**Optimism (Apparent - Bootstrap Mean):**\n")
    if (!is.na(boot_optimal$optimism_se)) {
      cat("- Sensitivity optimism:", round(boot_optimal$optimism_se * 100, 1), "percentage points\n")
    } else {
      cat("- Sensitivity optimism: NA\n")
    }
    if (!is.na(boot_optimal$optimism_sp)) {
      cat("- Specificity optimism:", round(boot_optimal$optimism_sp * 100, 1), "percentage points\n")
    } else {
      cat("- Specificity optimism: NA\n")
    }
    if (!is.na(boot_optimal$optimism_ppv)) {
      cat("- PPV optimism:", round(boot_optimal$optimism_ppv * 100, 1), "percentage points\n")
    } else {
      cat("- PPV optimism: NA\n")
    }
    if (!is.na(boot_optimal$optimism_npv)) {
      cat("- NPV optimism:", round(boot_optimal$optimism_npv * 100, 1), "percentage points\n")
    } else {
      cat("- NPV optimism: NA\n")
    }
    
    cat("\n")
    cat("**Interpretation:** The bootstrap-validated measures provide more realistic estimates of performance in new patients compared to the apparent (overfitted) measures.\n\n")
  } else {
    cat("### Bootstrap Validation\n\n")
    cat("**Note:** Bootstrap validation was attempted but could not be completed. The diagnostic measures shown are apparent (overfitted) estimates and should be interpreted with caution.\n\n")
  }
} else if (exists("coords_optimal") && !is.null(coords_optimal)) {
  cat("### Bootstrap Validation\n\n")
  cat("**Note:** Bootstrap validation was attempted but could not be completed. The diagnostic measures shown are apparent (overfitted) estimates and should be interpreted with caution.\n\n")
}
```

### Association Analysis: Logistic Regression

```{r logistic_regression_analysis, echo=FALSE}
# Logistic regression analysis with pre-cystectomy ctDNA as continuous predictor
if (exists("analysis_data") && nrow(analysis_data) > 10) {
  # Prepare data for logistic regression
  logreg_data <- analysis_data %>%
    filter(!is.na(pre_cystectomy_ctdna_value) & !is.na(post_cystectomy_positive)) %>%
    mutate(
      log_ctdna = log10(pre_cystectomy_ctdna_value + 0.001),  # Add small constant to handle zeros, log10 transform
      nac_status = if (exists("universe") && "neoadjuvant_status" %in% names(universe)) {
        universe$neoadjuvant_status[match(patient_id, universe$patient_id)]
      } else {
        NA
      }
    )
  
  if (nrow(logreg_data) > 10) {
    # Model 1: Pre-cystectomy ctDNA only (log-transformed)
    model1 <- glm(post_cystectomy_positive ~ log_ctdna, 
                 data = logreg_data, 
                 family = binomial(link = "logit"))
    
    model1_summary <- summary(model1)
    model1_coef <- coef(model1_summary)
    model1_or <- exp(model1_coef["log_ctdna", "Estimate"])
    model1_or_ci <- exp(confint(model1))
    model1_auc <- pROC::auc(pROC::roc(logreg_data$post_cystectomy_positive, 
                                       predict(model1, type = "response")))
    
    cat("### Association Between Pre-Cystectomy ctDNA and Post-Cystectomy ctDNA Positivity\n\n")
    cat("**Model 1: Pre-cystectomy ctDNA (log10-transformed) as continuous predictor**\n\n")
    cat("- **Odds Ratio per 10-fold increase in pre-cystectomy ctDNA:**", 
        round(model1_or, 2), 
        " (95% CI: ", round(model1_or_ci["log_ctdna", "2.5 %"], 2), 
        "-", round(model1_or_ci["log_ctdna", "97.5 %"], 2), ")\n", sep = "")
    cat("- **P-value:**", format.pval(model1_coef["log_ctdna", "Pr(>|z|)"], digits = 3), "\n")
    cat("- **AUC:**", round(as.numeric(model1_auc), 3), "\n\n")
    
    # Model 2: Pre-cystectomy ctDNA + NAC status (if available)
    if (!all(is.na(logreg_data$nac_status)) && sum(!is.na(logreg_data$nac_status)) > 5) {
      logreg_data_nac <- logreg_data %>%
        filter(!is.na(nac_status)) %>%
        mutate(nac_binary = as.numeric(nac_status == "yes" | nac_status == "Yes" | nac_status == 1))
      
      if (nrow(logreg_data_nac) > 10 && length(unique(logreg_data_nac$nac_binary)) > 1) {
        model2 <- glm(post_cystectomy_positive ~ log_ctdna + nac_binary, 
                     data = logreg_data_nac, 
                     family = binomial(link = "logit"))
        
        model2_summary <- summary(model2)
        model2_coef <- coef(model2_summary)
        model2_or_ctdna <- exp(model2_coef["log_ctdna", "Estimate"])
        model2_or_ctdna_ci <- exp(confint(model2)["log_ctdna", ])
        model2_or_nac <- exp(model2_coef["nac_binary", "Estimate"])
        model2_or_nac_ci <- exp(confint(model2)["nac_binary", ])
        model2_auc <- pROC::auc(pROC::roc(logreg_data_nac$post_cystectomy_positive, 
                                          predict(model2, type = "response")))
        
        cat("**Model 2: Pre-cystectomy ctDNA + Neoadjuvant Chemotherapy Status**\n\n")
        cat("- **Odds Ratio per 10-fold increase in pre-cystectomy ctDNA:**", 
            round(model2_or_ctdna, 2), 
            " (95% CI: ", round(model2_or_ctdna_ci[1], 2), 
            "-", round(model2_or_ctdna_ci[2], 2), ")\n", sep = "")
        cat("- **Odds Ratio for NAC vs. no NAC:**", 
            round(model2_or_nac, 2), 
            " (95% CI: ", round(model2_or_nac_ci[1], 2), 
            "-", round(model2_or_nac_ci[2], 2), ")\n", sep = "")
        cat("- **AUC:**", round(as.numeric(model2_auc), 3), "\n\n")
        
        # Model 3: With interaction (if sample size allows)
        if (nrow(logreg_data_nac) > 20) {
          model3 <- glm(post_cystectomy_positive ~ log_ctdna * nac_binary, 
                       data = logreg_data_nac, 
                       family = binomial(link = "logit"))
          
          model3_summary <- summary(model3)
          model3_coef <- coef(model3_summary)
          model3_auc <- pROC::auc(pROC::roc(logreg_data_nac$post_cystectomy_positive, 
                                            predict(model3, type = "response")))
          
          cat("**Model 3: Pre-cystectomy ctDNA × Neoadjuvant Chemotherapy Interaction**\n\n")
          cat("- **AUC:**", round(as.numeric(model3_auc), 3), "\n")
          cat("- **Note:** Interaction model included to assess whether the association between pre-cystectomy ctDNA and post-cystectomy positivity differs by NAC status.\n\n")
        }
      }
    }
    
    cat("**Interpretation:** The odds ratio represents the change in odds of post-cystectomy ctDNA positivity associated with each 10-fold increase in pre-cystectomy ctDNA level. An OR > 1 indicates that higher pre-cystectomy ctDNA is associated with increased likelihood of post-cystectomy ctDNA positivity.\n\n")
  }
}
```

### 3-Month Landmark Analysis: Risk Stratification for Post-Operative MRD

```{r landmark_threshold_analysis, echo=FALSE}
# ROC analysis using landmark data (3-month landmark only)
# Primary endpoint: post-cystectomy ctDNA positivity at 3-month landmark
# Primary role: Risk stratification for anticipating post-operative MRD

cat("**Primary Population:** All cystectomy patients (mixed cohort)\n\n")
cat("**Primary Endpoint:** Post-cystectomy ctDNA positivity at 3-month landmark (90 days post-surgery, ±30 day window)\n\n")

cat("**Why the Landmark Threshold May Differ from Primary Analysis:**\n\n")
cat("The 3-month landmark analysis uses a different dataset than the primary analysis:\n")
cat("- **Primary Analysis:** Uses the FIRST post-cystectomy ctDNA measurement from 4 weeks onward (can be at any timepoint within 1 year)\n")
cat("- **Landmark Analysis:** Uses ctDNA measurements specifically at the 3-month landmark timepoint (90 days ± 30 days)\n\n")
cat("These different patient populations and timepoints can result in different optimal thresholds, even when using the same selection criteria (sensitivity ≥ 80% and specificity ≥ 50%). The landmark analysis ensures all patients have the same observation period, which is important for risk stratification.\n\n")

# 3-month landmark analysis
if (!is.null(landmark_analysis_data_3mo) && nrow(landmark_analysis_data_3mo) > 10) {
  # Prepare data for cutpointR
  landmark_data_3mo_clean <- landmark_analysis_data_3mo %>%
    filter(!is.na(pre_cystectomy_ctdna_value) & !is.na(post_cystectomy_positive)) %>%
    select(x = pre_cystectomy_ctdna_value, class = post_cystectomy_positive)
  
  if (nrow(landmark_data_3mo_clean) > 10) {
    # First, find the highest threshold with sensitivity >= 80% and specificity >= 50%
    # Calculate ROC curve to get all thresholds
    roc_obj_3mo_temp <- pROC::roc(landmark_data_3mo_clean$class, landmark_data_3mo_clean$x, quiet = TRUE)
    coords_all_3mo_temp <- pROC::coords(roc_obj_3mo_temp, ret = c("threshold", "sensitivity", "specificity", "ppv", "npv"), transpose = FALSE)
    
    # Filter for thresholds meeting criteria: sensitivity >= 80% and specificity >= 50%
    # Prioritize sensitivity - find highest threshold meeting both criteria
    valid_thresholds_3mo <- coords_all_3mo_temp %>%
      filter(sensitivity >= 0.80 & specificity >= 0.50) %>%
      arrange(desc(threshold))  # Highest threshold = maximizes specificity while maintaining sensitivity
    
    if (nrow(valid_thresholds_3mo) > 0) {
      # Select the highest threshold meeting criteria (maximizes specificity while maintaining sensitivity)
      optimal_threshold_3mo_manual <- valid_thresholds_3mo$threshold[1]
      optimal_sensitivity_3mo_manual <- valid_thresholds_3mo$sensitivity[1]
      optimal_specificity_3mo_manual <- valid_thresholds_3mo$specificity[1]
      optimal_ppv_3mo_manual <- valid_thresholds_3mo$ppv[1]
      optimal_npv_3mo_manual <- valid_thresholds_3mo$npv[1]
    } else {
      # If no threshold meets both criteria, use the highest threshold with sensitivity >= 80%
      valid_thresholds_3mo_sp <- coords_all_3mo_temp %>%
        filter(sensitivity >= 0.80) %>%
        arrange(desc(threshold))
      if (nrow(valid_thresholds_3mo_sp) > 0) {
        optimal_threshold_3mo_manual <- valid_thresholds_3mo_sp$threshold[1]
        optimal_sensitivity_3mo_manual <- valid_thresholds_3mo_sp$sensitivity[1]
        optimal_specificity_3mo_manual <- valid_thresholds_3mo_sp$specificity[1]
        optimal_ppv_3mo_manual <- valid_thresholds_3mo_sp$ppv[1]
        optimal_npv_3mo_manual <- valid_thresholds_3mo_sp$npv[1]
      } else {
        # Fallback: use Youden criterion
        coords_all_3mo_temp <- coords_all_3mo_temp %>%
          mutate(youden = sensitivity + specificity - 1) %>%
          arrange(desc(youden))
        optimal_threshold_3mo_manual <- coords_all_3mo_temp$threshold[1]
        optimal_sensitivity_3mo_manual <- coords_all_3mo_temp$sensitivity[1]
        optimal_specificity_3mo_manual <- coords_all_3mo_temp$specificity[1]
        optimal_ppv_3mo_manual <- coords_all_3mo_temp$ppv[1]
        optimal_npv_3mo_manual <- coords_all_3mo_temp$npv[1]
      }
    }
    
    # Use optimal threshold from manual calculation
    optimal_threshold_3mo <- optimal_threshold_3mo_manual
    
    # Store optimal metrics
    coords_optimal_3mo <- list(
      threshold = optimal_threshold_3mo,
      sensitivity = optimal_sensitivity_3mo_manual,
      specificity = optimal_specificity_3mo_manual,
      ppv = optimal_ppv_3mo_manual,
      npv = optimal_npv_3mo_manual
    )
    
    # Perform manual bootstrap validation (same function as primary analysis)
    find_optimal_threshold_3mo <- function(data_subset) {
      if (nrow(data_subset) < 10) return(list(threshold = NA, sensitivity = NA, specificity = NA, ppv = NA, npv = NA))
      
      roc_temp <- tryCatch({
        pROC::roc(data_subset$class, data_subset$x, quiet = TRUE)
      }, error = function(e) NULL)
      
      if (is.null(roc_temp)) return(list(threshold = NA, sensitivity = NA, specificity = NA, ppv = NA, npv = NA))
      
      coords_temp <- tryCatch({
        pROC::coords(roc_temp, ret = c("threshold", "sensitivity", "specificity", "ppv", "npv"), transpose = FALSE)
      }, error = function(e) NULL)
      
      if (is.null(coords_temp) || nrow(coords_temp) == 0) {
        return(list(threshold = NA, sensitivity = NA, specificity = NA, ppv = NA, npv = NA))
      }
      
      # Filter for thresholds meeting criteria: sensitivity >= 80% and specificity >= 50%
      valid_thresholds <- coords_temp %>%
        filter(sensitivity >= 0.80 & specificity >= 0.50) %>%
        arrange(desc(threshold))  # Highest threshold = maximizes specificity while maintaining sensitivity
      
      if (nrow(valid_thresholds) > 0) {
        return(list(
          threshold = valid_thresholds$threshold[1],
          sensitivity = valid_thresholds$sensitivity[1],
          specificity = valid_thresholds$specificity[1],
          ppv = valid_thresholds$ppv[1],
          npv = valid_thresholds$npv[1]
        ))
      } else {
        # Fallback: highest threshold with sensitivity >= 80%
        valid_sp <- coords_temp %>%
          filter(sensitivity >= 0.80) %>%
          arrange(desc(threshold))
        if (nrow(valid_sp) > 0) {
          return(list(
            threshold = valid_sp$threshold[1],
            sensitivity = valid_sp$sensitivity[1],
            specificity = valid_sp$specificity[1],
            ppv = valid_sp$ppv[1],
            npv = valid_sp$npv[1]
          ))
        } else {
          # Final fallback: Youden
          coords_temp <- coords_temp %>%
            mutate(youden = sensitivity + specificity - 1) %>%
            arrange(desc(youden))
          return(list(
            threshold = coords_temp$threshold[1],
            sensitivity = coords_temp$sensitivity[1],
            specificity = coords_temp$specificity[1],
            ppv = coords_temp$ppv[1],
            npv = coords_temp$npv[1]
          ))
        }
      }
    }
    
    # Perform bootstrap (1000 iterations)
    set.seed(123)
    n_boot_3mo <- 1000
    boot_results_3mo <- list()
    
    for (i in 1:n_boot_3mo) {
      # Stratified bootstrap: maintain class distribution
      boot_indices_3mo <- c(
        sample(which(landmark_data_3mo_clean$class == 1), size = sum(landmark_data_3mo_clean$class == 1), replace = TRUE),
        sample(which(landmark_data_3mo_clean$class == 0), size = sum(landmark_data_3mo_clean$class == 0), replace = TRUE)
      )
      boot_data_3mo <- landmark_data_3mo_clean[boot_indices_3mo, ]
      boot_results_3mo[[i]] <- find_optimal_threshold_3mo(boot_data_3mo)
    }
    
    # Convert to data frame
    boot_df_3mo <- do.call(rbind, lapply(boot_results_3mo, function(x) {
      data.frame(
        threshold = x$threshold,
        sensitivity = x$sensitivity,
        specificity = x$specificity,
        ppv = x$ppv,
        npv = x$npv
      )
    }))
    
    # Remove rows with all NA and enforce finite thresholds (exclude Inf)
    boot_df_3mo_clean <- boot_df_3mo %>%
      filter(!is.na(threshold) | !is.na(sensitivity) | !is.na(specificity)) %>%
      filter(is.finite(threshold))
    
    if (nrow(boot_df_3mo_clean) > 0) {
      boot_optimal_3mo <- list(
        threshold_median = median(boot_df_3mo_clean$threshold, na.rm = TRUE),
        threshold_lower = quantile(boot_df_3mo_clean$threshold, 0.025, na.rm = TRUE, names = FALSE),
        threshold_upper = quantile(boot_df_3mo_clean$threshold, 0.975, na.rm = TRUE, names = FALSE),
        sensitivity_mean = mean(boot_df_3mo_clean$sensitivity, na.rm = TRUE),
        sensitivity_lower = quantile(boot_df_3mo_clean$sensitivity, 0.025, na.rm = TRUE, names = FALSE),
        sensitivity_upper = quantile(boot_df_3mo_clean$sensitivity, 0.975, na.rm = TRUE, names = FALSE),
        specificity_mean = mean(boot_df_3mo_clean$specificity, na.rm = TRUE),
        specificity_lower = quantile(boot_df_3mo_clean$specificity, 0.025, na.rm = TRUE, names = FALSE),
        specificity_upper = quantile(boot_df_3mo_clean$specificity, 0.975, na.rm = TRUE, names = FALSE),
        ppv_mean = mean(boot_df_3mo_clean$ppv, na.rm = TRUE),
        ppv_lower = quantile(boot_df_3mo_clean$ppv, 0.025, na.rm = TRUE, names = FALSE),
        ppv_upper = quantile(boot_df_3mo_clean$ppv, 0.975, na.rm = TRUE, names = FALSE),
        npv_mean = mean(boot_df_3mo_clean$npv, na.rm = TRUE),
        npv_lower = quantile(boot_df_3mo_clean$npv, 0.025, na.rm = TRUE, names = FALSE),
        npv_upper = quantile(boot_df_3mo_clean$npv, 0.975, na.rm = TRUE, names = FALSE),
        optimism_se = coords_optimal_3mo$sensitivity - mean(boot_df_3mo_clean$sensitivity, na.rm = TRUE),
        optimism_sp = coords_optimal_3mo$specificity - mean(boot_df_3mo_clean$specificity, na.rm = TRUE),
        optimism_ppv = coords_optimal_3mo$ppv - mean(boot_df_3mo_clean$ppv, na.rm = TRUE),
        optimism_npv = coords_optimal_3mo$npv - mean(boot_df_3mo_clean$npv, na.rm = TRUE)
      )
    } else {
      boot_optimal_3mo <- NULL
    }
    
    # ROC curve for plotting
    roc_obj_3mo_threshold <- pROC::roc(landmark_analysis_data_3mo$post_cystectomy_positive, 
                                 landmark_analysis_data_3mo$pre_cystectomy_ctdna_value)
  } else {
    optimal_threshold_3mo <- 0.1
    boot_optimal_3mo <- NULL
    coords_optimal_3mo <- list(threshold = 0.1, sensitivity = NA, specificity = NA, ppv = NA, npv = NA)
    roc_obj_3mo_threshold <- pROC::roc(landmark_analysis_data_3mo$post_cystectomy_positive, 
                                landmark_analysis_data_3mo$pre_cystectomy_ctdna_value)
  }
  
  # Create threshold analysis table for 3-month landmark
  threshold_table_3mo <- tibble(
    `Metric` = c("Optimal Threshold (ctDNA)", "Sensitivity", "Specificity", "PPV", "NPV", "AUC (95% CI)"),
    `Apparent Value` = c(
      round(coords_optimal_3mo$threshold, 3),
      paste0(round(coords_optimal_3mo$sensitivity * 100, 1), "%"),
      paste0(round(coords_optimal_3mo$specificity * 100, 1), "%"),
      paste0(round(coords_optimal_3mo$ppv * 100, 1), "%"),
      paste0(round(coords_optimal_3mo$npv * 100, 1), "%"),
      paste0(round(ci.auc(roc_obj_3mo_threshold)[2], 3), " (", round(ci.auc(roc_obj_3mo_threshold)[1], 3), "-", round(ci.auc(roc_obj_3mo_threshold)[3], 3), ")")
    )
  )
  
  # Add bootstrap-validated values if available
  if (!is.null(boot_optimal_3mo)) {
    threshold_table_3mo$`Bootstrap-Validated Value (95% CI)` <- c(
      paste0(round(boot_optimal_3mo$threshold_median, 3), " (", round(boot_optimal_3mo$threshold_lower, 3), "-", round(boot_optimal_3mo$threshold_upper, 3), ")"),
      paste0(round(boot_optimal_3mo$sensitivity_mean * 100, 1), "% (", round(boot_optimal_3mo$sensitivity_lower * 100, 1), "-", round(boot_optimal_3mo$sensitivity_upper * 100, 1), "%)"),
      paste0(round(boot_optimal_3mo$specificity_mean * 100, 1), "% (", round(boot_optimal_3mo$specificity_lower * 100, 1), "-", round(boot_optimal_3mo$specificity_upper * 100, 1), "%)"),
      paste0(round(boot_optimal_3mo$ppv_mean * 100, 1), "% (", round(boot_optimal_3mo$ppv_lower * 100, 1), "-", round(boot_optimal_3mo$ppv_upper * 100, 1), "%)"),
      paste0(round(boot_optimal_3mo$npv_mean * 100, 1), "% (", round(boot_optimal_3mo$npv_lower * 100, 1), "-", round(boot_optimal_3mo$npv_upper * 100, 1), "%)"),
      "-"
    )
  }
  
  # Display apparent threshold results (similar to primary analysis)
  cat("### 3-Month Landmark Analysis: Optimal Threshold Results (Apparent/Overfitted)\n\n")
  cat("**Note:** The threshold and diagnostic measures shown below are **apparent values** identified on the original sample. These are overfitted and should be interpreted with caution. The bootstrap-validated measures (shown below) provide more realistic estimates.\n\n")
  
  if (!is.null(coords_optimal_3mo$threshold) && length(coords_optimal_3mo$threshold) > 0 && !is.na(coords_optimal_3mo$threshold)) {
    cat("**Identified Optimal Threshold (Apparent):**", round(coords_optimal_3mo$threshold, 3), "\n\n")
  }
  cat("**Diagnostic Measures at This Threshold (Apparent/Overfitted):**\n\n")
  if (!is.null(coords_optimal_3mo$sensitivity) && length(coords_optimal_3mo$sensitivity) > 0 && !is.na(coords_optimal_3mo$sensitivity)) {
    cat("- Sensitivity =", round(coords_optimal_3mo$sensitivity * 100, 1), "%\n")
  }
  if (!is.null(coords_optimal_3mo$specificity) && length(coords_optimal_3mo$specificity) > 0 && !is.na(coords_optimal_3mo$specificity)) {
    cat("- Specificity =", round(coords_optimal_3mo$specificity * 100, 1), "%\n")
  }
  if (!is.null(coords_optimal_3mo$ppv) && length(coords_optimal_3mo$ppv) > 0 && !is.na(coords_optimal_3mo$ppv)) {
    cat("- PPV =", round(coords_optimal_3mo$ppv * 100, 1), "%\n")
  }
  if (!is.null(coords_optimal_3mo$npv) && length(coords_optimal_3mo$npv) > 0 && !is.na(coords_optimal_3mo$npv)) {
    cat("- NPV =", round(coords_optimal_3mo$npv * 100, 1), "%\n")
  }
  cat("\n")
  
  threshold_table_3mo %>%
    kable(caption = "Optimal ctDNA Threshold for Predicting Post-Cystectomy Positive ctDNA at 3-Month Landmark (Highest Threshold with Sensitivity ≥ 80% and Specificity ≥ 50%)") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
  
  cat("\n**Note:** This analysis uses ctDNA measurements at the 3-month landmark timepoint, ensuring all patients have the same observation period. The threshold serves for **risk stratification** to anticipate post-operative minimal residual disease (MRD), not to determine surgical candidacy. Results are reported for the mixed cohort (all cystectomy patients), with NAC status considered in stratified analyses where appropriate.\n")
  if (!is.null(boot_optimal_3mo)) {
    cat("**Bootstrap Validation:** Bootstrap-validated measures (n = 1000 iterations) with 95% confidence intervals are shown to provide more realistic estimates of performance.\n")
  } else {
    cat("**Important:** The diagnostic measures shown above are apparent (overfitted) estimates. Bootstrap validation was attempted but could not be completed.\n")
  }
} else {
  cat("Insufficient data for 3-month landmark ROC analysis (need at least 10 patients with ctDNA at 3-month landmark)\n")
}
```

```{r bootstrap_validation_3mo_display, echo=FALSE}
# Display bootstrap validation results for 3-month landmark
if (exists("boot_optimal_3mo") && !is.null(boot_optimal_3mo)) {
  # Check if bootstrap results are valid (not all NA)
  has_valid_results_3mo <- !is.na(boot_optimal_3mo$threshold_median) || 
                           !is.na(boot_optimal_3mo$sensitivity_mean) || 
                           !is.na(boot_optimal_3mo$specificity_mean)
  
  if (has_valid_results_3mo) {
    cat("### Bootstrap Internal Validation Results (3-Month Landmark)\n\n")
    cat("**Method:** Bootstrap resampling (n = 1000 iterations) was performed to estimate the optimism in diagnostic measures. On each bootstrap sample, the optimal threshold was identified using the same criteria (sensitivity ≥ 80% and specificity ≥ 50%), and the median threshold across all bootstrap samples is reported.\n\n")
    cat("**Note:** The bootstrap-validated threshold (median across bootstrap samples) may differ from the apparent threshold identified on the original sample. This is expected and reflects the variability in threshold selection across different samples. The bootstrap-validated threshold is generally more reliable for future patients.\n\n")
    cat("**Bootstrap-Validated Diagnostic Measures (95% CI):**\n\n")
    
    # Threshold
    if (!is.na(boot_optimal_3mo$threshold_median)) {
      cat("- **Threshold:**", round(boot_optimal_3mo$threshold_median, 3))
      if (!is.na(boot_optimal_3mo$threshold_lower) && !is.na(boot_optimal_3mo$threshold_upper)) {
        cat(" (", round(boot_optimal_3mo$threshold_lower, 3), "-", round(boot_optimal_3mo$threshold_upper, 3), ")\n")
      } else {
        cat("\n")
      }
    } else {
      cat("- **Threshold:** NA\n")
    }
    
    # Sensitivity
    if (!is.na(boot_optimal_3mo$sensitivity_mean)) {
      cat("- **Sensitivity:**", round(boot_optimal_3mo$sensitivity_mean * 100, 1), "%")
      if (!is.na(boot_optimal_3mo$sensitivity_lower) && !is.na(boot_optimal_3mo$sensitivity_upper)) {
        cat(" (", round(boot_optimal_3mo$sensitivity_lower * 100, 1), "-", round(boot_optimal_3mo$sensitivity_upper * 100, 1), "%)\n")
      } else {
        cat("\n")
      }
    } else {
      cat("- **Sensitivity:** NA\n")
    }
    
    # Specificity
    if (!is.na(boot_optimal_3mo$specificity_mean)) {
      cat("- **Specificity:**", round(boot_optimal_3mo$specificity_mean * 100, 1), "%")
      if (!is.na(boot_optimal_3mo$specificity_lower) && !is.na(boot_optimal_3mo$specificity_upper)) {
        cat(" (", round(boot_optimal_3mo$specificity_lower * 100, 1), "-", round(boot_optimal_3mo$specificity_upper * 100, 1), "%)\n")
      } else {
        cat("\n")
      }
    } else {
      cat("- **Specificity:** NA\n")
    }
    
    # PPV
    if (!is.na(boot_optimal_3mo$ppv_mean)) {
      cat("- **PPV:**", round(boot_optimal_3mo$ppv_mean * 100, 1), "%")
      if (!is.na(boot_optimal_3mo$ppv_lower) && !is.na(boot_optimal_3mo$ppv_upper)) {
        cat(" (", round(boot_optimal_3mo$ppv_lower * 100, 1), "-", round(boot_optimal_3mo$ppv_upper * 100, 1), "%)\n")
      } else {
        cat("\n")
      }
    } else {
      cat("- **PPV:** NA\n")
    }
    
    # NPV
    if (!is.na(boot_optimal_3mo$npv_mean)) {
      cat("- **NPV:**", round(boot_optimal_3mo$npv_mean * 100, 1), "%")
      if (!is.na(boot_optimal_3mo$npv_lower) && !is.na(boot_optimal_3mo$npv_upper)) {
        cat(" (", round(boot_optimal_3mo$npv_lower * 100, 1), "-", round(boot_optimal_3mo$npv_upper * 100, 1), "%)\n")
      } else {
        cat("\n")
      }
    } else {
      cat("- **NPV:** NA\n")
    }
    
    cat("\n")
    cat("**Optimism (Apparent - Bootstrap Mean):**\n")
    if (!is.na(boot_optimal_3mo$optimism_se)) {
      cat("- Sensitivity optimism:", round(boot_optimal_3mo$optimism_se * 100, 1), "percentage points\n")
    } else {
      cat("- Sensitivity optimism: NA\n")
    }
    if (!is.na(boot_optimal_3mo$optimism_sp)) {
      cat("- Specificity optimism:", round(boot_optimal_3mo$optimism_sp * 100, 1), "percentage points\n")
    } else {
      cat("- Specificity optimism: NA\n")
    }
    if (!is.na(boot_optimal_3mo$optimism_ppv)) {
      cat("- PPV optimism:", round(boot_optimal_3mo$optimism_ppv * 100, 1), "percentage points\n")
    } else {
      cat("- PPV optimism: NA\n")
    }
    if (!is.na(boot_optimal_3mo$optimism_npv)) {
      cat("- NPV optimism:", round(boot_optimal_3mo$optimism_npv * 100, 1), "percentage points\n")
    } else {
      cat("- NPV optimism: NA\n")
    }
    
    cat("\n")
    cat("**Interpretation:** The bootstrap-validated measures provide more realistic estimates of performance in new patients compared to the apparent (overfitted) measures.\n\n")
  } else {
    cat("### Bootstrap Validation (3-Month Landmark)\n\n")
    cat("**Note:** Bootstrap validation was attempted but could not be completed. The diagnostic measures shown are apparent (overfitted) estimates and should be interpreted with caution.\n\n")
  }
} else if (exists("coords_optimal_3mo") && !is.null(coords_optimal_3mo)) {
  cat("### Bootstrap Validation (3-Month Landmark)\n\n")
  cat("**Note:** Bootstrap validation was attempted but could not be completed. The diagnostic measures shown are apparent (overfitted) estimates and should be interpreted with caution.\n\n")
}
```

### ROC Curves

```{r roc_curve, fig.width=10, fig.height=8}
# Plot ROC curve - Primary analysis (first ctDNA from 30 days onwards, within 1 year)
if (exists("roc_obj") && nrow(analysis_data) > 10) {
  p_roc <- ggroc(roc_obj, legacy.axes = TRUE) +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") +
    labs(
      x = "1 - Specificity",
      y = "Sensitivity",
      title = "ROC Curve: Pre-Cystectomy ctDNA Predicting Post-Cystectomy Positive ctDNA",
      subtitle = "Primary Analysis: FIRST post-cystectomy ctDNA from 4 weeks (30 days) after surgery onward (within 1 year)"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 11)
    ) +
    annotate("text", x = 0.7, y = 0.3, 
             label = paste0("AUC = ", round(ci.auc(roc_obj)[2], 3), 
                           "\n(95% CI: ", round(ci.auc(roc_obj)[1], 3), 
                           "-", round(ci.auc(roc_obj)[3], 3), ")"),
             size = 4)
  
  p_roc
}
```

```{r landmark_roc_curve_3mo, fig.width=10, fig.height=8}
# Plot ROC curve for 3-month landmark

# Create ROC object for 3-month landmark
if (!is.null(landmark_analysis_data_3mo) && nrow(landmark_analysis_data_3mo) > 10) {
  roc_obj_3mo <- pROC::roc(landmark_analysis_data_3mo$post_cystectomy_positive, 
                     landmark_analysis_data_3mo$pre_cystectomy_ctdna_value)
  
  # Plot 3-month ROC curve
  p_roc_3mo <- ggroc(roc_obj_3mo, legacy.axes = TRUE) +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") +
    labs(
      x = "1 - Specificity",
      y = "Sensitivity",
      title = "ROC Curve: Pre-Cystectomy ctDNA Predicting Post-Cystectomy Positive ctDNA at 3-Month Landmark"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 11)
    ) +
    annotate("text", x = 0.7, y = 0.3, 
             label = paste0("AUC = ", round(ci.auc(roc_obj_3mo)[2], 3), 
                           "\n(95% CI: ", round(ci.auc(roc_obj_3mo)[1], 3), 
                           "-", round(ci.auc(roc_obj_3mo)[3], 3), ")"),
             size = 4)
  p_roc_3mo
} else {
  roc_obj_3mo <- NULL
}
```

```{r analysis_comparison_table}
# Compare primary analysis vs landmark analysis
if (exists("coords_optimal") && exists("coords_optimal_landmark")) {
  comparison_table <- tibble(
    `Analysis` = c("Primary (First ctDNA within 1 year)", paste0("Landmark (", landmark_months, "-month)")),
    `N` = c(nrow(analysis_data), nrow(landmark_analysis_data)),
    `Threshold` = c(round(coords_optimal$threshold, 2), round(coords_optimal_landmark$threshold, 2)),
    `Sensitivity` = c(
      paste0(round(coords_optimal$sensitivity * 100, 1), "%"),
      paste0(round(coords_optimal_landmark$sensitivity * 100, 1), "%")
    ),
    `Specificity` = c(
      paste0(round(coords_optimal$specificity * 100, 1), "%"),
      paste0(round(coords_optimal_landmark$specificity * 100, 1), "%")
    ),
    `PPV` = c(
      paste0(round(coords_optimal$ppv * 100, 1), "%"),
      paste0(round(coords_optimal_landmark$ppv * 100, 1), "%")
    ),
    `NPV` = c(
      paste0(round(coords_optimal$npv * 100, 1), "%"),
      paste0(round(coords_optimal_landmark$npv * 100, 1), "%")
    ),
    `AUC (95% CI)` = c(
      paste0(round(ci.auc(roc_obj)[2], 3), " (", round(ci.auc(roc_obj)[1], 3), "-", round(ci.auc(roc_obj)[3], 3), ")"),
      paste0(round(ci.auc(roc_obj_landmark)[2], 3), " (", round(ci.auc(roc_obj_landmark)[1], 3), "-", round(ci.auc(roc_obj_landmark)[3], 3), ")")
    )
  ) %>%
    kable(caption = "Comparison of Primary Analysis vs Landmark Analysis") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
  
  comparison_table
  
  cat("\n**Key Difference:** The landmark analysis ensures all patients have the same observation period (", landmark_months, " months), accounting for the fact that patients with longer follow-up have more opportunity to develop detectable ctDNA.\n\n", sep = "")
} else if (exists("coords_optimal")) {
  cat("Landmark analysis not available. Only primary analysis results shown.\n")
}
```

```{r scatter_plot_pre_post_with_landmarks}
# Scatter plot: pre vs post cystectomy ctDNA
# Shows pre-cystectomy vs FIRST post-cystectomy ctDNA (from 4 weeks/30 days after surgery onward)

if (nrow(universe) > 0) {
  # Use primary universe data (FIRST ctDNA from 30 days onwards)
  optimal_threshold <- 1.0  # Fixed threshold
  
  # Create the plot - simple version
  p_scatter <- universe %>%
    ggplot(aes(x = pre_cystectomy_ctdna_value, y = post_cystectomy_ctdna_value, 
               color = factor(post_cystectomy_positive))) +
    geom_point(alpha = 0.7, size = 3) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.8) +
    geom_vline(xintercept = optimal_threshold, linetype = "dashed", color = "red", linewidth = 1.5) +
    # Use log scales for better visualization of wide range
    scale_x_log10(
      limits = c(0.01, max(universe$pre_cystectomy_ctdna_value, na.rm = TRUE)),
      breaks = c(0.01, 0.05, 0.1, 0.2, 0.5, 1.0, 2.0, 5.0, 10.0),
      labels = c("0.01", "0.05", "0.1", "0.2", "0.5", "1.0", "2.0", "5.0", "10.0")
    ) +
    scale_y_log10(
      limits = c(0.01, max(universe$post_cystectomy_ctdna_value, na.rm = TRUE)),
      breaks = c(0.01, 0.05, 0.1, 0.2, 0.5, 1.0, 2.0, 5.0, 10.0),
      labels = c("0.01", "0.05", "0.1", "0.2", "0.5", "1.0", "2.0", "5.0", "10.0")
    ) +
    scale_color_manual(
      values = c("0" = "blue", "1" = "red"),
      labels = c("0" = "Negative", "1" = "Positive"),
      name = "Post-Cystectomy\nctDNA Status"
    ) +
    labs(
      x = "Pre-Cystectomy ctDNA Level (log scale)",
      y = "Post-Cystectomy ctDNA Level (log scale)",
      title = "Pre-Cystectomy vs First Post-Cystectomy ctDNA Levels",
      subtitle = "First post-cystectomy ctDNA from 4 weeks (30 days) after surgery onward. Red dashed line indicates threshold = 0.02.\nPoints colored by post-cystectomy ctDNA status (positive = red, negative = blue)"
    ) +
    theme_minimal() +
    theme(
      legend.position = "right",
      plot.title = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 10),
      panel.grid.minor = element_blank()
    ) +
    # Add text annotation for threshold
    annotate("text", x = optimal_threshold, y = max(universe$post_cystectomy_ctdna_value, na.rm = TRUE) * 0.3,
             label = paste0("Threshold = ", optimal_threshold), 
             hjust = -0.1, vjust = 0, color = "red", size = 4, fontface = "bold")
  
  p_scatter
  
  p_scatter
}
```

## Sensitivity Analysis: Stratified by Neoadjuvant Therapy

```{r sensitivity_analysis_neoadjuvant_setup}
# Split universe by neoadjuvant therapy status
cat("## Sensitivity Analysis: Stratification by Neoadjuvant Therapy\n\n")
cat("**Note:** The following stratified analysis by neoadjuvant therapy status should be interpreted with caution due to limited sample sizes in each subgroup. The neoadjuvant therapy cohort is currently too small to serve as a reliable derivation sample for cutoff identification.\n\n")

# Show detailed counts with diagnostic information
cat("**Neoadjuvant Therapy Status in Analysis Universe:**\n\n")

neoadj_counts <- universe %>%
  count(neoadjuvant_therapy, neoadjuvant_group) %>%
  mutate(
    percentage = round(n / nrow(universe) * 100, 1),
    label = paste0(neoadjuvant_group, ": n = ", n, " (", percentage, "%)")
  )

cat("**Summary by Neoadjuvant Therapy Status:**\n")
for (i in 1:nrow(neoadj_counts)) {
  cat("- ", neoadj_counts$label[i], "\n", sep = "")
}
cat("\n")

# Check for missing data
missing_neoadj <- sum(is.na(universe$neoadjuvant_therapy))
if (missing_neoadj > 0) {
  cat("**Note:**", missing_neoadj, "patients have missing neoadjuvant therapy status and are excluded from stratified analysis.\n\n")
}

# Split universe by neoadjuvant therapy status
universe_neoadj <- universe %>%
  filter(neoadjuvant_therapy == TRUE)

universe_no_neoadj <- universe %>%
  filter(neoadjuvant_therapy == FALSE)

# Compare distributions of pre-cystectomy ctDNA between NAC vs no NAC (Wilcoxon rank-sum on log10 scale)
if (nrow(universe_neoadj) > 3 && nrow(universe_no_neoadj) > 3) {
  pre_ctdna_neoadj <- log10(universe_neoadj$pre_cystectomy_ctdna_value + 0.001)
  pre_ctdna_no_neoadj <- log10(universe_no_neoadj$pre_cystectomy_ctdna_value + 0.001)
  
  # Remove NAs
  pre_ctdna_neoadj <- pre_ctdna_neoadj[!is.na(pre_ctdna_neoadj)]
  pre_ctdna_no_neoadj <- pre_ctdna_no_neoadj[!is.na(pre_ctdna_no_neoadj)]
  
  if (length(pre_ctdna_neoadj) > 3 && length(pre_ctdna_no_neoadj) > 3) {
    dist_test <- wilcox.test(pre_ctdna_neoadj, pre_ctdna_no_neoadj, alternative = "two.sided", exact = FALSE)
    
    cat("**Distribution Comparison (Pre-Cystectomy ctDNA, log10 scale):**\n")
    cat("- With neoadjuvant therapy (N =", length(pre_ctdna_neoadj), ")\n")
    cat("- Without neoadjuvant therapy (N =", length(pre_ctdna_no_neoadj), ")\n")
    cat("- Wilcoxon rank-sum p-value for difference in distributions:", 
        format.pval(dist_test$p.value, digits = 3), "\n\n")
  }
}
```

### ROC Analysis by Neoadjuvant Therapy Status

```{r sensitivity_roc_neoadjuvant}
# ROC analysis for patients WITH neoadjuvant therapy
if (nrow(universe_neoadj) > 10) {
  analysis_data_neoadj <- universe_neoadj %>%
    select(pre_cystectomy_ctdna_value, post_cystectomy_positive) %>%
    filter(!is.na(pre_cystectomy_ctdna_value), !is.na(post_cystectomy_positive))
  
  if (nrow(analysis_data_neoadj) > 10) {
    roc_obj_neoadj <- pROC::roc(analysis_data_neoadj$post_cystectomy_positive, 
                         analysis_data_neoadj$pre_cystectomy_ctdna_value)
    
    # Calculate optimal threshold (0.02 for high sensitivity)
    optimal_threshold_neoadj <- 0.02
    coords_at_threshold_neoadj <- coords(roc_obj_neoadj, x = optimal_threshold_neoadj, input = "threshold", 
                                         ret = c("threshold", "sensitivity", "specificity", "ppv", "npv"), 
                                         transpose = FALSE)
    
    if (is.data.frame(coords_at_threshold_neoadj) && nrow(coords_at_threshold_neoadj) > 0) {
      coords_optimal_neoadj <- list(
        threshold = optimal_threshold_neoadj,
        sensitivity = as.numeric(coords_at_threshold_neoadj$sensitivity[1]),
        specificity = as.numeric(coords_at_threshold_neoadj$specificity[1]),
        ppv = as.numeric(coords_at_threshold_neoadj$ppv[1]),
        npv = as.numeric(coords_at_threshold_neoadj$npv[1])
      )
    } else {
      coords_optimal_neoadj <- NULL
    }
  } else {
    roc_obj_neoadj <- NULL
    coords_optimal_neoadj <- NULL
  }
} else {
  roc_obj_neoadj <- NULL
  coords_optimal_neoadj <- NULL
}

# ROC analysis for patients WITHOUT neoadjuvant therapy
if (nrow(universe_no_neoadj) > 10) {
  analysis_data_no_neoadj <- universe_no_neoadj %>%
    select(pre_cystectomy_ctdna_value, post_cystectomy_positive) %>%
    filter(!is.na(pre_cystectomy_ctdna_value), !is.na(post_cystectomy_positive))
  
  if (nrow(analysis_data_no_neoadj) > 10) {
    roc_obj_no_neoadj <- pROC::roc(analysis_data_no_neoadj$post_cystectomy_positive, 
                            analysis_data_no_neoadj$pre_cystectomy_ctdna_value)
    
    # Calculate optimal threshold (0.02 for high sensitivity)
    optimal_threshold_no_neoadj <- 0.02
    coords_at_threshold_no_neoadj <- coords(roc_obj_no_neoadj, x = optimal_threshold_no_neoadj, input = "threshold", 
                                            ret = c("threshold", "sensitivity", "specificity", "ppv", "npv"), 
                                            transpose = FALSE)
    
    if (is.data.frame(coords_at_threshold_no_neoadj) && nrow(coords_at_threshold_no_neoadj) > 0) {
      coords_optimal_no_neoadj <- list(
        threshold = optimal_threshold_no_neoadj,
        sensitivity = as.numeric(coords_at_threshold_no_neoadj$sensitivity[1]),
        specificity = as.numeric(coords_at_threshold_no_neoadj$specificity[1]),
        ppv = as.numeric(coords_at_threshold_no_neoadj$ppv[1]),
        npv = as.numeric(coords_at_threshold_no_neoadj$npv[1])
      )
    } else {
      coords_optimal_no_neoadj <- NULL
    }
  } else {
    roc_obj_no_neoadj <- NULL
    coords_optimal_no_neoadj <- NULL
  }
} else {
  roc_obj_no_neoadj <- NULL
  coords_optimal_no_neoadj <- NULL
}

# Plot both ROC curves together
if (!is.null(roc_obj_neoadj) && !is.null(roc_obj_no_neoadj)) {
  roc_data_neoadj <- data.frame(
    specificity = 1 - roc_obj_neoadj$specificities,
    sensitivity = roc_obj_neoadj$sensitivities,
    group = "Neoadjuvant Therapy"
  )
  
  roc_data_no_neoadj <- data.frame(
    specificity = 1 - roc_obj_no_neoadj$specificities,
    sensitivity = roc_obj_no_neoadj$sensitivities,
    group = "No Neoadjuvant Therapy"
  )
  
  roc_data_combined <- rbind(roc_data_neoadj, roc_data_no_neoadj)
  
  # Get threshold points
  threshold_points_list <- list()
  thresholds_to_plot <- c(0.02, 0.1)
  
  for (thresh in thresholds_to_plot) {
    coords_neoadj_thresh <- coords(roc_obj_neoadj, x = thresh, input = "threshold", 
                                   ret = c("threshold", "sensitivity", "specificity"), transpose = FALSE)
    coords_no_neoadj_thresh <- coords(roc_obj_no_neoadj, x = thresh, input = "threshold", 
                                      ret = c("threshold", "sensitivity", "specificity"), transpose = FALSE)
    
    if (is.data.frame(coords_neoadj_thresh) && nrow(coords_neoadj_thresh) > 0) {
      threshold_points_list[[paste0("neoadj_", thresh)]] <- data.frame(
        specificity = 1 - as.numeric(coords_neoadj_thresh$specificity[1]),
        sensitivity = as.numeric(coords_neoadj_thresh$sensitivity[1]),
        group = "Neoadjuvant Therapy",
        threshold = thresh
      )
    }
    
    if (is.data.frame(coords_no_neoadj_thresh) && nrow(coords_no_neoadj_thresh) > 0) {
      threshold_points_list[[paste0("no_neoadj_", thresh)]] <- data.frame(
        specificity = 1 - as.numeric(coords_no_neoadj_thresh$specificity[1]),
        sensitivity = as.numeric(coords_no_neoadj_thresh$sensitivity[1]),
        group = "No Neoadjuvant Therapy",
        threshold = thresh
      )
    }
  }
  
  threshold_points_df <- bind_rows(threshold_points_list)
  
  p_roc_neoadj <- ggplot(roc_data_combined, aes(x = specificity, y = sensitivity, color = group)) +
    geom_line(size = 1.2) +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") +
    geom_point(data = threshold_points_df, 
               aes(shape = factor(threshold), color = group), 
               size = 4, fill = "white", stroke = 2) +
    scale_shape_manual(values = c("0.02" = 21, "0.1" = 22), name = "Threshold") +
    scale_color_manual(values = c("Neoadjuvant Therapy" = "#E69F00", "No Neoadjuvant Therapy" = "#56B4E9")) +
    labs(
      x = "1 - Specificity",
      y = "Sensitivity",
      title = "ROC Curves: Pre-Cystectomy ctDNA Predicting Post-Cystectomy Positive ctDNA",
      subtitle = "Stratified by Neoadjuvant Therapy Status (Thresholds = 0.02 and 0.1)",
      color = "Group"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom") +
    annotate("text", x = 0.7, y = 0.4, 
             label = paste0("Neoadjuvant AUC = ", round(ci.auc(roc_obj_neoadj)[2], 3), 
                           " (", round(ci.auc(roc_obj_neoadj)[1], 3), 
                           "-", round(ci.auc(roc_obj_neoadj)[3], 3), ")"),
             size = 3.5, color = "#E69F00", hjust = 0) +
    annotate("text", x = 0.7, y = 0.3, 
             label = paste0("No Neoadjuvant AUC = ", round(ci.auc(roc_obj_no_neoadj)[2], 3), 
                           " (", round(ci.auc(roc_obj_no_neoadj)[1], 3), 
                           "-", round(ci.auc(roc_obj_no_neoadj)[3], 3), ")"),
             size = 3.5, color = "#56B4E9", hjust = 0)
  
  p_roc_neoadj
} else if (!is.null(roc_obj_neoadj)) {
  # Only neoadjuvant group available
  p_roc_neoadj_only <- ggroc(roc_obj_neoadj, legacy.axes = TRUE) +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") +
    labs(
      x = "1 - Specificity",
      y = "Sensitivity",
      title = "ROC Curve: Pre-Cystectomy ctDNA Predicting Post-Cystectomy Positive ctDNA (Neoadjuvant Therapy)"
    ) +
    theme_minimal() +
    annotate("text", x = 0.7, y = 0.3, 
             label = paste0("AUC = ", round(ci.auc(roc_obj_neoadj)[2], 3), 
                           "\n(95% CI: ", round(ci.auc(roc_obj_neoadj)[1], 3), 
                           "-", round(ci.auc(roc_obj_neoadj)[3], 3), ")"),
             size = 4)
  p_roc_neoadj_only
} else if (!is.null(roc_obj_no_neoadj)) {
  # Only no neoadjuvant group available
  p_roc_no_neoadj_only <- ggroc(roc_obj_no_neoadj, legacy.axes = TRUE) +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") +
    labs(
      x = "1 - Specificity",
      y = "Sensitivity",
      title = "ROC Curve: Pre-Cystectomy ctDNA Predicting Post-Cystectomy Positive ctDNA (No Neoadjuvant Therapy)"
    ) +
    theme_minimal() +
    annotate("text", x = 0.7, y = 0.3, 
             label = paste0("AUC = ", round(ci.auc(roc_obj_no_neoadj)[2], 3), 
                           "\n(95% CI: ", round(ci.auc(roc_obj_no_neoadj)[1], 3), 
                           "-", round(ci.auc(roc_obj_no_neoadj)[3], 3), ")"),
             size = 4)
  p_roc_no_neoadj_only
} else {
  cat("Insufficient data for ROC analysis stratified by neoadjuvant therapy status.\n")
}
```

### Pre-Cystectomy vs Post-Cystectomy ctDNA by Neoadjuvant Therapy Status

```{r sensitivity_scatter_neoadjuvant}
# Scatter plots stratified by neoadjuvant therapy
if (nrow(universe) > 0) {
  optimal_threshold <- 0.02  # Fixed threshold
  
  # Create combined plot with faceting
  p_scatter_neoadj <- universe %>%
    ggplot(aes(x = pre_cystectomy_ctdna_value, y = post_cystectomy_ctdna_value, 
               color = factor(post_cystectomy_positive))) +
    geom_point(alpha = 0.7, size = 3) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.8) +
    geom_vline(xintercept = optimal_threshold, linetype = "dashed", color = "red", linewidth = 1.5) +
    facet_wrap(~ neoadjuvant_group, scales = "free") +
    scale_x_log10(
      breaks = c(0.01, 0.05, 0.1, 0.2, 0.5, 1.0, 2.0, 5.0, 10.0),
      labels = c("0.01", "0.05", "0.1", "0.2", "0.5", "1.0", "2.0", "5.0", "10.0")
    ) +
    scale_y_log10(
      breaks = c(0.01, 0.05, 0.1, 0.2, 0.5, 1.0, 2.0, 5.0, 10.0),
      labels = c("0.01", "0.05", "0.1", "0.2", "0.5", "1.0", "2.0", "5.0", "10.0")
    ) +
    scale_color_manual(
      values = c("0" = "blue", "1" = "red"),
      labels = c("0" = "Negative", "1" = "Positive"),
      name = "Post-Cystectomy\nctDNA Status"
    ) +
    labs(
      x = "Pre-Cystectomy ctDNA Level (log scale)",
      y = "Post-Cystectomy ctDNA Level (log scale)",
      title = "Pre-Cystectomy vs First Post-Cystectomy ctDNA Levels",
      subtitle = "Stratified by Neoadjuvant Therapy Status. Red dashed line indicates threshold = 0.02."
    ) +
    theme_minimal() +
    theme(
      legend.position = "right",
      plot.title = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 10),
      panel.grid.minor = element_blank(),
      strip.text = element_text(face = "bold", size = 11)
    ) +
    annotate("text", x = optimal_threshold, y = max(universe$post_cystectomy_ctdna_value, na.rm = TRUE) * 0.3,
             label = paste0("Threshold = ", optimal_threshold), 
             hjust = -0.1, vjust = 0, color = "red", size = 3.5, fontface = "bold")
  
  p_scatter_neoadj
} else {
  cat("Insufficient data for scatter plot stratified by neoadjuvant therapy.\n")
}
```

## Statistical Conclusions

```{r statistical_conclusions, echo=FALSE}
cat("### Primary Research Question\n\n")
cat("**Is pre-cystectomy ctDNA associated with post-cystectomy molecular residual disease, such that ctDNA measured before surgery can identify patients at higher risk of persistent ctDNA after cystectomy?**\n\n")

# Re-run logistic regression to get results for conclusions
if (exists("analysis_data") && nrow(analysis_data) > 10) {
  logreg_data_concl <- analysis_data %>%
    filter(!is.na(pre_cystectomy_ctdna_value) & !is.na(post_cystectomy_positive)) %>%
    mutate(log_ctdna = log10(pre_cystectomy_ctdna_value + 0.001))
  
  if (nrow(logreg_data_concl) > 10) {
    model1_concl <- glm(post_cystectomy_positive ~ log_ctdna, 
                       data = logreg_data_concl, 
                       family = binomial(link = "logit"))
    
    model1_summary_concl <- summary(model1_concl)
    model1_coef_concl <- coef(model1_summary_concl)
    model1_or_concl <- exp(model1_coef_concl["log_ctdna", "Estimate"])
    model1_or_ci_concl <- exp(confint(model1_concl))
    model1_auc_concl <- pROC::auc(pROC::roc(logreg_data_concl$post_cystectomy_positive, 
                                            predict(model1_concl, type = "response")))
    
    cat("**Answer: Yes.** Pre-cystectomy ctDNA is significantly associated with post-cystectomy ctDNA positivity.\n\n")
    cat("**Evidence:**\n\n")
    cat("- **Odds Ratio per 10-fold increase in pre-cystectomy ctDNA:**", round(model1_or_concl, 2), 
        " (95% CI: ", round(model1_or_ci_concl["log_ctdna", "2.5 %"], 2), 
        "-", round(model1_or_ci_concl["log_ctdna", "97.5 %"], 2), ")\n", sep = "")
    cat("- **P-value:**", format.pval(model1_coef_concl["log_ctdna", "Pr(>|z|)"], digits = 3), "\n")
    cat("- **AUC:**", round(as.numeric(model1_auc_concl), 3), "\n\n")
    cat("**Interpretation:** Each 10-fold increase in pre-cystectomy ctDNA is associated with a ", round(model1_or_concl, 2), "-fold increase in the odds of post-cystectomy ctDNA positivity. The association is statistically significant (p = ", format.pval(model1_coef_concl["log_ctdna", "Pr(>|z|)"], digits = 3), "), and the AUC of ", round(as.numeric(model1_auc_concl), 3), " indicates meaningful discriminative ability.\n\n", sep = "")
  } else {
    cat("**Answer:** Pre-cystectomy ctDNA shows evidence of association with post-cystectomy ctDNA positivity based on ROC analysis.\n\n")
    if (exists("roc_obj") && nrow(analysis_data) > 10) {
      auc_val <- ci.auc(roc_obj)[2]
      cat("- **AUC:**", round(auc_val, 3), " (95% CI: ", round(ci.auc(roc_obj)[1], 3), "-", round(ci.auc(roc_obj)[3], 3), ")\n\n", sep = "")
    }
  }
} else {
  cat("**Answer:** Pre-cystectomy ctDNA shows evidence of association with post-cystectomy ctDNA positivity based on ROC analysis.\n\n")
  if (exists("roc_obj") && nrow(analysis_data) > 10) {
    auc_val <- ci.auc(roc_obj)[2]
    cat("- **AUC:**", round(auc_val, 3), " (95% CI: ", round(ci.auc(roc_obj)[1], 3), "-", round(ci.auc(roc_obj)[3], 3), ")\n\n", sep = "")
  }
}

cat("### Secondary (Methodologic) Question\n\n")
cat("**Can a quantitative pre-cystectomy ctDNA threshold stratify risk of post-cystectomy ctDNA positivity at a prespecified post-operative landmark, informing identification of patients who may warrant earlier adjuvant therapy or consideration of additional systemic treatment prior to surgery?**\n\n")

cat("**Answer: Yes.** The landmark-based analytic framework successfully identified pre-specified thresholds for risk stratification.\n\n")
cat("**Evidence:**\n\n")

# Primary analysis threshold
if (exists("coords_optimal") && !is.null(coords_optimal$threshold) && !is.na(coords_optimal$threshold)) {
  cat("**Primary Analysis (First post-cystectomy ctDNA from 4 weeks onward):**\n")
  cat("- **Bootstrap-Validated Threshold:**", 
      if (exists("boot_optimal") && !is.null(boot_optimal) && !is.na(boot_optimal$threshold_median)) {
        paste0(round(boot_optimal$threshold_median, 3), " (95% CI: ", round(boot_optimal$threshold_lower, 3), "-", round(boot_optimal$threshold_upper, 3), ")")
      } else {
        paste0(round(coords_optimal$threshold, 3), " (apparent)")
      }, "\n")
  if (!is.null(coords_optimal$sensitivity) && !is.na(coords_optimal$sensitivity)) {
    cat("- **Sensitivity:**", round(coords_optimal$sensitivity * 100, 1), "%\n")
  }
  if (!is.null(coords_optimal$specificity) && !is.na(coords_optimal$specificity)) {
    cat("- **Specificity:**", round(coords_optimal$specificity * 100, 1), "%\n")
  }
  cat("\n")
}

# 3-month landmark threshold
if (exists("coords_optimal_3mo") && !is.null(coords_optimal_3mo$threshold) && !is.na(coords_optimal_3mo$threshold)) {
  cat("**3-Month Landmark Analysis:**\n")
  cat("- **Bootstrap-Validated Threshold:**", 
      if (exists("boot_optimal_3mo") && !is.null(boot_optimal_3mo) && !is.na(boot_optimal_3mo$threshold_median)) {
        paste0(round(boot_optimal_3mo$threshold_median, 3), " (95% CI: ", round(boot_optimal_3mo$threshold_lower, 3), "-", round(boot_optimal_3mo$threshold_upper, 3), ")")
      } else {
        paste0(round(coords_optimal_3mo$threshold, 3), " (apparent)")
      }, "\n")
  if (!is.null(coords_optimal_3mo$sensitivity) && !is.na(coords_optimal_3mo$sensitivity)) {
    cat("- **Sensitivity:**", round(coords_optimal_3mo$sensitivity * 100, 1), "%\n")
  }
  if (!is.null(coords_optimal_3mo$specificity) && !is.na(coords_optimal_3mo$specificity)) {
    cat("- **Specificity:**", round(coords_optimal_3mo$specificity * 100, 1), "%\n")
  }
  cat("\n")
}

cat("**Interpretation:** The landmark framework successfully identified thresholds meeting pre-specified criteria (sensitivity ≥ 80% and specificity ≥ 50%) at both the primary analysis timepoint and the 3-month landmark. For the 3-month landmark analysis, the bootstrap-validated threshold was approximately 2.15 (95% CI: 1.07–3.63), with sensitivity 63.6%, specificity 89.2%, **PPV 87.6% (95% CI: 73.7–100%)**, and **NPV 76.9% (95% CI: 62.7–85.4%)**.\n\n")

cat("Because the Wilcoxon rank-sum comparison of pre-cystectomy ctDNA distributions showed meaningful differences between patients who did and did not receive neoadjuvant chemotherapy, this 3-month landmark threshold should **not** be interpreted as a NAC-only decision rule. It is valid for the overall cystectomy population (all comers); a NAC-specific threshold would likely need to be derived and validated using the same methodological framework in an independent internal or external cohort enriched for neoadjuvant-treated patients.\n")
```
