---
title: "Metastatic Disease - ctDNA Level at Diagnosis and On-treatment"
date: "`r Sys.Date()`"
output: html_document
---

## Research Question

**What is the association between circulating tumor DNA (ctDNA) level before starting a new line of therapy and duration of complete response?**

This analysis examines whether ctDNA levels measured before starting each new line of metastatic therapy are prognostic for duration of complete response in patients with metastatic urothelial carcinoma. The analysis includes all lines of therapy (1L, 2L, 3L+, not limited to first-line), with each line of therapy as a separate observation and time 0 defined as the start of each new line of therapy.

**Note on Endpoint Definition:** We analyze "Duration of Complete Response" (DCR), which represents the time from therapy initiation until loss of complete response or death. **Events** occur when: (1) the patient dies, OR (2) the patient is alive but has obvious disease present on most recent imaging (i.e., loss of complete response, similar to RECIST progression from CR). **Patients are censored** when they are alive and have no obvious disease on most recent imaging (i.e., maintaining complete response, similar to RECIST CR). This endpoint captures both survival and maintenance of complete response, which is clinically meaningful in the metastatic setting. Note: Patients with stable disease or partial response (but with disease still present on imaging) are considered events, as they have lost or never achieved complete response. **Important:** Patients who died do NOT require imaging data for inclusion (death is an event). Only alive patients require imaging data to determine complete response status.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "")
knitr::opts_knit$set(root.dir = "~/R")
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Load shared setup
source("ctDNA_Research_Questions_Shared_Setup.R")

# Load required libraries
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(stringr)
  library(readxl)
  library(janitor)
  library(lubridate)
  library(ggplot2)
  library(survival)
  library(survminer)
  library(knitr)
  library(kableExtra)
})
```

```{r load_data, include=FALSE}
# Load minimal data
data_list <- load_minimal_data(include_survival = TRUE)
dat_combined <- data_list$dat_combined
ctdna_long <- data_list$ctdna_long
survival_data <- data_list$survival_data
```

```{r load_bajorin_data, include=FALSE}
# ====================================================================================
# Load Bajorin Risk Factors Data (KPS and Visceral Metastases)
# ====================================================================================
# Load data from ctDNA_Bajorin_Lexi.xlsx
# KPS: >=80% = 1, <80% = 0
# Visceral Mets: Yes/No

bajorin_data <- tryCatch({
  # Read Excel file - skip first row which is header
  bajorin_raw <- read_excel("ctDNA_Bajorin_Lexi.xlsx", sheet = 1, skip = 1)
  
  # Based on structure: Screener, Patient_ID, MRN, Therapy_Line, Baseline_ctDNA_Date, KPS, Visceral_Mets, Column8
  # Use column positions
  if (nrow(bajorin_raw) > 0 && ncol(bajorin_raw) >= 7) {
    # Extract columns by position
    bajorin_clean <- tibble(
      patient_id_raw = as.character(bajorin_raw[[2]]),  # Column 2: Patient ID
      therapy_line_raw = as.character(bajorin_raw[[4]]),  # Column 4: Therapy Line
      kps_raw = as.character(bajorin_raw[[6]]),  # Column 6: KPS
      visceral_mets_raw = as.character(bajorin_raw[[7]])  # Column 7: Visceral Mets
    ) %>%
      # Remove header rows (rows that contain "Patient", "ID", "Screener", etc.)
      filter(!str_detect(tolower(patient_id_raw), "patient|id|screener|mrn|therapy")) %>%
      filter(!is.na(patient_id_raw), patient_id_raw != "", patient_id_raw != "NA") %>%
      # Convert patient_id - may be numeric or character
      mutate(
        patient_id = as.character(patient_id_raw),
        # Convert therapy line: "1L" -> 1, "2L" -> 2, "3L" -> 3, "4L+" -> 4
        therapy_line = case_when(
          str_detect(toupper(therapy_line_raw), "^1L$") ~ 1,
          str_detect(toupper(therapy_line_raw), "^2L$") ~ 2,
          str_detect(toupper(therapy_line_raw), "^3L$") ~ 3,
          str_detect(toupper(therapy_line_raw), "^4L\\+") ~ 4,
          str_detect(therapy_line_raw, "^\\d+") ~ as.numeric(str_extract(therapy_line_raw, "^\\d+")),
          TRUE ~ NA_real_
        ),
        # KPS: 1 = >=80%, 0 = <80%
        kps_binary = case_when(
          kps_raw == "1" | kps_raw == 1 ~ 1,
          kps_raw == "0" | kps_raw == 0 ~ 0,
          TRUE ~ NA_real_
        ),
        # Visceral metastases: Yes = 1, No = 0
        visceral_mets_binary = case_when(
          str_detect(tolower(visceral_mets_raw), "^yes$|^y$") ~ 1,
          str_detect(tolower(visceral_mets_raw), "^no$|^n$") ~ 0,
          TRUE ~ NA_real_
        )
      ) %>%
      filter(!is.na(patient_id), !is.na(therapy_line), patient_id != "") %>%
      select(patient_id, therapy_line, kps_binary, visceral_mets_binary)
    
    cat("Loaded Bajorin risk factors data:\n")
    cat("  Total observations:", nrow(bajorin_clean), "\n")
    cat("  Unique patients:", length(unique(bajorin_clean$patient_id)), "\n")
    cat("  KPS available:", sum(!is.na(bajorin_clean$kps_binary)), "\n")
    cat("  Visceral mets available:", sum(!is.na(bajorin_clean$visceral_mets_binary)), "\n")
    
    bajorin_clean
  } else {
    cat("Warning: Bajorin data file is empty or has insufficient columns.\n")
    tibble(patient_id = character(0), therapy_line = numeric(0), 
           kps_binary = numeric(0), visceral_mets_binary = numeric(0))
  }
}, error = function(e) {
  cat("Warning: Could not load Bajorin data file:", e$message, "\n")
  tibble(patient_id = character(0), therapy_line = numeric(0), 
         kps_binary = numeric(0), visceral_mets_binary = numeric(0))
})

# Check patient_id format matching
if (nrow(bajorin_data) > 0 && exists("universe") && nrow(universe) > 0) {
  sample_outcome_ids <- head(unique(universe$patient_id), 5)
  sample_bajorin_ids <- head(unique(bajorin_data$patient_id), 5)
  cat("\nSample patient IDs from outcomes:", paste(sample_outcome_ids, collapse = ", "), "\n")
  cat("Sample patient IDs from Bajorin data:", paste(sample_bajorin_ids, collapse = ", "), "\n")
  
  # Check if formats match
  if (any(sample_bajorin_ids %in% sample_outcome_ids)) {
    cat("Patient ID formats appear to match.\n")
  } else {
    cat("Note: Patient ID formats may differ. Will attempt to match.\n")
  }
}
```

```{r define_universe, include=FALSE}
# ==============================
# Universe: Patients at initial diagnosis of metastatic disease
# Either: (1) without prior treatment, OR (2) >6 months from anti-cancer directed therapy (radiation or systemic)
# ==============================

# Note: We no longer use "ctDNA at diagnosis" as a single anchor point
# Instead, we find ctDNA before each line of therapy (see universe definition below)

# Extract all therapy start dates and settings (systemic)
therapy_start_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("systemic.*therapy.*\\d+.*start.*date", ignore_case = TRUE))]
therapy_setting_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("systemic.*therapy.*\\d+.*treatment.*setting", ignore_case = TRUE))]

# Get all systemic therapy dates and identify palliative therapy
all_systemic_therapy_dates <- tibble(patient_id = character(0), therapy_date = as.Date(character(0)), is_palliative = logical(0))
if (length(therapy_start_cols) > 0) {
  # Get therapy dates
  therapy_dates_long <- dat_combined %>%
    select(patient_id, any_of(therapy_start_cols)) %>%
    pivot_longer(-patient_id, names_to = "therapy_var", values_to = "start_date_raw") %>%
    mutate(
      start_date = as_date_safe(start_date_raw),
      therapy_num = str_extract(therapy_var, "\\d+")
    ) %>%
    filter(!is.na(start_date)) %>%
    select(patient_id, therapy_num, therapy_date = start_date)
  
  # Get therapy settings if available
  if (length(therapy_setting_cols) > 0) {
    therapy_settings_long <- dat_combined %>%
      select(patient_id, any_of(therapy_setting_cols)) %>%
      pivot_longer(-patient_id, names_to = "setting_var", values_to = "setting_raw") %>%
      mutate(
        setting = as.character(setting_raw),
        therapy_num = str_extract(setting_var, "\\d+")
      ) %>%
      select(patient_id, therapy_num, setting)
    
    # Join and identify palliative therapy
    all_systemic_therapy_dates <- therapy_dates_long %>%
      left_join(therapy_settings_long, by = c("patient_id", "therapy_num")) %>%
      mutate(
        setting_lower = tolower(setting),
        is_palliative = str_detect(setting_lower, "palliative|metastatic|1l.*metastatic|2l.*metastatic")
      ) %>%
      select(patient_id, therapy_date, is_palliative)
  } else {
    # If no setting columns, assume all are potentially palliative for metastatic patients
    all_systemic_therapy_dates <- therapy_dates_long %>%
      mutate(is_palliative = TRUE) %>%
      select(patient_id, therapy_date, is_palliative)
  }
}

# Get ALL palliative therapy lines with their line numbers
# We need to identify which therapy number each palliative therapy corresponds to
all_palliative_therapy_lines <- tibble(patient_id = character(0), therapy_line = integer(0), therapy_start_date = as.Date(character(0)))

if (length(therapy_start_cols) > 0 && length(therapy_setting_cols) > 0) {
  # Get therapy dates with line numbers
  therapy_dates_with_lines <- dat_combined %>%
    select(patient_id, any_of(therapy_start_cols)) %>%
    pivot_longer(-patient_id, names_to = "therapy_var", values_to = "start_date_raw") %>%
    mutate(
      start_date = as_date_safe(start_date_raw),
      therapy_line = as.integer(str_extract(therapy_var, "\\d+"))
    ) %>%
    filter(!is.na(start_date), !is.na(therapy_line)) %>%
    select(patient_id, therapy_line, therapy_start_date = start_date)
  
  # Get therapy settings with line numbers
  therapy_settings_with_lines <- dat_combined %>%
    select(patient_id, any_of(therapy_setting_cols)) %>%
    pivot_longer(-patient_id, names_to = "setting_var", values_to = "setting_raw") %>%
    mutate(
      setting = as.character(setting_raw),
      therapy_line = as.integer(str_extract(setting_var, "\\d+"))
    ) %>%
    filter(!is.na(therapy_line)) %>%
    select(patient_id, therapy_line, setting)
  
  # Join and identify palliative therapy lines
  all_palliative_therapy_lines <- therapy_dates_with_lines %>%
    left_join(therapy_settings_with_lines, by = c("patient_id", "therapy_line")) %>%
    mutate(
      setting_lower = tolower(setting),
      is_palliative = str_detect(setting_lower, "palliative|metastatic|1l.*metastatic|2l.*metastatic|3l.*metastatic")
    ) %>%
    filter(is_palliative) %>%
    select(patient_id, therapy_line, therapy_start_date) %>%
    arrange(patient_id, therapy_line)
}

# Also get first palliative therapy for metastatic status check
first_palliative_therapy <- all_palliative_therapy_lines %>%
  group_by(patient_id) %>%
  summarise(first_palliative_start = min(therapy_start_date, na.rm = TRUE), .groups = "drop")

# Extract radiation dates
radiation_date_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("radiation.*date|radiotherapy.*date|rt.*date", ignore_case = TRUE))]
therapy_regimen_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("systemic.*therapy.*\\d+.*regimen|treatment.*regimen", ignore_case = TRUE))]

all_radiation_dates <- tibble(patient_id = character(0), radiation_date = as.Date(character(0)))
if (length(radiation_date_cols) > 0) {
  all_radiation_dates <- dat_combined %>%
    select(patient_id, all_of(radiation_date_cols[1])) %>%
    rename(radiation_date_raw = all_of(radiation_date_cols[1])) %>%
    mutate(radiation_date = as_date_safe(radiation_date_raw)) %>%
    filter(!is.na(radiation_date)) %>%
    select(patient_id, radiation_date)
} else if (length(therapy_regimen_cols) > 0 && length(therapy_start_cols) > 0) {
  # Check therapy regimens for radiation
  all_radiation_dates <- dat_combined %>%
    select(patient_id, all_of(therapy_regimen_cols), all_of(therapy_start_cols)) %>%
    pivot_longer(-patient_id, names_to = "var", values_to = "value") %>%
    mutate(
      var_type = if_else(str_detect(var, "regimen"), "regimen", "start_date"),
      therapy_num = str_extract(var, "\\d+")
    ) %>%
    pivot_wider(id_cols = c(patient_id, therapy_num), names_from = var_type, values_from = value) %>%
    mutate(
      regimen_lower = tolower(as.character(regimen)),
      start_date = as_date_safe(start_date),
      is_radiation = str_detect(regimen_lower, "radiation|radiotherapy|rt|sbrt|stereotactic")
    ) %>%
    filter(is_radiation, !is.na(start_date)) %>%
    group_by(patient_id) %>%
    summarise(radiation_date = min(start_date, na.rm = TRUE), .groups = "drop")
}

# Combine all anti-cancer therapy dates (systemic or radiation)
all_anti_cancer_therapy <- bind_rows(
  all_systemic_therapy_dates %>% rename(therapy_date = therapy_date),
  all_radiation_dates %>% rename(therapy_date = radiation_date)
) %>%
  group_by(patient_id) %>%
  summarise(earliest_therapy_date = min(therapy_date, na.rm = TRUE), .groups = "drop")

# Identify patients with confirmed metastatic disease
# Criteria (any of the following):
# 1. Clinical disease state at first ctDNA contains "metastatic" or "advanced" OR
# 2. Patient started palliative therapy (palliative therapy is for metastatic disease) OR
# 3. Pathology M stage = M1 OR
# 4. metastatic_disease_ever_diagnosed_yes_no = "Yes"
clinical_state_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("clinical_disease_state.*metastatic|clinical_disease_state.*first.*ct", ignore_case = TRUE))]
metastatic_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("metastatic_disease_ever_diagnosed|metastatic.*yes.*no", ignore_case = TRUE))]
m_stage_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("pathology_m_stage", ignore_case = TRUE))]

# Get metastatic status based on multiple criteria
metastatic_status <- dat_combined %>%
  select(patient_id, 
         any_of(clinical_state_cols),
         any_of(metastatic_cols),
         any_of(m_stage_cols)) %>%
  mutate(
    # Check clinical disease state at first ctDNA
    metastatic_at_ctdna = if (length(clinical_state_cols) > 0) {
      str_detect(tolower(as.character(.data[[clinical_state_cols[1]]])), "metastatic|advanced")
    } else {
      NA
    },
    # Check metastatic_disease_ever_diagnosed_yes_no
    metastatic_ever_diagnosed = if (length(metastatic_cols) > 0) {
      case_when(
        str_detect(tolower(as.character(.data[[metastatic_cols[1]]])), "yes|y|1") ~ TRUE,
        str_detect(tolower(as.character(.data[[metastatic_cols[1]]])), "no|n|0") ~ FALSE,
        TRUE ~ NA
      )
    } else {
      NA
    },
    # Check M stage (M1 = metastatic)
    m_stage_metastatic = if (length(m_stage_cols) > 0) {
      str_detect(toupper(as.character(.data[[m_stage_cols[1]]])), "M1")
    } else {
      NA
    }
  ) %>%
  select(patient_id, metastatic_at_ctdna, metastatic_ever_diagnosed, m_stage_metastatic)

# Also check if patient has palliative therapy (indicates metastatic disease)
# We already have first_palliative_therapy from earlier, so we can use that
# Patient has confirmed metastatic disease if ANY of the criteria are met:
# - Clinical state at first ctDNA indicates metastatic/advanced, OR
# - Patient has palliative therapy, OR
# - Pathology M stage = M1, OR
# - metastatic_disease_ever_diagnosed_yes_no = "Yes"
metastatic_status <- metastatic_status %>%
  left_join(first_palliative_therapy, by = "patient_id") %>%
  mutate(
    has_palliative_therapy = !is.na(first_palliative_start),
    # Patient has confirmed metastatic disease if any criterion is met
    has_metastatic_disease = case_when(
      !is.na(metastatic_at_ctdna) & metastatic_at_ctdna == TRUE ~ TRUE,
      has_palliative_therapy == TRUE ~ TRUE,
      !is.na(m_stage_metastatic) & m_stage_metastatic == TRUE ~ TRUE,
      !is.na(metastatic_ever_diagnosed) & metastatic_ever_diagnosed == TRUE ~ TRUE,
      # If we have negative indicators for all criteria, mark as FALSE
      !is.na(metastatic_at_ctdna) & metastatic_at_ctdna == FALSE & 
      has_palliative_therapy == FALSE & 
      !is.na(m_stage_metastatic) & m_stage_metastatic == FALSE & 
      !is.na(metastatic_ever_diagnosed) & metastatic_ever_diagnosed == FALSE ~ FALSE,
      TRUE ~ NA
    )
  ) %>%
  select(patient_id, has_metastatic_disease, metastatic_at_ctdna, has_palliative_therapy, m_stage_metastatic, metastatic_ever_diagnosed)

# Define universe: ALL lines of metastatic therapy
# For each line of palliative therapy, find the most recent ctDNA before that line starts (within 21 days)
# Each patient can contribute multiple observations (one per line of therapy)

# First, identify patients with confirmed metastatic disease
patients_with_metastatic <- metastatic_status %>%
  filter(has_metastatic_disease == TRUE) %>%
  select(patient_id)

# For each palliative therapy line, find the most recent ctDNA before that line starts
universe_by_line <- all_palliative_therapy_lines %>%
  inner_join(patients_with_metastatic, by = "patient_id") %>%
  # For each therapy line, find ctDNA before that line starts (within 21 days)
  left_join(ctdna_long, by = "patient_id") %>%
  filter(
    !is.na(ctdna_date),
    ctdna_date <= therapy_start_date,  # ctDNA must be before or at therapy start
    ctdna_date >= (therapy_start_date - days(21))  # ctDNA within 21 days before therapy start
  ) %>%
  # For each patient-line combination, take the most recent ctDNA before that line
  arrange(patient_id, therapy_line, desc(ctdna_date)) %>%
  group_by(patient_id, therapy_line, therapy_start_date) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  select(patient_id, therapy_line, therapy_start_date, 
         ctdna_date, ctdna_value) %>%
  rename(
    ctdna_before_line_date = ctdna_date,
    ctdna_before_line_value = ctdna_value
  )

# Create universe dataset with one row per patient-line combination
universe <- universe_by_line %>%
  mutate(
    # Time 0 is the therapy start date (not ctDNA date)
    time_zero_date = therapy_start_date,
    # Calculate days from ctDNA to therapy start
    days_from_ctdna_to_therapy = as.numeric(therapy_start_date - ctdna_before_line_date)
  ) %>%
  select(patient_id, therapy_line, time_zero_date, 
         ctdna_before_line_date, ctdna_before_line_value, 
         days_from_ctdna_to_therapy)

# Diagnostic output
cat("Universe size: n =", nrow(universe), " patient-line observations\n")
cat("Unique patients: n =", length(unique(universe$patient_id)), "\n")
cat("Population studied: Patients with metastatic disease starting a new line of therapy\n")
cat("  - Includes all lines of palliative/metastatic therapy (not just first-line)\n")
cat("  - ctDNA measured within 21 days before starting each new line of therapy\n")
cat("  - Time 0 = start of each new line of therapy\n")
cat("  - With confirmed metastatic disease diagnosis\n")

# OPTIMIZATION: Cache universe for faster subsequent runs
if (!dir.exists("cache")) dir.create("cache")
saveRDS(universe, "cache/universe.rds")
```

```{r population_description}
# Display population description at top
# Note: N shown here is universe size; final analysis N (after survival data filtering) is shown in descriptive table below
cat("## Population Studied\n\n")
cat("**Universe:** Patients with metastatic disease starting a new line of therapy\n\n")
cat("**N =", nrow(universe), " patient-line observations (", length(unique(universe$patient_id)), " unique patients)\n\n")
cat("**Note:** Final analysis N (after excluding observations due to missing survival data) is shown in the exclusion and descriptive tables below.\n\n")
cat("**Inclusion Criteria:**\n")
cat("- Confirmed metastatic disease\n")
cat("- Starting a new line of palliative/metastatic therapy (includes 1L, 2L, 3L+, not limited to first-line)\n")
cat("- Had ctDNA level measured within 21 days before starting that line of therapy\n")
cat("- Time 0 = start of each new line of therapy\n\n")
cat("**Note:** Each patient can contribute multiple observations (one per line of therapy). Line of therapy is included as a covariate in the analysis.\n\n")
```

```{r patient_ids_table}
# Display patient IDs with baseline ctDNA dates and therapy lines
# Show all patient-line combinations with baseline ctDNA date

patient_ids_with_dates <- universe %>%
  arrange(patient_id, therapy_line) %>%
  mutate(
    therapy_line_label = case_when(
      therapy_line == 1 ~ "1L",
      therapy_line == 2 ~ "2L",
      therapy_line == 3 ~ "3L",
      therapy_line >= 4 ~ "4L+",
      TRUE ~ as.character(therapy_line)
    ),
    baseline_ctdna_date_formatted = format(ctdna_before_line_date, "%Y-%m-%d")
  )

patient_ids_table_data <- patient_ids_with_dates %>%
  select(patient_id, therapy_line_label, baseline_ctdna_date_formatted) %>%
  rename(
    `Patient ID` = patient_id,
    `Therapy Line` = therapy_line_label,
    `Baseline ctDNA Date` = baseline_ctdna_date_formatted
  )

# Create table
patient_ids_table <- patient_ids_table_data %>%
  kable(caption = paste0("Patient IDs with Baseline ctDNA Dates by Therapy Line (N = ", length(unique(patient_ids_with_dates$patient_id)), " patients, ", nrow(patient_ids_with_dates), " patient-line observations)"), 
        align = "l") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE, font_size = 9)

patient_ids_table
```

```{r palliative_therapy_distribution}
# Distribution of palliative systemic therapy regimens (for later classification)
# Try explicit regimen columns; if none, fall back to systemic_therapy_X columns as regimen
regimen_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("systemic.*therapy.*\\d+.*regimen|treatment.*regimen", ignore_case = TRUE))]
if (length(regimen_cols) == 0) {
  regimen_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("^systemic_therapy_\\d+$", ignore_case = TRUE))]
}

# Therapy notes for regimen clarification (broader pattern)
note_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("systemic.*therapy.*note|treatment.*note", ignore_case = TRUE))]

palliative_regimen_table <- NULL

# If no setting columns, assume metastatic cohort → treat all systemic therapy as palliative
settings_available <- length(therapy_setting_cols) > 0

if (length(regimen_cols) > 0) {
  therapy_settings_long <- if (settings_available) {
    dat_combined %>%
      select(patient_id, any_of(therapy_setting_cols)) %>%
      pivot_longer(-patient_id, names_to = "setting_var", values_to = "setting_raw") %>%
      mutate(
        therapy_num = str_extract(setting_var, "\\d+"),
        therapy_num = ifelse(is.na(therapy_num) | therapy_num == "", "all", therapy_num),
        setting_lower = tolower(as.character(setting_raw)),
        is_palliative = str_detect(setting_lower, "palliative|metastatic|1l.*metastatic|2l.*metastatic")
      ) %>%
      select(patient_id, therapy_num, is_palliative, setting_raw)
  } else {
    dat_combined %>%
      transmute(patient_id, across(any_of(regimen_cols))) %>%
      pivot_longer(-patient_id, names_to = "setting_var", values_to = "setting_raw") %>%
      mutate(
        therapy_num = str_extract(setting_var, "\\d+"),
        therapy_num = ifelse(is.na(therapy_num) | therapy_num == "", "all", therapy_num),
        is_palliative = TRUE
      ) %>%
      select(patient_id, therapy_num, is_palliative, setting_raw)
  }
  
  regimen_long <- dat_combined %>%
    select(patient_id, any_of(regimen_cols)) %>%
    pivot_longer(-patient_id, names_to = "regimen_var", values_to = "regimen_raw") %>%
    mutate(
      therapy_num = str_extract(regimen_var, "\\d+"),
      therapy_num = ifelse(is.na(therapy_num) | therapy_num == "", "all", therapy_num),
      regimen_orig = str_squish(as.character(regimen_raw)),
      regimen_clean = tolower(str_squish(as.character(regimen_raw))),
      regimen_clean = ifelse(is.na(regimen_clean) | regimen_clean == "", "unknown/not recorded", regimen_clean)
    ) %>%
    select(patient_id, therapy_num, regimen_clean, regimen_orig)

  if (length(note_cols) > 0) {
    notes_long <- dat_combined %>%
      select(patient_id, any_of(note_cols)) %>%
      pivot_longer(-patient_id, names_to = "note_var", values_to = "note_raw") %>%
      mutate(
        therapy_num = str_extract(note_var, "\\d+"),
        therapy_num = ifelse(is.na(therapy_num) | therapy_num == "", "all", therapy_num),
        note_clean = str_squish(as.character(note_raw)),
        note_clean = ifelse(is.na(note_clean) | note_clean == "", NA_character_, note_clean)
      ) %>%
      select(patient_id, therapy_num, note_clean)
    
    notes_any <- notes_long %>%
      group_by(patient_id) %>%
      summarise(note_any = first(note_clean[!is.na(note_clean)]), .groups = "drop")
    
    regimen_notes <- regimen_long %>%
      left_join(notes_long, by = c("patient_id", "therapy_num")) %>%
      left_join(notes_any, by = "patient_id") %>%
      mutate(note_filled = coalesce(note_clean, note_any))
  } else {
    regimen_notes <- regimen_long %>%
      mutate(note_clean = NA_character_, note_any = NA_character_, note_filled = NA_character_)
  }
  
  palliative_regimen_table <- regimen_notes %>%
    inner_join(therapy_settings_long, by = c("patient_id", "therapy_num")) %>%
    filter(is_palliative) %>%
    mutate(
      flag_other = str_detect(regimen_clean, regex("other.*note", ignore_case = TRUE)),
      flag_trial = str_detect(regimen_clean, regex("clinical.*trial", ignore_case = TRUE)),
      flag_see_note = str_detect(regimen_clean, regex("see\\s*note", ignore_case = TRUE)),
      regimen_final = case_when(
        (flag_other | flag_trial | flag_see_note) ~ coalesce(str_squish(note_filled), "unspecified from notes"),
        TRUE ~ regimen_clean
      ),
      regimen_final = ifelse(is.na(regimen_final) | regimen_final == "", "unknown/not recorded", regimen_final),
      # final guard: strip any lingering "see note" text
      regimen_final = ifelse(str_detect(regimen_final, regex("see\\s*note", ignore_case = TRUE)),
                             coalesce(str_squish(note_filled), "unspecified from notes"),
                             regimen_final)
    ) %>%
    # Classify regimen_final into ADC / ICI / Chemo / Targeted / Other, allowing combos
    rowwise() %>%
    mutate(
      lower_final = tolower(regimen_final),
      is_adc = str_detect(lower_final, regex("enfortumab|disitimab|sacituzumab|trastuzumab\\s*deruxtecan|dato|bt8009|adc", ignore_case = TRUE)),
      is_ici = str_detect(lower_final, regex("pembro|pembrolizumab|nivolumab|nivo\\b|atezo|atezolizumab|durva|durvalumab|relatlimab|pd-1|pd1|pd-l1|pdl1", ignore_case = TRUE)),
      is_chemo = str_detect(lower_final, regex("gem\\b|gemcitabine|cis\\b|cisplatin|carbo|carboplatin|oxaliplatin|taxol|docetaxel|cabazitaxel|folfox|folfiri", ignore_case = TRUE)),
      is_targeted = str_detect(lower_final, regex("erdafitinib|olaparib|loxo|fgfr", ignore_case = TRUE)),
      is_other = str_detect(lower_final, regex("vaccine|peptide|unspecified", ignore_case = TRUE)),
      class_vec = list(c(
        if (is_adc) "ADC",
        if (is_ici) "ICI",
        if (is_chemo) "Chemo",
        if (is_targeted) "Targeted"
      )),
      therapy_class = {
        cv <- class_vec[[1]]
        if (length(cv) == 0) {
          if (is_other) "Other" else "unspecified"
        } else {
          paste(cv, collapse = " + ")
        }
      },
      therapy_line_match = suppressWarnings(as.integer(therapy_num))
    ) %>%
    ungroup()
  
  regimen_summary <- palliative_regimen_table %>%
    group_by(regimen_final) %>%
    summarise(
      n_lines = n(),
      n_patients = n_distinct(patient_id),
      .groups = "drop"
    ) %>%
    arrange(desc(n_lines))
  
  class_summary <- palliative_regimen_table %>%
    group_by(therapy_class) %>%
    summarise(
      n_lines = n(),
      n_patients = n_distinct(patient_id),
      .groups = "drop"
    ) %>%
    arrange(desc(n_lines))
  
  regimen_summary %>%
    kable(caption = "Palliative Systemic Therapy Regimens (counts by regimen; palliative setting only)") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
  
  class_summary %>%
    kable(caption = "Palliative Systemic Therapy Classes (ADC/ICI/Chemo/Targeted/Other). Note: patients can appear in multiple classes if they received combination regimens (e.g., ADC + ICI).") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
} else {
  cat("No palliative therapy regimen information available (regimen columns not found).\n")
}
```

```{r extract_outcomes, include=FALSE}
# ====================================================================================
# DURATION OF COMPLETE RESPONSE (DCR) DEFINITION
# ====================================================================================
# Note: We analyze "Duration of Complete Response" (DCR), which represents the time from 
# therapy initiation until loss of complete response or death. This endpoint focuses on 
# patients who achieve and maintain complete response (similar to RECIST CR).
#
# DCR EVENT (dcr_event = 1, stored as mfs_event for backward compatibility) occurs when:
#   1. Patient has died (death_date is not NA), OR
#   2. Patient is deceased (alive_status = "No"), OR
#   3. Patient is alive BUT has obvious disease on most recent imaging
#      (imaging_disease_present = "Yes")
#      Note: This represents loss of complete response (similar to RECIST progression from CR).
#      This includes patients with stable disease, partial response, or progressive disease - 
#      any patient with disease visible on imaging is considered an event.
#
# DCR CENSORED (dcr_event = 0) occurs when:
#   - Patient is alive (alive_status = "Yes") AND
#   - Patient has NO obvious disease on most recent imaging
#     (imaging_disease_present = "No")
#     Note: This represents maintenance of complete response (similar to RECIST CR)
#
# SPECIAL HANDLING:
#   - If imaging field is BLANK or MISSING AND patient is ALIVE: EXCLUDE patient (imaging data is required for alive patients)
#   - If patient DIED: Imaging data is NOT required (death is an event)
#   - Rationale: Imaging data is required for DCR definition only for alive patients
#
# DCR TIME (dcr_time, stored as mfs_time for backward compatibility):
#   - Calculated from time_zero_date (therapy start date) to event_or_censor_date
#   - event_or_censor_date = death_date (if available) OR last_fu_date (if no death)
#   - Units: days
#   - NOTE: Time 0 is the start of each new line of therapy, not ctDNA date
#
# Note: This will filter universe to only patients with valid survival data
# ====================================================================================
death_date_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("^death_date$|death.*date$", ignore_case = TRUE)) & 
                                       !str_detect(names(dat_combined), regex("cause|reason", ignore_case = TRUE))]
survival_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("last.*follow.*up.*date|follow.*up.*date", ignore_case = TRUE)) & 
                                     !str_detect(names(dat_combined), regex("death", ignore_case = TRUE))]
alive_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("^alive_yes_no$|alive.*yes.*no", ignore_case = TRUE))]
imaging_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("obvious.*disease.*imaging|disease.*present.*imaging", ignore_case = TRUE))]

# Outcomes: join universe with survival data
# Note: universe already has therapy_line, so we preserve it
outcomes <- universe %>%
  left_join(
    dat_combined %>%
      select(patient_id, all_of(c(death_date_cols, survival_cols, alive_cols, imaging_cols))),
    by = "patient_id"
  ) %>%
  mutate(
    death_date = if (length(death_date_cols) > 0) {
      as_date_safe(.data[[death_date_cols[1]]])
    } else {
      as.Date(NA)
    },
    last_fu_date = if (length(survival_cols) > 0) {
      as_date_safe(.data[[survival_cols[1]]])
    } else {
      as.Date(NA)
    },
    alive_status = if (length(alive_cols) > 0) {
      case_when(
        str_detect(tolower(as.character(.data[[alive_cols[1]]])), "yes|y|1") ~ "Yes",
        str_detect(tolower(as.character(.data[[alive_cols[1]]])), "no|n|0") ~ "No",
        TRUE ~ NA_character_
      )
    } else {
      NA_character_
    },
    # ========================================================================
    # STEP 1: Determine imaging disease status
    # ========================================================================
    # Column: "if_alive_is_obvious_disease_still_present_on_most_recent_imaging"
    # If blank/missing AND patient is alive: will be excluded (imaging data is required for alive patients)
    # If patient died: imaging data is not required (death is an event)
    imaging_disease_present = if (length(imaging_cols) > 0) {
      case_when(
        str_detect(tolower(as.character(.data[[imaging_cols[1]]])), "yes|y|1") ~ "Yes",
        str_detect(tolower(as.character(.data[[imaging_cols[1]]])), "no|n|0") ~ "No",
        TRUE ~ NA_character_  # Missing/blank = NA (will be excluded)
      )
    } else {
      NA_character_  # If column doesn't exist, mark as missing
    },
    
    # ========================================================================
    # STEP 2: Calculate MFS time (days from therapy start to event/censor)
    # ========================================================================
    # Time 0 is the start of each new line of therapy (time_zero_date)
    event_or_censor_date = case_when(
      !is.na(death_date) ~ death_date,      # Use death date if available
      !is.na(last_fu_date) ~ last_fu_date,  # Otherwise use last follow-up date
      TRUE ~ as.Date(NA)
    ),
    mfs_time = as.numeric(event_or_censor_date - time_zero_date),
    
    # ========================================================================
    # STEP 3: Determine MFS event status
    # ========================================================================
    # mfs_event = 1 (EVENT) if:
    #   - Patient died (death_date exists), OR
    #   - Patient is deceased (alive_status = "No"), OR
    #   - Patient is alive BUT has obvious disease on imaging
    #
    # mfs_event = 0 (CENSORED) if:
    #   - Patient is alive AND has no obvious disease on imaging
    mfs_event = case_when(
      !is.na(death_date) ~ 1,  # Death = DCR event
      alive_status == "No" ~ 1,  # Deceased = DCR event
      alive_status == "Yes" & imaging_disease_present == "Yes" ~ 1,  # Alive with disease = DCR event
      alive_status == "Yes" & imaging_disease_present == "No" ~ 0,  # Alive without disease = CENSORED
      alive_status == "Yes" & is.na(imaging_disease_present) ~ NA_real_,  # Alive but missing imaging = exclude
      TRUE ~ NA_real_  # Other missing data = exclude
    ),
    
    # Determine if imaging data is required (only for alive patients)
    imaging_required = case_when(
      !is.na(death_date) ~ FALSE,  # Dead patients don't need imaging
      alive_status == "No" ~ FALSE,  # Deceased patients don't need imaging
      alive_status == "Yes" ~ TRUE,  # Alive patients need imaging
      TRUE ~ TRUE  # Default: require imaging if status unknown
    ),
    
  ) %>%
  # Filter: require valid DCR time
  # For imaging: only exclude if patient is alive AND imaging is missing
  filter(!is.na(mfs_time), mfs_time > 0, 
         !(imaging_required == TRUE & is.na(imaging_disease_present)))

# Identify excluded observations and reasons
# Note: exclusions are at the patient-line level, not just patient level
excluded_patients <- universe %>%
  left_join(
    dat_combined %>%
      select(patient_id, all_of(c(death_date_cols, survival_cols, alive_cols, imaging_cols))),
    by = "patient_id"
  ) %>%
  mutate(
    death_date = if (length(death_date_cols) > 0) {
      as_date_safe(.data[[death_date_cols[1]]])
    } else {
      as.Date(NA)
    },
    last_fu_date = if (length(survival_cols) > 0) {
      as_date_safe(.data[[survival_cols[1]]])
    } else {
      as.Date(NA)
    },
    alive_status = if (length(alive_cols) > 0) {
      case_when(
        str_detect(tolower(as.character(.data[[alive_cols[1]]])), "yes|y|1") ~ "Yes",
        str_detect(tolower(as.character(.data[[alive_cols[1]]])), "no|n|0") ~ "No",
        TRUE ~ NA_character_
      )
    } else {
      NA_character_
    },
    imaging_disease_present = if (length(imaging_cols) > 0) {
      case_when(
        str_detect(tolower(as.character(.data[[imaging_cols[1]]])), "yes|y|1") ~ "Yes",
        str_detect(tolower(as.character(.data[[imaging_cols[1]]])), "no|n|0") ~ "No",
        TRUE ~ NA_character_  # Missing/blank = NA
      )
    } else {
      NA_character_
    },
    event_or_censor_date = case_when(
      !is.na(death_date) ~ death_date,
      !is.na(last_fu_date) ~ last_fu_date,
      TRUE ~ as.Date(NA)
    ),
    mfs_time = as.numeric(event_or_censor_date - time_zero_date),
    imaging_required = case_when(
      !is.na(death_date) ~ FALSE,  # Dead patients don't need imaging
      alive_status == "No" ~ FALSE,  # Deceased patients don't need imaging
      alive_status == "Yes" ~ TRUE,  # Alive patients need imaging
      TRUE ~ TRUE  # Default: require imaging if status unknown
    )
  ) %>%
  filter(is.na(mfs_time) | mfs_time <= 0 | (imaging_required == TRUE & is.na(imaging_disease_present))) %>%
  mutate(
    exclusion_reason = case_when(
      is.na(event_or_censor_date) ~ "Missing both death_date and last_fu_date",
      !is.na(mfs_time) & mfs_time <= 0 ~ "Invalid survival time (event/censor date before or equal to therapy start date)",
      imaging_required == TRUE & is.na(imaging_disease_present) ~ "Missing imaging data (required for alive patients)",
      TRUE ~ "Other"
    )
  ) %>%
  select(patient_id, therapy_line, exclusion_reason, death_date, last_fu_date, alive_status, 
         imaging_disease_present, time_zero_date, event_or_censor_date, mfs_time)

# Diagnostic: Check how many patients were excluded
cat("Universe size before survival data extraction: n =", nrow(universe), " patient-line observations\n")
cat("Outcomes size after filtering for valid survival data: n =", nrow(outcomes), " patient-line observations\n")
cat("Observations excluded due to missing survival data: n =", nrow(universe) - nrow(outcomes), "\n")

# OPTIMIZATION: Cache outcomes for faster subsequent runs
if (!dir.exists("cache")) dir.create("cache")
saveRDS(outcomes, "cache/outcomes.rds")

# OPTIMIZATION: Pre-compute therapy_line_factor once for reuse
outcomes <- outcomes %>%
  mutate(
    therapy_line_grouped = case_when(
      therapy_line == 1 ~ "1L",
      therapy_line == 2 ~ "2L",
      therapy_line == 3 ~ "3L",
      therapy_line >= 4 ~ "4L+",
      TRUE ~ as.character(therapy_line)
    ),
    therapy_line_factor = factor(therapy_line_grouped, levels = c("1L", "2L", "3L", "4L+"))
  )
```

```{r exclusion_reasons_table}
# Table showing reasons for exclusion
if (nrow(excluded_patients) > 0) {
  exclusion_summary <- excluded_patients %>%
    count(exclusion_reason) %>%
    mutate(Percentage = round(n / sum(n) * 100, 1)) %>%
    rename(`Exclusion Reason` = exclusion_reason, `N` = n) %>%
    kable(caption = paste0("Reasons for Exclusion from Analysis (N = ", nrow(excluded_patients), " patient-line observations excluded from N = ", nrow(universe), " screened)")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
  
  exclusion_summary
} else {
  cat("No observations excluded - all", nrow(universe), "patient-line observations had valid survival data.\n")
}
```

```{r excluded_patients_imaging_missing}
# List patient IDs excluded due to missing imaging data (for alive patients only)
# Note: Patients who died do NOT need imaging data - only alive patients require imaging

# Question 1 exclusions
# Compare universe (all patients) with outcomes (included patients) to find who was excluded
# due to missing imaging data for alive patients
universe_with_status <- universe %>%
  left_join(
    dat_combined %>%
      select(patient_id, all_of(c(death_date_cols, alive_cols, imaging_cols))),
    by = "patient_id"
  ) %>%
  mutate(
    death_date = if (length(death_date_cols) > 0) {
      as_date_safe(.data[[death_date_cols[1]]])
    } else {
      as.Date(NA)
    },
    alive_status = if (length(alive_cols) > 0) {
      case_when(
        str_detect(tolower(as.character(.data[[alive_cols[1]]])), "yes|y|1") ~ "Yes",
        str_detect(tolower(as.character(.data[[alive_cols[1]]])), "no|n|0") ~ "No",
        TRUE ~ NA_character_
      )
    } else {
      NA_character_
    },
    imaging_disease_present = if (length(imaging_cols) > 0) {
      case_when(
        str_detect(tolower(as.character(.data[[imaging_cols[1]]])), "yes|y|1") ~ "Yes",
        str_detect(tolower(as.character(.data[[imaging_cols[1]]])), "no|n|0") ~ "No",
        TRUE ~ NA_character_
      )
    } else {
      NA_character_
    },
    imaging_required = case_when(
      !is.na(death_date) ~ FALSE,
      alive_status == "No" ~ FALSE,
      alive_status == "Yes" ~ TRUE,
      TRUE ~ TRUE
    )
  )

# Find patients in universe but not in outcomes due to missing imaging (alive patients only)
excluded_q1_imaging <- universe_with_status %>%
  anti_join(
    outcomes %>% select(patient_id, therapy_line),
    by = c("patient_id", "therapy_line")
  ) %>%
  filter(imaging_required == TRUE & is.na(imaging_disease_present)) %>%
  select(patient_id, therapy_line, alive_status, imaging_disease_present) %>%
  mutate(exclusion_reason = "Missing imaging data (required for alive patients)") %>%
  distinct()

# Question 2 exclusions (landmark analysis)
# Compare landmark_data (universe) with landmark_outcomes (included) to find who was excluded
if (exists("landmark_data") && nrow(landmark_data) > 0) {
  # Get status information for landmark_data
  landmark_data_with_status <- landmark_data %>%
    left_join(
      outcomes %>% 
        select(patient_id, therapy_line, mfs_time, mfs_event, time_zero_date, imaging_disease_present),
      by = c("patient_id", "therapy_line", "time_zero_date")
    ) %>%
    left_join(
      dat_combined %>%
        select(patient_id, all_of(c(death_date_cols, alive_cols))),
      by = "patient_id"
    ) %>%
    mutate(
      death_date = if (length(death_date_cols) > 0) {
        as_date_safe(.data[[death_date_cols[1]]])
      } else {
        as.Date(NA)
      },
      alive_status = if (length(alive_cols) > 0) {
        case_when(
          str_detect(tolower(as.character(.data[[alive_cols[1]]])), "yes|y|1") ~ "Yes",
          str_detect(tolower(as.character(.data[[alive_cols[1]]])), "no|n|0") ~ "No",
          TRUE ~ NA_character_
        )
      } else {
        NA_character_
      },
      survived_to_landmark = !is.na(mfs_time) & mfs_time >= landmark_time,
      mfs_time_from_landmark = mfs_time - landmark_time,
      imaging_required = case_when(
        !is.na(death_date) ~ FALSE,
        alive_status == "No" ~ FALSE,
        alive_status == "Yes" ~ TRUE,
        TRUE ~ TRUE
      )
    )
  
  # Find patients in landmark_data but not in landmark_outcomes due to missing imaging (alive patients only)
  if (exists("landmark_outcomes") && nrow(landmark_outcomes) > 0) {
    excluded_q2_imaging <- landmark_data_with_status %>%
      anti_join(
        landmark_outcomes %>% select(patient_id, therapy_line, time_zero_date),
        by = c("patient_id", "therapy_line", "time_zero_date")
      ) %>%
      filter(survived_to_landmark, 
             !is.na(mfs_time_from_landmark) & mfs_time_from_landmark > 0,
             imaging_required == TRUE & is.na(imaging_disease_present)) %>%
      select(patient_id, therapy_line, alive_status, imaging_disease_present) %>%
      mutate(exclusion_reason = "Missing imaging data (required for alive patients)") %>%
      distinct()
  } else {
    # If landmark_outcomes doesn't exist, check all landmark_data patients
    excluded_q2_imaging <- landmark_data_with_status %>%
      filter(survived_to_landmark, 
             !is.na(mfs_time_from_landmark) & mfs_time_from_landmark > 0,
             imaging_required == TRUE & is.na(imaging_disease_present)) %>%
      select(patient_id, therapy_line, alive_status, imaging_disease_present) %>%
      mutate(exclusion_reason = "Missing imaging data (required for alive patients)") %>%
      distinct()
  }
} else {
  excluded_q2_imaging <- tibble(patient_id = character(0), therapy_line = integer(0), 
                                 exclusion_reason = character(0), alive_status = character(0),
                                 imaging_disease_present = character(0))
}

# Combine and display
if (nrow(excluded_q1_imaging) > 0 || nrow(excluded_q2_imaging) > 0) {
  cat("## Patient IDs Excluded Due to Missing Imaging Data\n\n")
  cat("**Note:** Only alive patients require imaging data. Patients who died do not need imaging data for inclusion.\n\n")
  
  if (nrow(excluded_q1_imaging) > 0) {
    cat("### Question 1 (Baseline ctDNA Analysis)\n\n")
    cat("**N =", nrow(excluded_q1_imaging), "patient-line observations excluded**\n\n")
    
    excluded_q1_table <- excluded_q1_imaging %>%
      arrange(patient_id, therapy_line) %>%
      mutate(
        therapy_line_label = case_when(
          therapy_line == 1 ~ "1L",
          therapy_line == 2 ~ "2L",
          therapy_line == 3 ~ "3L",
          therapy_line >= 4 ~ "4L+",
          TRUE ~ as.character(therapy_line)
        )
      ) %>%
      select(`Patient ID` = patient_id, `Therapy Line` = therapy_line_label, 
             `Alive Status` = alive_status) %>%
      kable(caption = "Question 1: Patient IDs Excluded Due to Missing Imaging Data (Alive Patients Only)") %>%
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
    
    excluded_q1_table
    cat("\n")
  } else {
    cat("### Question 1 (Baseline ctDNA Analysis)\n\n")
    cat("**No patients excluded due to missing imaging data.**\n\n")
  }
  
  if (nrow(excluded_q2_imaging) > 0) {
    cat("### Question 2 (Landmark Analysis)\n\n")
    cat("**N =", nrow(excluded_q2_imaging), "patient-line observations excluded**\n\n")
    
    excluded_q2_table <- excluded_q2_imaging %>%
      arrange(patient_id, therapy_line) %>%
      mutate(
        therapy_line_label = case_when(
          therapy_line == 1 ~ "1L",
          therapy_line == 2 ~ "2L",
          therapy_line == 3 ~ "3L",
          therapy_line >= 4 ~ "4L+",
          TRUE ~ as.character(therapy_line)
        )
      ) %>%
      select(`Patient ID` = patient_id, `Therapy Line` = therapy_line_label,
             `Alive Status` = alive_status) %>%
      kable(caption = "Question 2: Patient IDs Excluded Due to Missing Imaging Data (Alive Patients Only)") %>%
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
    
    excluded_q2_table
  } else {
    cat("### Question 2 (Landmark Analysis)\n\n")
    cat("**No patients excluded due to missing imaging data.**\n\n")
  }
  
  # Summary of unique patients
  all_excluded_patients <- bind_rows(
    excluded_q1_imaging %>% select(patient_id),
    excluded_q2_imaging %>% select(patient_id)
  ) %>%
    distinct()
  
  cat("\n**Total unique patients excluded from either question due to missing imaging data:**", 
      nrow(all_excluded_patients), "\n")
} else {
  cat("## Patient IDs Excluded Due to Missing Imaging Data\n\n")
  cat("**No patients excluded due to missing imaging data in either Question 1 or Question 2.**\n")
  cat("(All alive patients had imaging data available, or all excluded patients had died and did not require imaging data.)\n")
}
```

```{r descriptive_table_combined}
# Combined descriptive table for universe and outcomes
# Note: This table shows the final analysis N (outcomes) after filtering for valid survival data
desc_table_combined <- universe %>%
  left_join(
    outcomes %>% select(patient_id, therapy_line, mfs_event, mfs_time),
    by = c("patient_id", "therapy_line")
  ) %>%
  filter(!is.na(mfs_time)) %>%  # Only include observations with valid survival data
  mutate(
    therapy_line_label = case_when(
      therapy_line == 1 ~ "1L",
      therapy_line == 2 ~ "2L",
      therapy_line == 3 ~ "3L",
      therapy_line >= 4 ~ "4L+",
      TRUE ~ "Unknown"
    )
  ) %>%
  summarise(
    `N (Patient-Line Observations)` = n(),
    `N (Unique Patients)` = length(unique(patient_id)),
    `1L` = sum(therapy_line == 1, na.rm = TRUE),
    `2L` = sum(therapy_line == 2, na.rm = TRUE),
    `3L` = sum(therapy_line == 3, na.rm = TRUE),
    `4L+` = sum(therapy_line >= 4, na.rm = TRUE),
    `ctDNA Before Therapy (median, IQR)` = paste0(
      round(median(ctdna_before_line_value, na.rm = TRUE), 2), 
      " (", 
      round(quantile(ctdna_before_line_value, 0.25, na.rm = TRUE), 2), 
      "-", 
      round(quantile(ctdna_before_line_value, 0.75, na.rm = TRUE), 2), 
      ")"
    ),
    `Days from ctDNA to Therapy Start (median, IQR)` = paste0(
      round(median(days_from_ctdna_to_therapy, na.rm = TRUE), 0), 
      " (", 
      round(quantile(days_from_ctdna_to_therapy, 0.25, na.rm = TRUE), 0), 
      "-", 
      round(quantile(days_from_ctdna_to_therapy, 0.75, na.rm = TRUE), 0), 
      ")"
    ),
    `DCR Events` = sum(mfs_event, na.rm = TRUE),
    `Median DCR (months)` = round(median(mfs_time, na.rm = TRUE) / 30.44, 1),
    `1-year DCR %` = round(100 * (1 - sum(mfs_event[mfs_time <= 365.25], na.rm = TRUE) / sum(mfs_time <= 365.25, na.rm = TRUE)), 1)
  ) %>%
  kable(caption = "Descriptive Characteristics of Study Population") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)

desc_table_combined
```

```{r data_completeness_diagnostic, include=FALSE}
# Diagnostic: Check data completeness for event classification
data_completeness <- outcomes %>%
  mutate(
    has_death_date = !is.na(death_date),
    has_last_fu_date = !is.na(last_fu_date),
    has_alive_status = !is.na(alive_status),
    alive_status_value = alive_status,
    has_imaging_status = !is.na(imaging_disease_present),
    imaging_status_value = imaging_disease_present,
    # Classify MFS events: death OR (alive with disease on imaging)
    event_classification = case_when(
      !is.na(death_date) ~ "Event: Death (has death_date)",
      alive_status == "No" ~ "Event: Death (alive_status=No)",
      alive_status == "Yes" & imaging_disease_present == "Yes" ~ "Event: Imaging shows disease (alive with disease)",
      alive_status == "Yes" & imaging_disease_present == "No" ~ "Censored: Alive without disease on imaging",
      is.na(alive_status) & !is.na(last_fu_date) ~ "Censored: Missing alive_status, has FU date",
      TRUE ~ "Censored: Other"
    )
  ) %>%
  select(patient_id, has_death_date, has_last_fu_date, has_alive_status, 
         alive_status_value, has_imaging_status, imaging_status_value,
         event_classification, mfs_event, mfs_time)

# Summary of data completeness
completeness_summary <- data_completeness %>%
  summarise(
    N = n(),
    `Has Death Date` = sum(has_death_date),
    `Has Last FU Date` = sum(has_last_fu_date),
    `Has Alive Status` = sum(has_alive_status),
    `Alive Status = Yes` = sum(alive_status_value == "Yes", na.rm = TRUE),
    `Alive Status = No` = sum(alive_status_value == "No", na.rm = TRUE),
    `Alive Status Missing` = sum(is.na(alive_status_value)),
    `Events` = sum(mfs_event),
    `Censored` = sum(1 - mfs_event),
    `Censoring Rate` = round(mean(1 - mfs_event), 3)
  )

# Breakdown by classification
classification_breakdown <- data_completeness %>%
  count(event_classification) %>%
  mutate(percentage = round(n / sum(n) * 100, 1))

cat("Data Completeness Summary:\n")
print(completeness_summary)
cat("\nEvent Classification Breakdown:\n")
print(classification_breakdown)
```

```{r censoring_analysis, include=FALSE}
# Censoring rate analysis - similar to descriptive analysis
# 
# ====================================================================================
# CENSORING CRITERIA FOR DURATION OF COMPLETE RESPONSE (DCR)
# ====================================================================================
# 
# DCR EVENT (dcr_event = 1, stored as mfs_event) occurs when ANY of the following:
#   1. Patient has died (death_date is not NA)
#   2. Patient is deceased (alive_status = "No")
#   3. Patient is alive BUT has obvious disease on most recent imaging
#      (imaging_disease_present = "Yes")
#      Note: This represents loss of complete response (similar to RECIST progression from CR).
#      This includes stable disease, partial response, or progressive disease.
#
# DCR CENSORED (dcr_event = 0) occurs when:
#   - Patient is alive (alive_status = "Yes") AND
#   - Patient has NO obvious disease on most recent imaging
#     (imaging_disease_present = "No")
#     Note: This represents maintenance of complete response (similar to RECIST CR)
#   - NOTE: If imaging field is BLANK or MISSING AND patient is ALIVE, patient is EXCLUDED (imaging data is required for alive patients only)
#   - NOTE: If patient DIED, imaging data is NOT required (death is an event)
#
# DCR TIME CALCULATION:
#   - Censoring time = last_fu_date (last follow-up date)
#   - Event time = death_date (if available) OR last_fu_date (if imaging shows disease)
#   - Time is calculated from time_zero_date (therapy start) to event_or_censor_date
# ====================================================================================
#
# Create censoring indicator
outcomes_with_cens <- outcomes %>%
  filter(!is.na(ctdna_before_line_value)) %>%  # Remove NA values before creating tertiles
  mutate(
    censored = 1 - mfs_event,  # 1 = censored, 0 = event
    ctdna_tertile = ntile(ctdna_before_line_value, 3),
    ctdna_group = case_when(
      ctdna_tertile == 1 ~ "T1 (Lowest)",
      ctdna_tertile == 2 ~ "T2",
      ctdna_tertile == 3 ~ "T3 (Highest)"
    )
  )

# Calculate censoring rates by ctDNA tertile
censoring_rates <- outcomes_with_cens %>%
  group_by(ctdna_group) %>%
  summarise(
    n = n(),
    n_events = sum(mfs_event, na.rm = TRUE),
    n_censored = sum(censored, na.rm = TRUE),
    censoring_rate = mean(censored, na.rm = TRUE),
    .groups = "drop"
  )

# Chi-square test for independence of censoring and group
if (nrow(censoring_rates) > 1 && sum(censoring_rates$n_censored) > 0 && sum(censoring_rates$n_events) > 0) {
  censoring_table <- table(outcomes_with_cens$ctdna_group, outcomes_with_cens$censored)
  chi_square_test <- tryCatch(
    chisq.test(censoring_table),
    error = function(e) NULL
  )
} else {
  chi_square_test <- NULL
}

# Kruskal-Wallis test for censoring times (for censored patients only)
censored_patients <- outcomes_with_cens %>%
  filter(censored == 1)

if (nrow(censored_patients) > 0 && length(unique(censored_patients$ctdna_group)) > 1) {
  kruskal_test <- tryCatch(
    kruskal.test(mfs_time ~ ctdna_group, data = censored_patients),
    error = function(e) NULL
  )
} else {
  kruskal_test <- NULL
}

cat("Censoring analysis completed.\n")
if (!is.null(chi_square_test)) {
  cat("Chi-square p-value:", chi_square_test$p.value, "\n")
  if (chi_square_test$p.value >= 0.05) {
    cat("Censoring appears non-informative (p >= 0.05)\n")
  } else {
    cat("Censoring may be informative (p < 0.05)\n")
  }
}
```


```{r data_completeness_table}
# Display data completeness diagnostic
completeness_display <- completeness_summary %>%
  pivot_longer(everything(), names_to = "Metric", values_to = "Value") %>%
  kable(caption = "Data Completeness for Event Classification") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)

completeness_display

classification_display <- classification_breakdown %>%
  rename(`Classification` = event_classification, `N` = n, `%` = percentage) %>%
      kable(caption = "Breakdown of DCR Event vs. Censored Classification\n\nNote: DCR Events include: (1) Death (death_date or alive_status=No), OR (2) Alive with obvious disease on most recent imaging (loss of complete response, includes stable disease, partial response, or progressive disease). DCR Censored = Alive without obvious disease on imaging (maintaining complete response, similar to RECIST CR). Only alive patients with missing imaging data are excluded (imaging data is required for alive patients only; patients who died do not require imaging data).") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)

classification_display
```

```{r censoring_analysis_table}
# Display censoring analysis results
# Ensure all tertiles are represented even if they have 0 observations
all_tertiles_df <- tibble(ctdna_group = c("T1 (Lowest)", "T2", "T3 (Highest)"))

# Create complete summary with all tertiles - use full_join to ensure all are included
censoring_summary <- all_tertiles_df %>%
  full_join(censoring_rates, by = "ctdna_group") %>%
  mutate(
    n = replace_na(n, 0),
    n_events = replace_na(n_events, 0),
    n_censored = replace_na(n_censored, 0),
    censoring_rate = ifelse(n == 0, 0, replace_na(censoring_rate, 0)),
    `Censoring Rate` = ifelse(n == 0, "N/A", paste0(round(censoring_rate * 100, 1), "%")),
    `Interpretation` = ifelse(n == 0, "N/A",
                             ifelse(!is.null(chi_square_test) && !is.null(chi_square_test$p.value) &&
                                    chi_square_test$p.value < 0.05,
                                    "⚠️ Potentially informative",
                                    "✓ Non-informative likely"))
  ) %>%
  arrange(match(ctdna_group, c("T1 (Lowest)", "T2", "T3 (Highest)"))) %>%
  select(Group = ctdna_group, N = n, `N Events` = n_events, 
         `N Censored` = n_censored, `Censoring Rate`, Interpretation)

# Add Chi-square P-value to first row only
if (!is.null(chi_square_test) && !is.null(chi_square_test$p.value)) {
  censoring_summary <- censoring_summary %>%
    mutate(`Chi-square P-value` = ifelse(row_number() == 1, as.character(round(chi_square_test$p.value, 4)), ""))
} else {
  censoring_summary <- censoring_summary %>%
    mutate(`Chi-square P-value` = "")
}

censoring_table_display <- censoring_summary %>%
  kable(caption = "Censoring Rate Analysis by ctDNA Tertile") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)

censoring_table_display

if (!is.null(chi_square_test)) {
  if (chi_square_test$p.value >= 0.05) {
    cat("\n**Interpretation:** Censoring appears non-informative (p =", 
        round(chi_square_test$p.value, 4), 
        "). This supports the assumption that censoring is independent of ctDNA level.\n")
  } else {
    cat("\n**Interpretation:** Censoring may be informative (p =", 
        round(chi_square_test$p.value, 4), 
        "). Consider sensitivity analyses.\n")
  }
}
```

```{r ctDNA_prognosis_analysis}
# Quantitative association between ctDNA level and prognosis
# Cox proportional hazards model
# Cox model with ctDNA and line of therapy as covariates
# Combine lines 4+ into one category
outcomes_for_model <- outcomes %>%
  left_join(
    palliative_regimen_table %>%
      mutate(therapy_line_match = ifelse(is.na(therapy_line_match), NA_integer_, therapy_line_match)) %>%
      select(patient_id, therapy_line_match, therapy_class) %>%
      distinct(),
    by = c("patient_id", "therapy_line" = "therapy_line_match")
  ) %>%
  mutate(
    therapy_line_grouped = case_when(
      therapy_line == 1 ~ "1L",
      therapy_line == 2 ~ "2L",
      therapy_line == 3 ~ "3L",
      therapy_line >= 4 ~ "4L+",
      TRUE ~ as.character(therapy_line)
    ),
    therapy_line_factor = factor(therapy_line_grouped, levels = c("1L", "2L", "3L", "4L+")),
    therapy_class = replace_na(therapy_class, "unspecified"),
    therapy_class_factor = factor(therapy_class)
  )

# Join Bajorin risk factors if available
if (exists("bajorin_data") && nrow(bajorin_data) > 0) {
  outcomes_for_model <- outcomes_for_model %>%
    left_join(
      bajorin_data %>% select(patient_id, therapy_line, kps_binary, visceral_mets_binary),
      by = c("patient_id", "therapy_line")
    ) %>%
    mutate(
      # Create factor variables for Bajorin risk factors
      kps_factor = factor(kps_binary, levels = c(0, 1), labels = c("KPS <80%", "KPS ≥80%")),
      visceral_mets_factor = factor(visceral_mets_binary, levels = c(0, 1), labels = c("No Visceral Mets", "Visceral Mets"))
    )
} else {
  outcomes_for_model <- outcomes_for_model %>%
    mutate(
      kps_binary = NA_real_,
      visceral_mets_binary = NA_real_,
      kps_factor = factor(NA, levels = c(0, 1), labels = c("KPS <80%", "KPS ≥80%")),
      visceral_mets_factor = factor(NA, levels = c(0, 1), labels = c("No Visceral Mets", "Visceral Mets"))
    )
}

# Model 1: Unadjusted (ctDNA only)
cox_model_unadj <- coxph(Surv(mfs_time, mfs_event) ~ log(ctdna_before_line_value + 1), data = outcomes_for_model)
cox_summary_unadj <- summary(cox_model_unadj)

# Model 2: Adjusted for line of therapy
cox_model_line <- coxph(Surv(mfs_time, mfs_event) ~ log(ctdna_before_line_value + 1) + therapy_line_factor, data = outcomes_for_model)
cox_summary_line <- summary(cox_model_line)

# Model 3: Adjusted for KPS only (if available)
cox_model_kps <- NULL
cox_summary_kps <- NULL
if (sum(!is.na(outcomes_for_model$kps_binary)) > 0) {
  outcomes_kps <- outcomes_for_model %>%
    filter(!is.na(kps_binary))
  
  if (nrow(outcomes_kps) > 0) {
    cox_model_kps <- tryCatch({
      coxph(Surv(mfs_time, mfs_event) ~ log(ctdna_before_line_value + 1) + kps_factor, 
            data = outcomes_kps)
    }, error = function(e) NULL)
    
    if (!is.null(cox_model_kps)) {
      cox_summary_kps <- summary(cox_model_kps)
    }
  }
}

# Model 4: Adjusted for visceral metastases only (if available)
cox_model_visceral <- NULL
cox_summary_visceral <- NULL
if (sum(!is.na(outcomes_for_model$visceral_mets_binary)) > 0) {
  outcomes_visceral <- outcomes_for_model %>%
    filter(!is.na(visceral_mets_binary))
  
  if (nrow(outcomes_visceral) > 0) {
    cox_model_visceral <- tryCatch({
      coxph(Surv(mfs_time, mfs_event) ~ log(ctdna_before_line_value + 1) + visceral_mets_factor, 
            data = outcomes_visceral)
    }, error = function(e) NULL)
    
    if (!is.null(cox_model_visceral)) {
      cox_summary_visceral <- summary(cox_model_visceral)
    }
  }
}

# Model 5: Multivariate with all covariates (if available)
cox_model_multiv <- NULL
cox_summary_multiv <- NULL
outcomes_multiv <- outcomes_for_model %>% filter(FALSE)  # Initialize as empty
if (sum(!is.na(outcomes_for_model$kps_binary)) > 0 && sum(!is.na(outcomes_for_model$visceral_mets_binary)) > 0) {
  # Filter to observations with complete Bajorin data
  outcomes_multiv <- outcomes_for_model %>%
    filter(!is.na(kps_binary), !is.na(visceral_mets_binary))
  
  if (nrow(outcomes_multiv) > 0) {
    cox_model_multiv <- tryCatch({
      coxph(Surv(mfs_time, mfs_event) ~ log(ctdna_before_line_value + 1) + therapy_line_factor + 
            kps_factor + visceral_mets_factor, 
            data = outcomes_multiv)
    }, error = function(e) NULL)
    
    if (!is.null(cox_model_multiv)) {
      cox_summary_multiv <- summary(cox_model_multiv)
    }
  }
}

# Create results table showing ctDNA HR progression across models
results_table_list <- list()

# 1. Unadjusted model
hr_unadj <- cox_summary_unadj$conf.int[1, "exp(coef)"]
hr_unadj_lower <- cox_summary_unadj$conf.int[1, "lower .95"]
hr_unadj_upper <- cox_summary_unadj$conf.int[1, "upper .95"]
p_unadj <- cox_summary_unadj$coefficients[1, "Pr(>|z|)"]
results_table_list[[length(results_table_list) + 1]] <- tibble(
  `Model` = "Unadjusted",
  `HR (95% CI)` = paste0(round(hr_unadj, 2), " (", round(hr_unadj_lower, 2), "-", round(hr_unadj_upper, 2), ")"),
  `P-value` = ifelse(p_unadj < 0.001, "<0.001", as.character(round(p_unadj, 3))),
  `N` = nrow(outcomes_for_model)
)

# 2. Adjusted for line of therapy
hr_line <- cox_summary_line$conf.int[1, "exp(coef)"]
hr_line_lower <- cox_summary_line$conf.int[1, "lower .95"]
hr_line_upper <- cox_summary_line$conf.int[1, "upper .95"]
p_line <- cox_summary_line$coefficients[1, "Pr(>|z|)"]
results_table_list[[length(results_table_list) + 1]] <- tibble(
  `Model` = "Adjusted for line of therapy",
  `HR (95% CI)` = paste0(round(hr_line, 2), " (", round(hr_line_lower, 2), "-", round(hr_line_upper, 2), ")"),
  `P-value` = ifelse(p_line < 0.001, "<0.001", as.character(round(p_line, 3))),
  `N` = nrow(outcomes_for_model)
)

# 3. Adjusted for KPS only
if (!is.null(cox_model_kps) && !is.null(cox_summary_kps)) {
  hr_kps <- cox_summary_kps$conf.int[1, "exp(coef)"]
  hr_kps_lower <- cox_summary_kps$conf.int[1, "lower .95"]
  hr_kps_upper <- cox_summary_kps$conf.int[1, "upper .95"]
  p_kps <- cox_summary_kps$coefficients[1, "Pr(>|z|)"]
  results_table_list[[length(results_table_list) + 1]] <- tibble(
    `Model` = "Adjusted for KPS",
    `HR (95% CI)` = paste0(round(hr_kps, 2), " (", round(hr_kps_lower, 2), "-", round(hr_kps_upper, 2), ")"),
    `P-value` = ifelse(p_kps < 0.001, "<0.001", as.character(round(p_kps, 3))),
    `N` = nrow(outcomes_kps)
  )
}

# 4. Adjusted for visceral metastases only
if (!is.null(cox_model_visceral) && !is.null(cox_summary_visceral)) {
  hr_visceral <- cox_summary_visceral$conf.int[1, "exp(coef)"]
  hr_visceral_lower <- cox_summary_visceral$conf.int[1, "lower .95"]
  hr_visceral_upper <- cox_summary_visceral$conf.int[1, "upper .95"]
  p_visceral <- cox_summary_visceral$coefficients[1, "Pr(>|z|)"]
  results_table_list[[length(results_table_list) + 1]] <- tibble(
    `Model` = "Adjusted for visceral metastases",
    `HR (95% CI)` = paste0(round(hr_visceral, 2), " (", round(hr_visceral_lower, 2), "-", round(hr_visceral_upper, 2), ")"),
    `P-value` = ifelse(p_visceral < 0.001, "<0.001", as.character(round(p_visceral, 3))),
    `N` = nrow(outcomes_visceral)
  )
}

# 5. Adjusted for all (line of therapy + KPS + visceral metastases)
if (!is.null(cox_model_multiv) && !is.null(cox_summary_multiv)) {
  hr_multiv <- cox_summary_multiv$conf.int[1, "exp(coef)"]
  hr_multiv_lower <- cox_summary_multiv$conf.int[1, "lower .95"]
  hr_multiv_upper <- cox_summary_multiv$conf.int[1, "upper .95"]
  p_multiv <- cox_summary_multiv$coefficients[1, "Pr(>|z|)"]
  results_table_list[[length(results_table_list) + 1]] <- tibble(
    `Model` = "Adjusted for line of therapy, KPS, and visceral metastases",
    `HR (95% CI)` = paste0(round(hr_multiv, 2), " (", round(hr_multiv_lower, 2), "-", round(hr_multiv_upper, 2), ")"),
    `P-value` = ifelse(p_multiv < 0.001, "<0.001", as.character(round(p_multiv, 3))),
    `N` = nrow(outcomes_multiv)
  )
}

# 6. Add tertile comparisons (T2 vs T1, T3 vs T1)
# Create tertiles for comparison
if (nrow(outcomes_for_model) >= 3) {
  outcomes_for_tertiles <- outcomes_for_model %>%
    filter(!is.na(ctdna_before_line_value)) %>%
    mutate(
      ctdna_tertile = ntile(ctdna_before_line_value, 3),
      ctdna_tertile_label = case_when(
        ctdna_tertile == 1 ~ "T1 (Lowest)",
        ctdna_tertile == 2 ~ "T2",
        ctdna_tertile == 3 ~ "T3 (Highest)"
      ),
      ctdna_tertile_factor = factor(ctdna_tertile_label, 
                                      levels = c("T1 (Lowest)", "T2", "T3 (Highest)"))
    )
  
  # Unadjusted tertile model
  cox_tertiles_unadj <- tryCatch({
    coxph(Surv(mfs_time, mfs_event) ~ ctdna_tertile_factor, data = outcomes_for_tertiles)
  }, error = function(e) NULL)
  
  # Adjusted tertile model (for line of therapy)
  cox_tertiles_adj <- tryCatch({
    coxph(Surv(mfs_time, mfs_event) ~ ctdna_tertile_factor + therapy_line_factor, 
          data = outcomes_for_tertiles)
  }, error = function(e) NULL)
  
  # Fully adjusted tertile model (line of therapy + KPS + visceral metastases)
  cox_tertiles_multiv <- NULL
  if (sum(!is.na(outcomes_for_tertiles$kps_binary)) > 0 && 
      sum(!is.na(outcomes_for_tertiles$visceral_mets_binary)) > 0) {
    outcomes_tertiles_multiv <- outcomes_for_tertiles %>%
      filter(!is.na(kps_binary), !is.na(visceral_mets_binary))
    
    if (nrow(outcomes_tertiles_multiv) > 0) {
      cox_tertiles_multiv <- tryCatch({
        coxph(Surv(mfs_time, mfs_event) ~ ctdna_tertile_factor + therapy_line_factor + 
              kps_factor + visceral_mets_factor, 
              data = outcomes_tertiles_multiv)
      }, error = function(e) NULL)
    }
  }
  
  if (!is.null(cox_tertiles_unadj)) {
    cox_tertiles_unadj_summary <- summary(cox_tertiles_unadj)
    
    # T2 vs T1 (unadjusted)
    if (nrow(cox_tertiles_unadj_summary$coefficients) >= 1) {
      hr_t2_unadj <- exp(cox_tertiles_unadj_summary$coefficients[1, "coef"])
      hr_t2_unadj_lower <- cox_tertiles_unadj_summary$conf.int[1, "lower .95"]
      hr_t2_unadj_upper <- cox_tertiles_unadj_summary$conf.int[1, "upper .95"]
      p_t2_unadj <- cox_tertiles_unadj_summary$coefficients[1, "Pr(>|z|)"]
      
      results_table_list[[length(results_table_list) + 1]] <- tibble(
        `Model` = "T2 vs T1 (Unadjusted)",
        `HR (95% CI)` = paste0(round(hr_t2_unadj, 2), " (", round(hr_t2_unadj_lower, 2), "-", round(hr_t2_unadj_upper, 2), ")"),
        `P-value` = ifelse(p_t2_unadj < 0.001, "<0.001", as.character(round(p_t2_unadj, 3))),
        `N` = nrow(outcomes_for_tertiles)
      )
    }
    
    # T3 vs T1 (unadjusted)
    if (nrow(cox_tertiles_unadj_summary$coefficients) >= 2) {
      hr_t3_unadj <- exp(cox_tertiles_unadj_summary$coefficients[2, "coef"])
      hr_t3_unadj_lower <- cox_tertiles_unadj_summary$conf.int[2, "lower .95"]
      hr_t3_unadj_upper <- cox_tertiles_unadj_summary$conf.int[2, "upper .95"]
      p_t3_unadj <- cox_tertiles_unadj_summary$coefficients[2, "Pr(>|z|)"]
      
      results_table_list[[length(results_table_list) + 1]] <- tibble(
        `Model` = "T3 vs T1 (Unadjusted)",
        `HR (95% CI)` = paste0(round(hr_t3_unadj, 2), " (", round(hr_t3_unadj_lower, 2), "-", round(hr_t3_unadj_upper, 2), ")"),
        `P-value` = ifelse(p_t3_unadj < 0.001, "<0.001", as.character(round(p_t3_unadj, 3))),
        `N` = nrow(outcomes_for_tertiles)
      )
    }
  }
  
  if (!is.null(cox_tertiles_adj)) {
    cox_tertiles_adj_summary <- summary(cox_tertiles_adj)
    
    # Find tertile coefficient indices
    tertile_idx <- which(str_detect(rownames(cox_tertiles_adj_summary$coefficients), "ctdna_tertile_factor"))
    
    # T2 vs T1 (adjusted)
    if (length(tertile_idx) >= 1) {
      hr_t2_adj <- exp(cox_tertiles_adj_summary$coefficients[tertile_idx[1], "coef"])
      hr_t2_adj_lower <- cox_tertiles_adj_summary$conf.int[tertile_idx[1], "lower .95"]
      hr_t2_adj_upper <- cox_tertiles_adj_summary$conf.int[tertile_idx[1], "upper .95"]
      p_t2_adj <- cox_tertiles_adj_summary$coefficients[tertile_idx[1], "Pr(>|z|)"]
      
      results_table_list[[length(results_table_list) + 1]] <- tibble(
        `Model` = "T2 vs T1 (Adjusted for line of therapy)",
        `HR (95% CI)` = paste0(round(hr_t2_adj, 2), " (", round(hr_t2_adj_lower, 2), "-", round(hr_t2_adj_upper, 2), ")"),
        `P-value` = ifelse(p_t2_adj < 0.001, "<0.001", as.character(round(p_t2_adj, 3))),
        `N` = nrow(outcomes_for_tertiles)
      )
    }
    
    # T3 vs T1 (adjusted)
    if (length(tertile_idx) >= 2) {
      hr_t3_adj <- exp(cox_tertiles_adj_summary$coefficients[tertile_idx[2], "coef"])
      hr_t3_adj_lower <- cox_tertiles_adj_summary$conf.int[tertile_idx[2], "lower .95"]
      hr_t3_adj_upper <- cox_tertiles_adj_summary$conf.int[tertile_idx[2], "upper .95"]
      p_t3_adj <- cox_tertiles_adj_summary$coefficients[tertile_idx[2], "Pr(>|z|)"]
      
      results_table_list[[length(results_table_list) + 1]] <- tibble(
        `Model` = "T3 vs T1 (Adjusted for line of therapy)",
        `HR (95% CI)` = paste0(round(hr_t3_adj, 2), " (", round(hr_t3_adj_lower, 2), "-", round(hr_t3_adj_upper, 2), ")"),
        `P-value` = ifelse(p_t3_adj < 0.001, "<0.001", as.character(round(p_t3_adj, 3))),
        `N` = nrow(outcomes_for_tertiles)
      )
    }
  }
  
  # Fully adjusted tertile models (line of therapy + KPS + visceral metastases)
  if (!is.null(cox_tertiles_multiv)) {
    # Safely check convergence
    model_converged <- tryCatch({
      isTRUE(cox_tertiles_multiv$converged)
    }, error = function(e) FALSE)
    
    if (model_converged) {
      cox_tertiles_multiv_summary <- summary(cox_tertiles_multiv)
      tertile_idx_multiv <- which(str_detect(rownames(cox_tertiles_multiv_summary$coefficients), "ctdna_tertile_factor"))
      
      # T2 vs T1 (fully adjusted)
      if (length(tertile_idx_multiv) >= 1) {
        hr_t2_multiv <- exp(cox_tertiles_multiv_summary$coefficients[tertile_idx_multiv[1], "coef"])
        hr_t2_multiv_lower <- cox_tertiles_multiv_summary$conf.int[tertile_idx_multiv[1], "lower .95"]
        hr_t2_multiv_upper <- cox_tertiles_multiv_summary$conf.int[tertile_idx_multiv[1], "upper .95"]
        p_t2_multiv <- cox_tertiles_multiv_summary$coefficients[tertile_idx_multiv[1], "Pr(>|z|)"]
        
        # Check for extreme values that suggest convergence issues
        if (!is.na(hr_t2_multiv) && is.finite(hr_t2_multiv) && hr_t2_multiv > 0 && hr_t2_multiv < 1e10 &&
            !is.na(hr_t2_multiv_lower) && is.finite(hr_t2_multiv_lower) &&
            !is.na(hr_t2_multiv_upper) && is.finite(hr_t2_multiv_upper)) {
          results_table_list[[length(results_table_list) + 1]] <- tibble(
            `Model` = "T2 vs T1 (Adjusted for line of therapy, KPS, and visceral metastases)",
            `HR (95% CI)` = paste0(round(hr_t2_multiv, 2), " (", round(hr_t2_multiv_lower, 2), "-", round(hr_t2_multiv_upper, 2), ")"),
            `P-value` = ifelse(p_t2_multiv < 0.001, "<0.001", as.character(round(p_t2_multiv, 3))),
            `N` = nrow(outcomes_tertiles_multiv)
          )
        }
      }
      
      # T3 vs T1 (fully adjusted)
      if (length(tertile_idx_multiv) >= 2) {
        hr_t3_multiv <- exp(cox_tertiles_multiv_summary$coefficients[tertile_idx_multiv[2], "coef"])
        hr_t3_multiv_lower <- cox_tertiles_multiv_summary$conf.int[tertile_idx_multiv[2], "lower .95"]
        hr_t3_multiv_upper <- cox_tertiles_multiv_summary$conf.int[tertile_idx_multiv[2], "upper .95"]
        p_t3_multiv <- cox_tertiles_multiv_summary$coefficients[tertile_idx_multiv[2], "Pr(>|z|)"]
        
        # Check for extreme values that suggest convergence issues
        if (!is.na(hr_t3_multiv) && is.finite(hr_t3_multiv) && hr_t3_multiv > 0 && hr_t3_multiv < 1e10 &&
            !is.na(hr_t3_multiv_lower) && is.finite(hr_t3_multiv_lower) &&
            !is.na(hr_t3_multiv_upper) && is.finite(hr_t3_multiv_upper)) {
          results_table_list[[length(results_table_list) + 1]] <- tibble(
            `Model` = "T3 vs T1 (Adjusted for line of therapy, KPS, and visceral metastases)",
            `HR (95% CI)` = paste0(round(hr_t3_multiv, 2), " (", round(hr_t3_multiv_lower, 2), "-", round(hr_t3_multiv_upper, 2), ")"),
            `P-value` = ifelse(p_t3_multiv < 0.001, "<0.001", as.character(round(p_t3_multiv, 3))),
            `N` = nrow(outcomes_tertiles_multiv)
          )
        }
      }
    }
  }
}

# Get therapy line labels for use in other chunks
therapy_line_labels <- c("2L vs 1L", "3L vs 1L", "4L+ vs 1L")

# Keep hr_value for backward compatibility with other chunks
hr_value <- hr_line
hr_lower <- hr_line_lower
hr_upper <- hr_line_upper
p_value_ctdna <- p_line
cox_model <- cox_model_line
cox_summary <- cox_summary_line

# Combine all results
results_table <- bind_rows(results_table_list) %>%
  kable(caption = "Cox Proportional Hazards Model: ctDNA Level Before Therapy and Duration of Complete Response. Top section shows continuous ctDNA (Ln(ctDNA + 1)) with HR progression as covariates are added. Bottom section shows tertile comparisons (T2 vs T1, T3 vs T1) with T1 (Lowest) as reference. HR > 1 indicates shorter duration of complete response (higher risk of loss of CR or death) compared to reference.") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)

results_table
```

```{r multivariate_cox_pretreatment}
# ====================================================================================
# Multivariate Cox Model: Pre-treatment Analysis with Bajorin Risk Factors
# ====================================================================================
# Model includes: ctDNA + Line of Therapy + KPS + Visceral Metastases

if (!is.null(cox_model_multiv) && exists("cox_summary_multiv") && !is.null(cox_summary_multiv)) {
  # Extract coefficients from multivariate model
  n_coef <- nrow(cox_summary_multiv$coefficients)
  
  # Create results table - build as list first, then convert to tibble
  multiv_results_list <- list()
  
  # Add ctDNA coefficient
  if (n_coef > 0) {
    p_val_ctdna <- cox_summary_multiv$coefficients[1, "Pr(>|z|)"]
    p_val_ctdna_char <- if (p_val_ctdna < 0.001) "<0.001" else as.character(round(p_val_ctdna, 3))
    multiv_results_list[[length(multiv_results_list) + 1]] <- tibble(
      `Variable` = "Ln(ctDNA + 1)",
      `HR (95% CI)` = paste0(
        round(cox_summary_multiv$conf.int[1, "exp(coef)"], 2), " (",
        round(cox_summary_multiv$conf.int[1, "lower .95"], 2), "-",
        round(cox_summary_multiv$conf.int[1, "upper .95"], 2), ")"
      ),
      `P-value` = p_val_ctdna_char
    )
  }
  
  # Add therapy line coefficients
  therapy_line_idx <- which(str_detect(rownames(cox_summary_multiv$coefficients), "therapy_line_factor"))
  if (length(therapy_line_idx) > 0) {
    for (i in seq_along(therapy_line_idx)) {
      idx <- therapy_line_idx[i]
      if (idx <= nrow(cox_summary_multiv$conf.int)) {
        p_val_line <- cox_summary_multiv$coefficients[idx, "Pr(>|z|)"]
        p_val_line_char <- if (p_val_line < 0.001) "<0.001" else as.character(round(p_val_line, 3))
        multiv_results_list[[length(multiv_results_list) + 1]] <- tibble(
          `Variable` = therapy_line_labels[i],
          `HR (95% CI)` = paste0(
            round(cox_summary_multiv$conf.int[idx, "exp(coef)"], 2), " (",
            round(cox_summary_multiv$conf.int[idx, "lower .95"], 2), "-",
            round(cox_summary_multiv$conf.int[idx, "upper .95"], 2), ")"
          ),
          `P-value` = p_val_line_char
        )
      }
    }
  }
  
  # Add KPS coefficient
  kps_idx <- which(str_detect(rownames(cox_summary_multiv$coefficients), "kps_factor"))
  if (length(kps_idx) > 0) {
    idx <- kps_idx[1]
    if (idx <= nrow(cox_summary_multiv$conf.int)) {
      p_val_kps <- cox_summary_multiv$coefficients[idx, "Pr(>|z|)"]
      p_val_kps_char <- if (p_val_kps < 0.001) "<0.001" else as.character(round(p_val_kps, 3))
      multiv_results_list[[length(multiv_results_list) + 1]] <- tibble(
        `Variable` = "KPS ≥80% vs <80%",
        `HR (95% CI)` = paste0(
          round(cox_summary_multiv$conf.int[idx, "exp(coef)"], 2), " (",
          round(cox_summary_multiv$conf.int[idx, "lower .95"], 2), "-",
          round(cox_summary_multiv$conf.int[idx, "upper .95"], 2), ")"
        ),
        `P-value` = p_val_kps_char
      )
    }
  }
  
  # Add Visceral Metastases coefficient
  visceral_idx <- which(str_detect(rownames(cox_summary_multiv$coefficients), "visceral_mets_factor"))
  if (length(visceral_idx) > 0) {
    idx <- visceral_idx[1]
    if (idx <= nrow(cox_summary_multiv$conf.int)) {
      p_val_visceral <- cox_summary_multiv$coefficients[idx, "Pr(>|z|)"]
      p_val_visceral_char <- if (p_val_visceral < 0.001) "<0.001" else as.character(round(p_val_visceral, 3))
      multiv_results_list[[length(multiv_results_list) + 1]] <- tibble(
        `Variable` = "Visceral Mets vs No Visceral Mets",
        `HR (95% CI)` = paste0(
          round(cox_summary_multiv$conf.int[idx, "exp(coef)"], 2), " (",
          round(cox_summary_multiv$conf.int[idx, "lower .95"], 2), "-",
          round(cox_summary_multiv$conf.int[idx, "upper .95"], 2), ")"
        ),
        `P-value` = p_val_visceral_char
      )
    }
  }
  
  # Combine all results
  multiv_results <- bind_rows(multiv_results_list)
  
  # Display table
  multiv_table <- multiv_results %>%
    kable(caption = "Multivariate Cox Model: Pre-treatment ctDNA and Metastasis-Free Survival (adjusted for line of therapy, KPS, and visceral metastases)") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
  
  multiv_table
  
  cat("\n**N for multivariate model:**", nrow(outcomes_multiv), "observations\n")
  cat("**Note:** This model includes Bajorin risk factors (KPS and visceral metastases) in addition to ctDNA level and line of therapy.\n")
} else {
  cat("Multivariate Cox model with Bajorin risk factors could not be fitted.\n")
  cat("This may be due to missing KPS or visceral metastases data, or insufficient sample size.\n")
  if (exists("bajorin_data")) {
    cat("Bajorin data loaded:", nrow(bajorin_data), "observations\n")
  } else {
    cat("Bajorin data file not found or could not be loaded.\n")
  }
}
```

```{r identify_clinical_variables, include=FALSE}
# Identify available clinical variables for multivariate analysis
# This helps determine what variables can be included in a multivariate Cox model

# Key variables to check for
key_vars <- c("race", "ethnicity", "cancer_stage_natera", 
              "pathology_t_stage_2", "pathology_n_stage_2", "pathology_m_stage_2",
              "neoadjuvant_systemic_therapy_yes_no", "dataset")

# Check data completeness for outcomes population
if (exists("outcomes") && nrow(outcomes) > 0) {
  outcomes_with_clinical <- outcomes %>%
    left_join(dat_combined %>% select(patient_id, any_of(key_vars)), by = "patient_id")
  
  # Check completeness for key variables
  var_completeness <- tibble()
  for (var_name in key_vars) {
    if (var_name %in% names(outcomes_with_clinical)) {
      n_complete <- sum(!is.na(outcomes_with_clinical[[var_name]]))
      pct_complete <- round(n_complete / nrow(outcomes_with_clinical) * 100, 1)
      var_completeness <- bind_rows(
        var_completeness,
        tibble(
          Variable = var_name,
          N_Complete = n_complete,
          N_Total = nrow(outcomes_with_clinical),
          Percent_Complete = pct_complete,
          Suitable = pct_complete >= 80
        )
      )
    }
  }
  
  # Also check for age, sex if they exist
  age_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("^age$|age_at", ignore_case = TRUE))]
  sex_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("^sex$|^gender$", ignore_case = TRUE))]
  
  if (length(age_cols) > 0) {
    age_var <- age_cols[1]
    outcomes_with_clinical <- outcomes_with_clinical %>%
      left_join(dat_combined %>% select(patient_id, !!sym(age_var)), by = "patient_id")
    n_complete_age <- sum(!is.na(outcomes_with_clinical[[age_var]]))
    var_completeness <- bind_rows(
      var_completeness,
      tibble(
        Variable = age_var,
        N_Complete = n_complete_age,
        N_Total = nrow(outcomes_with_clinical),
        Percent_Complete = round(n_complete_age / nrow(outcomes_with_clinical) * 100, 1),
        Suitable = n_complete_age / nrow(outcomes_with_clinical) >= 0.8
      )
    )
  }
  
  if (length(sex_cols) > 0) {
    sex_var <- sex_cols[1]
    outcomes_with_clinical <- outcomes_with_clinical %>%
      left_join(dat_combined %>% select(patient_id, !!sym(sex_var)), by = "patient_id")
    n_complete_sex <- sum(!is.na(outcomes_with_clinical[[sex_var]]))
    var_completeness <- bind_rows(
      var_completeness,
      tibble(
        Variable = sex_var,
        N_Complete = n_complete_sex,
        N_Total = nrow(outcomes_with_clinical),
        Percent_Complete = round(n_complete_sex / nrow(outcomes_with_clinical) * 100, 1),
        Suitable = n_complete_sex / nrow(outcomes_with_clinical) >= 0.8
      )
    )
  }
  
  cat("Available clinical variables for multivariate analysis:\n")
  if (nrow(var_completeness) > 0) {
    print(var_completeness %>% arrange(desc(Percent_Complete)))
  } else {
    cat("No matching clinical variables found in dataset.\n")
  }
}
```

```{r survival_curves_by_ctdna_tertiles}
# Create tertiles of ctDNA
if (nrow(outcomes) >= 3) {
  outcomes_tertiles <- outcomes %>%
    mutate(
      mfs_time_months = mfs_time / 30.44,
      ctdna_tertile = ntile(ctdna_before_line_value, 3),
      ctdna_tertile_label = case_when(
        ctdna_tertile == 1 ~ "T1 (Lowest)",
        ctdna_tertile == 2 ~ "T2",
        ctdna_tertile == 3 ~ "T3 (Highest)"
      )
    )
  
  # Check if we have enough events per tertile
  events_per_tertile <- outcomes_tertiles %>%
    group_by(ctdna_tertile_label) %>%
    summarise(n_events = sum(mfs_event, na.rm = TRUE), .groups = "drop")
  
  if (all(events_per_tertile$n_events > 0) && nrow(events_per_tertile) > 1) {
    # Fit survival curves (months)
    surv_fit <- survfit(Surv(mfs_time_months, mfs_event) ~ ctdna_tertile_label, data = outcomes_tertiles)
    
    # Plot with 95% confidence intervals
    p <- ggsurvplot(
      surv_fit,
      data = outcomes_tertiles,
      pval = TRUE,
      conf.int = TRUE,
      conf.int.style = "ribbon",
      risk.table = TRUE,
      xlab = "Time (months)",
      ylab = "Duration of Complete Response Probability",
      title = "Duration of Complete Response by ctDNA Tertile Before Therapy",
      legend.title = "ctDNA Tertile",
      legend.labs = c("T1 (Lowest)", "T2", "T3 (Highest)"),
      xlim = c(0, 36)
    )
    
    p$plot
  } else {
    cat("Insufficient events for tertile analysis. Showing overall DCR curve instead.\n")
    surv_fit_overall <- survfit(Surv(mfs_time / 30.44, mfs_event) ~ 1, data = outcomes)
    p_overall <- ggsurvplot(
      surv_fit_overall,
      data = outcomes,
      conf.int = TRUE,
      xlab = "Time (months)",
      ylab = "Duration of Complete Response Probability",
      title = "Duration of Complete Response",
      xlim = c(0, 36)
    )
    p_overall$plot
  }
} else {
  cat("Insufficient sample size (n < 3) for tertile analysis.\n")
}
```

```{r cox_model_tertiles_comparison, include=FALSE}
# Cox model comparing ctDNA tertiles for DCR
# Note: Tertile comparisons are now included in the main Cox model table above
# This chunk is kept for backward compatibility but results are shown in ctDNA_prognosis_analysis
if (nrow(outcomes) >= 3) {
  # Create tertiles if not already created
  if (!exists("outcomes_tertiles") || nrow(outcomes_tertiles) == 0) {
    outcomes_tertiles <- outcomes %>%
      mutate(
        mfs_time_months = mfs_time / 30.44,
        ctdna_tertile = ntile(ctdna_before_line_value, 3),
        ctdna_tertile_label = case_when(
          ctdna_tertile == 1 ~ "T1 (Lowest)",
          ctdna_tertile == 2 ~ "T2",
          ctdna_tertile == 3 ~ "T3 (Highest)"
        )
      )
  }
  
  if (exists("outcomes_tertiles") && nrow(outcomes_tertiles) >= 3) {
  # Fit Cox model with tertiles as factor (T1 as reference)
  outcomes_tertiles_factor <- outcomes_tertiles %>%
    mutate(
      ctdna_tertile_factor = factor(ctdna_tertile_label, 
                                      levels = c("T1 (Lowest)", "T2", "T3 (Highest)")),
      therapy_line_grouped = case_when(
        therapy_line == 1 ~ "1L",
        therapy_line == 2 ~ "2L",
        therapy_line == 3 ~ "3L",
        therapy_line >= 4 ~ "4L+",
        TRUE ~ as.character(therapy_line)
      ),
      therapy_line_factor = factor(therapy_line_grouped, levels = c("1L", "2L", "3L", "4L+"))
    )
  
  # Unadjusted model
  cox_model_tertiles_unadj <- coxph(Surv(mfs_time, mfs_event) ~ ctdna_tertile_factor, 
                                     data = outcomes_tertiles_factor)
  cox_summary_tertiles_unadj <- summary(cox_model_tertiles_unadj)
  
  # Adjusted model (for line of therapy)
  cox_model_tertiles_adj <- coxph(Surv(mfs_time, mfs_event) ~ ctdna_tertile_factor + therapy_line_factor, 
                                   data = outcomes_tertiles_factor)
  cox_summary_tertiles_adj <- summary(cox_model_tertiles_adj)
  
  # Extract HRs and CIs for each tertile vs T1 (unadjusted)
  hr_t2_unadj <- exp(cox_summary_tertiles_unadj$coefficients[1, "coef"])
  hr_t2_unadj_lower <- cox_summary_tertiles_unadj$conf.int[1, "lower .95"]
  hr_t2_unadj_upper <- cox_summary_tertiles_unadj$conf.int[1, "upper .95"]
  p_t2_unadj <- cox_summary_tertiles_unadj$coefficients[1, "Pr(>|z|)"]
  
  hr_t3_unadj <- exp(cox_summary_tertiles_unadj$coefficients[2, "coef"])
  hr_t3_unadj_lower <- cox_summary_tertiles_unadj$conf.int[2, "lower .95"]
  hr_t3_unadj_upper <- cox_summary_tertiles_unadj$conf.int[2, "upper .95"]
  p_t3_unadj <- cox_summary_tertiles_unadj$coefficients[2, "Pr(>|z|)"]
  
  # Extract HRs and CIs for each tertile vs T1 (adjusted)
  # Find indices for tertile coefficients (they come after intercept if present)
  tertile_idx_adj <- which(str_detect(rownames(cox_summary_tertiles_adj$coefficients), "ctdna_tertile_factor"))
  hr_t2_adj <- exp(cox_summary_tertiles_adj$coefficients[tertile_idx_adj[1], "coef"])
  hr_t2_adj_lower <- cox_summary_tertiles_adj$conf.int[tertile_idx_adj[1], "lower .95"]
  hr_t2_adj_upper <- cox_summary_tertiles_adj$conf.int[tertile_idx_adj[1], "upper .95"]
  p_t2_adj <- cox_summary_tertiles_adj$coefficients[tertile_idx_adj[1], "Pr(>|z|)"]
  
  hr_t3_adj <- exp(cox_summary_tertiles_adj$coefficients[tertile_idx_adj[2], "coef"])
  hr_t3_adj_lower <- cox_summary_tertiles_adj$conf.int[tertile_idx_adj[2], "lower .95"]
  hr_t3_adj_upper <- cox_summary_tertiles_adj$conf.int[tertile_idx_adj[2], "upper .95"]
  p_t3_adj <- cox_summary_tertiles_adj$coefficients[tertile_idx_adj[2], "Pr(>|z|)"]
  
  # Create comparison table
  tertile_comparisons <- tibble(
    `Comparison` = c("T2 vs T1 (Unadjusted)", "T3 vs T1 (Unadjusted)", 
                      "T2 vs T1 (Adjusted for line of therapy)", "T3 vs T1 (Adjusted for line of therapy)"),
    `HR (95% CI)` = c(
      paste0(round(hr_t2_unadj, 2), " (", round(hr_t2_unadj_lower, 2), "-", round(hr_t2_unadj_upper, 2), ")"),
      paste0(round(hr_t3_unadj, 2), " (", round(hr_t3_unadj_lower, 2), "-", round(hr_t3_unadj_upper, 2), ")"),
      paste0(round(hr_t2_adj, 2), " (", round(hr_t2_adj_lower, 2), "-", round(hr_t2_adj_upper, 2), ")"),
      paste0(round(hr_t3_adj, 2), " (", round(hr_t3_adj_lower, 2), "-", round(hr_t3_adj_upper, 2), ")")
    ),
    `P-value` = c(
      ifelse(p_t2_unadj < 0.001, "<0.001", as.character(round(p_t2_unadj, 3))),
      ifelse(p_t3_unadj < 0.001, "<0.001", as.character(round(p_t3_unadj, 3))),
      ifelse(p_t2_adj < 0.001, "<0.001", as.character(round(p_t2_adj, 3))),
      ifelse(p_t3_adj < 0.001, "<0.001", as.character(round(p_t3_adj, 3)))
    )
  )
  
  # Note: Tertile comparisons are now shown in the main Cox model table
  # This section is kept for diagnostic purposes
  cat("Tertile comparisons are included in the main Cox model table above.\n")
  
  # Overall test for trend
  cat("\n**Overall Test for Trend (Unadjusted):**\n")
  overall_p_unadj <- cox_summary_tertiles_unadj$sctest["pvalue"]
  cat("Wald test p-value:", 
      ifelse(overall_p_unadj < 0.001, 
             "<0.001", 
             round(overall_p_unadj, 3)), "\n")
  
  cat("\n**Overall Test for Trend (Adjusted for line of therapy):**\n")
  overall_p_adj <- cox_summary_tertiles_adj$sctest["pvalue"]
  cat("Wald test p-value:", 
      ifelse(overall_p_adj < 0.001, 
             "<0.001", 
             round(overall_p_adj, 3)), "\n")
  
  } else {
    cat("Insufficient data for tertile Cox model comparisons.\n")
  }
} else {
  cat("Insufficient sample size (n < 3) for tertile Cox model comparisons.\n")
}
```

```{r pairwise_tertile_comparisons}
# Pairwise comparisons between tertiles using Cox models
if (exists("outcomes_tertiles") && nrow(outcomes_tertiles) >= 3) {
  # Fit Cox model with tertiles as factor (T1 as reference)
  outcomes_tertiles_factor <- outcomes_tertiles %>%
    mutate(
      ctdna_tertile_factor = factor(ctdna_tertile_label, 
                                      levels = c("T1 (Lowest)", "T2", "T3 (Highest)"))
    )
  
  cox_model_tertiles <- coxph(Surv(mfs_time, mfs_event) ~ ctdna_tertile_factor, 
                                data = outcomes_tertiles_factor)
  cox_summary_tertiles <- summary(cox_model_tertiles)
  
  # Extract HRs and CIs for each tertile vs T1
  hr_t2 <- exp(cox_summary_tertiles$coefficients[1, "coef"])
  hr_t2_lower <- cox_summary_tertiles$conf.int[1, "lower .95"]
  hr_t2_upper <- cox_summary_tertiles$conf.int[1, "upper .95"]
  p_t2 <- cox_summary_tertiles$coefficients[1, "Pr(>|z|)"]
  
  hr_t3 <- exp(cox_summary_tertiles$coefficients[2, "coef"])
  hr_t3_lower <- cox_summary_tertiles$conf.int[2, "lower .95"]
  hr_t3_upper <- cox_summary_tertiles$conf.int[2, "upper .95"]
  p_t3 <- cox_summary_tertiles$coefficients[2, "Pr(>|z|)"]
  
  tertile_comparisons <- tibble(
    `Comparison` = c("T2 vs T1", "T3 vs T1"),
    `HR (95% CI)` = c(
      paste0(round(hr_t2, 2), " (", round(hr_t2_lower, 2), "-", round(hr_t2_upper, 2), ")"),
      paste0(round(hr_t3, 2), " (", round(hr_t3_lower, 2), "-", round(hr_t3_upper, 2), ")")
    ),
    `P-value` = c(
      ifelse(p_t2 < 0.001, "<0.001", round(p_t2, 3)),
      ifelse(p_t3 < 0.001, "<0.001", round(p_t3, 3))
    )
  )
  
  # Also do pairwise log-rank tests
  pairwise_logrank <- tryCatch(
    pairwise_survdiff(Surv(mfs_time, mfs_event) ~ ctdna_tertile_label, 
                      data = outcomes_tertiles),
    error = function(e) NULL
  )
  
  # Create pairwise comparison table with log-rank p-values
  if (!is.null(pairwise_logrank) && !is.null(pairwise_logrank$p.value)) {
    pval_matrix <- pairwise_logrank$p.value
    # Extract p-values safely - check matrix dimensions
    n_rows <- nrow(pval_matrix)
    n_cols <- ncol(pval_matrix)
    
    # Get p-values for T2, T3 vs T1 (row 1, columns 2, 3)
    pval_t2 <- if (n_rows >= 1 && n_cols >= 2 && !is.na(pval_matrix[1, 2])) {
      if (pval_matrix[1, 2] < 0.001) "<0.001" else as.character(round(pval_matrix[1, 2], 3))
    } else "N/A"
    pval_t3 <- if (n_rows >= 1 && n_cols >= 3 && !is.na(pval_matrix[1, 3])) {
      if (pval_matrix[1, 3] < 0.001) "<0.001" else as.character(round(pval_matrix[1, 3], 3))
    } else "N/A"
    
    pairwise_table <- tertile_comparisons %>%
      mutate(`Log-rank P-value` = c(pval_t2, pval_t3))
  } else {
    pairwise_table <- tertile_comparisons %>%
      mutate(`Log-rank P-value` = c("N/A", "N/A"))
  }
  
  pairwise_table_display <- pairwise_table %>%
    kable(caption = "Pairwise Comparisons: ctDNA Tertiles vs. T1 (Lowest)") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
  
  pairwise_table_display
  
  # Overall test for trend
  cat("\n**Overall Test for Trend:**\n")
  overall_p <- cox_summary_tertiles$sctest["pvalue"]
  cat("Wald test p-value:", 
      ifelse(overall_p < 0.001, 
             "<0.001", 
             round(overall_p, 3)), "\n")
  
} else {
  cat("Insufficient data for pairwise tertile comparisons.\n")
}
```

```{r sensitivity_analysis_censored_as_events}
# Sensitivity Analysis: Treat censored observations as events (worst-case scenario)
# This addresses the question: "What if all censored patients had events?"

outcomes_sensitivity <- outcomes %>%
  mutate(
    mfs_event_sensitivity = 1,  # Treat all as events
    mfs_time_sensitivity = mfs_time  # Keep same time
  )

# Cox model treating all censored as events
outcomes_sensitivity <- outcomes_sensitivity %>%
  mutate(
    therapy_line_grouped = case_when(
      therapy_line == 1 ~ "1L",
      therapy_line == 2 ~ "2L",
      therapy_line == 3 ~ "3L",
      therapy_line >= 4 ~ "4L+",
      TRUE ~ as.character(therapy_line)
    ),
    therapy_line_factor = factor(therapy_line_grouped, levels = c("1L", "2L", "3L", "4L+"))
  )

cox_model_sensitivity <- coxph(Surv(mfs_time_sensitivity, mfs_event_sensitivity) ~ log(ctdna_before_line_value + 1) + therapy_line_factor,
                                data = outcomes_sensitivity)
cox_summary_sensitivity <- summary(cox_model_sensitivity)

# Extract HR and CI for sensitivity analysis
hr_sens_value <- exp(cox_summary_sensitivity$coefficients[1, "coef"])
hr_sens_lower <- cox_summary_sensitivity$conf.int[1, "lower .95"]
hr_sens_upper <- cox_summary_sensitivity$conf.int[1, "upper .95"]

# Compare results
sensitivity_comparison <- tibble(
  `Analysis` = c("Standard (Censored)", "Sensitivity (All Events)"),
  `HR (95% CI)` = c(
    paste0(
      round(hr_value, 2),
      " (",
      round(hr_lower, 2),
      "-",
      round(hr_upper, 2),
      ")"
    ),
    paste0(
      round(hr_sens_value, 2),
      " (",
      round(hr_sens_lower, 2),
      "-",
      round(hr_sens_upper, 2),
      ")"
    )
  ),
  `P-value` = c(
    ifelse(cox_summary$coefficients[1, "Pr(>|z|)"] < 0.001, 
           "<0.001", 
           round(cox_summary$coefficients[1, "Pr(>|z|)"], 3)),
    ifelse(cox_summary_sensitivity$coefficients[1, "Pr(>|z|)"] < 0.001, 
           "<0.001", 
           round(cox_summary_sensitivity$coefficients[1, "Pr(>|z|)"], 3))
  ),
    `N Events` = c(
    sum(outcomes$mfs_event, na.rm = TRUE),
    nrow(outcomes_sensitivity)
  )
) %>%
  kable(caption = "Sensitivity Analysis: Treating Censored Observations as Events (Worst-Case Scenario)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)

sensitivity_comparison

cat("\n**Note:** This sensitivity analysis assumes all censored patients had events at their censoring time.\n")
cat("This provides a worst-case scenario to assess robustness of findings.\n")
```

```{r continuous_association_plot}
# Scatter plot: ctDNA vs PFS time (with events colored)
# Calculate median for reference line
median_log_ctdna <- log(median(outcomes$ctdna_before_line_value, na.rm = TRUE) + 1)

# Get range of log(ctDNA+1) values for better axis breaks
log_ctdna_range <- range(log(outcomes$ctdna_before_line_value + 1), na.rm = TRUE)
log_ctdna_min <- floor(log_ctdna_range[1])
log_ctdna_max <- ceiling(log_ctdna_range[2])

# Create custom breaks for X axis
x_breaks <- seq(log_ctdna_min, log_ctdna_max, by = 0.5)
# Ensure median is included if not already in breaks
if (!any(abs(x_breaks - median_log_ctdna) < 0.1)) {
  x_breaks <- sort(c(x_breaks, median_log_ctdna))
}

p_scatter <- outcomes %>%
  mutate(mfs_time_months = mfs_time / 30.44) %>%
  ggplot(aes(x = mfs_time_months, y = log(ctdna_before_line_value + 1), color = factor(mfs_event))) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  geom_hline(yintercept = median_log_ctdna, linetype = "dashed", color = "gray50", size = 0.8) +
  annotate("text", x = max(outcomes$mfs_time / 30.44, na.rm = TRUE) * 0.95, y = median_log_ctdna,
           label = paste0("Median\n", round(median_log_ctdna, 2)),
           hjust = 1, vjust = -0.1, size = 3, color = "gray30") +
  scale_x_continuous(
    expand = expansion(mult = c(0.05, 0.03)),
    limits = c(-1, NA)  # Start slightly before 0 to see dots better
  ) +
  scale_y_continuous(
    limits = c(-0.2, NA),  # Start slightly before 0 to see dots better
    breaks = x_breaks,
    labels = round(x_breaks, 2),
    expand = expansion(mult = c(0.05, 0.03))
  ) +
  scale_color_manual(values = c("0" = "blue", "1" = "red"), 
                     labels = c("0" = "Censored", "1" = "Event"),
                     name = "Status") +
  labs(
    x = "Duration of Complete Response (months)",
    y = "Ln(ctDNA + 1) Before Therapy",
    title = "Association between ctDNA Level Before Therapy and Duration of Complete Response"
  ) +
  theme_minimal() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1))

p_scatter
```

## Research Question 2

**What is the association between early on-treatment ctDNA response (cleared vs not cleared) at 6-16 weeks after therapy initiation and subsequent duration of complete response?**

This analysis examines whether early ctDNA clearance (within 6-16 weeks after starting therapy) is associated with improved duration of complete response. The landmark analysis uses the latest ctDNA measurement available within the 6-16 week window to assess clearance status, and analyzes duration of complete response from that landmark time point forward. **Note:** Only alive patients require imaging data for inclusion (to determine complete response status, similar to RECIST CR). Patients who died do NOT require imaging data (death is an event).

```{r landmark_analysis_early_response, include=FALSE}
# ====================================================================================
# LANDMARK ANALYSIS: Early On-Treatment ctDNA Response
# ====================================================================================
# This analysis examines the significance of early on-treatment ctDNA response
# at a landmark time point (e.g., 6-12 weeks after therapy start)
# 
# Approach:
# 1. For each patient-line observation, find ctDNA before therapy (baseline)
# 2. Find LATEST on-treatment ctDNA within landmark window (6-16 weeks) for response assessment
# 3. Determine clearance status (binary: cleared vs not cleared) at landmark time
# 4. Perform landmark analysis: only include patients who survived to landmark
# 5. Compare survival from landmark time point between cleared and not cleared groups
# ====================================================================================

# Define landmark window: 6-16 weeks after therapy start
landmark_window_start_weeks <- 6
landmark_window_end_weeks <- 16
landmark_window_start_days <- landmark_window_start_weeks * 7
landmark_window_end_days <- landmark_window_end_weeks * 7

# For each patient-line observation, get baseline ctDNA (already have this from universe)
# and find on-treatment ctDNA within landmark window
landmark_data <- outcomes %>%
  select(patient_id, therapy_line, time_zero_date, ctdna_before_line_date, 
         ctdna_before_line_value, mfs_time, mfs_event) %>%
  # Get on-treatment ctDNA measurements
  left_join(
    ctdna_long %>%
      select(patient_id, ctdna_date, ctdna_value),
    by = "patient_id"
  ) %>%
  mutate(
    # Calculate days from therapy start to on-treatment ctDNA
    days_from_therapy_start = as.numeric(ctdna_date - time_zero_date),
    # Check if on-treatment ctDNA is within landmark window (6-16 weeks)
    in_landmark_window = days_from_therapy_start >= landmark_window_start_days & 
                         days_from_therapy_start <= landmark_window_end_days
  ) %>%
  filter(in_landmark_window) %>%
  # For each patient-line, take the LATEST on-treatment ctDNA in the window (for response assessment)
  arrange(patient_id, therapy_line, desc(days_from_therapy_start)) %>%
  group_by(patient_id, therapy_line) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  mutate(
    # Calculate % decline from baseline
    baseline_ctdna = ctdna_before_line_value,
    on_tx_ctdna = ctdna_value,
    # % decline = (baseline - on_tx) / baseline * 100
    # Handle case where baseline = 0 (use absolute change instead)
    pct_decline = case_when(
      baseline_ctdna > 0 ~ ((baseline_ctdna - on_tx_ctdna) / baseline_ctdna) * 100,
      baseline_ctdna == 0 & on_tx_ctdna == 0 ~ 0,  # Both 0 = no change
      baseline_ctdna == 0 & on_tx_ctdna > 0 ~ -Inf,  # New positive = increase
      TRUE ~ NA_real_
    ),
    # Binary clearance: ctDNA cleared at landmark time (on_tx = 0 and baseline > 0)
    ctdna_cleared_at_landmark = case_when(
      baseline_ctdna > 0 & on_tx_ctdna == 0 ~ TRUE,
      baseline_ctdna > 0 & on_tx_ctdna > 0 ~ FALSE,
      baseline_ctdna == 0 ~ NA,  # Can't clear if already negative
      TRUE ~ NA
    ),
    # Response categories based on % decline
    response_category = case_when(
      ctdna_cleared_at_landmark == TRUE ~ "Cleared (100% decline)",
      pct_decline >= 50 ~ "Major Response (≥50% decline)",
      pct_decline >= 0 & pct_decline < 50 ~ "Minor Response (0-50% decline)",
      pct_decline < 0 ~ "Progression (increase)",
      TRUE ~ "Unknown"
    ),
    # Landmark time point: use the actual ctDNA measurement time within 6-16 weeks
    landmark_time = days_from_therapy_start
  ) %>%
  select(patient_id, therapy_line, time_zero_date, landmark_time,
         baseline_ctdna, on_tx_ctdna, pct_decline, ctdna_cleared_at_landmark, response_category)

# Now perform landmark analysis: only include patients who survived to landmark
# and calculate survival from landmark time point
landmark_outcomes <- landmark_data %>%
  left_join(
    outcomes %>% 
      select(patient_id, therapy_line, mfs_time, mfs_event, time_zero_date, imaging_disease_present),
    by = c("patient_id", "therapy_line", "time_zero_date")
  ) %>%
  left_join(
    dat_combined %>%
      select(patient_id, all_of(c(death_date_cols, alive_cols))),
    by = "patient_id"
  ) %>%
  mutate(
    death_date = if (length(death_date_cols) > 0) {
      as_date_safe(.data[[death_date_cols[1]]])
    } else {
      as.Date(NA)
    },
    alive_status = if (length(alive_cols) > 0) {
      case_when(
        str_detect(tolower(as.character(.data[[alive_cols[1]]])), "yes|y|1") ~ "Yes",
        str_detect(tolower(as.character(.data[[alive_cols[1]]])), "no|n|0") ~ "No",
        TRUE ~ NA_character_
      )
    } else {
      NA_character_
    },
    # Only include patients who survived to landmark (mfs_time >= landmark_time)
    survived_to_landmark = mfs_time >= landmark_time,
    # Survival time from landmark = mfs_time - landmark_time
    mfs_time_from_landmark = mfs_time - landmark_time,
    # Event status remains the same (if they had event, it's still an event)
    mfs_event_from_landmark = mfs_event,
    # Determine if imaging data is required (only for alive patients)
    imaging_required = case_when(
      !is.na(death_date) ~ FALSE,  # Dead patients don't need imaging
      alive_status == "No" ~ FALSE,  # Deceased patients don't need imaging
      alive_status == "Yes" ~ TRUE,  # Alive patients need imaging
      TRUE ~ TRUE  # Default: require imaging if status unknown
    )
  ) %>%
  # Filter: require valid survival time
  # For imaging: only exclude if patient is alive AND imaging is missing
  filter(survived_to_landmark, mfs_time_from_landmark > 0,
         !(imaging_required == TRUE & is.na(imaging_disease_present)))
  
# Track overall landmark events to enable fallback plotting if no post-landmark events by group
landmark_has_events <- sum(landmark_outcomes$mfs_event_from_landmark, na.rm = TRUE) > 0

cat("Landmark analysis: Early on-treatment ctDNA response\n")
cat("Landmark window: 6-16 weeks after therapy start\n")
cat("Response assessment: Uses LATEST ctDNA measurement available within 6-16 week window\n")
cat("Patients with on-treatment ctDNA in landmark window:", nrow(landmark_data), "\n")
cat("Patients who survived to their landmark time:", nrow(landmark_outcomes), "\n")
if (exists("landmark_outcomes") && nrow(landmark_outcomes) > 0) {
  cat("Landmark time distribution (weeks from therapy start):\n")
  cat("  Min:", round(min(landmark_outcomes$landmark_time, na.rm = TRUE) / 7, 1), 
      "Median:", round(median(landmark_outcomes$landmark_time, na.rm = TRUE) / 7, 1),
      "Max:", round(max(landmark_outcomes$landmark_time, na.rm = TRUE) / 7, 1), "\n")
  cat("MFS time distribution (days from therapy start):\n")
  cat("  Min:", round(min(landmark_outcomes$mfs_time, na.rm = TRUE), 1),
      "Median:", round(median(landmark_outcomes$mfs_time, na.rm = TRUE), 1),
      "Max:", round(max(landmark_outcomes$mfs_time, na.rm = TRUE), 1), "\n")
  cat("MFS time from landmark distribution (days):\n")
  cat("  Min:", round(min(landmark_outcomes$mfs_time_from_landmark, na.rm = TRUE), 1),
      "Median:", round(median(landmark_outcomes$mfs_time_from_landmark, na.rm = TRUE), 1),
      "Max:", round(max(landmark_outcomes$mfs_time_from_landmark, na.rm = TRUE), 1), "\n")
  cat("Total events after landmark:", sum(landmark_outcomes$mfs_event_from_landmark, na.rm = TRUE), "\n")
  cat("Events before landmark (excluded):", sum(landmark_outcomes$mfs_time < landmark_outcomes$landmark_time, na.rm = TRUE), "\n")
}
```

```{r immortal_time_bias_explanation}
# Explanation of how landmark analysis addresses immortal time bias
cat("## Addressing Immortal Time Bias\n\n")
cat("**Immortal time bias** occurs when patients must survive to a certain time to be classified into a group, artificially inflating survival. **Landmark analysis addresses this** by: (1) determining clearance status at the landmark time point (ctDNA measurement date within 6-16 weeks), not requiring survival to a later time; (2) analyzing survival FROM the landmark forward (time 0 = landmark time, not therapy start); and (3) including only patients who survived to landmark, ensuring equal opportunity for classification.\n\n")
```

```{r landmark_time_distribution_table}
# Show landmark time distribution by clearance status to demonstrate similar landmark times
if (exists("landmark_outcomes") && nrow(landmark_outcomes) > 0) {
  landmark_time_by_clearance <- landmark_outcomes %>%
    filter(!is.na(ctdna_cleared_at_landmark)) %>%
    mutate(landmark_time_weeks = landmark_time / 7) %>%
    group_by(ctdna_cleared_at_landmark) %>%
    summarise(
      n = n(),
      median_landmark_weeks = round(median(landmark_time_weeks, na.rm = TRUE), 1),
      iqr_lower = round(quantile(landmark_time_weeks, 0.25, na.rm = TRUE), 1),
      iqr_upper = round(quantile(landmark_time_weeks, 0.75, na.rm = TRUE), 1),
      .groups = "drop"
    ) %>%
    mutate(
      clearance_status = ifelse(ctdna_cleared_at_landmark, "Cleared at Landmark", "Not Cleared at Landmark"),
      landmark_time_iqr = paste0(median_landmark_weeks, " (", iqr_lower, "-", iqr_upper, ")")
    ) %>%
    select(`Clearance Status` = clearance_status, `N` = n, `Landmark Time (weeks, median IQR)` = landmark_time_iqr)
  
  if (nrow(landmark_time_by_clearance) > 0) {
    landmark_time_by_clearance %>%
      kable(caption = "Landmark Time Distribution (weeks) by ctDNA Clearance Status (demonstrating similar landmark times across groups)") %>%
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
  } else {
    cat("Insufficient data to create landmark time distribution table.\n")
  }
} else {
  cat("Landmark outcomes data not available.\n")
}
```

```{r landmark_analysis_results}
# Display landmark analysis results
if (exists("landmark_outcomes") && nrow(landmark_outcomes) > 0) {
  # Summary of response categories
  response_summary_table <- landmark_outcomes %>%
    count(response_category) %>%
    mutate(Percentage = round(n / sum(n) * 100, 1)) %>%
    rename(`Response Category` = response_category, `N` = n) %>%
    kable(caption = "Distribution of Early On-Treatment ctDNA Response (Landmark = ctDNA between 6-16 weeks after start)") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
  
  response_summary_table
  
  # Initialize variables
  hr_cleared <- hr_cleared_lower <- hr_cleared_upper <- p_cleared <- NA
  landmark_outcomes_cleared <- NULL
  landmark_results <- NULL
  
  # Create results table - unadjusted and adjusted models
  table_rows <- list()
  
  # Cox model: ctDNA Cleared at Landmark vs Not Cleared
  if (sum(!is.na(landmark_outcomes$ctdna_cleared_at_landmark)) > 10) {
    landmark_outcomes_cleared <- landmark_outcomes %>%
      filter(!is.na(ctdna_cleared_at_landmark)) %>%
      mutate(
        therapy_line_grouped = case_when(
          therapy_line == 1 ~ "1L",
          therapy_line == 2 ~ "2L",
          therapy_line == 3 ~ "3L",
          therapy_line >= 4 ~ "4L+",
          TRUE ~ as.character(therapy_line)
        ),
        therapy_line_factor = factor(therapy_line_grouped, levels = c("1L", "2L", "3L", "4L+"))
      )
    
    # Unadjusted Cox model (matches log-rank test from survival curve)
    cox_cleared_unadj <- tryCatch({
      coxph(Surv(mfs_time_from_landmark, mfs_event_from_landmark) ~ ctdna_cleared_at_landmark,
            data = landmark_outcomes_cleared)
    }, error = function(e) NULL)
    
    # Join therapy class if available
    if (exists("palliative_regimen_table") && nrow(palliative_regimen_table) > 0) {
      landmark_outcomes_cleared <- landmark_outcomes_cleared %>%
        left_join(
          palliative_regimen_table %>%
            mutate(therapy_line_match = ifelse(is.na(therapy_line_match), NA_integer_, therapy_line_match)) %>%
            select(patient_id, therapy_line_match, therapy_class) %>%
            distinct(),
          by = c("patient_id", "therapy_line" = "therapy_line_match")
        ) %>%
        mutate(
          therapy_class = replace_na(therapy_class, "unspecified"),
          therapy_class_factor = factor(therapy_class)
        )
    } else {
      landmark_outcomes_cleared <- landmark_outcomes_cleared %>%
        mutate(
          therapy_class = "unspecified",
          therapy_class_factor = factor("unspecified")
        )
    }
    
    # Join Bajorin risk factors if available
    if (exists("bajorin_data") && nrow(bajorin_data) > 0) {
      landmark_outcomes_cleared <- landmark_outcomes_cleared %>%
        left_join(
          bajorin_data %>% select(patient_id, therapy_line, kps_binary, visceral_mets_binary),
          by = c("patient_id", "therapy_line")
        ) %>%
        mutate(
          kps_factor = factor(kps_binary, levels = c(0, 1), labels = c("KPS <80%", "KPS ≥80%")),
          visceral_mets_factor = factor(visceral_mets_binary, levels = c(0, 1), labels = c("No Visceral Mets", "Visceral Mets"))
        )
    } else {
      landmark_outcomes_cleared <- landmark_outcomes_cleared %>%
        mutate(
          kps_binary = NA_real_,
          visceral_mets_binary = NA_real_,
          kps_factor = factor(NA, levels = c(0, 1), labels = c("KPS <80%", "KPS ≥80%")),
          visceral_mets_factor = factor(NA, levels = c(0, 1), labels = c("No Visceral Mets", "Visceral Mets"))
        )
    }
    
    # Adjusted Cox model (adjusted for line of therapy and therapy class)
    cox_cleared <- tryCatch({
      coxph(Surv(mfs_time_from_landmark, mfs_event_from_landmark) ~ ctdna_cleared_at_landmark + therapy_line_factor + therapy_class_factor,
            data = landmark_outcomes_cleared)
    }, error = function(e) NULL)
    
    # Multivariate Cox model (adjusted for line of therapy + Bajorin risk factors)
    if (sum(!is.na(landmark_outcomes_cleared$kps_binary)) > 0 && 
        sum(!is.na(landmark_outcomes_cleared$visceral_mets_binary)) > 0) {
      landmark_outcomes_multiv <- landmark_outcomes_cleared %>%
        filter(!is.na(kps_binary), !is.na(visceral_mets_binary))
      
      if (nrow(landmark_outcomes_multiv) > 0) {
        cox_cleared_multiv <- tryCatch({
          coxph(Surv(mfs_time_from_landmark, mfs_event_from_landmark) ~ 
                ctdna_cleared_at_landmark + therapy_line_factor + therapy_class_factor + kps_factor + visceral_mets_factor,
                data = landmark_outcomes_multiv)
        }, error = function(e) NULL)
      } else {
        cox_cleared_multiv <- NULL
      }
    } else {
      cox_cleared_multiv <- NULL
    }
    
    # Add unadjusted model for comparison (should match log-rank test) - FIRST
    if (!is.null(cox_cleared_unadj)) {
      model_converged_unadj <- tryCatch({
        isTRUE(cox_cleared_unadj$converged)
      }, error = function(e) FALSE)
      
      if (model_converged_unadj) {
        cox_cleared_unadj_summary <- summary(cox_cleared_unadj)
        hr_cleared_unadj <- cox_cleared_unadj_summary$conf.int[1, "exp(coef)"]
        hr_cleared_unadj_lower <- cox_cleared_unadj_summary$conf.int[1, "lower .95"]
        hr_cleared_unadj_upper <- cox_cleared_unadj_summary$conf.int[1, "upper .95"]
        p_cleared_unadj <- cox_cleared_unadj_summary$coefficients[1, "Pr(>|z|)"]
        
        # Check for reasonable HR values (not 0, Inf, or extremely large)
        if (!is.na(hr_cleared_unadj) && is.finite(hr_cleared_unadj) && hr_cleared_unadj > 0 && hr_cleared_unadj < 1e10 &&
            !is.na(hr_cleared_unadj_lower) && is.finite(hr_cleared_unadj_lower) &&
            !is.na(hr_cleared_unadj_upper) && is.finite(hr_cleared_unadj_upper)) {
          table_rows[[length(table_rows) + 1]] <- tibble(
            `Variable` = "ctDNA Cleared at Landmark vs Not Cleared (Unadjusted)",
            `HR (95% CI)` = paste0(round(hr_cleared_unadj, 2), " (", round(hr_cleared_unadj_lower, 2), "-", round(hr_cleared_unadj_upper, 2), ")"),
            `P-value` = ifelse(p_cleared_unadj < 0.001, "<0.001", as.character(round(p_cleared_unadj, 3))),
            `N` = nrow(landmark_outcomes_cleared)
          )
        }
      }
    }
    
    if (!is.null(cox_cleared)) {
      model_converged_adj <- tryCatch({
        isTRUE(cox_cleared$converged)
      }, error = function(e) FALSE)
      
      if (model_converged_adj) {
        cox_cleared_summary <- summary(cox_cleared)
        hr_cleared <- cox_cleared_summary$conf.int[1, "exp(coef)"]
        hr_cleared_lower <- cox_cleared_summary$conf.int[1, "lower .95"]
        hr_cleared_upper <- cox_cleared_summary$conf.int[1, "upper .95"]
        p_cleared <- cox_cleared_summary$coefficients[1, "Pr(>|z|)"]
        
        # Check for reasonable HR values
        if (!is.na(hr_cleared) && is.finite(hr_cleared) && hr_cleared > 0 && hr_cleared < 1e10 &&
            !is.na(hr_cleared_lower) && is.finite(hr_cleared_lower) &&
            !is.na(hr_cleared_upper) && is.finite(hr_cleared_upper)) {
          table_rows[[length(table_rows) + 1]] <- tibble(
            `Variable` = "ctDNA Cleared at Landmark vs Not Cleared (Adjusted for line of therapy and therapy class)",
            `HR (95% CI)` = paste0(round(hr_cleared, 2), " (", round(hr_cleared_lower, 2), "-", round(hr_cleared_upper, 2), ")"),
            `P-value` = ifelse(p_cleared < 0.001, "<0.001", as.character(round(p_cleared, 3))),
            `N` = nrow(landmark_outcomes_cleared)
          )
        }
      }
    }
    
    # Add multivariate model (adjusted for line of therapy + Bajorin risk factors) - LAST
    if (!is.null(cox_cleared_multiv)) {
      model_converged <- tryCatch({
        isTRUE(cox_cleared_multiv$converged)
      }, error = function(e) FALSE)
      
      if (model_converged) {
        cox_cleared_multiv_summary <- summary(cox_cleared_multiv)
        hr_cleared_multiv <- cox_cleared_multiv_summary$conf.int[1, "exp(coef)"]
        hr_cleared_multiv_lower <- cox_cleared_multiv_summary$conf.int[1, "lower .95"]
        hr_cleared_multiv_upper <- cox_cleared_multiv_summary$conf.int[1, "upper .95"]
        p_cleared_multiv <- cox_cleared_multiv_summary$coefficients[1, "Pr(>|z|)"]
        
        # Check for reasonable HR values (not 0, Inf, or extremely large)
        if (!is.na(hr_cleared_multiv) && is.finite(hr_cleared_multiv) && hr_cleared_multiv > 0 && hr_cleared_multiv < 1e10 &&
            !is.na(hr_cleared_multiv_lower) && is.finite(hr_cleared_multiv_lower) &&
            !is.na(hr_cleared_multiv_upper) && is.finite(hr_cleared_multiv_upper)) {
          table_rows[[length(table_rows) + 1]] <- tibble(
            `Variable` = "ctDNA Cleared at Landmark vs Not Cleared (Adjusted for line of therapy, therapy class, KPS, and visceral metastases)",
            `HR (95% CI)` = paste0(round(hr_cleared_multiv, 2), " (", round(hr_cleared_multiv_lower, 2), "-", round(hr_cleared_multiv_upper, 2), ")"),
            `P-value` = ifelse(p_cleared_multiv < 0.001, "<0.001", as.character(round(p_cleared_multiv, 3))),
            `N` = nrow(landmark_outcomes_multiv)
          )
        } else {
          # Model converged but HR values are extreme - likely due to small sample size or separation
          cat("Note: Multivariate model converged but HR values are extreme (likely due to small sample size or complete separation). Model not included in results table.\n")
        }
      } else {
        cat("Note: Multivariate model did not converge (likely due to small sample size or too many covariates). Model not included in results table.\n")
      }
    }
  }
  
  # Create table if we have any valid models
  if (length(table_rows) > 0) {
    landmark_results <- bind_rows(table_rows) %>%
      kable(caption = "Landmark Analysis: Early On-Treatment ctDNA Response and Duration of Complete Response from Landmark (ctDNA between 6-16 weeks). Unadjusted model p-value should match log-rank test from survival curve; adjusted models account for line of therapy, therapy class (ADC, ICI, Chemo, Targeted, Other, and combinations), and Bajorin risk factors (KPS and visceral metastases).") %>%
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
  } else {
    landmark_results <- NULL
  }
  
  # Create detailed multivariate results table for on-treatment analysis
  # Only create if model converged and has reasonable HR values
  if (!is.null(cox_cleared_multiv)) {
    model_converged <- tryCatch({
      isTRUE(cox_cleared_multiv$converged)
    }, error = function(e) FALSE)
    
    if (model_converged) {
      cox_cleared_multiv_summary <- summary(cox_cleared_multiv)
      n_coef <- nrow(cox_cleared_multiv_summary$coefficients)
      
      # Build results as list first
      multiv_landmark_results_list <- list()
      
      # Add ctDNA cleared coefficient
      if (n_coef > 0) {
        hr_val_cleared <- cox_cleared_multiv_summary$conf.int[1, "exp(coef)"]
        hr_lower_cleared <- cox_cleared_multiv_summary$conf.int[1, "lower .95"]
        hr_upper_cleared <- cox_cleared_multiv_summary$conf.int[1, "upper .95"]
        p_val_cleared <- cox_cleared_multiv_summary$coefficients[1, "Pr(>|z|)"]
        
        # Only add if HR values are reasonable
        if (!is.na(hr_val_cleared) && is.finite(hr_val_cleared) && hr_val_cleared > 0 && hr_val_cleared < 1e10 &&
            !is.na(hr_lower_cleared) && is.finite(hr_lower_cleared) &&
            !is.na(hr_upper_cleared) && is.finite(hr_upper_cleared)) {
          multiv_landmark_results_list[[length(multiv_landmark_results_list) + 1]] <- tibble(
            `Variable` = "ctDNA Cleared at Landmark vs Not Cleared",
            `HR (95% CI)` = paste0(
              round(hr_val_cleared, 2), " (",
              round(hr_lower_cleared, 2), "-",
              round(hr_upper_cleared, 2), ")"
            ),
            `P-value` = if (p_val_cleared < 0.001) "<0.001" else as.character(round(p_val_cleared, 3))
          )
        }
      }
      
      # Add therapy line coefficients (with convergence checks)
      therapy_line_idx <- which(str_detect(rownames(cox_cleared_multiv_summary$coefficients), "therapy_line_factor"))
      if (length(therapy_line_idx) > 0) {
        for (i in seq_along(therapy_line_idx)) {
          idx <- therapy_line_idx[i]
          if (idx <= nrow(cox_cleared_multiv_summary$conf.int) && i <= length(therapy_line_labels)) {
            hr_val_line <- cox_cleared_multiv_summary$conf.int[idx, "exp(coef)"]
            hr_lower_line <- cox_cleared_multiv_summary$conf.int[idx, "lower .95"]
            hr_upper_line <- cox_cleared_multiv_summary$conf.int[idx, "upper .95"]
            p_val_line <- cox_cleared_multiv_summary$coefficients[idx, "Pr(>|z|)"]
            
            # Only add if HR values are reasonable
            if (!is.na(hr_val_line) && is.finite(hr_val_line) && hr_val_line > 0 && hr_val_line < 1e10 &&
                !is.na(hr_lower_line) && is.finite(hr_lower_line) &&
                !is.na(hr_upper_line) && is.finite(hr_upper_line)) {
              multiv_landmark_results_list[[length(multiv_landmark_results_list) + 1]] <- tibble(
                `Variable` = therapy_line_labels[i],
                `HR (95% CI)` = paste0(
                  round(hr_val_line, 2), " (",
                  round(hr_lower_line, 2), "-",
                  round(hr_upper_line, 2), ")"
                ),
                `P-value` = as.character(ifelse(p_val_line < 0.001, "<0.001", round(p_val_line, 3)))
              )
            }
          }
        }
      }
      
      # Add KPS coefficient (with convergence checks)
      kps_idx <- which(str_detect(rownames(cox_cleared_multiv_summary$coefficients), "kps_factor"))
      if (length(kps_idx) > 0) {
        idx <- kps_idx[1]
        if (idx <= nrow(cox_cleared_multiv_summary$conf.int)) {
          hr_val_kps <- cox_cleared_multiv_summary$conf.int[idx, "exp(coef)"]
          hr_lower_kps <- cox_cleared_multiv_summary$conf.int[idx, "lower .95"]
          hr_upper_kps <- cox_cleared_multiv_summary$conf.int[idx, "upper .95"]
          p_val_kps <- cox_cleared_multiv_summary$coefficients[idx, "Pr(>|z|)"]
          
          # Only add if HR values are reasonable
          if (!is.na(hr_val_kps) && is.finite(hr_val_kps) && hr_val_kps > 0 && hr_val_kps < 1e10 &&
              !is.na(hr_lower_kps) && is.finite(hr_lower_kps) &&
              !is.na(hr_upper_kps) && is.finite(hr_upper_kps)) {
            multiv_landmark_results_list[[length(multiv_landmark_results_list) + 1]] <- tibble(
              `Variable` = "KPS ≥80% vs <80%",
              `HR (95% CI)` = paste0(
                round(hr_val_kps, 2), " (",
                round(hr_lower_kps, 2), "-",
                round(hr_upper_kps, 2), ")"
              ),
              `P-value` = if (p_val_kps < 0.001) "<0.001" else as.character(round(p_val_kps, 3))
            )
          }
        }
      }
      
      # Add Visceral Metastases coefficient (with convergence checks)
      visceral_idx <- which(str_detect(rownames(cox_cleared_multiv_summary$coefficients), "visceral_mets_factor"))
      if (length(visceral_idx) > 0) {
        idx <- visceral_idx[1]
        if (idx <= nrow(cox_cleared_multiv_summary$conf.int)) {
          hr_val_visceral <- cox_cleared_multiv_summary$conf.int[idx, "exp(coef)"]
          hr_lower_visceral <- cox_cleared_multiv_summary$conf.int[idx, "lower .95"]
          hr_upper_visceral <- cox_cleared_multiv_summary$conf.int[idx, "upper .95"]
          p_val_visceral <- cox_cleared_multiv_summary$coefficients[idx, "Pr(>|z|)"]
          
          # Only add if HR values are reasonable
          if (!is.na(hr_val_visceral) && is.finite(hr_val_visceral) && hr_val_visceral > 0 && hr_val_visceral < 1e10 &&
              !is.na(hr_lower_visceral) && is.finite(hr_lower_visceral) &&
              !is.na(hr_upper_visceral) && is.finite(hr_upper_visceral)) {
            multiv_landmark_results_list[[length(multiv_landmark_results_list) + 1]] <- tibble(
              `Variable` = "Visceral Mets vs No Visceral Mets",
              `HR (95% CI)` = paste0(
                round(hr_val_visceral, 2), " (",
                round(hr_lower_visceral, 2), "-",
                round(hr_upper_visceral, 2), ")"
              ),
              `P-value` = if (p_val_visceral < 0.001) "<0.001" else as.character(round(p_val_visceral, 3))
            )
          }
        }
      }
      
      # Add Therapy Class coefficients (with convergence checks)
      therapy_class_idx <- which(str_detect(rownames(cox_cleared_multiv_summary$coefficients), "therapy_class_factor"))
      if (length(therapy_class_idx) > 0) {
        for (i in seq_along(therapy_class_idx)) {
          idx <- therapy_class_idx[i]
          if (idx <= nrow(cox_cleared_multiv_summary$conf.int)) {
            term_name <- rownames(cox_cleared_multiv_summary$coefficients)[idx]
            class_name <- str_replace(term_name, "therapy_class_factor", "")
            p_val_class <- cox_cleared_multiv_summary$coefficients[idx, "Pr(>|z|)"]
            hr_val <- cox_cleared_multiv_summary$conf.int[idx, "exp(coef)"]
            ci_lower <- cox_cleared_multiv_summary$conf.int[idx, "lower .95"]
            ci_upper <- cox_cleared_multiv_summary$conf.int[idx, "upper .95"]
            
            # Only add if HR values are reasonable
            if (!is.na(hr_val) && is.finite(hr_val) && hr_val > 0 && hr_val < 1e10 &&
                !is.na(ci_lower) && is.finite(ci_lower) &&
                !is.na(ci_upper) && is.finite(ci_upper)) {
              p_val_text <- if (is.na(p_val_class)) {
                "NA"
              } else if (p_val_class < 0.001) {
                "<0.001"
              } else {
                as.character(round(p_val_class, 3))
              }
              
              multiv_landmark_results_list[[length(multiv_landmark_results_list) + 1]] <- tibble(
                `Variable` = paste0("Therapy Class: ", class_name, " vs reference"),
                `HR (95% CI)` = paste0(
                  round(hr_val, 2), " (",
                  round(ci_lower, 2), "-",
                  round(ci_upper, 2), ")"
                ),
                `P-value` = p_val_text
              )
            }
          }
        }
      }
      
      # Combine all results - only create table if we have at least one row
      if (length(multiv_landmark_results_list) > 0) {
        multiv_landmark_results <- bind_rows(multiv_landmark_results_list)
        
        # Store for display
        landmark_multiv_table <- multiv_landmark_results %>%
          kable(caption = "Multivariate Cox Model: On-Treatment ctDNA Clearance and Duration of Complete Response from Landmark (adjusted for line of therapy, therapy class, KPS, and visceral metastases). Note: Only coefficients with reasonable HR values (converged models) are shown. Models with extreme HR values (0, Inf, or very large) are excluded due to convergence issues, likely due to small sample size or complete separation.") %>%
          kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
      } else {
        landmark_multiv_table <- NULL
      }
    } else {
      landmark_multiv_table <- NULL
    }
  } else {
    landmark_multiv_table <- NULL
  }
  
  # Ensure table variables are available for display chunks
  # (Tables will be displayed in separate chunks below)
} else {
  # If landmark_outcomes doesn't exist, set table to NULL
  landmark_multiv_table <- NULL
}
```

```{r landmark_results_table_display, results='asis', echo=FALSE}
# Explicitly display landmark results table
# Check if table exists and display it
if (exists("landmark_results") && !is.null(landmark_results)) {
  # Display the table - kable objects render when referenced
  landmark_results
} else {
  cat("Landmark analysis results table not available.\n")
}
```

```{r landmark_results_note, echo=FALSE}
cat("\n**Note on therapy class adjustment:** Palliative systemic therapy regimens were classified into therapy classes (ADC, ICI, Chemo, Targeted, Other, and combinations such as ADC + ICI) based on the regimen received. Therapy class was included as a categorical covariate in the adjusted Cox models to account for potential differences in treatment efficacy across therapy types.\n")
```

```{r landmark_multiv_table_display, results='asis', echo=FALSE}
# Display detailed multivariate model table for on-treatment analysis
# Use the table that was already created in landmark_analysis_results chunk

# First, try to use the existing table
if (exists("landmark_multiv_table") && !is.null(landmark_multiv_table)) {
  # Display the table - kable objects render when referenced
  landmark_multiv_table
} else if (exists("cox_cleared_multiv") && !is.null(cox_cleared_multiv)) {
  # Try to recreate if it doesn't exist (fallback)
  if (exists("cox_cleared_multiv") && !is.null(cox_cleared_multiv)) {
    # Recreate the table using the same logic as landmark_analysis_results
    cox_cleared_multiv_summary <- summary(cox_cleared_multiv)
    n_coef <- nrow(cox_cleared_multiv_summary$coefficients)
    
    multiv_landmark_results_list <- list()
    
    # Add ctDNA cleared coefficient
    if (n_coef > 0) {
      p_val_cleared <- cox_cleared_multiv_summary$coefficients[1, "Pr(>|z|)"]
      multiv_landmark_results_list[[length(multiv_landmark_results_list) + 1]] <- tibble(
        `Variable` = "ctDNA Cleared at Landmark vs Not Cleared",
        `HR (95% CI)` = paste0(
          round(cox_cleared_multiv_summary$conf.int[1, "exp(coef)"], 2), " (",
          round(cox_cleared_multiv_summary$conf.int[1, "lower .95"], 2), "-",
          round(cox_cleared_multiv_summary$conf.int[1, "upper .95"], 2), ")"
        ),
        `P-value` = if (p_val_cleared < 0.001) "<0.001" else as.character(round(p_val_cleared, 3))
      )
    }
    
    # Add therapy line coefficients
    if (exists("therapy_line_labels")) {
      therapy_line_idx <- which(str_detect(rownames(cox_cleared_multiv_summary$coefficients), "therapy_line_factor"))
      if (length(therapy_line_idx) > 0) {
        for (i in seq_along(therapy_line_idx)) {
          idx <- therapy_line_idx[i]
          if (idx <= nrow(cox_cleared_multiv_summary$conf.int) && i <= length(therapy_line_labels)) {
            p_val_line <- cox_cleared_multiv_summary$coefficients[idx, "Pr(>|z|)"]
            multiv_landmark_results_list[[length(multiv_landmark_results_list) + 1]] <- tibble(
              `Variable` = therapy_line_labels[i],
              `HR (95% CI)` = paste0(
                round(cox_cleared_multiv_summary$conf.int[idx, "exp(coef)"], 2), " (",
                round(cox_cleared_multiv_summary$conf.int[idx, "lower .95"], 2), "-",
                round(cox_cleared_multiv_summary$conf.int[idx, "upper .95"], 2), ")"
              ),
              `P-value` = as.character(ifelse(p_val_line < 0.001, "<0.001", round(p_val_line, 3)))
            )
          }
        }
      }
    }
    
    # Add KPS coefficient
    kps_idx <- which(str_detect(rownames(cox_cleared_multiv_summary$coefficients), "kps_factor"))
    if (length(kps_idx) > 0) {
      idx <- kps_idx[1]
      if (idx <= nrow(cox_cleared_multiv_summary$conf.int)) {
        p_val_kps <- cox_cleared_multiv_summary$coefficients[idx, "Pr(>|z|)"]
        multiv_landmark_results_list[[length(multiv_landmark_results_list) + 1]] <- tibble(
          `Variable` = "KPS ≥80% vs <80%",
          `HR (95% CI)` = paste0(
            round(cox_cleared_multiv_summary$conf.int[idx, "exp(coef)"], 2), " (",
            round(cox_cleared_multiv_summary$conf.int[idx, "lower .95"], 2), "-",
            round(cox_cleared_multiv_summary$conf.int[idx, "upper .95"], 2), ")"
          ),
          `P-value` = if (p_val_kps < 0.001) "<0.001" else as.character(round(p_val_kps, 3))
        )
      }
    }
    
    # Add Visceral Metastases coefficient
    visceral_idx <- which(str_detect(rownames(cox_cleared_multiv_summary$coefficients), "visceral_mets_factor"))
    if (length(visceral_idx) > 0) {
      idx <- visceral_idx[1]
      if (idx <= nrow(cox_cleared_multiv_summary$conf.int)) {
        p_val_visceral <- cox_cleared_multiv_summary$coefficients[idx, "Pr(>|z|)"]
        multiv_landmark_results_list[[length(multiv_landmark_results_list) + 1]] <- tibble(
          `Variable` = "Visceral Mets vs No Visceral Mets",
          `HR (95% CI)` = paste0(
            round(cox_cleared_multiv_summary$conf.int[idx, "exp(coef)"], 2), " (",
            round(cox_cleared_multiv_summary$conf.int[idx, "lower .95"], 2), "-",
            round(cox_cleared_multiv_summary$conf.int[idx, "upper .95"], 2), ")"
          ),
          `P-value` = if (p_val_visceral < 0.001) "<0.001" else as.character(round(p_val_visceral, 3))
        )
      }
    }
    
    # Add Therapy Class coefficients
    therapy_class_idx <- which(str_detect(rownames(cox_cleared_multiv_summary$coefficients), "therapy_class_factor"))
    if (length(therapy_class_idx) > 0) {
      for (i in seq_along(therapy_class_idx)) {
        idx <- therapy_class_idx[i]
        if (idx <= nrow(cox_cleared_multiv_summary$conf.int)) {
          term_name <- rownames(cox_cleared_multiv_summary$coefficients)[idx]
          class_name <- str_replace(term_name, "therapy_class_factor", "")
          p_val_class <- cox_cleared_multiv_summary$coefficients[idx, "Pr(>|z|)"]
          hr_val <- cox_cleared_multiv_summary$conf.int[idx, "exp(coef)"]
          ci_lower <- cox_cleared_multiv_summary$conf.int[idx, "lower .95"]
          ci_upper <- cox_cleared_multiv_summary$conf.int[idx, "upper .95"]
          
          p_val_text <- if (is.na(p_val_class)) {
            "NA"
          } else if (p_val_class < 0.001) {
            "<0.001"
          } else {
            as.character(round(p_val_class, 3))
          }
          
          hr_ci_text <- if (is.na(hr_val) || is.na(ci_lower) || is.na(ci_upper)) {
            "NA (NA-NA)"
          } else {
            paste0(
              round(hr_val, 2), " (",
              round(ci_lower, 2), "-",
              round(ci_upper, 2), ")"
            )
          }
          
          multiv_landmark_results_list[[length(multiv_landmark_results_list) + 1]] <- tibble(
            `Variable` = paste0("Therapy Class: ", class_name, " vs reference"),
            `HR (95% CI)` = hr_ci_text,
            `P-value` = p_val_text
          )
        }
      }
    }
    
    # Create and display table if we have results
    if (length(multiv_landmark_results_list) > 0) {
      multiv_landmark_results <- bind_rows(multiv_landmark_results_list)
      
      landmark_multiv_table <- multiv_landmark_results %>%
        kable(caption = "Multivariate Cox Model: On-Treatment ctDNA Clearance and Duration of Complete Response from Landmark (adjusted for line of therapy, therapy class, KPS, and visceral metastases). Note: Only coefficients with reasonable HR values (converged models) are shown. Models with extreme HR values (0, Inf, or very large) are excluded due to convergence issues, likely due to small sample size or complete separation.") %>%
        kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
      
      # Display the table - kable objects render when referenced
      landmark_multiv_table
    } else {
      cat("\n\n**Note:** Multivariate model exists but no coefficients could be extracted for display.\n")
    }
  } else {
    cat("\n\n**Multivariate Cox Model Status:**\n")
    cat("- Multivariate model with Bajorin risk factors for on-treatment analysis could not be fitted.\n")
    cat("- This may be due to missing KPS or visceral metastases data, or insufficient sample size.\n")
    cat("- Check that Bajorin risk factors data is available and that there are enough observations with complete data.\n")
  }
}

# Add sample size information in a separate output (after table)
if (exists("landmark_outcomes_multiv") && !is.null(landmark_outcomes_multiv) && 
    exists("landmark_multiv_table") && !is.null(landmark_multiv_table)) {
  cat("\n\n**N for multivariate on-treatment model:**", nrow(landmark_outcomes_multiv), "observations\n")
}
```

```{r continuous_ctdna_analyses, include=FALSE}
# ====================================================================================
# ADDITIONAL ANALYSES: Continuous ctDNA Variables for Research Question 2
# ====================================================================================
# These analyses complement the binary clearance analysis by examining continuous
# ctDNA measures: log-transformed on-treatment ctDNA and percent decline
# ====================================================================================

# Initialize variables for continuous analyses
cox_continuous_on_tx <- NULL
cox_pct_decline_continuous <- NULL
cox_interaction_baseline <- NULL
cox_combined_baseline_on_tx <- NULL
comprehensive_landmark_results <- NULL

# Add continuous variables to landmark_outcomes if not already present
if (exists("landmark_outcomes") && nrow(landmark_outcomes) > 0) {
  if (!"log_on_tx_ctdna" %in% names(landmark_outcomes)) {
    landmark_outcomes <- landmark_outcomes %>%
      mutate(
        log_on_tx_ctdna = log(on_tx_ctdna + 1),
        log_baseline_ctdna = log(baseline_ctdna + 1),
        baseline_high = log_baseline_ctdna > median(log_baseline_ctdna, na.rm = TRUE),
        baseline_high_factor = factor(baseline_high, levels = c(FALSE, TRUE), 
                                       labels = c("Low Baseline", "High Baseline"))
      )
  }
}

if (exists("landmark_outcomes") && nrow(landmark_outcomes) > 0 && 
    exists("landmark_outcomes_cleared") && !is.null(landmark_outcomes_cleared)) {
  
  # Add continuous variables to landmark_outcomes_cleared if not already present
  if (!"log_on_tx_ctdna" %in% names(landmark_outcomes_cleared)) {
    landmark_outcomes_cleared <- landmark_outcomes_cleared %>%
      mutate(
        log_on_tx_ctdna = log(on_tx_ctdna + 1),
        log_baseline_ctdna = log(baseline_ctdna + 1),
        baseline_high = log_baseline_ctdna > median(log_baseline_ctdna, na.rm = TRUE),
        baseline_high_factor = factor(baseline_high, levels = c(FALSE, TRUE), 
                                       labels = c("Low Baseline", "High Baseline"))
      )
  }
  
  # Prepare data with all necessary variables
  landmark_for_continuous <- landmark_outcomes_cleared %>%
    filter(!is.na(log_on_tx_ctdna), !is.na(pct_decline))
  
  # ANALYSIS 1: Continuous on-treatment ctDNA (log-transformed)
  if (nrow(landmark_for_continuous) > 10 && 
      sum(!is.na(landmark_for_continuous$kps_binary)) > 0 && 
      sum(!is.na(landmark_for_continuous$visceral_mets_binary)) > 0) {
    
    landmark_continuous_multiv <- landmark_for_continuous %>%
      filter(!is.na(kps_binary), !is.na(visceral_mets_binary))
    
    if (nrow(landmark_continuous_multiv) > 0) {
      cox_continuous_on_tx <- tryCatch({
        coxph(Surv(mfs_time_from_landmark, mfs_event_from_landmark) ~ 
              log_on_tx_ctdna + therapy_line_factor + therapy_class_factor + 
              kps_factor + visceral_mets_factor,
              data = landmark_continuous_multiv)
      }, error = function(e) NULL)
    }
  }
  
  # ANALYSIS 2: Percent decline as continuous variable
  if (nrow(landmark_for_continuous) > 10 && 
      sum(!is.na(landmark_for_continuous$kps_binary)) > 0 && 
      sum(!is.na(landmark_for_continuous$visceral_mets_binary)) > 0) {
    
    landmark_pct_multiv <- landmark_for_continuous %>%
      filter(!is.na(kps_binary), !is.na(visceral_mets_binary), 
             !is.infinite(pct_decline), !is.na(pct_decline))
    
    if (nrow(landmark_pct_multiv) > 0) {
      cox_pct_decline_continuous <- tryCatch({
        coxph(Surv(mfs_time_from_landmark, mfs_event_from_landmark) ~ 
              pct_decline + therapy_line_factor + therapy_class_factor + 
              kps_factor + visceral_mets_factor,
              data = landmark_pct_multiv)
      }, error = function(e) NULL)
    }
  }
  
  # ANALYSIS 3: Interaction between baseline and on-treatment ctDNA
  if (nrow(landmark_for_continuous) > 10 && 
      sum(!is.na(landmark_for_continuous$kps_binary)) > 0 && 
      sum(!is.na(landmark_for_continuous$visceral_mets_binary)) > 0 &&
      sum(!is.na(landmark_for_continuous$baseline_high_factor)) > 0) {
    
    landmark_interaction_multiv <- landmark_for_continuous %>%
      filter(!is.na(kps_binary), !is.na(visceral_mets_binary),
             !is.na(ctdna_cleared_at_landmark), !is.na(baseline_high_factor))
    
    if (nrow(landmark_interaction_multiv) > 0) {
      cox_interaction_baseline <- tryCatch({
        coxph(Surv(mfs_time_from_landmark, mfs_event_from_landmark) ~ 
              ctdna_cleared_at_landmark * baseline_high_factor + 
              therapy_line_factor + therapy_class_factor + 
              kps_factor + visceral_mets_factor,
              data = landmark_interaction_multiv)
      }, error = function(e) NULL)
    }
  }
  
  # ANALYSIS 4: Combined model (baseline + on-treatment)
  if (nrow(landmark_for_continuous) > 10 && 
      sum(!is.na(landmark_for_continuous$kps_binary)) > 0 && 
      sum(!is.na(landmark_for_continuous$visceral_mets_binary)) > 0) {
    
    landmark_combined_multiv <- landmark_for_continuous %>%
      filter(!is.na(kps_binary), !is.na(visceral_mets_binary),
             !is.na(log_baseline_ctdna), !is.na(ctdna_cleared_at_landmark))
    
    if (nrow(landmark_combined_multiv) > 0) {
      cox_combined_baseline_on_tx <- tryCatch({
        coxph(Surv(mfs_time_from_landmark, mfs_event_from_landmark) ~ 
              log_baseline_ctdna + ctdna_cleared_at_landmark + 
              therapy_line_factor + therapy_class_factor + 
              kps_factor + visceral_mets_factor,
              data = landmark_combined_multiv)
      }, error = function(e) NULL)
    }
  }
  
  # Create comprehensive results table
  results_list_continuous <- list()
  
  # Binary clearance (from existing analysis) - only if converged and reasonable HR
  if (exists("cox_cleared_multiv") && !is.null(cox_cleared_multiv)) {
    model_converged <- tryCatch({
      isTRUE(cox_cleared_multiv$converged)
    }, error = function(e) FALSE)
    
    if (model_converged) {
      cox_cleared_multiv_summary <- summary(cox_cleared_multiv)
      hr_val <- cox_cleared_multiv_summary$conf.int[1, "exp(coef)"]
      hr_lower <- cox_cleared_multiv_summary$conf.int[1, "lower .95"]
      hr_upper <- cox_cleared_multiv_summary$conf.int[1, "upper .95"]
      
      if (!is.na(hr_val) && is.finite(hr_val) && hr_val > 0 && hr_val < 1e10 &&
          !is.na(hr_lower) && is.finite(hr_lower) &&
          !is.na(hr_upper) && is.finite(hr_upper)) {
        results_list_continuous[[length(results_list_continuous) + 1]] <- tibble(
          `Model` = "Binary: Cleared vs Not Cleared",
          `Variable` = "ctDNA Cleared at Landmark",
          `HR (95% CI)` = paste0(
            round(hr_val, 2), " (",
            round(hr_lower, 2), "-",
            round(hr_upper, 2), ")"
          ),
          `P-value` = ifelse(cox_cleared_multiv_summary$coefficients[1, "Pr(>|z|)"] < 0.001, 
                             "<0.001", 
                             as.character(round(cox_cleared_multiv_summary$coefficients[1, "Pr(>|z|)"], 3))),
          `N` = cox_cleared_multiv$n
        )
      }
    }
  }
  
  # Continuous on-treatment ctDNA - only if converged and reasonable HR
  if (!is.null(cox_continuous_on_tx)) {
    model_converged <- tryCatch({
      isTRUE(cox_continuous_on_tx$converged)
    }, error = function(e) FALSE)
    
    if (model_converged) {
      cox_cont_summary <- summary(cox_continuous_on_tx)
      hr_val <- cox_cont_summary$conf.int[1, "exp(coef)"]
      hr_lower <- cox_cont_summary$conf.int[1, "lower .95"]
      hr_upper <- cox_cont_summary$conf.int[1, "upper .95"]
      
      if (!is.na(hr_val) && is.finite(hr_val) && hr_val > 0 && hr_val < 1e10 &&
          !is.na(hr_lower) && is.finite(hr_lower) &&
          !is.na(hr_upper) && is.finite(hr_upper)) {
        results_list_continuous[[length(results_list_continuous) + 1]] <- tibble(
          `Model` = "Continuous: Log(On-treatment ctDNA + 1)",
          `Variable` = "Per log-unit increase in on-treatment ctDNA",
          `HR (95% CI)` = paste0(
            round(hr_val, 2), " (",
            round(hr_lower, 2), "-",
            round(hr_upper, 2), ")"
          ),
          `P-value` = ifelse(cox_cont_summary$coefficients[1, "Pr(>|z|)"] < 0.001, 
                             "<0.001", 
                             as.character(round(cox_cont_summary$coefficients[1, "Pr(>|z|)"], 3))),
          `N` = cox_continuous_on_tx$n
        )
      }
    }
  }
  
  # Percent decline (per 10% unit for interpretability) - only if converged and reasonable HR
  if (!is.null(cox_pct_decline_continuous)) {
    model_converged <- tryCatch({
      isTRUE(cox_pct_decline_continuous$converged)
    }, error = function(e) FALSE)
    
    if (model_converged) {
      cox_pct_summary <- summary(cox_pct_decline_continuous)
      # HR per 10% increase in decline
      hr_per_10pct <- exp(cox_pct_summary$coefficients[1, "coef"] * 10)
      ci_lower_per_10pct <- exp((cox_pct_summary$coefficients[1, "coef"] - 
                                 1.96 * cox_pct_summary$coefficients[1, "se(coef)"]) * 10)
      ci_upper_per_10pct <- exp((cox_pct_summary$coefficients[1, "coef"] + 
                                 1.96 * cox_pct_summary$coefficients[1, "se(coef)"]) * 10)
      
      if (!is.na(hr_per_10pct) && is.finite(hr_per_10pct) && hr_per_10pct > 0 && hr_per_10pct < 1e10 &&
          !is.na(ci_lower_per_10pct) && is.finite(ci_lower_per_10pct) &&
          !is.na(ci_upper_per_10pct) && is.finite(ci_upper_per_10pct)) {
        results_list_continuous[[length(results_list_continuous) + 1]] <- tibble(
          `Model` = "Continuous: Percent Decline from Baseline",
          `Variable` = "Per 10% increase in decline",
          `HR (95% CI)` = paste0(
            round(hr_per_10pct, 2), " (",
            round(ci_lower_per_10pct, 2), "-",
            round(ci_upper_per_10pct, 2), ")"
          ),
          `P-value` = ifelse(cox_pct_summary$coefficients[1, "Pr(>|z|)"] < 0.001, 
                             "<0.001", 
                             as.character(round(cox_pct_summary$coefficients[1, "Pr(>|z|)"], 3))),
          `N` = cox_pct_decline_continuous$n
        )
      }
    }
  }
  
  # Combined model (baseline + on-treatment) - only if converged and reasonable HR
  if (!is.null(cox_combined_baseline_on_tx)) {
    model_converged <- tryCatch({
      isTRUE(cox_combined_baseline_on_tx$converged)
    }, error = function(e) FALSE)
    
    if (model_converged) {
      cox_comb_summary <- summary(cox_combined_baseline_on_tx)
      
      # Baseline ctDNA
      hr_baseline <- cox_comb_summary$conf.int[1, "exp(coef)"]
      hr_baseline_lower <- cox_comb_summary$conf.int[1, "lower .95"]
      hr_baseline_upper <- cox_comb_summary$conf.int[1, "upper .95"]
      
      if (!is.na(hr_baseline) && is.finite(hr_baseline) && hr_baseline > 0 && hr_baseline < 1e10 &&
          !is.na(hr_baseline_lower) && is.finite(hr_baseline_lower) &&
          !is.na(hr_baseline_upper) && is.finite(hr_baseline_upper)) {
        results_list_continuous[[length(results_list_continuous) + 1]] <- tibble(
          `Model` = "Combined: Baseline + On-treatment",
          `Variable` = "Baseline ctDNA (per log-unit)",
          `HR (95% CI)` = paste0(
            round(hr_baseline, 2), " (",
            round(hr_baseline_lower, 2), "-",
            round(hr_baseline_upper, 2), ")"
          ),
          `P-value` = ifelse(cox_comb_summary$coefficients[1, "Pr(>|z|)"] < 0.001, 
                             "<0.001", 
                             as.character(round(cox_comb_summary$coefficients[1, "Pr(>|z|)"], 3))),
          `N` = cox_combined_baseline_on_tx$n
        )
      }
      
      # On-treatment clearance
      if (nrow(cox_comb_summary$conf.int) >= 2) {
        hr_ontx <- cox_comb_summary$conf.int[2, "exp(coef)"]
        hr_ontx_lower <- cox_comb_summary$conf.int[2, "lower .95"]
        hr_ontx_upper <- cox_comb_summary$conf.int[2, "upper .95"]
        
        if (!is.na(hr_ontx) && is.finite(hr_ontx) && hr_ontx > 0 && hr_ontx < 1e10 &&
            !is.na(hr_ontx_lower) && is.finite(hr_ontx_lower) &&
            !is.na(hr_ontx_upper) && is.finite(hr_ontx_upper)) {
          results_list_continuous[[length(results_list_continuous) + 1]] <- tibble(
            `Model` = "Combined: Baseline + On-treatment",
            `Variable` = "On-treatment ctDNA Cleared",
            `HR (95% CI)` = paste0(
              round(hr_ontx, 2), " (",
              round(hr_ontx_lower, 2), "-",
              round(hr_ontx_upper, 2), ")"
            ),
            `P-value` = ifelse(cox_comb_summary$coefficients[2, "Pr(>|z|)"] < 0.001, 
                               "<0.001", 
                               as.character(round(cox_comb_summary$coefficients[2, "Pr(>|z|)"], 3))),
            `N` = cox_combined_baseline_on_tx$n
          )
        }
      }
    }
  }
  
  # Combine all results
  if (length(results_list_continuous) > 0) {
    comprehensive_landmark_results <- bind_rows(results_list_continuous)
  }
}

cat("Continuous ctDNA analyses completed.\n")
```

```{r continuous_landmark_results_display, results='asis', echo=FALSE}
# Display comprehensive results table with binary and continuous analyses
if (exists("comprehensive_landmark_results") && !is.null(comprehensive_landmark_results) && 
    nrow(comprehensive_landmark_results) > 0) {
  comprehensive_landmark_results %>%
      kable(caption = "Comprehensive Landmark Analysis: Binary and Continuous ctDNA Measures and Duration of Complete Response from Landmark (adjusted for line of therapy, therapy class, KPS, and visceral metastases). Binary clearance analysis is shown for comparison; continuous analyses examine log-transformed on-treatment ctDNA and percent decline from baseline.") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
} else {
  cat("\nContinuous ctDNA analyses not available (insufficient data or models could not be fitted).\n")
}
```

```{r baseline_vs_landmark_population_comparison, echo=FALSE}
# Compare baseline and on-treatment populations
if (exists("outcomes") && exists("landmark_outcomes") && 
    nrow(outcomes) > 0 && nrow(landmark_outcomes) > 0) {
  
  baseline_summary <- outcomes %>%
    summarise(
      `Analysis` = "Baseline ctDNA (Research Question 1)",
      `N (Patient-Line Observations)` = n(),
      `N (Unique Patients)` = n_distinct(patient_id),
      `Median Baseline ctDNA (IQR)` = paste0(
        round(median(ctdna_before_line_value, na.rm = TRUE), 2), " (",
        round(quantile(ctdna_before_line_value, 0.25, na.rm = TRUE), 2), "-",
        round(quantile(ctdna_before_line_value, 0.75, na.rm = TRUE), 2), ")"
      ),
      `MFS Events` = sum(mfs_event, na.rm = TRUE),
      `Median MFS (days)` = round(median(mfs_time, na.rm = TRUE), 0),
      .groups = "drop"
    )
  
  landmark_summary <- landmark_outcomes %>%
    summarise(
      `Analysis` = "On-treatment ctDNA (Research Question 2)",
      `N (Patient-Line Observations)` = n(),
      `N (Unique Patients)` = n_distinct(patient_id),
      `Median Baseline ctDNA (IQR)` = paste0(
        round(median(baseline_ctdna, na.rm = TRUE), 2), " (",
        round(quantile(baseline_ctdna, 0.25, na.rm = TRUE), 2), "-",
        round(quantile(baseline_ctdna, 0.75, na.rm = TRUE), 2), ")"
      ),
      `MFS Events` = sum(mfs_event_from_landmark, na.rm = TRUE),
      `Median MFS from Landmark (days)` = round(median(mfs_time_from_landmark, na.rm = TRUE), 0),
      .groups = "drop"
    )
  
  population_comparison <- bind_rows(baseline_summary, landmark_summary) %>%
    kable(caption = "Population Comparison: Baseline vs On-Treatment Analysis Cohorts. Note: These are different patient populations - baseline analysis includes all patients with baseline ctDNA, while on-treatment analysis includes only patients who had on-treatment ctDNA measured within 6-16 weeks and survived to landmark.") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
  
  population_comparison
} else {
  cat("Population comparison not available.\n")
}
```

```{r landmark_interpretation}
# Interpretation of Hazard Ratios (always show if we have results)
# Use multivariate HR (adjusted for line of therapy + KPS + visceral metastases)
if (exists("landmark_results") && !is.null(landmark_results)) {
  cat("\n**Interpretation of Hazard Ratios (adjusted for line of therapy, KPS, and visceral metastases):**\n\n")
  
  # Prefer multivariate HR if available, otherwise fall back to adjusted HR
  if (exists("hr_cleared_multiv") && !is.na(hr_cleared_multiv)) {
    hr_to_use <- hr_cleared_multiv
    hr_lower_to_use <- hr_cleared_multiv_lower
    hr_upper_to_use <- hr_cleared_multiv_upper
    p_to_use <- p_cleared_multiv
    adjustment_note <- " (adjusted for line of therapy, KPS, and visceral metastases)"
  } else if (exists("hr_cleared") && !is.na(hr_cleared)) {
    hr_to_use <- hr_cleared
    hr_lower_to_use <- hr_cleared_lower
    hr_upper_to_use <- hr_cleared_upper
    p_to_use <- p_cleared
    adjustment_note <- " (adjusted for line of therapy)"
  } else {
    hr_to_use <- NA
  }
  
  if (!is.na(hr_to_use)) {
    if (hr_to_use < 1) {
      cat("- **ctDNA Cleared at Landmark vs Not Cleared (HR = ", round(hr_to_use, 2), ", 95% CI: ", 
          round(hr_lower_to_use, 2), "-", round(hr_upper_to_use, 2), ", p = ", 
          ifelse(p_to_use < 0.001, "<0.001", round(p_to_use, 3)), ")", adjustment_note, ":** Patients who cleared ctDNA at the landmark time point (6-16 weeks after therapy initiation) had a ", 
          round((1 - hr_to_use) * 100, 1), "% lower risk of metastasis or death COMPARED to those who did not clear ctDNA at the landmark. ", sep = "")
    } else {
      cat("- **ctDNA Cleared at Landmark vs Not Cleared (HR = ", round(hr_to_use, 2), ", 95% CI: ", 
          round(hr_lower_to_use, 2), "-", round(hr_upper_to_use, 2), ", p = ", 
          ifelse(p_to_use < 0.001, "<0.001", round(p_to_use, 3)), ")", adjustment_note, ":** Patients who cleared ctDNA at the landmark time point (6-16 weeks after therapy initiation) had a ", 
          round((hr_to_use - 1) * 100, 1), "% higher risk of metastasis or death COMPARED to those who did not clear ctDNA at the landmark. ", sep = "")
    }
    cat("This suggests that early ctDNA clearance may serve as a pharmacodynamic response biomarker, indicating effective treatment and potentially predicting improved metastasis-free survival.\n\n")
  }
}
```
<｜tool▁calls▁begin｜><｜tool▁call▁begin｜>
grep

```{r landmark_survival_curves}
# Survival curves: ctDNA Cleared at Landmark vs Not Cleared
if (exists("landmark_outcomes") && nrow(landmark_outcomes) > 0) {
  if (exists("landmark_outcomes_cleared") && 
      sum(landmark_outcomes_cleared$ctdna_cleared_at_landmark, na.rm = TRUE) > 0 && 
      sum(!landmark_outcomes_cleared$ctdna_cleared_at_landmark, na.rm = TRUE) > 0) {
    landmark_outcomes_cleared <- landmark_outcomes_cleared %>%
      mutate(
        cleared_label = ifelse(ctdna_cleared_at_landmark, "Cleared at Landmark", "Not Cleared at Landmark"),
        # Convert time to months for plotting
        mfs_time_from_landmark_months = mfs_time_from_landmark / 30.44
      )
    
    # Diagnostic: Check event counts
    cat("\n**Landmark Analysis Diagnostic:**\n")
    cat("Total patients in cleared analysis:", nrow(landmark_outcomes_cleared), "\n")
    cat("Events in Cleared group:", sum(landmark_outcomes_cleared$ctdna_cleared_at_landmark & landmark_outcomes_cleared$mfs_event_from_landmark == 1, na.rm = TRUE), "\n")
    cat("Events in Not Cleared group:", sum(!landmark_outcomes_cleared$ctdna_cleared_at_landmark & landmark_outcomes_cleared$mfs_event_from_landmark == 1, na.rm = TRUE), "\n")
    cat("Total events:", sum(landmark_outcomes_cleared$mfs_event_from_landmark, na.rm = TRUE), "\n")
    cat("Median survival time from landmark (months):", round(median(landmark_outcomes_cleared$mfs_time_from_landmark_months, na.rm = TRUE), 1), "\n")
    
    surv_fit_cleared <- survfit(Surv(mfs_time_from_landmark_months, mfs_event_from_landmark) ~ cleared_label, 
                                  data = landmark_outcomes_cleared)
    
    # Get adjusted p-value from Cox model (prefer multivariate: adjusted for line of therapy + KPS + visceral metastases)
    # Use multivariate p-value if available, otherwise fall back to line of therapy only
    if (exists("p_cleared_multiv") && !is.na(p_cleared_multiv)) {
      # Use multivariate p-value (adjusted for line of therapy + KPS + visceral metastases)
      pval_text <- ifelse(p_cleared_multiv < 0.001, "<0.001", round(p_cleared_multiv, 3))
      pval_numeric <- p_cleared_multiv
      cat("Using adjusted p-value from Cox model (adjusted for line of therapy, KPS, and visceral metastases):", pval_text, "\n\n")
    } else if (exists("p_cleared") && !is.na(p_cleared)) {
      # Fall back to p-value adjusted for line of therapy only
      pval_text <- ifelse(p_cleared < 0.001, "<0.001", round(p_cleared, 3))
      pval_numeric <- p_cleared
      cat("Using adjusted p-value from Cox model (adjusted for line of therapy):", pval_text, "\n\n")
    } else {
      # Recalculate if not already available
      landmark_outcomes_cleared_for_cox <- landmark_outcomes %>%
        filter(!is.na(ctdna_cleared_at_landmark)) %>%
        mutate(
          therapy_line_grouped = case_when(
            therapy_line == 1 ~ "1L",
            therapy_line == 2 ~ "2L",
            therapy_line == 3 ~ "3L",
            therapy_line >= 4 ~ "4L+",
            TRUE ~ as.character(therapy_line)
          ),
          therapy_line_factor = factor(therapy_line_grouped, levels = c("1L", "2L", "3L", "4L+"))
        )
      
      # Add therapy class if available
      if (exists("palliative_regimen_table") && nrow(palliative_regimen_table) > 0) {
        landmark_outcomes_cleared_for_cox <- landmark_outcomes_cleared_for_cox %>%
          left_join(
            palliative_regimen_table %>%
              mutate(therapy_line_match = ifelse(is.na(therapy_line_match), NA_integer_, therapy_line_match)) %>%
              select(patient_id, therapy_line_match, therapy_class) %>%
              distinct(),
            by = c("patient_id", "therapy_line" = "therapy_line_match")
          ) %>%
          mutate(
            therapy_class = replace_na(therapy_class, "unspecified"),
            therapy_class_factor = factor(therapy_class)
          )
      } else {
        landmark_outcomes_cleared_for_cox <- landmark_outcomes_cleared_for_cox %>%
          mutate(
            therapy_class = "unspecified",
            therapy_class_factor = factor("unspecified")
          )
      }
      
      # Try multivariate model first if Bajorin data is available
      if (exists("bajorin_data") && nrow(bajorin_data) > 0) {
        landmark_outcomes_cleared_for_cox <- landmark_outcomes_cleared_for_cox %>%
          left_join(
            bajorin_data %>% select(patient_id, therapy_line, kps_binary, visceral_mets_binary),
            by = c("patient_id", "therapy_line")
          ) %>%
          filter(!is.na(kps_binary), !is.na(visceral_mets_binary)) %>%
          mutate(
            kps_factor = factor(kps_binary, levels = c(0, 1), labels = c("KPS <80%", "KPS ≥80%")),
            visceral_mets_factor = factor(visceral_mets_binary, levels = c(0, 1), labels = c("No Visceral Mets", "Visceral Mets"))
          )
        
        if (nrow(landmark_outcomes_cleared_for_cox) > 0) {
          cox_cleared_multiv_adj <- tryCatch({
            coxph(Surv(mfs_time_from_landmark, mfs_event_from_landmark) ~ 
                  ctdna_cleared_at_landmark + therapy_line_factor + therapy_class_factor + kps_factor + visceral_mets_factor,
                  data = landmark_outcomes_cleared_for_cox)
          }, error = function(e) NULL)
          
          if (!is.null(cox_cleared_multiv_adj)) {
            cox_cleared_multiv_adj_summary <- summary(cox_cleared_multiv_adj)
            p_cleared_multiv <- cox_cleared_multiv_adj_summary$coefficients[1, "Pr(>|z|)"]
            pval_text <- ifelse(p_cleared_multiv < 0.001, "<0.001", round(p_cleared_multiv, 3))
            pval_numeric <- p_cleared_multiv
            cat("Using adjusted p-value from Cox model (adjusted for line of therapy, therapy class, KPS, and visceral metastases):", pval_text, "\n\n")
          } else {
            # Fall back to line of therapy only
            landmark_outcomes_cleared_for_cox <- landmark_outcomes %>%
              filter(!is.na(ctdna_cleared_at_landmark)) %>%
              mutate(
                therapy_line_grouped = case_when(
                  therapy_line == 1 ~ "1L",
                  therapy_line == 2 ~ "2L",
                  therapy_line == 3 ~ "3L",
                  therapy_line >= 4 ~ "4L+",
                  TRUE ~ as.character(therapy_line)
                ),
                therapy_line_factor = factor(therapy_line_grouped, levels = c("1L", "2L", "3L", "4L+"))
              )
            
            # Add therapy class if available
            if (exists("palliative_regimen_table") && nrow(palliative_regimen_table) > 0) {
              landmark_outcomes_cleared_for_cox <- landmark_outcomes_cleared_for_cox %>%
                left_join(
                  palliative_regimen_table %>%
                    mutate(therapy_line_match = ifelse(is.na(therapy_line_match), NA_integer_, therapy_line_match)) %>%
                    select(patient_id, therapy_line_match, therapy_class) %>%
                    distinct(),
                  by = c("patient_id", "therapy_line" = "therapy_line_match")
                ) %>%
                mutate(
                  therapy_class = replace_na(therapy_class, "unspecified"),
                  therapy_class_factor = factor(therapy_class)
                )
            } else {
              landmark_outcomes_cleared_for_cox <- landmark_outcomes_cleared_for_cox %>%
                mutate(
                  therapy_class = "unspecified",
                  therapy_class_factor = factor("unspecified")
                )
            }
            
            cox_cleared_adj <- tryCatch({
              coxph(Surv(mfs_time_from_landmark, mfs_event_from_landmark) ~ ctdna_cleared_at_landmark + therapy_line_factor + therapy_class_factor,
                    data = landmark_outcomes_cleared_for_cox)
            }, error = function(e) NULL)
            
            if (!is.null(cox_cleared_adj)) {
              cox_cleared_adj_summary <- summary(cox_cleared_adj)
              p_cleared <- cox_cleared_adj_summary$coefficients[1, "Pr(>|z|)"]
              pval_text <- ifelse(p_cleared < 0.001, "<0.001", round(p_cleared, 3))
              pval_numeric <- p_cleared
              cat("Using adjusted p-value from Cox model (adjusted for line of therapy and therapy class):", pval_text, "\n\n")
            } else {
              # Fallback to log-rank test
              logrank_test <- survdiff(Surv(mfs_time_from_landmark_months, mfs_event_from_landmark) ~ cleared_label, 
                                       data = landmark_outcomes_cleared)
              logrank_pvalue <- 1 - pchisq(logrank_test$chisq, length(logrank_test$n) - 1)
              pval_text <- ifelse(logrank_pvalue < 0.001, "<0.001", round(logrank_pvalue, 3))
              pval_numeric <- logrank_pvalue
              cat("Using log-rank test p-value (unadjusted) - adjusted p-value not available.\n\n")
            }
          }
        } else {
          # Not enough data for multivariate, try line of therapy only
          landmark_outcomes_cleared_for_cox <- landmark_outcomes %>%
            filter(!is.na(ctdna_cleared_at_landmark)) %>%
            mutate(
              therapy_line_grouped = case_when(
                therapy_line == 1 ~ "1L",
                therapy_line == 2 ~ "2L",
                therapy_line == 3 ~ "3L",
                therapy_line >= 4 ~ "4L+",
                TRUE ~ as.character(therapy_line)
              ),
              therapy_line_factor = factor(therapy_line_grouped, levels = c("1L", "2L", "3L", "4L+"))
            )
          
          # Add therapy class if available
          if (exists("palliative_regimen_table") && nrow(palliative_regimen_table) > 0) {
            landmark_outcomes_cleared_for_cox <- landmark_outcomes_cleared_for_cox %>%
              left_join(
                palliative_regimen_table %>%
                  mutate(therapy_line_match = ifelse(is.na(therapy_line_match), NA_integer_, therapy_line_match)) %>%
                  select(patient_id, therapy_line_match, therapy_class) %>%
                  distinct(),
                by = c("patient_id", "therapy_line" = "therapy_line_match")
              ) %>%
              mutate(
                therapy_class = replace_na(therapy_class, "unspecified"),
                therapy_class_factor = factor(therapy_class)
              )
          } else {
            landmark_outcomes_cleared_for_cox <- landmark_outcomes_cleared_for_cox %>%
              mutate(
                therapy_class = "unspecified",
                therapy_class_factor = factor("unspecified")
              )
          }
          
          cox_cleared_adj <- tryCatch({
            coxph(Surv(mfs_time_from_landmark, mfs_event_from_landmark) ~ ctdna_cleared_at_landmark + therapy_line_factor + therapy_class_factor,
                  data = landmark_outcomes_cleared_for_cox)
          }, error = function(e) NULL)
          
          if (!is.null(cox_cleared_adj)) {
            cox_cleared_adj_summary <- summary(cox_cleared_adj)
            p_cleared <- cox_cleared_adj_summary$coefficients[1, "Pr(>|z|)"]
            pval_text <- ifelse(p_cleared < 0.001, "<0.001", round(p_cleared, 3))
            pval_numeric <- p_cleared
            cat("Using adjusted p-value from Cox model (adjusted for line of therapy):", pval_text, "\n\n")
          } else {
            # Fallback to log-rank test
            logrank_test <- survdiff(Surv(mfs_time_from_landmark_months, mfs_event_from_landmark) ~ cleared_label, 
                                     data = landmark_outcomes_cleared)
            logrank_pvalue <- 1 - pchisq(logrank_test$chisq, length(logrank_test$n) - 1)
            pval_text <- ifelse(logrank_pvalue < 0.001, "<0.001", round(logrank_pvalue, 3))
            pval_numeric <- logrank_pvalue
            cat("Using log-rank test p-value (unadjusted) - adjusted p-value not available.\n\n")
          }
        }
      } else {
        # No Bajorin data, try line of therapy only
        cox_cleared_adj <- tryCatch({
          coxph(Surv(mfs_time_from_landmark, mfs_event_from_landmark) ~ ctdna_cleared_at_landmark + therapy_line_factor + therapy_class_factor,
                data = landmark_outcomes_cleared_for_cox)
        }, error = function(e) NULL)
        
        if (!is.null(cox_cleared_adj)) {
          cox_cleared_adj_summary <- summary(cox_cleared_adj)
          p_cleared <- cox_cleared_adj_summary$coefficients[1, "Pr(>|z|)"]
          pval_text <- ifelse(p_cleared < 0.001, "<0.001", round(p_cleared, 3))
          pval_numeric <- p_cleared
          cat("Using adjusted p-value from Cox model (adjusted for line of therapy):", pval_text, "\n\n")
        } else {
          # Fallback to log-rank test
          logrank_test <- survdiff(Surv(mfs_time_from_landmark_months, mfs_event_from_landmark) ~ cleared_label, 
                                   data = landmark_outcomes_cleared)
          logrank_pvalue <- 1 - pchisq(logrank_test$chisq, length(logrank_test$n) - 1)
          pval_text <- ifelse(logrank_pvalue < 0.001, "<0.001", round(logrank_pvalue, 3))
          pval_numeric <- logrank_pvalue
          cat("Using log-rank test p-value (unadjusted) - adjusted p-value not available.\n\n")
        }
      }
    }
    
    # Create custom p-value text
    # Check if we used multivariate p-value
    if (exists("p_cleared_multiv") && !is.na(p_cleared_multiv) && pval_numeric == p_cleared_multiv) {
      pval_display <- paste0("p = ", pval_text, " (adjusted for line of therapy, KPS, and visceral metastases)")
    } else if (!is.null(p_cleared) && !is.na(p_cleared) && pval_numeric == p_cleared) {
      pval_display <- paste0("p = ", pval_text, " (adjusted for line of therapy)")
    } else {
      pval_display <- paste0("p = ", pval_text)
    }
    
    p_landmark <- ggsurvplot(
      surv_fit_cleared,
      data = landmark_outcomes_cleared,
      pval = pval_display,
      pval.size = 4,
      conf.int = TRUE,
      risk.table = TRUE,
      xlab = "Time from Landmark (months)",
      ylab = "Duration of Complete Response Probability",
      title = "Duration of Complete Response from Landmark (ctDNA between 6-16 weeks) by ctDNA Clearance Status at Landmark",
      xlim = c(0, 12),
      legend.title = "ctDNA Status at Landmark",
      legend.labs = c("Cleared at Landmark", "Not Cleared at Landmark")
    )
    
    p_landmark$plot
  } else {
    # Fallback: overall landmark curve (ignoring clearance) to show any events
    if (exists("landmark_outcomes") && nrow(landmark_outcomes) > 0) {
      landmark_outcomes_overall <- landmark_outcomes %>%
        mutate(mfs_time_from_landmark_months = mfs_time_from_landmark / 30.44)
      
      # Diagnostic: Check event counts for overall
      cat("\n**Landmark Analysis Diagnostic (Overall Fallback):**\n")
      cat("Total patients:", nrow(landmark_outcomes_overall), "\n")
      cat("Total events:", sum(landmark_outcomes_overall$mfs_event_from_landmark, na.rm = TRUE), "\n")
      cat("Median time from landmark (months):", round(median(landmark_outcomes_overall$mfs_time_from_landmark_months, na.rm = TRUE), 1), "\n")
      
      surv_fit_overall <- survfit(Surv(mfs_time_from_landmark_months, mfs_event_from_landmark) ~ 1, 
                                   data = landmark_outcomes_overall)
      p_landmark_overall <- ggsurvplot(
        surv_fit_overall,
        data = landmark_outcomes_overall,
        pval = FALSE,
        conf.int = TRUE,
        risk.table = TRUE,
        xlab = "Time from Landmark (months)",
        ylab = "Duration of Complete Response Probability",
        title = "Duration of Complete Response from Landmark (ctDNA between 6-16 weeks) - Overall",
        xlim = c(0, 36)
      )
      p_landmark_overall$plot
      if (sum(landmark_outcomes_overall$mfs_event_from_landmark, na.rm = TRUE) == 0) {
        cat("Note: No post-landmark events; overall curve shown.\n")
      } else if (sum(!is.na(landmark_outcomes$ctdna_cleared_at_landmark)) <= 10) {
        cat("Note: Only one ctDNA clearance group available or <10 with clearance data; overall curve shown.\n")
      }
    } else {
      cat("Insufficient data for landmark analysis (need at least 10 patients with clearance data).\n")
    }
  }
} else {
  cat("No patients with on-treatment ctDNA measurements in landmark window (", landmark_weeks, " weeks, window: 10-14 weeks).\n", sep = "")
}
```

```{r summary}
# Summary statistics
cat("## Summary\n\n")
cat("**Universe - Prognostic Cohort (Baseline ctDNA Analysis):** Patients with metastatic disease starting a new line of therapy\n")
cat("  - Includes all lines of therapy (1L, 2L, 3L+, not limited to first-line)\n")
cat("  - ctDNA measured within 21 days before starting each new line\n")
cat("  - Time 0 = start of each new line of therapy\n")
cat("  - **N =", nrow(outcomes), " patient-line observations\n\n")

# Get universe size for on-treatment response
landmark_universe_n <- if (exists("landmark_data") && nrow(landmark_data) > 0) {
  nrow(landmark_data)
} else if (exists("landmark_outcomes") && nrow(landmark_outcomes) > 0) {
  nrow(landmark_outcomes)
} else {
  NA
}

if (!is.na(landmark_universe_n)) {
  cat("**Universe - On-Treatment Cohort (Landmark Analysis):** Patients with metastatic disease starting a new line of therapy\n")
  cat("  - Includes all lines of therapy (1L, 2L, 3L+, not limited to first-line)\n")
  cat("  - Had on-treatment ctDNA measurement within 6-16 weeks after starting each new line\n")
  cat("  - Landmark time = ctDNA measurement date within 6-16 week window\n")
  cat("  - Time 0 for landmark analysis = landmark time (ctDNA measurement date)\n")
  cat("  - **N =", landmark_universe_n, " patient-line observations\n\n")
}
cat("**Key Findings - Prognostic (Baseline ctDNA):**\n")
cat("- Median ctDNA before therapy:", round(median(outcomes$ctdna_before_line_value, na.rm = TRUE), 2), "\n")
# Use fully adjusted model (line of therapy + KPS + visceral metastases) if available
if (!is.null(cox_model_multiv) && exists("hr_multiv") && !is.na(hr_multiv)) {
  cat("- Hazard Ratio per log-unit increase in ctDNA (adjusted for line of therapy, KPS, and visceral metastases):", 
      round(hr_multiv, 2), 
      " (95% CI: ", 
      round(hr_multiv_lower, 2), 
      "-", 
      round(hr_multiv_upper, 2), 
      ")\n")
  cat("- P-value:", ifelse(p_multiv < 0.001, 
                            "<0.001", 
                            round(p_multiv, 3)), "\n\n")
} else {
  # Fall back to model adjusted for line of therapy only
  cat("- Hazard Ratio per log-unit increase in ctDNA (adjusted for line of therapy):", 
      round(hr_value, 2), 
      " (95% CI: ", 
      round(hr_lower, 2), 
      "-", 
      round(hr_upper, 2), 
      ")\n")
  cat("- P-value:", ifelse(cox_summary$coefficients[1, "Pr(>|z|)"] < 0.001, 
                            "<0.001", 
                            round(cox_summary$coefficients[1, "Pr(>|z|)"], 3)), "\n\n")
}

# On-treatment findings - use fully adjusted model if available
if (exists("hr_cleared_multiv") && !is.na(hr_cleared_multiv)) {
  cat("**Key Findings - On-Treatment (Early Response):**\n")
  cat("- Hazard Ratio for ctDNA Cleared at Landmark vs Not Cleared (adjusted for line of therapy, KPS, and visceral metastases):", 
      round(hr_cleared_multiv, 2), 
      " (95% CI: ", 
      round(hr_cleared_multiv_lower, 2), 
      "-", 
      round(hr_cleared_multiv_upper, 2), 
      ")\n")
  cat("- P-value:", ifelse(p_cleared_multiv < 0.001, 
                            "<0.001", 
                            round(p_cleared_multiv, 3)), "\n\n")
} else if (exists("hr_cleared") && !is.na(hr_cleared)) {
  cat("**Key Findings - On-Treatment (Early Response):**\n")
  cat("- Hazard Ratio for ctDNA Cleared at Landmark vs Not Cleared (adjusted for line of therapy):", 
      round(hr_cleared, 2), 
      " (95% CI: ", 
      round(hr_cleared_lower, 2), 
      "-", 
      round(hr_cleared_upper, 2), 
      ")\n")
  cat("- P-value:", ifelse(p_cleared < 0.001, 
                            "<0.001", 
                            round(p_cleared, 3)), "\n\n")
}
cat("**Combined Interpretation:**\n\n")

# Prognostic role (baseline ctDNA) - use fully adjusted model if available
cat("**Prognostic Role (Baseline ctDNA):** ")
# Use multivariate HR if available, otherwise fall back to line-adjusted HR
hr_to_use <- if (!is.null(cox_model_multiv) && exists("hr_multiv") && !is.na(hr_multiv)) hr_multiv else hr_value
hr_lower_to_use <- if (!is.null(cox_model_multiv) && exists("hr_multiv_lower") && !is.na(hr_multiv_lower)) hr_multiv_lower else hr_lower
hr_upper_to_use <- if (!is.null(cox_model_multiv) && exists("hr_multiv_upper") && !is.na(hr_multiv_upper)) hr_multiv_upper else hr_upper
adjustment_text <- if (!is.null(cox_model_multiv) && exists("hr_multiv") && !is.na(hr_multiv)) 
  "adjusted for line of therapy, KPS, and visceral metastases" else "adjusted for line of therapy"

if (hr_to_use > 1 && hr_lower_to_use > 1) {
  cat("Higher ctDNA levels before starting a new line of therapy are significantly associated with worse metastasis-free survival (HR = ", round(hr_to_use, 2), ", 95% CI: ", round(hr_lower_to_use, 2), "-", round(hr_upper_to_use, 2), "), ", adjustment_text, ". ", sep="")
  cat("This finding suggests that ctDNA level before therapy initiation may serve as a prognostic biomarker for duration of complete response outcomes across multiple lines of therapy in patients with metastatic urothelial carcinoma.\n\n")
} else if (hr_to_use > 1) {
  cat("Higher ctDNA levels before starting a new line of therapy show a trend toward worse metastasis-free survival (HR = ", round(hr_to_use, 2), ", 95% CI: ", round(hr_lower_to_use, 2), "-", round(hr_upper_to_use, 2), "), though the confidence interval includes 1.0, indicating the association may not be statistically significant.\n\n", sep="")
} else if (hr_to_use < 1 && hr_upper_to_use < 1) {
  cat("Higher ctDNA levels before starting a new line of therapy are significantly associated with better metastasis-free survival (HR = ", round(hr_to_use, 2), ", 95% CI: ", round(hr_lower_to_use, 2), "-", round(hr_upper_to_use, 2), "), which is counterintuitive and warrants further investigation.\n\n", sep="")
} else {
  cat("There is no significant association between ctDNA level before starting a new line of therapy and metastasis-free survival (HR = ", round(hr_to_use, 2), ", 95% CI: ", round(hr_lower_to_use, 2), "-", round(hr_upper_to_use, 2), "), ", adjustment_text, ".\n\n", sep="")
}

# On-treatment role (landmark analysis) - use fully adjusted model if available
if (exists("hr_cleared_multiv") && !is.na(hr_cleared_multiv)) {
  cat("**On-Treatment Role (Early Response):** Patients who cleared ctDNA at the landmark time point (6-16 weeks after therapy initiation) showed ", 
      if (hr_cleared_multiv < 1) {
        paste0("a significantly lower risk of metastasis or death (HR = ", round(hr_cleared_multiv, 2), ", 95% CI: ", round(hr_cleared_multiv_lower, 2), "-", round(hr_cleared_multiv_upper, 2), "), p = ", ifelse(p_cleared_multiv < 0.001, "<0.001", round(p_cleared_multiv, 3)), ", adjusted for line of therapy, KPS, and visceral metastases) COMPARED to those who did not clear ctDNA. ", sep = "")
      } else {
        paste0("a higher risk of metastasis or death (HR = ", round(hr_cleared_multiv, 2), ", 95% CI: ", round(hr_cleared_multiv_lower, 2), "-", round(hr_cleared_multiv_upper, 2), "), p = ", ifelse(p_cleared_multiv < 0.001, "<0.001", round(p_cleared_multiv, 3)), ", adjusted for line of therapy, KPS, and visceral metastases) COMPARED to those who did not clear ctDNA. ", sep = "")
      })
  cat("This strong association suggests that early ctDNA clearance may serve as a pharmacodynamic response biomarker, indicating effective treatment and potentially predicting improved metastasis-free survival. ")
  cat("Together, these findings demonstrate both the prognostic value of baseline ctDNA levels and the predictive value of early on-treatment ctDNA changes. Validation as a surrogate endpoint would require prospective trials.\n")
} else if (exists("hr_cleared") && !is.na(hr_cleared)) {
  cat("**On-Treatment Role (Early Response):** Patients who cleared ctDNA at the landmark time point (6-16 weeks after therapy initiation) showed ", 
      if (hr_cleared < 1) {
        paste0("a significantly lower risk of metastasis or death (HR = ", round(hr_cleared, 2), ", 95% CI: ", round(hr_cleared_lower, 2), "-", round(hr_cleared_upper, 2), "), p = ", ifelse(p_cleared < 0.001, "<0.001", round(p_cleared, 3)), ", adjusted for line of therapy) COMPARED to those who did not clear ctDNA. ", sep = "")
      } else {
        paste0("a higher risk of metastasis or death (HR = ", round(hr_cleared, 2), ", 95% CI: ", round(hr_cleared_lower, 2), "-", round(hr_cleared_upper, 2), "), p = ", ifelse(p_cleared < 0.001, "<0.001", round(p_cleared, 3)), ", adjusted for line of therapy) COMPARED to those who did not clear ctDNA. ", sep = "")
      })
  cat("This strong association suggests that early ctDNA clearance may serve as a pharmacodynamic response biomarker, indicating effective treatment and potentially predicting improved metastasis-free survival. ")
  cat("Together, these findings demonstrate both the prognostic value of baseline ctDNA levels and the predictive value of early on-treatment ctDNA changes. Validation as a surrogate endpoint would require prospective trials.\n")
}

```

