---
title: "Research Question Universes and Swimmer Plots"
author: "Exploratory Analysis"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

This document calculates patient universes and generates swimmer plots for four specific research questions:

1. **Question 1**: Association of on-treatment clinical events and pre-treatment factors on ctDNA profiles
2. **Question 2**: Case series of patients with undetectable ctDNA following TURBT alone
3. **Question 3**: ctDNA thresholds/kinetics for organ-confined vs metastatic recurrence
4. **Question 4**: Long-term outcomes of persistently low-level detectable ctDNA

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "")
knitr::opts_knit$set(root.dir = "~/R")
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Load shared setup
source("ctDNA_Research_Questions_Shared_Setup.R")

# Load required libraries
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(stringr)
  library(readxl)
  library(janitor)
  library(lubridate)
  library(ggplot2)
  library(patchwork)
  library(survival)
  library(survminer)
  library(broom)
  library(knitr)
  library(kableExtra)
})
```

```{r load_data, include=FALSE}
# Load data
data_list <- load_minimal_data(include_survival = TRUE)
dat_combined <- data_list$dat_combined
ctdna_long <- data_list$ctdna_long
survival_data <- data_list$survival_data

# Build ctDNA long if not available
if (is.null(ctdna_long) || nrow(ctdna_long) == 0) {
  ctdna_long <- build_ctdna_long_minimal(dat_combined, "patient_id")
}
```

```{r research_questions_universes}
# ==============================
# Research Questions - Universe Calculations
# ==============================

# Question 1: Association of on-treatment clinical events and pre-treatment factors on ctDNA profiles
# Universe: Patients who underwent cystectomy

# Extract cystectomy information
surgery_type_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("surgery.*type|surgery.*procedure|surgery_\\d+_type|surgery_\\d+_procedure", ignore_case = TRUE))]
surgery_date_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("surgery.*date|surgery_\\d+_date", ignore_case = TRUE))]

# Get patients who underwent cystectomy
if (length(surgery_type_cols) > 0 && length(surgery_date_cols) > 0) {
  cystectomy_patients <- dat_combined %>%
    select(patient_id, all_of(surgery_type_cols), all_of(surgery_date_cols)) %>%
    pivot_longer(cols = all_of(surgery_type_cols), names_to = "surgery_type_var", values_to = "surgery_type") %>%
    pivot_longer(cols = all_of(surgery_date_cols), names_to = "surgery_date_var", values_to = "surgery_date_raw") %>%
    mutate(
      surgery_type = as.character(surgery_type),
      surgery_date = as_date_safe(surgery_date_raw),
      is_cystectomy = str_detect(tolower(surgery_type), "cystectomy|robotic.*cystectomy|open.*cystectomy")
    ) %>%
    filter(is_cystectomy, !is.na(surgery_date)) %>%
    distinct(patient_id) %>%
    pull(patient_id)
  
  n_cystectomy <- length(cystectomy_patients)
} else {
  # Fallback: assume surgery_1_date is cystectomy if no type column
  if (length(surgery_date_cols) > 0) {
    cystectomy_patients <- dat_combined %>%
      select(patient_id, all_of(surgery_date_cols[1])) %>%
      rename(surgery_date_raw = all_of(surgery_date_cols[1])) %>%
      mutate(surgery_date = as_date_safe(surgery_date_raw)) %>%
      filter(!is.na(surgery_date)) %>%
      pull(patient_id)
    
    n_cystectomy <- length(cystectomy_patients)
  } else {
    cystectomy_patients <- character(0)
    n_cystectomy <- 0
  }
}

# Question 2: Case series of patients with undetectable ctDNA following TURBT alone
# Universe: Patients who underwent TURBT alone and had undetectable ctDNA afterwards

ctdna_long_local <- ctdna_long %>%
  mutate(patient_id = as.character(patient_id))

if (nrow(ctdna_long_local) > 0) {
  # Find patients with ctDNA clearance (detectable to undetectable)
  patients_with_clearance <- ctdna_long_local %>%
    arrange(patient_id, ctdna_date) %>%
    group_by(patient_id) %>%
    summarise(
      first_detectable_date = if (any(ctdna_value > 0, na.rm = TRUE)) {
        min(ctdna_date[ctdna_value > 0], na.rm = TRUE)
      } else {
        as.Date(NA)
      },
      first_undetectable_date = if (any(ctdna_value == 0, na.rm = TRUE)) {
        min(ctdna_date[ctdna_value == 0], na.rm = TRUE)
      } else {
        as.Date(NA)
      },
      .groups = "drop"
    ) %>%
    mutate(
      had_clearance_pattern = !is.na(first_detectable_date) & 
                             !is.na(first_undetectable_date) &
                             first_undetectable_date > first_detectable_date
    ) %>%
    filter(had_clearance_pattern) %>%
    pull(patient_id)
  
  # Exclude patients with cystectomy
  if (length(surgery_type_cols) > 0 && length(surgery_date_cols) > 0) {
    patients_with_cystectomy <- dat_combined %>%
      mutate(patient_id = as.character(patient_id)) %>%
      select(patient_id, all_of(surgery_type_cols), all_of(surgery_date_cols)) %>%
      pivot_longer(cols = all_of(surgery_type_cols), names_to = "surgery_type_var", values_to = "surgery_type") %>%
      pivot_longer(cols = all_of(surgery_date_cols), names_to = "surgery_date_var", values_to = "surgery_date_raw") %>%
      mutate(
        surgery_type = as.character(surgery_type),
        surgery_date = as_date_safe(surgery_date_raw),
        is_cystectomy = str_detect(tolower(surgery_type), "cystectomy|robotic.*cystectomy|open.*cystectomy")
      ) %>%
      filter(!is.na(surgery_date), is_cystectomy) %>%
      distinct(patient_id) %>%
      pull(patient_id)
  } else {
    patients_with_cystectomy <- character(0)
  }
  
  # Calculate when ctDNA became undetectable
  clearance_dates <- ctdna_long_local %>%
    filter(patient_id %in% patients_with_clearance) %>%
    arrange(patient_id, ctdna_date) %>%
    group_by(patient_id) %>%
    summarise(
      first_detectable_date = if (any(ctdna_value > 0, na.rm = TRUE)) {
        min(ctdna_date[ctdna_value > 0], na.rm = TRUE)
      } else {
        as.Date(NA)
      },
      .groups = "drop"
    ) %>%
    left_join(
      ctdna_long_local %>%
        filter(patient_id %in% patients_with_clearance) %>%
        select(patient_id, ctdna_date, ctdna_value),
      by = "patient_id"
    ) %>%
    filter(
      !is.na(first_detectable_date),
      ctdna_value == 0,
      ctdna_date > first_detectable_date
    ) %>%
    group_by(patient_id) %>%
    summarise(
      clearance_date = min(ctdna_date, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    filter(!is.na(clearance_date))
  
  # Exclude patients with systemic therapy that started less than 1 year after clearance
  therapy_start_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("systemic.*therapy.*\\d+.*start.*date", ignore_case = TRUE))]
  patients_with_therapy_too_soon <- character(0)
  if (length(therapy_start_cols) > 0 && nrow(clearance_dates) > 0) {
    therapy_dates <- dat_combined %>%
      mutate(patient_id = as.character(patient_id)) %>%
      filter(patient_id %in% clearance_dates$patient_id) %>%
      select(patient_id, all_of(therapy_start_cols)) %>%
      pivot_longer(-patient_id, names_to = "var", values_to = "therapy_start_raw") %>%
      mutate(therapy_start = as_date_safe(therapy_start_raw)) %>%
      filter(!is.na(therapy_start)) %>%
      left_join(clearance_dates %>% select(patient_id, clearance_date), by = "patient_id") %>%
      filter(
        !is.na(clearance_date),
        as.numeric(therapy_start - clearance_date) < 365
      ) %>%
      distinct(patient_id) %>%
      pull(patient_id)
    patients_with_therapy_too_soon <- therapy_dates
  }
  
  # Exclude patients with radiation
  radiation_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("radiation|rt|radiotherapy", ignore_case = TRUE))]
  patients_with_radiation_too_soon <- character(0)
  if (length(radiation_cols) > 0 && nrow(clearance_dates) > 0) {
    radiation_date_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("radiation.*date|rt.*date", ignore_case = TRUE))]
    if (length(radiation_date_cols) > 0) {
      radiation_dates <- dat_combined %>%
        mutate(patient_id = as.character(patient_id)) %>%
        filter(patient_id %in% clearance_dates$patient_id) %>%
        select(patient_id, all_of(radiation_date_cols)) %>%
        pivot_longer(-patient_id, names_to = "var", values_to = "radiation_date_raw") %>%
        mutate(radiation_date = as_date_safe(radiation_date_raw)) %>%
        filter(!is.na(radiation_date)) %>%
        left_join(clearance_dates %>% select(patient_id, clearance_date), by = "patient_id") %>%
        filter(
          !is.na(clearance_date),
          radiation_date < clearance_date | as.numeric(radiation_date - clearance_date) < 365
        ) %>%
        distinct(patient_id) %>%
        pull(patient_id)
      patients_with_radiation_too_soon <- unique(c(patients_with_radiation_too_soon, radiation_dates))
    }
  }
  
  # Final list: patients with clearance, no cystectomy, no therapy/radiation within 1 year
  turbt_undetectable <- setdiff(
    setdiff(
      setdiff(patients_with_clearance, patients_with_cystectomy),
      patients_with_therapy_too_soon
    ),
    patients_with_radiation_too_soon
  )
  
  n_turbt_undetectable <- length(turbt_undetectable)
} else {
  turbt_undetectable <- character(0)
  n_turbt_undetectable <- 0
}

# Question 3: ctDNA thresholds/kinetics for organ-confined vs metastatic recurrence
# Universe: Patients who underwent cystectomy and had ctDNA levels available beforehand

if (exists("cystectomy_patients") && length(cystectomy_patients) > 0 && exists("ctdna_long")) {
  if (length(surgery_date_cols) > 0) {
    cystectomy_dates <- dat_combined %>%
      filter(patient_id %in% cystectomy_patients) %>%
      select(patient_id, all_of(surgery_date_cols)) %>%
      pivot_longer(-patient_id, names_to = "var", values_to = "surgery_date_raw") %>%
      mutate(surgery_date = as_date_safe(surgery_date_raw)) %>%
      filter(!is.na(surgery_date)) %>%
      group_by(patient_id) %>%
      summarise(cystectomy_date = min(surgery_date, na.rm = TRUE), .groups = "drop")
    
    patients_ctdna_before_cyst <- ctdna_long %>%
      left_join(cystectomy_dates, by = "patient_id") %>%
      filter(
        !is.na(cystectomy_date),
        ctdna_date < cystectomy_date,
        !is.na(ctdna_value)
      ) %>%
      distinct(patient_id) %>%
      pull(patient_id)
    
    n_ctdna_before_cyst <- length(patients_ctdna_before_cyst)
  } else {
    n_ctdna_before_cyst <- 0
  }
} else {
  n_ctdna_before_cyst <- 0
}

# Question 4: Long-term outcomes of persistently low-level detectable ctDNA
# Universe: Patients who underwent cystectomy and had persistently low-level detectable ctDNA

# Get baseline ctDNA
baseline_ctdna <- ctdna_long %>%
  group_by(patient_id) %>%
  arrange(ctdna_date) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  select(patient_id, baseline_value = ctdna_value, baseline_date = ctdna_date)

# Get survival outcomes for follow-up duration
if (exists("survival_data")) {
  survival_outcomes <- dat_combined %>%
    select(patient_id) %>%
    distinct() %>%
    left_join(survival_data, by = "patient_id") %>%
    mutate(
      os_time = ifelse(!is.na(explicit_death_date) | (alive_yes_no == "No"), 
                      as.numeric(survival_date - baseline_date) / 30.44, 
                      NA_real_)
    )
  
  if (exists("cystectomy_patients") && length(cystectomy_patients) > 0) {
    cyst_fu <- survival_outcomes %>%
      filter(patient_id %in% cystectomy_patients) %>%
      mutate(
        fu_duration_months = ifelse(!is.na(os_time), os_time, NA_real_)
      ) %>%
      select(patient_id, fu_duration_months)
    
    median_baseline <- median(baseline_ctdna$baseline_value[baseline_ctdna$baseline_value > 0], na.rm = TRUE)
    median_fu <- median(cyst_fu$fu_duration_months, na.rm = TRUE)
    
    if (is.finite(median_baseline) && is.finite(median_fu) && median_fu > 0) {
      persistent_low <- ctdna_long %>%
        left_join(baseline_ctdna, by = "patient_id") %>%
        left_join(cyst_fu, by = "patient_id") %>%
        filter(
          patient_id %in% cystectomy_patients,
          !is.na(baseline_value),
          baseline_value > 0,
          baseline_value < median_baseline,
          !is.na(fu_duration_months),
          fu_duration_months > median_fu
        ) %>%
        group_by(patient_id) %>%
        summarise(
          n_positive = sum(ctdna_value > 0, na.rm = TRUE),
          n_measurements = n(),
          .groups = "drop"
        ) %>%
        filter(
          n_positive >= 2,
          n_measurements >= 3
        ) %>%
        pull(patient_id)
      
      n_persistent_low <- length(persistent_low)
      persistent_low_definition <- paste0(
        "Baseline ctDNA > 0 and < ", round(median_baseline, 2), 
        " (median); Follow-up > ", round(median_fu, 1), " months (median); ",
        "≥2 positive measurements; ≥3 total measurements"
      )
    } else {
      persistent_low <- character(0)
      n_persistent_low <- 0
      persistent_low_definition <- "Criteria not met"
    }
  } else {
    persistent_low <- character(0)
    n_persistent_low <- 0
    persistent_low_definition <- "No cystectomy patients"
  }
} else {
  persistent_low <- character(0)
  n_persistent_low <- 0
  persistent_low_definition <- "No survival data"
}

# Create summary table
questions_summary <- tibble(
  Question = c(
    "Question 1: Association of on-treatment clinical events and pre-treatment factors on ctDNA profiles",
    "Question 2: Case series of patients with undetectable ctDNA following TURBT alone",
    "Question 3: ctDNA thresholds/kinetics for organ-confined vs metastatic recurrence",
    "Question 4: Long-term outcomes of persistently low-level detectable ctDNA"
  ),
  Universe = c(
    "Patients who underwent cystectomy",
    "Patients who underwent TURBT alone and had undetectable ctDNA afterwards without cancer directed therapy",
    "Patients who underwent cystectomy and had ctDNA levels available beforehand",
    "Patients who underwent cystectomy and had persistently low-level detectable ctDNA over extended follow-up"
  ),
  `N` = c(n_cystectomy, n_turbt_undetectable, n_ctdna_before_cyst, n_persistent_low),
  `Patient IDs` = c(
    NA_character_,
    if (length(turbt_undetectable) > 0) paste(sort(turbt_undetectable), collapse = ", ") else "None",
    NA_character_,
    if (length(persistent_low) > 0) paste(sort(persistent_low), collapse = ", ") else "None"
  ),
  `Definition` = c(
    NA_character_,
    NA_character_,
    NA_character_,
    persistent_low_definition
  )
)

questions_summary %>%
  kable(caption = "Research Questions - Universe Calculations", 
        col.names = c("Question", "Universe", "N", "Patient IDs", "Definition")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r research_questions_swimmer_plots, fig.width=16, fig.height=12, out.width="100%"}
# ==============================
# Research Questions Swimmer Plots
# ==============================

# Helper function to create swimmer plot
create_research_question_swimmer <- function(patient_ids, title) {
  if (length(patient_ids) == 0) return(NULL)
  
  patient_ids_char <- as.character(patient_ids)
  patients_data <- dat_combined %>%
    mutate(patient_id = as.character(patient_id)) %>%
    filter(patient_id %in% patient_ids_char)
  
  if (nrow(patients_data) == 0) return(NULL)
  
  ctdna_long_char <- ctdna_long %>%
    mutate(patient_id = as.character(patient_id))
  
  first_ctdna_dates <- ctdna_long_char %>%
    filter(patient_id %in% patient_ids_char) %>%
    group_by(patient_id) %>%
    summarise(first_ctdna_date = min(ctdna_date, na.rm = TRUE), .groups = "drop")
  
  therapy_start_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("systemic.*therapy.*\\d+.*start.*date", ignore_case = TRUE))]
  therapy_setting_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("systemic.*therapy.*\\d+.*treatment.*setting", ignore_case = TRUE))]
  therapy_start_long <- if (length(therapy_start_cols) > 0) {
    dat_combined %>%
      mutate(patient_id = as.character(patient_id)) %>%
      filter(patient_id %in% patient_ids_char) %>%
      select(patient_id, all_of(therapy_start_cols)) %>%
      pivot_longer(-patient_id, names_to = "var", values_to = "therapy_start_raw") %>%
      mutate(
        therapy_start = as_date_safe(therapy_start_raw),
        therapy_idx = as.integer(str_extract(var, "\\d+"))
      ) %>%
      filter(!is.na(therapy_start))
  } else {
    tibble(patient_id = character(0), therapy_start = as.Date(character(0)), therapy_idx = integer(0))
  }

  first_therapy_dates <- therapy_start_long %>%
    group_by(patient_id) %>%
    summarise(first_therapy_date = min(therapy_start, na.rm = TRUE), .groups = "drop")

  # Identify first palliative therapy start (fallback marker for metastatic relapse)
  palliative_dates <- if (nrow(therapy_start_long) > 0 && length(therapy_setting_cols) > 0) {
    settings_long <- dat_combined %>%
      mutate(patient_id = as.character(patient_id)) %>%
      filter(patient_id %in% patient_ids_char) %>%
      select(patient_id, all_of(therapy_setting_cols)) %>%
      pivot_longer(-patient_id, names_to = "var", values_to = "setting_raw") %>%
      mutate(
        therapy_idx = as.integer(str_extract(var, "\\d+")),
        setting_clean = clean_setting(setting_raw)
      )
    therapy_start_long %>%
      inner_join(settings_long, by = c("patient_id", "therapy_idx")) %>%
      filter(setting_clean == "Palliative") %>%
      group_by(patient_id) %>%
      summarise(palliative_start = min(therapy_start, na.rm = TRUE), .groups = "drop")
  } else {
    tibble(patient_id = character(0), palliative_start = as.Date(character(0)))
  }

  # Metastatic relapse dates (direct), otherwise fallback to first palliative start
  relapse_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("metastatic.*relapse.*date|relapse.*date|metastasis.*date", ignore_case = TRUE))]
  relapse_dates <- if (length(relapse_cols) > 0) {
    dat_combined %>%
      mutate(patient_id = as.character(patient_id)) %>%
      filter(patient_id %in% patient_ids_char) %>%
      select(patient_id, all_of(relapse_cols)) %>%
      pivot_longer(-patient_id, names_to = "relapse_var", values_to = "relapse_date_raw") %>%
      mutate(relapse_date = as_date_safe(relapse_date_raw)) %>%
      filter(!is.na(relapse_date)) %>%
      group_by(patient_id) %>%
      summarise(relapse_date = min(relapse_date, na.rm = TRUE), .groups = "drop")
  } else {
    tibble(patient_id = character(0), relapse_date = as.Date(character(0)))
  }

  if (nrow(relapse_dates) == 0 && nrow(palliative_dates) > 0) {
    relapse_dates <- palliative_dates %>%
      rename(relapse_date = palliative_start) %>%
      filter(!is.na(relapse_date))
  }
  
  follow_up_cols <- names(dat_combined)[str_detect(names(dat_combined), regex("last.*follow.*up.*date|follow.*up.*date", ignore_case = TRUE)) & 
                                        !str_detect(names(dat_combined), regex("death", ignore_case = TRUE))]
  follow_up_dates <- if (length(follow_up_cols) > 0) {
    dat_combined %>%
      mutate(patient_id = as.character(patient_id)) %>%
      filter(patient_id %in% patient_ids_char) %>%
      select(patient_id, all_of(follow_up_cols[1])) %>%
      rename(follow_up_date_raw = all_of(follow_up_cols[1])) %>%
      mutate(follow_up_date = as_date_safe(follow_up_date_raw)) %>%
      filter(!is.na(follow_up_date)) %>%
      select(patient_id, follow_up_date)
  } else {
    tibble(patient_id = character(0), follow_up_date = as.Date(character(0)))
  }
  
  patient_timeline <- patients_data %>%
    select(patient_id) %>%
    left_join(first_ctdna_dates, by = "patient_id") %>%
    left_join(first_therapy_dates, by = "patient_id") %>%
    left_join(follow_up_dates, by = "patient_id") %>%
    rowwise() %>%
    mutate(
      first_known_event = min(c(first_ctdna_date, first_therapy_date, follow_up_date), na.rm = TRUE),
      time_anchor = coalesce(follow_up_date, first_ctdna_date, first_therapy_date),
      first_known_event = ifelse(is.na(first_known_event) | is.infinite(first_known_event), 
                                time_anchor, first_known_event),
      follow_up_interval_days = if (!is.na(time_anchor) && !is.na(first_known_event)) {
        as.numeric(time_anchor - first_known_event)
      } else {
        NA_real_
      },
      follow_up_interval_months = follow_up_interval_days / 30.44
    ) %>%
    ungroup() %>%
    filter(!is.na(time_anchor))
  
  if (nrow(patient_timeline) == 0) return(NULL)
  
  plot_ctdna <- ctdna_long_char %>%
    filter(patient_id %in% patient_timeline$patient_id) %>%
    inner_join(patient_timeline %>% select(patient_id, time_anchor, first_known_event), by = "patient_id") %>%
    mutate(
      months_from_anchor = as.numeric(ctdna_date - first_known_event) / 30.44,
      ctdna_detectable = ctdna_value > 0
    )

  plot_relapse <- relapse_dates %>%
    inner_join(patient_timeline %>% select(patient_id, first_known_event), by = "patient_id") %>%
    mutate(relapse_months = as.numeric(relapse_date - first_known_event) / 30.44) %>%
    filter(!is.na(relapse_months))
  
  patient_order <- patient_timeline %>%
    arrange(patient_id) %>%
    pull(patient_id)
  
  patient_timeline <- patient_timeline %>%
    mutate(patient_id_ordered = factor(patient_id, levels = patient_order))
  
  plot_ctdna <- plot_ctdna %>%
    mutate(patient_id_ordered = factor(patient_id, levels = patient_order))
  
  x_max <- ceiling(max(c(patient_timeline$follow_up_interval_months, plot_ctdna$months_from_anchor), na.rm = TRUE)) + 3
  x_breaks <- seq(0, x_max, by = 6)
  
  p <- ggplot() +
    geom_segment(
      data = patient_timeline,
      aes(x = 0, xend = follow_up_interval_months,
          y = patient_id_ordered, yend = patient_id_ordered),
      linewidth = 3, color = "gray85", alpha = 0.6
    ) +
    geom_point(
      data = plot_ctdna %>% filter(ctdna_detectable),
      aes(x = months_from_anchor, y = patient_id_ordered),
      size = 2, shape = 21, fill = "red", color = "white", stroke = 0.3
    ) +
    geom_point(
      data = plot_ctdna %>% filter(!ctdna_detectable),
      aes(x = months_from_anchor, y = patient_id_ordered),
      size = 2, shape = 21, fill = "green3", color = "white", stroke = 0.3
    ) +
    {if (nrow(plot_relapse) > 0) {
      geom_point(
        data = plot_relapse %>% mutate(patient_id_ordered = factor(patient_id, levels = patient_order)),
        aes(x = relapse_months, y = patient_id_ordered),
        size = 4.5, shape = 23, fill = "red3", color = "darkred", stroke = 1.1
      )
    }} +
    geom_vline(xintercept = 0, linetype = "dashed", color = "black", linewidth = 0.7) +
    scale_y_discrete(limits = levels(patient_timeline$patient_id_ordered)) +
    labs(
      title = title,
      x = "Time from First Known Event (months)",
      y = "Patient ID"
    ) +
    theme_minimal(base_size = 10) +
    theme(
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
      axis.text.y = element_text(size = 7),
      panel.grid.major.y = element_blank()
    ) +
    scale_x_continuous(breaks = x_breaks, limits = c(-3, x_max), expand = c(0.02, 0))
  
  return(p)
}

# Generate plots
plots_list <- list()

if (exists("turbt_undetectable") && length(turbt_undetectable) > 0) {
  p_anna <- create_research_question_swimmer(
    as.character(turbt_undetectable),
    paste0("Question 2: TURBT Alone with Undetectable ctDNA (n = ", length(turbt_undetectable), ")")
  )
  if (!is.null(p_anna)) plots_list[["Question2"]] <- p_anna
}

if (exists("patients_ctdna_before_cyst") && length(patients_ctdna_before_cyst) > 0) {
  p_lydia <- create_research_question_swimmer(
    as.character(patients_ctdna_before_cyst),
    paste0("Question 3: Cystectomy with ctDNA Before (n = ", length(patients_ctdna_before_cyst), ")")
  )
  if (!is.null(p_lydia)) plots_list[["Question3"]] <- p_lydia
}

if (exists("persistent_low") && length(persistent_low) > 0) {
  p_thomas <- create_research_question_swimmer(
    as.character(persistent_low),
    paste0("Question 4: Persistently Low-Level Detectable ctDNA (n = ", length(persistent_low), ")")
  )
  if (!is.null(p_thomas)) plots_list[["Question4"]] <- p_thomas
}

if (length(plots_list) > 0) {
  combined_plots <- wrap_plots(plots_list, ncol = 1) +
    plot_annotation(
      title = "Research Question Swimmer Plots",
      subtitle = "ctDNA trajectories and follow-up for selected patient subsets"
    )
  combined_plots
} else {
  cat("No swimmer plots available (no eligible patients found for any question).\n")
}

```

```{r question4_landmark_12mo_mfs, fig.width=10, fig.height=7, out.width="100%"}
# ==============================
# Question 4: 12-month landmark MFS by ctDNA pattern
# Groups: Persistently Undetectable vs Persistently Low Detectable vs Detectable/Other
# ==============================

if (exists("cystectomy_patients") && length(cystectomy_patients) > 0 && nrow(ctdna_long) > 0 && nrow(survival_data) > 0) {
  landmark_days <- 365
  
  # Cystectomy dates (reuse if already built)
  if (!exists("cystectomy_dates")) {
    if (length(surgery_date_cols) > 0) {
      cystectomy_dates <- dat_combined %>%
        filter(patient_id %in% cystectomy_patients) %>%
        select(patient_id, all_of(surgery_date_cols)) %>%
        pivot_longer(-patient_id, names_to = "var", values_to = "surgery_date_raw") %>%
        mutate(surgery_date = as_date_safe(surgery_date_raw)) %>%
        filter(!is.na(surgery_date)) %>%
        group_by(patient_id) %>%
        summarise(cystectomy_date = min(surgery_date, na.rm = TRUE), .groups = "drop")
    } else {
      cystectomy_dates <- tibble(patient_id = character(0), cystectomy_date = as.Date(character(0)))
    }
  }
  
  # ctDNA within 12 months post-cystectomy
  ctdna_12mo <- ctdna_long %>%
    mutate(patient_id = as.character(patient_id)) %>%
    inner_join(cystectomy_dates, by = "patient_id") %>%
    filter(!is.na(cystectomy_date),
           !is.na(ctdna_value),
           ctdna_date >= cystectomy_date,
           ctdna_date <= cystectomy_date + days(landmark_days))
  
  pos_thresh <- median(ctdna_12mo$ctdna_value[ctdna_12mo$ctdna_value > 0], na.rm = TRUE)
  if (!is.finite(pos_thresh) || is.na(pos_thresh)) pos_thresh <- 0.02
  
  ctdna_groups_12mo <- ctdna_12mo %>%
    group_by(patient_id) %>%
    summarise(
      n_meas = n(),
      n_pos = sum(ctdna_value > 0, na.rm = TRUE),
      max_val = max(ctdna_value, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      group_12mo = case_when(
        n_meas > 0 && n_pos == 0 ~ "Persistently Undetectable",
        n_meas >= 2 && n_pos >= 2 && max_val <= pos_thresh ~ "Persistently Low Detectable",
        n_meas > 0 ~ "Detectable/Other",
        TRUE ~ "No ctDNA in 12m"
      )
    )
  
  # Survival inputs: relapse/metastasis dates, death, follow-up
  relapse_cols <- names(survival_data)[str_detect(names(survival_data), regex("metastatic.*relapse.*date|relapse.*date|metastasis.*date|progression.*date", ignore_case = TRUE))]
  death_cols <- names(survival_data)[str_detect(names(survival_data), regex("death_date|explicit_death_date", ignore_case = TRUE))]
  followup_cols <- names(survival_data)[str_detect(names(survival_data), regex("last.*follow.*up.*date|follow.*up.*date|survival_date", ignore_case = TRUE))]
  
  survival_inputs <- survival_data %>%
    mutate(patient_id = as.character(patient_id))
  
  survival_inputs <- survival_inputs %>%
    mutate(
      relapse_any = if (length(relapse_cols) > 0) dplyr::coalesce(!!!syms(relapse_cols)) else as.Date(NA),
      death_date_any = if (length(death_cols) > 0) dplyr::coalesce(!!!syms(death_cols)) else as.Date(NA),
      follow_up_any = if (length(followup_cols) > 0) dplyr::coalesce(!!!syms(followup_cols)) else as.Date(NA)
    ) %>%
    select(patient_id, relapse_any, death_date_any, follow_up_any)
  
  # Landmark cohort: require follow-up beyond 12 months from cystectomy and at least one ctDNA measure or censor info
  mfs_landmark <- cystectomy_dates %>%
    inner_join(survival_inputs, by = "patient_id") %>%
    left_join(ctdna_groups_12mo, by = "patient_id") %>%
    mutate(
      group_12mo = ifelse(is.na(group_12mo), "No ctDNA in 12m", group_12mo),
      landmark_date = cystectomy_date + days(landmark_days),
      censor_date = follow_up_any,
      event_date = dplyr::coalesce(relapse_any, death_date_any, follow_up_any),
      # Exclude if event before landmark (fail landmark eligibility)
      event_before_landmark = (!is.na(relapse_any) & relapse_any <= landmark_date) |
                              (!is.na(death_date_any) & death_date_any <= landmark_date),
      eligible = !event_before_landmark & !is.na(censor_date) & censor_date > landmark_date
    ) %>%
    filter(eligible) %>%
    mutate(
      mfs_time = as.numeric(event_date - landmark_date) / 30.44,
      mfs_event = ifelse(!is.na(relapse_any) | !is.na(death_date_any), 1, 0)
    ) %>%
    filter(!is.na(mfs_time), mfs_time >= 0)
  
  if (nrow(mfs_landmark) > 0 && length(unique(mfs_landmark$group_12mo)) > 1) {
    mfs_landmark <- mfs_landmark %>%
      mutate(group_12mo = factor(group_12mo,
                                 levels = c("Persistently Undetectable", "Persistently Low Detectable", "Detectable/Other", "No ctDNA in 12m")))
    
    surv_obj <- Surv(mfs_landmark$mfs_time, mfs_landmark$mfs_event)
    fit <- survfit(surv_obj ~ group_12mo, data = mfs_landmark)
    pval <- surv_pvalue(fit)$pval
    
    plot_mfs <- ggsurvplot(
      fit,
      data = mfs_landmark,
      conf.int = TRUE,
      risk.table = TRUE,
      pval = paste0("Log-rank p = ", format.pval(pval, digits = 3, eps = 0.001)),
      legend.title = "ctDNA pattern @12m",
      legend.labs = levels(mfs_landmark$group_12mo),
      xlab = "Months since 12m landmark",
      ylab = "Metastasis-free survival probability",
      palette = c("#1b9e77", "#d95f02", "#7570b3", "#999999")
    )
    
    # Cox model (reference: Persistently Undetectable)
    cox_fit <- coxph(surv_obj ~ group_12mo, data = mfs_landmark)
    cox_tbl <- broom::tidy(cox_fit, exponentiate = TRUE, conf.int = TRUE) %>%
      mutate(term = str_replace(term, "group_12mo", "")) %>%
      select(Term = term, HR = estimate, `95% CI Lower` = conf.low, `95% CI Upper` = conf.high, `p-value` = p.value) %>%
      mutate(across(c(HR, `95% CI Lower`, `95% CI Upper`), ~round(., 2)),
             `p-value` = ifelse(`p-value` < 0.001, "<0.001", round(`p-value`, 3)))
    
    print(plot_mfs$plot)
    
    cox_tbl %>%
      kable(caption = "12-month Landmark Cox Model: ctDNA pattern vs MFS",
            align = "lcccc") %>%
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
  } else {
    cat("Insufficient eligible patients to plot 12-month landmark MFS.\n")
  }
} else {
  cat("Insufficient data to perform 12-month landmark MFS analysis for Question 4.\n")
}
```
