---
title: "StemAI: Epigenetic Combination Co-Scientist (HMA+BCL2 in Stem-like Cancers)"
subtitle: "Parallel pipeline to HMA_BCL2_Expanded — Master + children"
author: "Alex Karol"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
always_allow_html: true
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    df_print: paged
    fig_caption: true
    self_contained: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 5,
  fig.show = "asis",
  dev = "png",
  dpi = 150,
  cache = TRUE,
  cache.path = if (requireNamespace("here", quietly = TRUE)) here::here("_cache", "StemAI_") else "_cache/StemAI_"
)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
if (!requireNamespace("here", quietly = TRUE)) install.packages("here")
if (requireNamespace("here", quietly = TRUE)) knitr::opts_knit$set(root.dir = here::here())
source(here::here("scripts", "stemai_config.R"), local = TRUE)
source(here::here("scripts", "stemai_helpers.R"), local = TRUE)
gdsc_dir <- if (dir.exists("data")) "data" else file.path("..", "data")
```

# Human input

**Query scope (fixed for this build):** `r STEMAI_QUERY_SCOPE`

Epigenetic-based combination therapy inquiry for stem-like cancers: identify biomarker-coordinated vulnerabilities and rational HMA + BCL2i (and related) combinations using open-source pharmacogenomics, with TCGA and literature as unseen validation layers.

## How to run

| Environment | Command |
|-------------|---------|
| **Cloud agent (recommended)** | `Rscript scripts/render_stemai_cloud.R` |
| Local | `Rscript scripts/render_stemai_cloud.R` or knit this file |

**Prerequisites:** `Rscript scripts/load_depmap_cache.R`; optional `Rscript scripts/download_tcga_prad.R` and `Rscript scripts/download_tcga_laml.R`. For extended stem-like cohort (DLBCL, neuroblastoma): knit `Cell_Line_Assignment_Verification_TNBC_Lymphoma.Rmd` first or use its RDS.

```{r workflow-schematic, fig.cap="StemAI workflow (PowerPoint-aligned). Inputs → integration → engineered features → mapping → ranked outputs + evidence."}
if (requireNamespace("DiagrammeR", quietly = TRUE)) {
  DiagrammeR::grViz("
digraph stemai {
  graph [layout = dot, rankdir = LR, bgcolor = white]
  node [shape = box, style = 'rounded,filled', fontname = Helvetica]
  human [label = 'Human input\\nEpigenetic combination inquiry', fillcolor = '#E8F4FF']
  subgraph cluster_in {
    label = 'Open-source inputs'
    depmap [label = 'DepMap\\nRNA + CRISPR + PRISM', fillcolor = '#DBEAFE']
    gdsc [label = 'GDSC\\nDose-response', fillcolor = '#DCFCE7']
    ctrp [label = 'CTRPv2\\nDose-response', fillcolor = '#EDE9FE']
    gse [label = 'GSE68379\\nMethylation', fillcolor = '#FEE2E2']
    tcga [label = 'TCGA\\nPRAD + LAML', fillcolor = '#FEF3C7']
  }
  integrate [label = 'Cross-platform integration\\nHarmonization + drug-class aggregation', fillcolor = '#FFF7ED']
  features [label = 'Engineered features\\nSensitivity strata, concordance, composite score', fillcolor = '#ECFDF5', style = 'rounded,dashed']
  map [label = 'Mechanistic vulnerability mapping', fillcolor = '#E0F2FE']
  out [label = 'Final output\\nRanked targets + combinations\\nPubMed evidence', fillcolor = '#D1FAE5']
  human -> depmap
  depmap -> integrate
  gdsc -> integrate
  ctrp -> integrate
  gse -> integrate
  tcga -> integrate
  integrate -> features
  features -> map
  map -> out
}
")
} else {
  svg_path <- here::here("outputs", "simple_hma_bcl2_workflow_schematic.svg")
  if (file.exists(svg_path)) {
    knitr::include_graphics(normalizePath(svg_path, winslash = "/"))
  } else {
    knitr::asis_output("Install DiagrammeR for inline workflow schematic, or add outputs/simple_hma_bcl2_workflow_schematic.svg.")
  }
}
```

```{r child-01, cache = FALSE, child = 'rmarkdown/StemAI_01_Data_Loading.Rmd'}
```

```{r child-02, cache = FALSE, child = 'rmarkdown/StemAI_02_CrossPlatform_Integration.Rmd'}
```

```{r child-03, cache = FALSE, child = 'rmarkdown/StemAI_03_Feature_Engineering.Rmd'}
```

```{r child-04, cache = FALSE, child = 'rmarkdown/StemAI_04_Mechanistic_Mapping.Rmd'}
```

```{r child-05, cache = FALSE, child = 'rmarkdown/StemAI_05_Evidence_Packages.Rmd'}
```

```{r child-06, cache = FALSE, child = 'rmarkdown/StemAI_06_Ranked_Outputs_and_Validation.Rmd'}
```

# Session info

```{r session-info}
knitr::kable(
  tibble::tibble(Item = c("Query scope", "Workspace RDS", "Features RDS", "Evidence RDS"),
                 Value = c(STEMAI_QUERY_SCOPE, STEMAI_WORKSPACE_RDS, STEMAI_FEATURES_RDS, STEMAI_EVIDENCE_RDS)),
  caption = "StemAI artifact paths."
)
if (requireNamespace("sessioninfo", quietly = TRUE)) {
  print(sessioninfo::session_info())
}
```
