---
title: "UTUC vs Bladder Cancer ctDNA Shedding Patterns and ctDNA vs Imaging for Relapse Detection"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "")
knitr::opts_knit$set(root.dir = "~/R")
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Load shared setup
source("ctDNA_Research_Questions_Shared_Setup.R")

# Load required libraries
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(stringr)
  library(readxl)
  library(janitor)
  library(lubridate)
  library(ggplot2)
  library(purrr)
  library(knitr)
  library(kableExtra)
  library(pROC)
})
```

```{r load_data, include=FALSE}
# Load minimal data
data_list <- load_minimal_data(include_survival = TRUE)
dat_combined <- data_list$dat_combined
ctdna_long <- data_list$ctdna_long
```

## Research Questions

1. **What are the differences in ctDNA shedding patterns between UTUC (upper tract urothelial carcinoma) and bladder cancer (UC)?**
2. **What are the sensitivity and NPV differences for ctDNA compared to imaging for relapse detection?**

## Methods

### Tumor Location Classification

```{r classify_tumor_location, include=FALSE}
# Classify patients as UTUC vs Bladder Cancer
# Look for primary tumor location columns
location_cols <- names(dat_combined)[str_detect(names(dat_combined), 
  regex("primary.*tumor.*location|tumor.*location|primary.*site", ignore_case = TRUE))]

cat("**Tumor Location Columns Found:**\n")
if (length(location_cols) > 0) {
  cat(paste(location_cols, collapse = ", "), "\n\n")
} else {
  cat("No tumor location columns found. Checking alternative column names...\n\n")
}

# Create tumor location classification
dat_combined <- dat_combined %>%
  mutate(
    # Try to identify tumor location from available columns
    primary_location_raw = if (length(location_cols) > 0) {
      as.character(.data[[location_cols[1]]])
    } else {
      NA_character_
    },
    primary_location_clean = tolower(str_trim(primary_location_raw)),
    
    # Classify as UTUC vs Bladder
    tumor_type = case_when(
      str_detect(primary_location_clean, "renal|pelvis|ureter|upper.*tract|utuc|upper.*urinary") ~ "UTUC",
      str_detect(primary_location_clean, "bladder|lower.*tract") ~ "Bladder",
      is.na(primary_location_clean) | primary_location_clean == "" ~ "Unknown",
      TRUE ~ "Other"
    )
  )

# Summary of classification
cat("**Tumor Type Classification:**\n")
tumor_type_summary <- dat_combined %>%
  count(tumor_type) %>%
  mutate(percentage = round(n / nrow(dat_combined) * 100, 1))

print(tumor_type_summary)
cat("\n")
```

### ctDNA Shedding Pattern Analysis

```{r ctdna_shedding_analysis, include=FALSE}
# Compare ctDNA levels and positivity rates between UTUC and Bladder
# For each patient, calculate:
# - First ctDNA value
# - Maximum ctDNA value
# - Mean ctDNA value
# - Proportion of positive ctDNA samples
# - Number of ctDNA samples

ctdna_summary_by_patient <- ctdna_long %>%
  group_by(patient_id) %>%
  summarise(
    n_samples = n(),
    first_ctdna_value = first(ctdna_value),
    max_ctdna_value = max(ctdna_value, na.rm = TRUE),
    mean_ctdna_value = mean(ctdna_value, na.rm = TRUE),
    median_ctdna_value = median(ctdna_value, na.rm = TRUE),
    n_positive = sum(ctdna_value > 0, na.rm = TRUE),
    prop_positive = n_positive / n_samples,
    ever_positive = any(ctdna_value > 0, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(
    dat_combined %>% select(patient_id, tumor_type),
    by = "patient_id"
  ) %>%
  filter(tumor_type %in% c("UTUC", "Bladder"))

# Summary statistics by tumor type
ctdna_comparison <- ctdna_summary_by_patient %>%
  group_by(tumor_type) %>%
  summarise(
    n_patients = n(),
    n_with_ctdna = sum(!is.na(first_ctdna_value)),
    median_first_ctdna = median(first_ctdna_value, na.rm = TRUE),
    iqr_first_ctdna_low = quantile(first_ctdna_value, 0.25, na.rm = TRUE),
    iqr_first_ctdna_high = quantile(first_ctdna_value, 0.75, na.rm = TRUE),
    median_max_ctdna = median(max_ctdna_value, na.rm = TRUE),
    iqr_max_ctdna_low = quantile(max_ctdna_value, 0.25, na.rm = TRUE),
    iqr_max_ctdna_high = quantile(max_ctdna_value, 0.75, na.rm = TRUE),
    median_mean_ctdna = median(mean_ctdna_value, na.rm = TRUE),
    prop_ever_positive = mean(ever_positive, na.rm = TRUE),
    mean_prop_positive = mean(prop_positive, na.rm = TRUE),
    median_n_samples = median(n_samples, na.rm = TRUE),
    .groups = "drop"
  )
```

## Results

### Question 1: ctDNA Shedding Patterns - UTUC vs Bladder Cancer

```{r display_shedding_comparison}
cat("### ctDNA Shedding Patterns: UTUC vs Bladder Cancer\n\n")

# Display comparison table
comparison_table <- ctdna_comparison %>%
  mutate(
    `First ctDNA (median, IQR)` = paste0(
      round(median_first_ctdna, 3), " (",
      round(iqr_first_ctdna_low, 3), "-",
      round(iqr_first_ctdna_high, 3), ")"
    ),
    `Max ctDNA (median, IQR)` = paste0(
      round(median_max_ctdna, 3), " (",
      round(iqr_max_ctdna_low, 3), "-",
      round(iqr_max_ctdna_high, 3), ")"
    ),
    `Mean ctDNA (median)` = round(median_mean_ctdna, 3),
    `% Ever Positive` = paste0(round(prop_ever_positive * 100, 1), "%"),
    `Mean % Positive Samples` = paste0(round(mean_prop_positive * 100, 1), "%"),
    `Median # Samples` = round(median_n_samples, 1)
  ) %>%
  select(
    `Tumor Type` = tumor_type,
    `N Patients` = n_patients,
    `N with ctDNA` = n_with_ctdna,
    `First ctDNA (median, IQR)`,
    `Max ctDNA (median, IQR)`,
    `Mean ctDNA (median)`,
    `% Ever Positive`,
    `Mean % Positive Samples`,
    `Median # Samples`
  ) %>%
  kable(caption = "Comparison of ctDNA Shedding Patterns: UTUC vs Bladder Cancer") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)

comparison_table

# Statistical tests
if (nrow(ctdna_summary_by_patient %>% filter(tumor_type == "UTUC")) > 5 &&
    nrow(ctdna_summary_by_patient %>% filter(tumor_type == "Bladder")) > 5) {
  
  # Wilcoxon rank-sum test for first ctDNA
  test_first <- wilcox.test(
    first_ctdna_value ~ tumor_type,
    data = ctdna_summary_by_patient,
    alternative = "two.sided"
  )
  
  # Wilcoxon rank-sum test for max ctDNA
  test_max <- wilcox.test(
    max_ctdna_value ~ tumor_type,
    data = ctdna_summary_by_patient,
    alternative = "two.sided"
  )
  
  # Chi-square test for ever positive
  test_positive <- chisq.test(
    table(ctdna_summary_by_patient$tumor_type, ctdna_summary_by_patient$ever_positive)
  )
  
  cat("\n**Statistical Tests:**\n\n")
  cat("- **First ctDNA value:** Wilcoxon rank-sum p-value =", 
      format.pval(test_first$p.value, digits = 3), "\n")
  cat("- **Max ctDNA value:** Wilcoxon rank-sum p-value =", 
      format.pval(test_max$p.value, digits = 3), "\n")
  cat("- **Ever positive:** Chi-square p-value =", 
      format.pval(test_positive$p.value, digits = 3), "\n\n")
}
```

```{r ctdna_distribution_plot, fig.width=12, fig.height=8}
# Box plots comparing ctDNA distributions
if (nrow(ctdna_summary_by_patient) > 0) {
  p1 <- ctdna_summary_by_patient %>%
    filter(!is.na(tumor_type)) %>%
    ggplot(aes(x = tumor_type, y = log10(first_ctdna_value + 0.001), fill = tumor_type)) +
    geom_boxplot(alpha = 0.7) +
    geom_jitter(width = 0.2, alpha = 0.3) +
    labs(
      x = "Tumor Type",
      y = "Log10(First ctDNA + 0.001)",
      title = "Distribution of First ctDNA Value: UTUC vs Bladder"
    ) +
    theme_minimal() +
    theme(legend.position = "none")
  
  p2 <- ctdna_summary_by_patient %>%
    filter(!is.na(tumor_type)) %>%
    ggplot(aes(x = tumor_type, y = log10(max_ctdna_value + 0.001), fill = tumor_type)) +
    geom_boxplot(alpha = 0.7) +
    geom_jitter(width = 0.2, alpha = 0.3) +
    labs(
      x = "Tumor Type",
      y = "Log10(Max ctDNA + 0.001)",
      title = "Distribution of Maximum ctDNA Value: UTUC vs Bladder"
    ) +
    theme_minimal() +
    theme(legend.position = "none")
  
  p3 <- ctdna_summary_by_patient %>%
    filter(!is.na(tumor_type)) %>%
    ggplot(aes(x = tumor_type, fill = factor(ever_positive))) +
    geom_bar(position = "fill", alpha = 0.7) +
    scale_fill_manual(
      values = c("FALSE" = "blue", "TRUE" = "red"),
      labels = c("FALSE" = "Never Positive", "TRUE" = "Ever Positive"),
      name = "ctDNA Status"
    ) +
    labs(
      x = "Tumor Type",
      y = "Proportion",
      title = "Proportion of Patients with Ever-Positive ctDNA: UTUC vs Bladder"
    ) +
    theme_minimal()
  
  # Combine plots
  library(gridExtra)
  grid.arrange(p1, p2, p3, ncol = 2)
}
```

### Question 2: ctDNA vs Imaging for Relapse Detection

```{r imaging_comparison_setup, include=FALSE}
# Identify imaging columns
imaging_cols <- names(dat_combined)[str_detect(names(dat_combined), 
  regex("obvious.*disease.*imaging|disease.*present.*imaging|imaging.*disease|relapse.*imaging|recurrence.*imaging", 
        ignore_case = TRUE))]

cat("**Imaging Columns Found:**\n")
if (length(imaging_cols) > 0) {
  cat(paste(imaging_cols, collapse = ", "), "\n\n")
} else {
  cat("No imaging columns found. Checking alternative column names...\n\n")
}

# Identify relapse/recurrence columns
relapse_cols <- names(dat_combined)[str_detect(names(dat_combined),
  regex("relapse|recurrence|metastatic.*disease|distant.*metastasis", ignore_case = TRUE))]

cat("**Relapse/Recurrence Columns Found:**\n")
if (length(relapse_cols) > 0) {
  cat(paste(head(relapse_cols, 5), collapse = ", "), "\n\n")
} else {
  cat("No relapse columns found.\n\n")
}

# Create imaging status variable
dat_combined <- dat_combined %>%
  mutate(
    imaging_disease_present = if (length(imaging_cols) > 0) {
      case_when(
        str_detect(tolower(as.character(.data[[imaging_cols[1]]])), "yes|y|1|positive") ~ "Yes",
        str_detect(tolower(as.character(.data[[imaging_cols[1]]])), "no|n|0|negative") ~ "No",
        TRUE ~ NA_character_
      )
    } else {
      NA_character_
    },
    imaging_positive = case_when(
      imaging_disease_present == "Yes" ~ 1,
      imaging_disease_present == "No" ~ 0,
      TRUE ~ NA_real_
    )
  )

# Get relapse status
dat_combined <- dat_combined %>%
  mutate(
    metastatic_ever = if ("metastatic_disease_ever_diagnosed_yes_no" %in% names(dat_combined)) {
      case_when(
        str_detect(tolower(as.character(metastatic_disease_ever_diagnosed_yes_no)), "yes|y|1") ~ 1,
        str_detect(tolower(as.character(metastatic_disease_ever_diagnosed_yes_no)), "no|n|0") ~ 0,
        TRUE ~ NA_real_
      )
    } else {
      NA_real_
    }
  )
```

```{r ctdna_vs_imaging_analysis, include=FALSE}
# For ctDNA vs imaging comparison, we need to:
# 1. Define a time window (e.g., 12 months from T0)
# 2. Determine ctDNA status within that window
# 3. Determine imaging status within that window
# 4. Calculate sensitivity, specificity, PPV, NPV

# Get first ctDNA date per patient (T0)
first_ctdna_dates <- ctdna_long %>%
  arrange(patient_id, ctdna_date) %>%
  group_by(patient_id) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  select(patient_id, t0_date = ctdna_date)

# Get ctDNA status within 12 months of T0
ctdna_within_12mo <- ctdna_long %>%
  left_join(first_ctdna_dates, by = "patient_id") %>%
  filter(
    !is.na(t0_date),
    !is.na(ctdna_date),
    as.numeric(ctdna_date - t0_date) <= 365,  # Within 12 months
    as.numeric(ctdna_date - t0_date) >= 0     # After T0
  ) %>%
  group_by(patient_id) %>%
  summarise(
    ctdna_ever_positive_12mo = any(ctdna_value > 0, na.rm = TRUE),
    ctdna_always_negative_12mo = all(ctdna_value == 0 | is.na(ctdna_value), na.rm = TRUE),
    latest_ctdna_date_12mo = max(ctdna_date, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    ctdna_status_12mo = case_when(
      ctdna_ever_positive_12mo ~ "Positive",
      ctdna_always_negative_12mo ~ "Negative",
      TRUE ~ "Unknown"
    )
  )

# For imaging, we'll use the imaging_disease_present field
# We need to check if imaging was done within 6 months of latest ctDNA (as per ctDNA_Power.Rmd definition)
imaging_comparison_data <- dat_combined %>%
  select(patient_id, imaging_positive, metastatic_ever, tumor_type) %>%
  left_join(ctdna_within_12mo, by = "patient_id") %>%
  filter(
    !is.na(ctdna_status_12mo),
    !is.na(imaging_positive) | !is.na(metastatic_ever)
  ) %>%
  mutate(
    # Use imaging if available, otherwise use metastatic_ever as proxy
    imaging_relapse = case_when(
      !is.na(imaging_positive) ~ imaging_positive,
      !is.na(metastatic_ever) ~ metastatic_ever,
      TRUE ~ NA_real_
    ),
    ctdna_positive = case_when(
      ctdna_status_12mo == "Positive" ~ 1,
      ctdna_status_12mo == "Negative" ~ 0,
      TRUE ~ NA_real_
    )
  ) %>%
  filter(!is.na(imaging_relapse), !is.na(ctdna_positive))

# Calculate diagnostic measures
if (nrow(imaging_comparison_data) > 10) {
  # Create 2x2 table
  diagnostic_table <- table(
    Imaging = imaging_comparison_data$imaging_relapse,
    ctDNA = imaging_comparison_data$ctdna_positive
  )
  
  # Calculate metrics
  # Sensitivity = P(ctDNA+ | Imaging+)
  # Specificity = P(ctDNA- | Imaging-)
  # PPV = P(Imaging+ | ctDNA+)
  # NPV = P(Imaging- | ctDNA-)
  
  if (nrow(diagnostic_table) == 2 && ncol(diagnostic_table) == 2) {
    TP <- diagnostic_table[2, 2]  # Imaging+, ctDNA+
    TN <- diagnostic_table[1, 1]  # Imaging-, ctDNA-
    FP <- diagnostic_table[1, 2]  # Imaging-, ctDNA+
    FN <- diagnostic_table[2, 1]  # Imaging+, ctDNA-
    
    sensitivity <- TP / (TP + FN)
    specificity <- TN / (TN + FP)
    ppv <- TP / (TP + FP)
    npv <- TN / (TN + FN)
    
    # Calculate 95% CI using Wilson score interval
    calc_ci <- function(p, n) {
      z <- 1.96
      denominator <- 1 + (z^2 / n)
      center <- (p + (z^2 / (2 * n))) / denominator
      margin <- (z / denominator) * sqrt((p * (1 - p) / n) + (z^2 / (4 * n^2)))
      c(max(0, center - margin), min(1, center + margin))
    }
    
    sens_ci <- calc_ci(sensitivity, TP + FN)
    spec_ci <- calc_ci(specificity, TN + FP)
    ppv_ci <- calc_ci(ppv, TP + FP)
    npv_ci <- calc_ci(npv, TN + FN)
    
    diagnostic_metrics <- tibble(
      Metric = c("Sensitivity", "Specificity", "PPV", "NPV"),
      Value = c(sensitivity, specificity, ppv, npv),
      `95% CI Lower` = c(sens_ci[1], spec_ci[1], ppv_ci[1], npv_ci[1]),
      `95% CI Upper` = c(sens_ci[2], spec_ci[2], ppv_ci[2], npv_ci[2]),
      Definition = c(
        "P(ctDNA+ | Imaging+)",
        "P(ctDNA- | Imaging-)",
        "P(Imaging+ | ctDNA+)",
        "P(Imaging- | ctDNA-)"
      )
    )
  } else {
    diagnostic_metrics <- NULL
  }
} else {
  diagnostic_metrics <- NULL
  diagnostic_table <- NULL
}
```

```{r display_imaging_comparison}
cat("### ctDNA vs Imaging for Relapse Detection\n\n")

if (!is.null(diagnostic_table) && nrow(imaging_comparison_data) > 10) {
  cat("**2x2 Contingency Table:**\n\n")
  print(diagnostic_table)
  cat("\n")
  
  cat("**N =", nrow(imaging_comparison_data), "patients with both ctDNA and imaging data**\n\n")
  
  if (!is.null(diagnostic_metrics)) {
    diagnostic_display <- diagnostic_metrics %>%
      mutate(
        `Value (95% CI)` = paste0(
          round(Value * 100, 1), "% (",
          round(`95% CI Lower` * 100, 1), "-",
          round(`95% CI Upper` * 100, 1), "%)"
        )
      ) %>%
      select(Metric, `Value (95% CI)`, Definition) %>%
      kable(caption = "Diagnostic Performance of ctDNA vs Imaging for Relapse Detection") %>%
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
    
    diagnostic_display
    
    cat("\n**Key Findings:**\n\n")
    cat("- **Sensitivity:**", round(sensitivity * 100, 1), "% of patients with imaging-detected relapse had positive ctDNA within 12 months\n")
    cat("- **NPV:**", round(npv * 100, 1), "% of patients with negative ctDNA did not have imaging-detected relapse\n")
    cat("- **Specificity:**", round(specificity * 100, 1), "% of patients without imaging-detected relapse had negative ctDNA\n")
    cat("- **PPV:**", round(ppv * 100, 1), "% of patients with positive ctDNA had imaging-detected relapse\n\n")
  }
} else {
  cat("**Insufficient data for ctDNA vs imaging comparison.**\n\n")
  cat("**N =", nrow(imaging_comparison_data), "patients with both ctDNA and imaging data**\n\n")
  cat("**Note:** Need at least 10 patients with both ctDNA and imaging data for reliable comparison.\n\n")
}
```

```{r stratified_by_tumor_type}
# Stratify by tumor type if sample size allows
if (nrow(imaging_comparison_data) > 20) {
  imaging_by_tumor_type <- imaging_comparison_data %>%
    filter(tumor_type %in% c("UTUC", "Bladder")) %>%
    group_by(tumor_type) %>%
    summarise(
      n = n(),
      TP = sum(imaging_relapse == 1 & ctdna_positive == 1, na.rm = TRUE),
      TN = sum(imaging_relapse == 0 & ctdna_positive == 0, na.rm = TRUE),
      FP = sum(imaging_relapse == 0 & ctdna_positive == 1, na.rm = TRUE),
      FN = sum(imaging_relapse == 1 & ctdna_positive == 0, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      Sensitivity = ifelse(TP + FN > 0, TP / (TP + FN), NA_real_),
      Specificity = ifelse(TN + FP > 0, TN / (TN + FP), NA_real_),
      PPV = ifelse(TP + FP > 0, TP / (TP + FP), NA_real_),
      NPV = ifelse(TN + FN > 0, TN / (TN + FN), NA_real_)
    )
  
  if (nrow(imaging_by_tumor_type) > 0) {
    cat("### ctDNA vs Imaging: Stratified by Tumor Type\n\n")
    
    stratified_table <- imaging_by_tumor_type %>%
      mutate(
        `Sensitivity (95% CI)` = ifelse(!is.na(Sensitivity),
          paste0(round(Sensitivity * 100, 1), "%"),
          "N/A"
        ),
        `NPV (95% CI)` = ifelse(!is.na(NPV),
          paste0(round(NPV * 100, 1), "%"),
          "N/A"
        ),
        `Specificity (95% CI)` = ifelse(!is.na(Specificity),
          paste0(round(Specificity * 100, 1), "%"),
          "N/A"
        ),
        `PPV (95% CI)` = ifelse(!is.na(PPV),
          paste0(round(PPV * 100, 1), "%"),
          "N/A"
        )
      ) %>%
      select(
        `Tumor Type` = tumor_type,
        N = n,
        `Sensitivity` = `Sensitivity (95% CI)`,
        `NPV` = `NPV (95% CI)`,
        `Specificity` = `Specificity (95% CI)`,
        `PPV` = `PPV (95% CI)`
      ) %>%
      kable(caption = "Diagnostic Performance of ctDNA vs Imaging, Stratified by Tumor Type") %>%
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
    
    stratified_table
  }
}
```

## Summary

```{r summary}
cat("## Summary\n\n")

cat("### Question 1: ctDNA Shedding Patterns - UTUC vs Bladder Cancer\n\n")
if (nrow(ctdna_comparison) >= 2) {
  cat("**Key Findings:**\n\n")
  for (i in 1:nrow(ctdna_comparison)) {
    cat("- **", ctdna_comparison$tumor_type[i], ":**\n", sep = "")
    cat("  - Median first ctDNA:", round(ctdna_comparison$median_first_ctdna[i], 3), "\n")
    cat("  - Median max ctDNA:", round(ctdna_comparison$median_max_ctdna[i], 3), "\n")
    cat("  - % ever positive:", round(ctdna_comparison$prop_ever_positive[i] * 100, 1), "%\n\n")
  }
}

cat("### Question 2: ctDNA vs Imaging for Relapse Detection\n\n")
if (!is.null(diagnostic_metrics) && nrow(diagnostic_metrics) > 0) {
  cat("**Key Findings:**\n\n")
  cat("- **Sensitivity:**", round(sensitivity * 100, 1), "% - ctDNA detects", 
      round(sensitivity * 100, 1), "% of imaging-detected relapses\n")
  cat("- **NPV:**", round(npv * 100, 1), "% -", round(npv * 100, 1), 
      "% of patients with negative ctDNA do not have imaging-detected relapse\n")
  cat("- **Specificity:**", round(specificity * 100, 1), "%\n")
  cat("- **PPV:**", round(ppv * 100, 1), "%\n\n")
  
  cat("**Clinical Interpretation:**\n\n")
  if (sensitivity >= 0.80) {
    cat("- ctDNA has **high sensitivity** for detecting relapse compared to imaging\n")
  } else {
    cat("- ctDNA has **moderate sensitivity** for detecting relapse compared to imaging\n")
  }
  
  if (npv >= 0.90) {
    cat("- ctDNA has **excellent NPV** - negative ctDNA strongly predicts absence of imaging-detected relapse\n")
  } else if (npv >= 0.80) {
    cat("- ctDNA has **good NPV** - negative ctDNA predicts absence of imaging-detected relapse\n")
  } else {
    cat("- ctDNA has **moderate NPV** - negative ctDNA may not reliably exclude relapse\n")
  }
}
```

